### 2012-02-01 PetersonGedeHubbard2012.tex
Changed to Springer journal format.

\documentclass{svjour3}                     % onecolumn (standard format)
\RequirePackage{fix-cm}

\usepackage{graphicx}
\usepackage{amsmath}
\begin{document}

\title{Open source symbolic multibody dynamics software}

\titlerunning{Open source symbolic multibody dynamics software}        % if too long for running head

\author{Dale L. Peterson\and Gilbert Gede\and Mont Hubbard}

\institute{Dale L. Peterson, Gilbert Gede, Mont Hubbard \at
           Sports Biomechanics Laboratory\\
           Department of Mechanical and Aerospace Engineering\\
           University of California Davis\\
           Davis, CA 95616-5294\\
           Tel.: +1 530 752 2163\\
           \email{\{dlpeterson,ggede,mhubbard\}@ucdavis.edu}}

\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}
  Abstract here.
\keywords{symbolic dynamics software, nonholonomic systems, control}
\end{abstract}

\section{Introduction}
\label{intro}

\section{Derivation of linearization algorithm}
\label{derivations}
\input{linearization}

\begin{acknowledgements}
  This material is based upon work partially supported by the National Science
  Foundation under award 0928339 and two Google Summer of Code projects (2009,
  2011).  Aaron Muerer, Brian Granger, Mateuz Paprocki, and Ondrej Certik
  provided valuable help with SymPy and Python.  Jason Moore, Thomas Johnston,
  Evan Sperber, and Andrew Kickertz provided valuable feedback during
  discussions of multibody dynamics and control related issues.
\end{acknowledgements}

%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliography{}   % name your BibTeX data base
\end{document}

### 2012-02-01 PetersonGedeHubbard2012.tex
Added .bst files, not sure which is actually needed.
First start on linearization section.

@@
 \usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{bm}
+
+%% Big-O notation.
+\providecommand{\OO}[1]{\ensuremath{\operatorname{O}\bigl(#1\bigr)}}
+
 \begin{document}
@@
 \label{derivations}
-\input{linearization}
+The goal of the following calculations is to concretely establish the first
+order relationships (about an arbitrary equilibrium point) between an arbitary
+choice of independent coordinates and speeds, and the time-derivatives of all
+coordinates and speeds (both independent and dependent).  This enables several
+common tasks: (1) stability analysis of arbitrary equilibria, (2) traditional
+control system design, (3) control system design with output feedback based
+upon quantities that are dependent (i.e., not state variables), (4)
+determination of measurement equations for use in Kalman filter design when
+measured quantities are at the acceleration level (i.e., accelerometers).  The
+formalism presented here is generic enough to cover most examples of
+time-varying constrained multibody systems with arbitrary external inputs and
+arbitrary specified quantities.
+
+The complete configuration of the system is assumed to be described by
+generalized coordinates $\bm{q}\in\bm{R}^n$, its velocity completely described
+by generalized speeds $\bm{u}\in\bm{R}^o$.  It is assumed that there are $l$
+holonomic constraints ($f_h : \mathbf{R}^n \times \mathbf{R} \to
+\mathbf{R}^l$), and $m$ nonholonomic constraints ($f_{nh} : \mathbf{R}^n \times
+\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^l$). Two functions $\bm{f}_0 :
+\mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R} \to \mathbf{R}^n$ and
+$\bm{f}_1 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R} \to
+\mathbf{R}^n$ relate the coordinate time derivatives to the generalized speeds,
+and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
+\mathbf{R}^p \times \mathbf{R} \to \mathbf{R}^n$ and $\bm{f}_3 : \mathbf{R}^n
+\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
+\to \mathbf{R}^n$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
+the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
+their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
+assumptions can be summarized by the vector equations
+\begin{align}
+  \label{eq:holonomic}
+  \bm{f}_{h}(\bm{q}, t) &= \bm{0}\\
+  \label{eq:nonholonomic}
+  \bm{f}_{nh}(\bm{q}, \bm{u}, t) &= \bm{0}\\
+  \label{eq:kindiffs}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
+    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
+  \label{eq:dyndiffs}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, \bm{r}, t)
+    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
+\end{align}
+where (\ref{eq:holonomic}) is the holonomic constraints,
+(\ref{eq:nonholonomic}) is the nonholonomic and time-differentiated holonomic
+constraints, (\ref{eq:kindiffs}) is the kinematic differential equations
+(linear in both $\bm{\dot{q}}$ and $\bm{u}$ and in the trivial case,
+$\bm{\dot{q}} = \bm{u}$, so $f_0 = -f_1$), and (\ref{eq:dyndiffs}) include both
+the dynamic differential equations (arising from application of Kane's method)
+and the time-differentiated version of equation \ref{eq:nonholonomic}.
+Equation (\ref{eq:dyndiffs}) will necessarily by linear in $\bm{\dot{u}}$, but
+generally nonlinear in all other quantities (though often linear in the
+external inputs $\bm{r}$).  In this formalism, both independent and dependent
+quantities (coordinates and speeds) appear in these equations; we will address
+this is shortly.
+
+The system described by equations \ref{eq:holonomic}-\ref{eq:dyndiffs} has
+$o-m$ velocity degrees of freedom.  In state space form, the trajectories of
+the system exist in a $p \triangleq n - l + o - m$ dimensional space, though
+conserved quantities (e.g., total mechanical energy) may restrict trajectories
+to a lower dimensional manifold embedded in this $p$ dimensional space.  In
+general this cannot be assumed, however, especially in the presence of external
+inputs which may change the total mechanical energy of the system.  While more
+than $p$ of the state variables may be of interest (i.e., dependent coordinates
+or speeds may be of interest in themselves), only $p$ quantities are truly
+independent.  Which of these $p$ coordinates and speeds should be chosen as
+independent out of the possible $n + o$ may be obvious for small systems where
+there is intuition about the system behavior, but in general the optimal choice
+depends on how the configuration $\bm{q}$ affects the gradients of equations
+\ref{eq:holonomic} and \ref{eq:nonholonomic}.  On the other hand, in systems
+with cyclic (ignorable) coordinates, fewer than $p$ state variables may be of
+interest; one example is the Whipple bicycle model where $p=10$ but 5 of these
+state variables are cyclic coordinates (coordinates of rear wheel contact
+point, heading, and both wheel angles) and are often not of much interest in
+themselves, nor are they neccessary for most analyses.
+
+We begin with a Taylor series expansion of equations
+\ref{eq:holonomic}-\ref{eq:dyndiffs} about an equilibrium $\bm{q}=\bm{q}^*$,
+$\bm{\dot{q}}=\bm{\dot{q}}^*$, $\bm{u}=\bm{u}^*$,
+$\bm{\dot{u}}=\bm{\dot{u}}^*$.  It assumed that $\bm{q}^*, \bm{\dot{q}}^*,
+\bm{u}^*, \bm{\dot{u}}^*$ satisify equations
+\ref{eq:holonomic}-\ref{eq:dyndiffs}.  The Taylor series expansion of the
+constraint equations (\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) about the
+equilibrium is
+\begin{align}
+  \label{eq:holonomic_expansion}
+  \bm{f}_{h}(\bm{q}, t) &\approx \underbrace{f_{h}(\bm{q}^*, t)}_{\bm{0}} +
+    \nabla_{\bm{q}}f_{h}(\bm{q}^*, t)\delta \bm{q}\\
+  \label{eq:nonholonomic_expansion}
+  \bm{f}_{nh}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{nh}(\bm{q}^*,
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{nh}(\bm{q}^*, \bm{u}^*, t)\delta \bm{q} + \nabla_{\bm{u}}f_{nh}(\bm{q}^*, \bm{u}^*, t)\delta \bm{u}
+\end{align}
+The
 
@@
 
+% Not sure which bibliography style we are supposed to use
+%\bibliographystyle{spbasic}      % basic style, author-year citations
 %\bibliographystyle{spmpsci}      % mathematics and physical sciences
+%\bibliographystyle{spphys}       % APS-like style for physics
 %\bibliography{}   % name your BibTeX data base

### 2012-02-02 PetersonGedeHubbard2012.tex
Finished unconstrained state equations, nearly done with permutation calculations.

@@
 and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
-\mathbf{R}^p \times \mathbf{R} \to \mathbf{R}^n$ and $\bm{f}_3 : \mathbf{R}^n
+\mathbf{R} \to \mathbf{R}^o$ and $\bm{f}_3 : \mathbf{R}^n
 \times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
-\to \mathbf{R}^n$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
+\to \mathbf{R}^o$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
 the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
@@
   \label{eq:dyndiffs}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, \bm{r}, t)
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
     + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
 \end{align}
-where (\ref{eq:holonomic}) is the holonomic constraints,
-(\ref{eq:nonholonomic}) is the nonholonomic and time-differentiated holonomic
-constraints, (\ref{eq:kindiffs}) is the kinematic differential equations
+where equations (\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) are the
+holonomic and nonholonomic constraints (and possibly time differentiated holonomic
+constraints), respectively. Equations (\ref{eq:kindiffs}) are the kinematic differential equations
 (linear in both $\bm{\dot{q}}$ and $\bm{u}$ and in the trivial case,
-$\bm{\dot{q}} = \bm{u}$, so $f_0 = -f_1$), and (\ref{eq:dyndiffs}) include both
-the dynamic differential equations (arising from application of Kane's method)
-and the time-differentiated version of equation \ref{eq:nonholonomic}.
+$\bm{\dot{q}} = \bm{u}$, so $f_0 = -f_1$).  Equations (\ref{eq:dyndiffs}) are
+the dynamic differential equations
+and the time differentiated version of equation \ref{eq:nonholonomic}.
 Equation (\ref{eq:dyndiffs}) will necessarily by linear in $\bm{\dot{u}}$, but
@@
 quantities (coordinates and speeds) appear in these equations; we will address
-this is shortly.
+this is shortly.  This following calculations formalism is exceedingly generic;
+in practice there is a lot of structure in equations
+(\ref{eq:nonholonomic}-\ref{eq:dyndiffs}) that can be leveraged to simplify the
+following calculations.  This will be expanded upon below.
 
@@
 
-We begin with a Taylor series expansion of equations
-\ref{eq:holonomic}-\ref{eq:dyndiffs} about an equilibrium $\bm{q}=\bm{q}^*$,
-$\bm{\dot{q}}=\bm{\dot{q}}^*$, $\bm{u}=\bm{u}^*$,
-$\bm{\dot{u}}=\bm{\dot{u}}^*$.  It assumed that $\bm{q}^*, \bm{\dot{q}}^*,
-\bm{u}^*, \bm{\dot{u}}^*$ satisify equations
-\ref{eq:holonomic}-\ref{eq:dyndiffs}.  The Taylor series expansion of the
-constraint equations (\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) about the
-equilibrium is
+We begin with a first order Taylor series expansion of equations
+(\ref{eq:holonomic})-(\ref{eq:dyndiffs}) about an equilibrium
+$\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$, $\bm{u}=\bm{u}^*$,
+$\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$ that satisfies equations
+(\ref{eq:holonomic})-(\ref{eq:dyndiffs}).  Determining these equilibrium values
+generally involves solving (\ref{eq:holonomic})-(\ref{eq:dyndiffs}).  In the interest of brevity, we omit writing this equilibrium after
+each gradient in the calculations below; all are evaluated at these equilibrium
+conditions.  Taylor series expansion of the constraint equations
+(\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) about the equilibrium is
 \begin{align}
@@
   \bm{f}_{h}(\bm{q}, t) &\approx \underbrace{f_{h}(\bm{q}^*, t)}_{\bm{0}} +
-    \nabla_{\bm{q}}f_{h}(\bm{q}^*, t)\delta \bm{q}\\
+    \nabla_{\bm{q}}f_{h} \cdot \delta \bm{q}\\
   \label{eq:nonholonomic_expansion}
   \bm{f}_{nh}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{nh}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{nh}(\bm{q}^*, \bm{u}^*, t)\delta \bm{q} + \nabla_{\bm{u}}f_{nh}(\bm{q}^*, \bm{u}^*, t)\delta \bm{u}
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{nh} \cdot \delta \bm{q} +
+  \nabla_{\bm{u}}f_{nh} \cdot \delta \bm{u}
 \end{align}
-The
+and the first terms are identically zero because of the assumption that the
+equilibrium point satisfies the constraints.  The Taylor series expansions of
+the functions appearing in the left hand side of equations (\ref{eq:kindiffs})
+are
+\begin{align}
+  \label{eq:f0_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
+  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \cdot \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \cdot \delta\bm{\dot{q}}\\
+  \label{eq:f1_expansion}
+  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
+  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \cdot \delta\bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{1} \cdot \delta\bm{u}
+\end{align}
+Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
+that the sum of the first term on the right hand side of each equation must
+equal zero, we obtain
+\begin{align}
+  \label{eq:f0_plus_f1_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
+  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \cdot \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \cdot \delta\bm{\dot{q}} +
+  \nabla_{\bm{u}}\bm{f}_{1} \cdot \delta\bm{u}
+\end{align}
+Similarly, a Taylor series expansion of the functions appearing in
+the left hand side of (\ref{eq:dyndiffs}) yields
+\begin{align}
+  \label{eq:f2_expansion}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
+      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
+      \nabla_{\bm{q}}\bm{f}_{2} \cdot \delta\bm{q}
+      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \cdot \delta\bm{\dot{u}}\\
+  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
+  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
+  \nabla_{\bm{q}}\bm{f}_{3} \cdot \delta\bm{q}\notag\\
+  \label{eq:f3_expansion}
+  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \cdot \delta\bm{\dot{q}}
+  + \nabla_{\bm{u}}\bm{f}_{3} \cdot \delta \bm{u}
+  + \nabla_{\bm{r}}\bm{f}_{3} \cdot \delta\bm{r}
+\end{align}
+Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
+that the sum of the first term on the right hand sides of these equations must
+equal zero, we obtain
+\begin{align}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3) \cdot
+  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \cdot \delta\bm{\dot{q}}\notag\\
+  \label{eq:f2_plus_f3_expansion}
+  &+ \nabla_{\bm{u}}\bm{f}_{3} \cdot \delta\bm{u} +
+  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \cdot \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \cdot \delta\bm{r}
+\end{align}
+
+Utilizing equations (\ref{eq:kindiffs}), (\ref{eq:f0_plus_f1_expansion}),
+(\ref{eq:dyndiffs}), and (\ref{eq:f2_plus_f3_expansion}), we introduce the
+following quantities
+\begin{align}
+  M_{00} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
+  M_{11} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
+  A_{00} &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+  A_{01} &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
+  A_{10} &\triangleq \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
+  -\nabla_{\bm{\dot{q}}}\bm{f}_3 (\nabla_{\bm{\dot{q}}}\bm{f}_{0})^{-1}
+  \nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+  A_{11} &\triangleq \nabla_{\bm{u}} \bm{f}_{3} - \nabla_{\bm{\dot{q}}}\bm{f}_{3}
+  (\nabla_{\bm{\dot{q}}}\bm{f}_0)^{-1} \nabla_{\bm{u}}\bm{f}_{1} \\
+  B_{1} &\triangleq \nabla_{\bm{r}}\bm{f}_{3}
+\end{align}
+
+Equating the right hand side of equations (\ref{eq:f0_plus_f1_expansion}) and
+(\ref{eq:f2_plus_f3_expansion}) each to zero, eliminating $\bm{\dot{q}}$ from
+(\ref{eq:f2_plus_f3_expansion}), and rewriting in matrix form, we obtain
+\begin{align}
+  \label{eq:state_space_unconstrained}
+  \left[
+    \begin{array}{cc}
+      M_{00} & \bm{0}_{n \times o} \\
+      \bm{0}_{o \times n} & M_{11}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       A_{00} & A_{01} \\
+       A_{10} & A_{11}
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q} \\
+        \delta \bm{u}
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{n \times q} \\
+        B_{1}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
++ o$, yet as was discussed earlier, only $p = n - l + o - m$ of these
+quantities are independent.  To address this issue, one must pick a particular
+set of independent coordinates and speeds and appeal to equations
+(\ref{eq:holonomic_expansion}) and (\ref{eq:nonholonomic_expansion}).  To this
+end, consider the following partitioning of the generalized coordinates and
+generalized speeds:
+\begin{equation*}
+  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
+      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
+      \qquad\qquad
+  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
+      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
+\end{equation*}
+where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
+are invertible permutation matrices which map an ordering which has the
+independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
+\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities
+($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the
+original ordering of the coordinates and speeds.
+
+Setting the right hand side of equation (\ref{eq:holonomic_expansion}) equal to
+zero, and making use of $P_q$, we have
+\begin{align}
+  \nabla_{\bm{q}}\bm{f}_{h} P_{q} \delta \bm{\tilde{q}} &=
+  \nabla_{\bm{q}}\bm{f}_{h} P_{qi} \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{h} P_{qd} \delta \bm{q_d}\notag\\
+  &= \bm{0}\notag\\
+  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{h} P_{qd})^{-1}
+  (\nabla_{\bm{q}}\bm{f}_{h} P_{qi}) \delta \bm{q}_i
+\end{align}
 

### 2012-02-07 PetersonGedeHubbard2012.tex
Added a bit about permutation matrices.

@@
 Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet as was discussed earlier, only $p = n - l + o - m$ of these
-quantities are independent.  To address this issue, one must pick a particular
-set of independent coordinates and speeds and appeal to equations
-(\ref{eq:holonomic_expansion}) and (\ref{eq:nonholonomic_expansion}).  To this
-end, consider the following partitioning of the generalized coordinates and
-generalized speeds:
++ o$, yet  only $p = n - l + o - m$ of these quantities are independent, due to
+the constraints (equations (\ref{eq:holonomic} and (\ref{eq:nonholonomic})).
+To address this issue, one must pick a particular set of independent
+coordinates and speeds and appeal to equations (\ref{eq:holonomic_expansion})
+and (\ref{eq:nonholonomic_expansion}).  To this end, consider the following
+partitioning of the generalized coordinates and generalized speeds:
 \begin{equation*}
@@
 ($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the
-original ordering of the coordinates and speeds.
+original ordering of the coordinates and speeds.  We use the notation $P_{qi}$
+and $P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$,
+respectively; similarly, $P_{ui}$ is the first $o-m$ colums of $P_{u}$ while
+$P_{ud}$ is the last $m$ columns of $P_u$.
 
@@
 \begin{align}
-  \nabla_{\bm{q}}\bm{f}_{h} P_{q} \delta \bm{\tilde{q}} &=
-  \nabla_{\bm{q}}\bm{f}_{h} P_{qi} \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{h} P_{qd} \delta \bm{q_d}\notag\\
+  \nabla_{\bm{q}}\bm{f}_{h} P_{q} \cdot \delta \bm{\tilde{q}} &=
+  \nabla_{\bm{q}}\bm{f}_{h} P_{qi} \cdot \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{h} P_{qd} \cdot \delta \bm{q_d}\notag\\
   &= \bm{0}\notag\\
   \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{h} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{h} P_{qi}) \delta \bm{q}_i
+  (\nabla_{\bm{q}}\bm{f}_{h} P_{qi}) \cdot \delta \bm{q}_i \notag\\
+  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{h} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{h} \right] P_{qi} \cdot \delta
+    \bm{q}_i
 \end{align}
+For convenience, we define
+\begin{equation}
+  \label{eq:C_0}
+  C_0 \triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{h} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{h} \right] P_{qi}
+\end{equation}
+which enables us to write
+\begin{align}
+  \label{eq:delta_q}
+  \delta \bm{q} &= C_0 \cdot \delta \bm{q}
+\end{align}
+
+Applying the same approach to (\ref{eq:nonholonomic_expansion}), we obtain
+\begin{align}
+\nabla_{\bm{q}}f_{nh} \cdot \delta \bm{q} +
+\nabla_{\bm{u}}f_{nh} \cdot \delta \bm{u} &= \nabla_{\bm{q}} \bm{f}_{nh} \cdot
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{nh} P_{ud} \cdot \delta \bm{u}_d \notag\\
+&= \bm{0}\notag\\
+\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{nh}
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{nh} \cdot \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i \right]\notag
+\end{align}
+We define
+\begin{align}
+  \label{eq:C_1}
+  C_1 &\triangleq -\left(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}\right)^{-1}
+  \nabla_{\bm{q}}\bm{f}_{nh} \\
+  \label{eq:C_2}
+  C_2 &\triangleq -\left(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}\right)^{-1} \nabla_{\bm{u}} \bm{f}_{nh} P_{ui}
+\end{align}
+which enables us to write
+\begin{align}
+  \label{eq:delta_u}
+  \delta \bm{u} &= C_1 \cdot \delta \bm{q} + C_2 \cdot \delta \bm{u}_i\notag\\
+  &= C_1 C_0 \cdot \delta \bm{q}_i + C_2 \cdot \delta \bm{u}_i
+\end{align}
+
+By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
+rewrite equation (\ref{eq:state_space_unconstrained}) as
+\begin{align}
+  \label{eq:state_space_constrained}
+  \left[
+    \begin{array}{cc}
+      M_{00} & \bm{0}_{n \times o} \\
+      \bm{0}_{o \times n} & M_{11}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       (A_{00} + A_{01} C_{1}) C_{0} & A_{01} C_{2} \\
+       (A_{10} + A_{11} C_{1}) C_{0} & A_{11} C_{2}
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q}_i \\
+        \delta \bm{u}_i
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{n \times q} \\
+        B_{1}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+
+There are two requirements for this method to work, namely, both
+$\nabla_{\bm{q}}\bm{f}_{h} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}$
+must be nonsingular.  Because of the presence of the permutation matrices, this
+implies proper selection of dependent coordinates and dependent speeds is
+critical.  The permutation matrices effectively select the columns of the
+constraint gradient matrices, and may be selected by examination of the
+constraint gradient matrix through a SVD or other numerical techniques; for
+small systems it may be possible to choose dependent speeds without resorting
+to numerical techniques if the system trajectories stay with the configuration
+space where these matrices are nonsingular.
+
+For stability analyses, one must examine the stability of the mapping $\delta
+\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
+$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
+the independent rows of these equations ($n-l+o-m$ of them) to form a square
+matrix whose eigenvalues govern the stability of the system about the point of
+linearization.
 

### 2012-02-10 PetersonGedeHubbard2012.tex
Fixed error in defintion of C1, C2, constrained state equations.
Made acceleration constraints distinct from dynamic equations.

@@
 The complete configuration of the system is assumed to be described by
-generalized coordinates $\bm{q}\in\bm{R}^n$, its velocity completely described
-by generalized speeds $\bm{u}\in\bm{R}^o$.  It is assumed that there are $l$
-holonomic constraints ($f_h : \mathbf{R}^n \times \mathbf{R} \to
-\mathbf{R}^l$), and $m$ nonholonomic constraints ($f_{nh} : \mathbf{R}^n \times
-\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^l$). Two functions $\bm{f}_0 :
-\mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R} \to \mathbf{R}^n$ and
-$\bm{f}_1 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R} \to
-\mathbf{R}^n$ relate the coordinate time derivatives to the generalized speeds,
-and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
-\mathbf{R} \to \mathbf{R}^o$ and $\bm{f}_3 : \mathbf{R}^n
-\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
-\to \mathbf{R}^o$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
-the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
-their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
-assumptions can be summarized by the vector equations
+generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
+described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
+All coordinates and speeds, not only the independent ones, are represented by
+$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
+relationships described in table \ref{table:assumptions}.  The system is
+described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
+quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
+independent.  This paper discusses the first order relationship between these
+independent quantities ($n-l$ independent coordinates, $o-m$ independent
+speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
+coordinates and $o$ speeds.
+%It is assumed that there are $l$
+%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
+%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
+%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
+%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
+%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
+%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
+%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
+%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
+%assumptions can be summarized by the following equations
+%\begin{align}
+%  \label{eq:configuration}
+%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
+%  \label{eq:velocity}
+%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
+%  \label{eq:acceleration}
+%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
+%  \label{eq:kindiffs}
+%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
+%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
+%  \label{eq:dyndiffs}
+%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
+%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
+%\end{align}
+\begin{table}[htbp]
+  \centering
+  \caption{Constrained multibody system governing equations}
+  \begin{tabular}[c]{r @{ $=$ } l l}
+    \multicolumn{2}{c}{Equation} & Mathematical definition \\
+    \hline
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}$ & $f_c : \mathbf{R}^n \times \mathbf{R} \to
+\mathbf{R}^l$ \\
+$\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}$ & $f_v : \mathbf{R}^n \times
+\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
+$\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}$ &
+$f_a : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R}^o \times
+\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
+$\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
+$\bm{0}$ & $\bm{f}_0 : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R} \to
+\mathbf{R}^n$\\
+\multicolumn{2}{c}{} & $\bm{f}_1 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}
+\to \mathbf{R}^n$ \\
+$\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t)$ &
+$\bm{0}$ & $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R} \to
+\mathbf{R}^{o-m}$,\\
+\multicolumn{2}{c}{} & $\bm{f}_3 : \mathbf{R}^n \times \mathbf{R}^n \times
+\mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R} \to \mathbf{R}^{o-m}$
+  \end{tabular}
+  \label{table:assumptions}
+\end{table}
+The first three equations in table \ref{table:assumptions} represent
+constraints derived purely from kinematic considerations at the configuration,
+velocity, and acceleration levels, respectively.  The velocity constraints may
+be nonholonomic or differentiated holonomic, while the acceleration constraints
+are assumed to be differentiated versions of the velocity constraints.  The
+fourth equation represents kinematic differential equations which relate
+$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
+The fifth equation represents the constraint free nonholonomic dynamic
+differential equations~\cite{Kane1985}, and will necessarily by linear in
+$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
+in all other quantities.  This formalism is exceedingly generic; in practice,
+the structure of the functions in table \ref{table:assumptions} can be
+leveraged to simplify the following calculations.  This will be discussed
+subsequently.
+
+%The system described by equations \ref{eq:configuration}-\ref{eq:dyndiffs} has
+%$o-m$ velocity degrees of freedom.
+Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
+manifold embedded in a $n + o$ dimensional space, though conserved quantities
+(e.g., total mechanical energy) may restrict trajectories to a lower
+dimensional manifold.  In general this cannot be assumed, however, especially
+in the presence of external control inputs $\bm{r}$ which may change the total
+mechanical energy of the system.  While more than $p$ of the independent state
+variables may be of interest (i.e., dependent coordinates or speeds may be of
+interest in themselves), only $p$ quantities are truly independent.  Which of
+these $p$ coordinates and speeds should be chosen as independent out of the
+possible $n + o$ may be obvious for small systems where there is intuition
+about the system behavior, but in general the optimal choice depends on how the
+configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
+hand, in systems with cyclic/ignorable coordinates (i.e., coordinates which do
+not appear in the dynamic equations of motion), fewer than $p$ state
+variables may be of interest; one example is the Whipple bicycle model where
+$p=10$ but 5 of these state variables are cyclic coordinates (coordinates of
+rear wheel contact point, heading, and both wheel angles) and are often not of
+much interest in themselves, nor are they necessary for most analyses.
+
+We begin with a first order Taylor series expansion of the equations in Table
+\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
+$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
+assumed that all of the equations in Table \ref{table:assumptions} are
+satisfied by these quantities.  In the interest of brevity, we omit writing
+this equilibrium after each gradient in the calculations below; all are
+evaluated at these equilibrium conditions.  Expansion of the three constraint
+equations yields
 \begin{align}
-  \label{eq:holonomic}
-  \bm{f}_{h}(\bm{q}, t) &= \bm{0}\\
-  \label{eq:nonholonomic}
-  \bm{f}_{nh}(\bm{q}, \bm{u}, t) &= \bm{0}\\
-  \label{eq:kindiffs}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
-    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
-  \label{eq:dyndiffs}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
-    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
+  \label{eq:configuration_expansion}
+  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{f_{c}(\bm{q}^*, t)}_{\bm{0}} +
+    \nabla_{\bm{q}}f_{c} \cdot \delta \bm{q}\\
+  \label{eq:velocity_expansion}
+  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{v} \cdot \delta \bm{q} +
+  \nabla_{\bm{u}}f_{v} \cdot \delta \bm{u} \\
+  \label{eq:acceleration_expansion}
+  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
+  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
+t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{a} \cdot \delta \bm{q} + \nabla_{\bm{\dot{q}}}f_{a}
+\cdot \delta \bm{\dot{q}} \notag\\
+&+ \nabla_{\bm{u}}f_{a} \cdot \delta \bm{u} + \nabla_{\bm{\dot{u}}}f_{a} \cdot
+\delta \bm{\dot{u}}
 \end{align}
-where equations (\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) are the
-holonomic and nonholonomic constraints (and possibly time differentiated holonomic
-constraints), respectively. Equations (\ref{eq:kindiffs}) are the kinematic differential equations
-(linear in both $\bm{\dot{q}}$ and $\bm{u}$ and in the trivial case,
-$\bm{\dot{q}} = \bm{u}$, so $f_0 = -f_1$).  Equations (\ref{eq:dyndiffs}) are
-the dynamic differential equations
-and the time differentiated version of equation \ref{eq:nonholonomic}.
-Equation (\ref{eq:dyndiffs}) will necessarily by linear in $\bm{\dot{u}}$, but
-generally nonlinear in all other quantities (though often linear in the
-external inputs $\bm{r}$).  In this formalism, both independent and dependent
-quantities (coordinates and speeds) appear in these equations; we will address
-this is shortly.  This following calculations formalism is exceedingly generic;
-in practice there is a lot of structure in equations
-(\ref{eq:nonholonomic}-\ref{eq:dyndiffs}) that can be leveraged to simplify the
-following calculations.  This will be expanded upon below.
-
-The system described by equations \ref{eq:holonomic}-\ref{eq:dyndiffs} has
-$o-m$ velocity degrees of freedom.  In state space form, the trajectories of
-the system exist in a $p \triangleq n - l + o - m$ dimensional space, though
-conserved quantities (e.g., total mechanical energy) may restrict trajectories
-to a lower dimensional manifold embedded in this $p$ dimensional space.  In
-general this cannot be assumed, however, especially in the presence of external
-inputs which may change the total mechanical energy of the system.  While more
-than $p$ of the state variables may be of interest (i.e., dependent coordinates
-or speeds may be of interest in themselves), only $p$ quantities are truly
-independent.  Which of these $p$ coordinates and speeds should be chosen as
-independent out of the possible $n + o$ may be obvious for small systems where
-there is intuition about the system behavior, but in general the optimal choice
-depends on how the configuration $\bm{q}$ affects the gradients of equations
-\ref{eq:holonomic} and \ref{eq:nonholonomic}.  On the other hand, in systems
-with cyclic (ignorable) coordinates, fewer than $p$ state variables may be of
-interest; one example is the Whipple bicycle model where $p=10$ but 5 of these
-state variables are cyclic coordinates (coordinates of rear wheel contact
-point, heading, and both wheel angles) and are often not of much interest in
-themselves, nor are they neccessary for most analyses.
-
-We begin with a first order Taylor series expansion of equations
-(\ref{eq:holonomic})-(\ref{eq:dyndiffs}) about an equilibrium
-$\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$, $\bm{u}=\bm{u}^*$,
-$\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$ that satisfies equations
-(\ref{eq:holonomic})-(\ref{eq:dyndiffs}).  Determining these equilibrium values
-generally involves solving (\ref{eq:holonomic})-(\ref{eq:dyndiffs}).  In the interest of brevity, we omit writing this equilibrium after
-each gradient in the calculations below; all are evaluated at these equilibrium
-conditions.  Taylor series expansion of the constraint equations
-(\ref{eq:holonomic}) and (\ref{eq:nonholonomic}) about the equilibrium is
-\begin{align}
-  \label{eq:holonomic_expansion}
-  \bm{f}_{h}(\bm{q}, t) &\approx \underbrace{f_{h}(\bm{q}^*, t)}_{\bm{0}} +
-    \nabla_{\bm{q}}f_{h} \cdot \delta \bm{q}\\
-  \label{eq:nonholonomic_expansion}
-  \bm{f}_{nh}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{nh}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{nh} \cdot \delta \bm{q} +
-  \nabla_{\bm{u}}f_{nh} \cdot \delta \bm{u}
-\end{align}
-and the first terms are identically zero because of the assumption that the
-equilibrium point satisfies the constraints.  The Taylor series expansions of
-the functions appearing in the left hand side of equations (\ref{eq:kindiffs})
-are
+The first terms are identically zero because of the assumption that the
+equilibrium point satisfies the constraints.  The Taylor series expansion of
+the kinematic differential equations is
 \begin{align}
@@
 \end{align}
-Similarly, a Taylor series expansion of the functions appearing in
-the left hand side of (\ref{eq:dyndiffs}) yields
+Similarly, a Taylor series expansion of the dynamic differential equations, we
+obtain
 \begin{align}
@@
 
-Utilizing equations (\ref{eq:kindiffs}), (\ref{eq:f0_plus_f1_expansion}),
-(\ref{eq:dyndiffs}), and (\ref{eq:f2_plus_f3_expansion}), we introduce the
-following quantities
+Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
+(\ref{eq:acceleration_expansion}),
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as per table
+\ref{table:assumptions}), and introducing the following definitions
 \begin{align}
-  M_{00} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
-  M_{11} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
-  A_{00} &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-  A_{01} &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
-  A_{10} &\triangleq \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
-  -\nabla_{\bm{\dot{q}}}\bm{f}_3 (\nabla_{\bm{\dot{q}}}\bm{f}_{0})^{-1}
-  \nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-  A_{11} &\triangleq \nabla_{\bm{u}} \bm{f}_{3} - \nabla_{\bm{\dot{q}}}\bm{f}_{3}
-  (\nabla_{\bm{\dot{q}}}\bm{f}_0)^{-1} \nabla_{\bm{u}}\bm{f}_{1} \\
-  B_{1} &\triangleq \nabla_{\bm{r}}\bm{f}_{3}
+\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
+\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a \\
+\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
+\tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
+\tilde{A}_{uqc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} \bm{f}_a\right] \\
+\tilde{A}_{uuc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_a\right] \\
+\tilde{A}_{uqd} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3)\right] \\
+\tilde{A}_{uud} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_u\right] \\
+\tilde{B}_{u}   &\triangleq \nabla_{\bm{r}}\bm{f}_{3}
 \end{align}
-
-Equating the right hand side of equations (\ref{eq:f0_plus_f1_expansion}) and
-(\ref{eq:f2_plus_f3_expansion}) each to zero, eliminating $\bm{\dot{q}}$ from
-(\ref{eq:f2_plus_f3_expansion}), and rewriting in matrix form, we obtain
+enables the unconstrained linear state space equations to be written as
 \begin{align}
@@
     \begin{array}{cc}
-      M_{00} & \bm{0}_{n \times o} \\
-      \bm{0}_{o \times n} & M_{11}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \bm{0}_{m \times n} & \tilde{M}_{uuc} \\
+      \bm{0}_{(o-m) \times n} & \tilde{M}_{uud}
     \end{array}
@@
      \begin{array}{cc}
-       A_{00} & A_{01} \\
-       A_{10} & A_{11}
+       \tilde{A}_{qq} & \tilde{A}_{qu} \\
+       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
+       \tilde{A}_{uqd} & \tilde{A}_{uud}
      \end{array}
@@
       \begin{array}{c}
-        \bm{0}_{n \times q} \\
-        B_{1}
+        \bm{0}_{(n + m) \times q} \\
+        \tilde{B}_{u}
       \end{array}
@@
 Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet  only $p = n - l + o - m$ of these quantities are independent, due to
-the constraints (equations (\ref{eq:holonomic} and (\ref{eq:nonholonomic})).
-To address this issue, one must pick a particular set of independent
-coordinates and speeds and appeal to equations (\ref{eq:holonomic_expansion})
-and (\ref{eq:nonholonomic_expansion}).  To this end, consider the following
++ o$, yet only $p = n - l + o - m$ of these quantities are independent.
+To address this issue, a particular set of independent
+coordinates and speeds must be selected. To this end, consider the following
 partitioning of the generalized coordinates and generalized speeds:
@@
 
-Setting the right hand side of equation (\ref{eq:holonomic_expansion}) equal to
+Setting the right hand side of equation (\ref{eq:configuration_expansion}) equal to
 zero, and making use of $P_q$, we have
@@
   \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \cdot \delta \bm{q}
+  \delta \bm{q} &= C_0 \cdot \delta \bm{q}_i
 \end{align}
 
-Applying the same approach to (\ref{eq:nonholonomic_expansion}), we obtain
+Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
 \begin{align}
@@
 P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{nh} \cdot \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i \right]\notag
+  \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i \right]\notag\\
+  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{nh} \cdot \delta \bm{q}\notag\\
+  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{nh} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{nh} \right] P_{ui} \cdot \delta \bm{u}_i
 \end{align}
@@
   \label{eq:C_1}
-  C_1 &\triangleq -\left(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}\right)^{-1}
-  \nabla_{\bm{q}}\bm{f}_{nh} \\
+  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{nh} \\
   \label{eq:C_2}
-  C_2 &\triangleq -\left(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}\right)^{-1} \nabla_{\bm{u}} \bm{f}_{nh} P_{ui}
+  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{nh} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{nh} \right] P_{ui}
 \end{align}
@@
      \begin{array}{cc}
-       (A_{00} + A_{01} C_{1}) C_{0} & A_{01} C_{2} \\
-       (A_{10} + A_{11} C_{1}) C_{0} & A_{11} C_{2}
+       (\tilde{A}_{qq} + A_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + A_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + A_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
      \end{array}
@@
       \begin{array}{c}
-        \bm{0}_{n \times q} \\
-        B_{1}
+        \bm{0}_{(n+m) \times q} \\
+        \tilde{B}_{u}
       \end{array}
@@
 % Not sure which bibliography style we are supposed to use
+%\bibliographystyle{spphys}       % APS-like style for physics
 %\bibliographystyle{spbasic}      % basic style, author-year citations
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliography{}   % name your BibTeX data base
+\bibliography{library}   % name your BibTeX data base
+\bibliographystyle{spmpsci}      % mathematics and physical sciences
 \end{document}

### 2012-02-13 PetersonGedeHubbard2012.tex
Fixed some typos, error in definition of B_u

@@
 The goal of the following calculations is to concretely establish the first
-order relationships (about an arbitrary equilibrium point) between an arbitary
+order relationships (about an arbitrary equilibrium point) between an arbitrary
 choice of independent coordinates and speeds, and the time-derivatives of all
 coordinates and speeds (both independent and dependent).  This enables several
-common tasks: (1) stability analysis of arbitrary equilibria, (2) traditional
-control system design, (3) control system design with output feedback based
+common tasks: (1) stability analysis, (2) control system design, (3) control system design with output feedback based
 upon quantities that are dependent (i.e., not state variables), (4)
@@
     \hline
-    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}$ & $f_c : \mathbf{R}^n \times \mathbf{R} \to
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}$ & $\bm{f}_c : \mathbf{R}^n \times \mathbf{R} \to
 \mathbf{R}^l$ \\
-$\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}$ & $f_v : \mathbf{R}^n \times
+$\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}$ & $\bm{f}_v : \mathbf{R}^n \times
 \mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
 $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}$ &
-$f_a : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R}^o \times
+$\bm{f}_a : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R}^o \times
 \mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
@@
 \tilde{A}_{uqd} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3)\right] \\
-\tilde{A}_{uud} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_u\right] \\
-\tilde{B}_{u}   &\triangleq \nabla_{\bm{r}}\bm{f}_{3}
+\tilde{A}_{uud} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_3\right] \\
+\tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
 \end{align}
@@
 and $P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$,
-respectively; similarly, $P_{ui}$ is the first $o-m$ colums of $P_{u}$ while
+respectively; similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while
 $P_{ud}$ is the last $m$ columns of $P_u$.
@@
 \begin{align}
-  \nabla_{\bm{q}}\bm{f}_{h} P_{q} \cdot \delta \bm{\tilde{q}} &=
-  \nabla_{\bm{q}}\bm{f}_{h} P_{qi} \cdot \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{h} P_{qd} \cdot \delta \bm{q_d}\notag\\
+  \nabla_{\bm{q}}\bm{f}_{c} P_{q} \cdot \delta \bm{\tilde{q}} &=
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \cdot \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \cdot \delta \bm{q_d}\notag\\
   &= \bm{0}\notag\\
-  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{h} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{h} P_{qi}) \cdot \delta \bm{q}_i \notag\\
+  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
+  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \cdot \delta \bm{q}_i \notag\\
   \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{h} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{h} \right] P_{qi} \cdot \delta
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \cdot \delta
     \bm{q}_i
@@
   C_0 \triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{h} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{h} \right] P_{qi}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}
 \end{equation}
@@
 \begin{align}
-\nabla_{\bm{q}}f_{nh} \cdot \delta \bm{q} +
-\nabla_{\bm{u}}f_{nh} \cdot \delta \bm{u} &= \nabla_{\bm{q}} \bm{f}_{nh} \cdot
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i +
-\nabla_{\bm{u}} \bm{f}_{nh} P_{ud} \cdot \delta \bm{u}_d \notag\\
+  \nabla_{\bm{q}}\bm{f}_{v} \cdot \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \cdot \delta \bm{u} &= \nabla_{\bm{q}} \bm{f}_{v} \cdot
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \cdot \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \cdot \delta \bm{u}_d \notag\\
 &= \bm{0}\notag\\
-\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{nh}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{nh} \cdot \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{nh} P_{ui} \cdot \delta \bm{u}_i \right]\notag\\
-  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{nh} \cdot \delta \bm{q}\notag\\
-  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{nh} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{nh} \right] P_{ui} \cdot \delta \bm{u}_i
+\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \cdot \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \cdot \delta \bm{u}_i \right]\notag\\
+  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \cdot \delta \bm{q}\notag\\
+  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui} \cdot \delta \bm{u}_i
 \end{align}
@@
   \label{eq:C_1}
-  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{nh} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{nh} \\
+  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \\
   \label{eq:C_2}
-  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{nh} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{nh} \right] P_{ui}
+  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui}
 \end{align}
@@
      \begin{array}{cc}
-       (\tilde{A}_{qq} + A_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + A_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + A_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
      \end{array}
@@
 There are two requirements for this method to work, namely, both
-$\nabla_{\bm{q}}\bm{f}_{h} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{nh} P_{ud}$
+$\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
 must be nonsingular.  Because of the presence of the permutation matrices, this

### 2012-02-20 PetersonGedeHubbard2012.tex
Removed qdot inverses; placed them in mass matrix
Distributed negative sign.
Started splitting up sections

@@
   Abstract here.
-\keywords{symbolic dynamics software, nonholonomic systems, control}
+\keywords{symbolic dynamics software, control, nonholonomic systems,
+          constrained systems}
 \end{abstract}
@@
 
-\section{Derivation of linearization algorithm}
-\label{derivations}
 The goal of the following calculations is to concretely establish the first
@@
   \centering
-  \caption{Constrained multibody system governing equations}
+  \caption{Constrained multibody system governing definitions and equations}
+
+  \begin{tabular}[c]{l l l}
+    Quantity & Shape & Description\\
+    \hline
+    $\bm{q}$ & $\mathbf{R}^n$ & Coordinates \\
+    $\bm{\dot{q}}$ & $\mathbf{R}^n$ & Time derivatives of coordinates \\
+    $\bm{u}$ & $\mathbf{R}^o$ & Speeds \\
+    $\bm{\dot{u}}$ & $\mathbf{R}^o$ & Time derivatives of speeds \\
+    $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
+  \end{tabular}
   \begin{tabular}[c]{r @{ $=$ } l l}
-    \multicolumn{2}{c}{Equation} & Mathematical definition \\
+    \multicolumn{3}{c}{ } \\
+    \multicolumn{2}{c}{Equation} & Description \\
     \hline
-    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}$ & $\bm{f}_c : \mathbf{R}^n \times \mathbf{R} \to
-\mathbf{R}^l$ \\
-$\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}$ & $\bm{f}_v : \mathbf{R}^n \times
-\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
-$\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}$ &
-$\bm{f}_a : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R}^o \times
-\mathbf{R}^o \times \mathbf{R} \to \mathbf{R}^m$ \\
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration (holonomic)
+    constraints \\
+    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity
+    (nonholonomic) constraints \\
+    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
+    \times 1}$ & Acceleration constraints \\
     $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
-$\bm{0}$ & $\bm{f}_0 : \mathbf{R}^n \times \mathbf{R}^n \times \mathbf{R} \to
-\mathbf{R}^n$\\
-\multicolumn{2}{c}{} & $\bm{f}_1 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}
-\to \mathbf{R}^n$ \\
-$\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t)$ &
-$\bm{0}$ & $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R} \to
-\mathbf{R}^{o-m}$,\\
-\multicolumn{2}{c}{} & $\bm{f}_3 : \mathbf{R}^n \times \mathbf{R}^n \times
-\mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R} \to \mathbf{R}^{o-m}$
+    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
+    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential
+    equations \\
   \end{tabular}
@@
 much interest in themselves, nor are they necessary for most analyses.
+
+\section{Derivation of linearization algorithm}
+\label{derivations}
 
@@
 \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
+\tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
 \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a \\
+\tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
 \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
@@
 \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
-\tilde{A}_{uqc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} \bm{f}_a\right] \\
-\tilde{A}_{uuc} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_a \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_a\right] \\
-\tilde{A}_{uqd} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qq} + \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3)\right] \\
-\tilde{A}_{uud} &\triangleq -\left[ \nabla_{\bm{\dot{q}}}\bm{f}_3 \tilde{M}_{qq}^{-1} \tilde{A}_{qu} + \nabla_{\bm{u}} \bm{f}_3\right] \\
+\tilde{A}_{uqc} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} \bm{f}_a \\
+\tilde{A}_{uuc} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_a \\
+\tilde{A}_{uqd} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
+\tilde{A}_{uud} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_3 \\
 \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
@@
       \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \bm{0}_{m \times n} & \tilde{M}_{uuc} \\
-      \bm{0}_{(o-m) \times n} & \tilde{M}_{uud}
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
     \end{array}
@@
 
-There are two requirements for this method to work, namely, both
+\section{Implementation and Examples}
+
+There are a number of additional steps needed to use these linear equations or
+form them more efficiently. There are also non-singularity requirements
+involved in selection of independent quantities.
+
+In order to use these equations in the state space representation, 2 things
+need to happen. First, matrices $\tilde{A}$ and $\tilde{B}$ need to be
+multiplied by the mass matrix inverse. We can then define
+
+\begin{align}
+    \prime{A} &\triangleq \tilde{M}^{-1} \tilde{A} \\
+    \prime{B} &\triangleq \tilde{M}^{-1} \tilde{B}
+\end{align}
+
+Now we will have matrices $\prime{A}$ and $\prime{B}$ which are of size $(o +
+n) \times (o - m + n - l)$ and $(o + n) \times p)$. We need to now extract only
+the rows which are associated with the independent quantities. We will define
+% Is there a better way to write transforming A' to A, B' to B?
+% Could we use the permutation matrices here?
+the reduced matrices as $A$ and $B$. We will also define the state space vector
+$x = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$; we now have the
+traditional state space representation.
+
+%For stability analyses, one must examine the stability of the mapping $\delta
+%\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
+%$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
+%the independent rows of these equations ($n-l+o-m$ of them) to form a square
+%matrix whose eigenvalues govern the stability of the system about the point of
+%linearization.
+
+Additionally, there are two requirements for this method to work, namely, both
 $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
@@
 
-For stability analyses, one must examine the stability of the mapping $\delta
-\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
-$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
-the independent rows of these equations ($n-l+o-m$ of them) to form a square
-matrix whose eigenvalues govern the stability of the system about the point of
-linearization.
+The computational cost of computing the jacobian matrices can also be reduced
+if the point of linearization is known ahead of time.
 

### 2012-02-21 PetersonGedeHubbard2012.tex
Updated table and made a few edits to permutation section.

@@
   \caption{Constrained multibody system governing definitions and equations}
-
+  \label{table:assumptions}
   \begin{tabular}[c]{l l l}
@@
     \hline
-    $\bm{q}$ & $\mathbf{R}^n$ & Coordinates \\
-    $\bm{\dot{q}}$ & $\mathbf{R}^n$ & Time derivatives of coordinates \\
-    $\bm{u}$ & $\mathbf{R}^o$ & Speeds \\
-    $\bm{\dot{u}}$ & $\mathbf{R}^o$ & Time derivatives of speeds \\
+    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^n$ & Coordinates and their time
+    derivatives\\
+    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^o$ & Speeds and their time derivatives\\
     $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
@@
     \hline
-    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration (holonomic)
-    constraints \\
-    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity
-    (nonholonomic) constraints \\
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration constraints \\
+    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity constraints \\
     $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
@@
     $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential
-    equations \\
+    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential equations
   \end{tabular}
-  \label{table:assumptions}
 \end{table}
@@
 
-Additionally, there are two requirements for this method to work, namely, both
-$\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$
-must be nonsingular.  Because of the presence of the permutation matrices, this
-implies proper selection of dependent coordinates and dependent speeds is
-critical.  The permutation matrices effectively select the columns of the
-constraint gradient matrices, and may be selected by examination of the
-constraint gradient matrix through a SVD or other numerical techniques; for
-small systems it may be possible to choose dependent speeds without resorting
-to numerical techniques if the system trajectories stay with the configuration
-space where these matrices are nonsingular.
+This method relies on the nonsingularity of $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$
+and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$.  The implication of this is that
+selection of dependent coordinates and dependent speeds is critical.  The
+permutation matrices effectively select the columns of the constraint gradient
+matrices, and may be selected by examination of the constraint gradient matrix
+through a SVD or other numerical techniques; for small systems it may be
+possible to choose dependent speeds without resorting to numerical techniques
+if the system trajectories stay with the configuration space where these
+matrices are nonsingular.
 

### 2012-02-21 PetersonGedeHubbard2012.tex
Fixed A' to A definition

--- 2012-02-20 PetersonGedeHubbard2012.tex
@@
 \begin{align}
-    \prime{A} &\triangleq \tilde{M}^{-1} \tilde{A} \\
-    \prime{B} &\triangleq \tilde{M}^{-1} \tilde{B}
+    A\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+    B\prime &\triangleq \tilde{M}^{-1} \tilde{B}
 \end{align}
 
-Now we will have matrices $\prime{A}$ and $\prime{B}$ which are of size $(o +
+Now we will have matrices $A\prime$ and $B\prime$ which are of size $(o +
 n) \times (o - m + n - l)$ and $(o + n) \times p)$. We need to now extract only
-the rows which are associated with the independent quantities. We will define
-% Is there a better way to write transforming A' to A, B' to B?
-% Could we use the permutation matrices here?
-the reduced matrices as $A$ and $B$. We will also define the state space vector
-$x = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$; we now have the
-traditional state space representation.
+the rows which are associated with the independent quantities. We will call the
+new matrices reduced matrices $A$ and $B$ and compute them with the following
 
-%For stability analyses, one must examine the stability of the mapping $\delta
-%\bm{q}_i \to \bm{\dot{q}}_i$.  This can be done by first solving for
-%$\delta\bm{\dot{x}} = \left[\delta\bm{\dot{q}},\,\delta\bm{\dot{u}}\right]^{T}$, then selecting
-%the independent rows of these equations ($n-l+o-m$ of them) to form a square
-%matrix whose eigenvalues govern the stability of the system about the point of
-%linearization.
+\begin{align}
+    P \prime &\triangleq \begin{matrix}
+        P_{qi} & \bm{O}_{(n - l) \times (o - m)} \\
+        \bm{O}_{(o - m) \times (n - l)} & P_{ui}
+    \end{matrix} \\
+    A &\triangleq P\prime^T A\prime \\
+    B &\triangleq P\prime^T B\prime \\
+\end{align}
+
+We will also define the state space vector $x =
+\left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$; we now have the traditional
+state space representation which only contains independent quantities.
 

### 2012-02-21 PetersonGedeHubbard2012.tex
Fixed stupid error I made in maths

@@
 \begin{align}
-    A\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
-    B\prime &\triangleq \tilde{M}^{-1} \tilde{B}
+    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
 \end{align}
 
-Now we will have matrices $A\prime$ and $B\prime$ which are of size $(o +
+Now we will have matrices $A^\prime$ and $B^\prime$ which are of size $(o +
 n) \times (o - m + n - l)$ and $(o + n) \times p)$. We need to now extract only
 the rows which are associated with the independent quantities. We will call the
-new matrices reduced matrices $A$ and $B$ and compute them with the following
+new reduced matrices $A$ and $B$ and compute them with the following
 
 \begin{align}
-    P \prime &\triangleq \begin{matrix}
+    P^\prime &\triangleq \begin{bmatrix}
         P_{qi} & \bm{O}_{(n - l) \times (o - m)} \\
         \bm{O}_{(o - m) \times (n - l)} & P_{ui}
-    \end{matrix} \\
-    A &\triangleq P\prime^T A\prime \\
-    B &\triangleq P\prime^T B\prime \\
+    \end{bmatrix} \\
+    A &\triangleq P^{\prime T} A^\prime \\
+    B &\triangleq P^{\prime T} B^\prime \\
 \end{align}

### 2012-02-21 PetersonGedeHubbard2012.tex
finished up substition/jacobian part

@@
 
-The computational cost of computing the jacobian matrices can also be reduced
-if the point of linearization is known ahead of time.
+If it is found that substatial computational time is spent in computing the
+jacobian matrices, it might be possible to reduce the time if the point of
+linearization is known. We are taking the partial derivatives with respect to
+$q$, $\dot{q}$, $u$, and $\dot{u}$ in a number of vectors. If we know the point
+of linearization (have numerical values for $q$, $\dot{q}$, $u$, $\dot{u}$), we
+can substitute all of the quantities except for those we are taking the
+derivative with respect to; e.g. when computing $\nabla_{\mb{u}} \bm{f}_3$ we
+substitute in the known values for $q$, $\dot{q}$, and $u$. In this situation,
+the cost of symbolic substitution needs to be considered. In practice, it has
+sometimes been computationally expensive. The potential speed-up needs to be
+considered next to the loss of generality of choosing a point of linearization
+after the fact.
 

### 2012-03-05 PetersonGedeHubbard2012.tex
Fixed a typo.
Fixed zero mat dimensions in Aprime Pprime
Shortened eq's 10-21 into 1 array
Also, fixed some typos/ clarified things in paragraph about linearing about a
point

@@
 \begin{align}
-\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 \\
+    \begin{array}{llcll}
+\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
 \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
-\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a \\
+\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
 \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
-\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 \\
+\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
 \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 \\
+\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
 \tilde{A}_{uqc} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} \bm{f}_a \\
-\tilde{A}_{uuc} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_a \\
+\tilde{A}_{uuc} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_a & \quad &
 \tilde{A}_{uqd} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
-\tilde{A}_{uud} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_3 \\
+\tilde{A}_{uud} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_3 & \quad &
 \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
+\end{array}
 \end{align}
@@
     P^\prime &\triangleq \begin{bmatrix}
-        P_{qi} & \bm{O}_{(n - l) \times (o - m)} \\
-        \bm{O}_{(o - m) \times (n - l)} & P_{ui}
+        P_{qi} & \bm{O}_{n \times (o - m)} \\
+        \bm{O}_{o \times (n - l)} & P_{ui}
     \end{bmatrix} \\
@@
 can substitute all of the quantities except for those we are taking the
-derivative with respect to; e.g. when computing $\nabla_{\mb{u}} \bm{f}_3$ we
-substitute in the known values for $q$, $\dot{q}$, and $u$. In this situation,
-the cost of symbolic substitution needs to be considered. In practice, it has
-sometimes been computationally expensive. The potential speed-up needs to be
-considered next to the loss of generality of choosing a point of linearization
-after the fact.
+derivative with respect to; e.g. when computing $\nabla_{\bm{u}} \bm{f}_3$ we
+substitute in the known values for $q$, $\dot{q}$, and $r$. In this situation,
+the cost of symbolic substitution needs to be considered. In practice,
+substituion is sometimes computationally expensive. A speed-up, which is not
+not guaranteed, needs to be weighed against the loss of flexibility to choose a
+point of linearization at a later time.
 

### 2012-03-05 PetersonGedeHubbard2012.tex
Added abstract and changed some keywords.

--- 2012-02-21 PetersonGedeHubbard2012.tex
@@
 \begin{abstract}
-  Abstract here.
-\keywords{symbolic dynamics software, control, nonholonomic systems,
-          constrained systems}
+  We present a systematic procedure for linearizing equations of motion
+  governing constrained multibody systems which have been derived with Kane's
+  method.  Systems with configuration and velocity level constraints are
+  address by the procedure.  The procedure avoids differential algebraic
+  equations and yields first order relationships between time derivatives of
+  generalized coordinates/speeds (both independent and dependent) and
+  independent generalized coordinates/speeds and exogenous inputs.  The
+  procedure enables stability analysis and control system design in a fashion
+  that ensures that all constraints are satisfied.  We present an
+  implementation of the algorithm and numerically validate the results on a
+  well studied model of a bicycle.
+\keywords{symbolic dynamics, control, constrained multibody systems}
 \end{abstract}
@@
 There are a number of additional steps needed to use these linear equations or
-form them more efficiently. There are also non-singularity requirements
+lorm them more efficiently. There are also non-singularity requirements
 involved in selection of independent quantities.
@@
 can substitute all of the quantities except for those we are taking the
-derivative with respect to; e.g. when computing $\nabla_{\mb{u}} \bm{f}_3$ we
+derivative with respect to; e.g. when computing $\nabla_{\bm{u}} \bm{f}_3$ we
 substitute in the known values for $q$, $\dot{q}$, and $u$. In this situation,

### 2012-03-05 PetersonGedeHubbard2012.tex
Wrote wrap up sentence for section 2
Finished abstract.

--- 2012-03-05 PetersonGedeHubbard2012.tex
@@
 \begin{abstract}
-  Abstract here.
-\keywords{symbolic dynamics software, control, nonholonomic systems,
-          constrained systems}
+  We present a systematic procedure for linearizing equations of motion
+  governing constrained multibody systems which have been derived with Kane's
+  method.  The procedure avoids differential algebraic equations and yields
+  linear relationships between time derivatives of generalized
+  coordinates/speeds (both independent and dependent) and independent
+  generalized coordinates/speeds and exogenous inputs.  These minimal
+  linearized equations effectively embed the constraints within them, and
+  enable stability analysis and control system design without the need to
+  explicitly address the constraints.  We numerically validate the results on a
+  well studied model of a bicycle which includes both configuration and
+  velocity constraints.
+\keywords{symbolic dynamics, control, constrained multibody systems}
 \end{abstract}
@@
 %\bibliographystyle{spbasic}      % basic style, author-year citations
+\section{Conclusion}
+
 \bibliography{library}   % name your BibTeX data base

### 2012-03-25 PetersonGedeHubbard2012.tex
Added some clarifying details; started conclusion.

@@
 
-\title{Open source symbolic multibody dynamics software}
+\title{A linearization procedure for constrained multibody systems}
 
-\titlerunning{Open source symbolic multibody dynamics software}        % if too long for running head
+\titlerunning{A linearization procedure for constrained multibody systems}        % if too long for running head
 
@@
 time-varying constrained multibody systems with arbitrary external inputs and
-arbitrary specified quantities.
+arbitrary specified quantities.  This procedure has been created for systems
+whose equations of motion have been found with Kane's Method; this ensures that
+only ordinary differential equations are generated and no differential
+algebraic equations arise.  While the use of Kane's Method is not required for
+this procedure, equations of motion which can be completely described by a series
+of ODEs is a requirement.
 
@@
 
+We see in Table \ref{table:assumptions} the differential equations which
+describe our system. The goal of this procedure is to create linear equations
+of motion with the following structure:
+\begin{align}
+    \label{eq:structure}
+    \tilde{M} \dot{\bm{x}} = \tilde{A} \bm{x}_i + \tilde{B} r
+\end{align}
+where $\bm{x}$ is the state vector of all the coordinates and speeds and
+$\bm{x}_i$ is the state vector of only the independent coordinates and speeds.
+From this equation we can perform numerical integration or stability analysis,
+or it can be reformulated into the traditional state space equations for
+controller design.
+
 We begin with a first order Taylor series expansion of the equations in Table
@@
 \begin{align}
+    \label{eq:quant_to_compute}
     \begin{array}{llcll}
@@
 
+We now have a linear system of equations which relates the time derivatives of all
+states, dependent or independent, to the independent coordinates, independent
+speeds, and exogenous inputs.
+
 \section{Implementation and Examples}
@@
     A &\triangleq P^{\prime T} A^\prime \\
-    B &\triangleq P^{\prime T} B^\prime \\
+    B &\triangleq P^{\prime T} B^\prime
 \end{align}
@@
 
+% Not sure which bibliography style we are supposed to use
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliographystyle{spbasic}      % basic style, author-year citations
+\section{Conclusion}
+In this paper, we have presented a procedure for creating linear equations of
+motion for a system with holonomic and nonholonomic constraints. We have also
+described some of the considerations that must be made when following this
+procedure as well as a numerical example showing the results of and validating
+our procedure.
+
+In summary, the procedure is:
+
+\begin{enumerate}
+    \item Describe the system in the form of the equations shown in Table
+        \ref{table:assumptions}.
+    \item Compute the quantities shown in equation (\ref{eq:quant_to_compute}).
+    \item Either through inspection or performing a SVD on
+        $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$
+        and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$, create the partition matrices
+        $P_u$ and $P_q$ which correctly identify the independent coordinates and
+        speeds.
+    \item Form the quantities in equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
+        (\ref{eq:C_2}); using these terms and those from equation
+        (\ref{eq:quant_to_compute}), form equation
+        (\ref{eq:state_space_constrained}).
+    \item Choose a point of linearization and:
+        \begin{enumerate}
+            \item integrate the equations of motion.
+            \item reformulate the equations into the traditional state space form.
+            \item perform stability analysis.
+            \item perform other analyses.
+        \end{enumerate}
+\end{enumerate}
+
 \begin{acknowledgements}
@@
 
-% Not sure which bibliography style we are supposed to use
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliographystyle{spbasic}      % basic style, author-year citations
-\section{Conclusion}
-
 \bibliography{library}   % name your BibTeX data base

### 2012-03-25 PetersonGedeHubbard2012.tex
Editted intro.

@@
 coordinates and speeds (both independent and dependent).  This enables several
-common tasks: (1) stability analysis, (2) control system design, (3) control system design with output feedback based
-upon quantities that are dependent (i.e., not state variables), (4)
-determination of measurement equations for use in Kalman filter design when
-measured quantities are at the acceleration level (i.e., accelerometers).  The
-formalism presented here is generic enough to cover most examples of
-time-varying constrained multibody systems with arbitrary external inputs and
-arbitrary specified quantities.  This procedure has been created for systems
-whose equations of motion have been found with Kane's Method; this ensures that
-only ordinary differential equations are generated and no differential
-algebraic equations arise.  While the use of Kane's Method is not required for
-this procedure, equations of motion which can be completely described by a series
-of ODEs is a requirement.
+common tasks: (1) stability analysis, (2) control system design, (3) control
+system design with output feedback based upon quantities that are dependent
+(i.e., not state variables), (4) determination of measurement equations for use
+in Kalman filter design when measured quantities are at the acceleration level
+(i.e., accelerometers).  The formalism presented here is generic enough to
+cover most examples of time-varying constrained multibody systems with
+arbitrary external inputs and arbitrary specified quantities.  This procedure
+has been created for systems whose equations of motion have been derived with
+Kane's Method; while this is not a strict requirement, systems
+which can be completely described by a set of ODEs is (i.e., no DAEs).
 

### 2012-03-26 PetersonGedeHubbard2012.tex
Added references.bib.
Changed wording in conclusion re pt. of lineariz
Minor changes
section 3 -> now titled discussion
fixed eq1 boldness
removed last paragraph in section 3

@@
     \label{eq:structure}
-    \tilde{M} \dot{\bm{x}} = \tilde{A} \bm{x}_i + \tilde{B} r
+    \tilde{M} \dot{\bm{x}} = \tilde{A} \bm{x}_i + \tilde{B} \bm{r}
 \end{align}
@@
 
-\section{Implementation and Examples}
+\section{Discussion}
 
@@
 
-If it is found that substatial computational time is spent in computing the
-jacobian matrices, it might be possible to reduce the time if the point of
-linearization is known. We are taking the partial derivatives with respect to
-$q$, $\dot{q}$, $u$, and $\dot{u}$ in a number of vectors. If we know the point
-of linearization (have numerical values for $q$, $\dot{q}$, $u$, $\dot{u}$), we
-can substitute all of the quantities except for those we are taking the
-derivative with respect to; e.g. when computing $\nabla_{\bm{u}} \bm{f}_3$ we
-substitute in the known values for $q$, $\dot{q}$, and $r$. In this situation,
-the cost of symbolic substitution needs to be considered. In practice,
-substituion is sometimes computationally expensive. A speed-up, which is not
-not guaranteed, needs to be weighed against the loss of flexibility to choose a
-point of linearization at a later time.
-
 % Not sure which bibliography style we are supposed to use
@@
         (\ref{eq:state_space_constrained}).
-    \item Choose a point of linearization and:
+    \item Choose a point of linearization which satisfies the equations in
+        Table \ref{table:assumptions} and:
         \begin{enumerate}
@@
 
-\bibliography{library}   % name your BibTeX data base
+\bibliography{references}   % name your BibTeX data base
 \bibliographystyle{spmpsci}      % mathematics and physical sciences

### 2012-03-26 PetersonGedeHubbard2012.tex
Made a few f's bold in equations 2-3.
Changes in conclusion

@@
   \label{eq:configuration_expansion}
-  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{f_{c}(\bm{q}^*, t)}_{\bm{0}} +
-    \nabla_{\bm{q}}f_{c} \cdot \delta \bm{q}\\
+  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
+  \nabla_{\bm{q}}\bm{f}_{c} \cdot \delta \bm{q}\\
   \label{eq:velocity_expansion}
   \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{v} \cdot \delta \bm{q} +
-  \nabla_{\bm{u}}f_{v} \cdot \delta \bm{u} \\
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \cdot \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \cdot \delta \bm{u} \\
   \label{eq:acceleration_expansion}
@@
   \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
-t)}_{\bm{0}} +  \nabla_{\bm{q}}f_{a} \cdot \delta \bm{q} + \nabla_{\bm{\dot{q}}}f_{a}
+t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \cdot \delta \bm{q} +
+\nabla_{\bm{\dot{q}}}\bm{f}_{a}
 \cdot \delta \bm{\dot{q}} \notag\\
-&+ \nabla_{\bm{u}}f_{a} \cdot \delta \bm{u} + \nabla_{\bm{\dot{u}}}f_{a} \cdot
+&+ \nabla_{\bm{u}}\bm{f}_{a} \cdot \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a} \cdot
 \delta \bm{\dot{u}}
@@
 There are a number of additional steps needed to use these linear equations or
-lorm them more efficiently. There are also non-singularity requirements
+form them more efficiently. There are also non-singularity requirements
 involved in selection of independent quantities.
@@
 \section{Conclusion}
-In this paper, we have presented a procedure for creating linear equations of
+We have presented a procedure for creating linear equations of
 motion for a system with holonomic and nonholonomic constraints. We have also
 described some of the considerations that must be made when following this
-procedure as well as a numerical example showing the results of and validating
-our procedure.
+procedure as well validating it with examples.
 
 In summary, the procedure is:
-
 \begin{enumerate}
@@
     \item Compute the quantities shown in equation (\ref{eq:quant_to_compute}).
-    \item Either through inspection or performing a SVD on
-        $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$
-        and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$, create the partition matrices
-        $P_u$ and $P_q$ which correctly identify the independent coordinates and
-        speeds.
+    \item Identify independent coordinates and speeds.
     \item Form the quantities in equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
-        (\ref{eq:C_2}); using these terms and those from equation
-        (\ref{eq:quant_to_compute}), form equation
-        (\ref{eq:state_space_constrained}).
-    \item Choose a point of linearization which satisfies the equations in
-        Table \ref{table:assumptions} and:
-        \begin{enumerate}
-            \item integrate the equations of motion.
-            \item reformulate the equations into the traditional state space form.
-            \item perform stability analysis.
-            \item perform other analyses.
-        \end{enumerate}
+        (\ref{eq:C_2}).
+    \item From the results of steps 2 and 4, for equation (\ref{eq:state_space_constrained}).
 \end{enumerate}

### 2012-03-30 PetersonGedeHubbard2012.tex
Improved intro.

@@
   coordinates/speeds (both independent and dependent) and independent
-  generalized coordinates/speeds and exogenous inputs.  These minimal
-  linearized equations effectively embed the constraints within them, and
-  enable stability analysis and control system design without the need to
-  explicitly address the constraints.  We numerically validate the results on a
-  well studied model of a bicycle which includes both configuration and
-  velocity constraints.
+  generalized coordinates/speeds and exogenous inputs.  These linear ODE's
+  effectively embed the constraints, and enable stability analysis and control
+  system design without the need to explicitly address the constraints.  We
+  numerically validate the results on a well studied model of a bicycle which
+  includes both configuration and velocity constraints.
 \keywords{symbolic dynamics, control, constrained multibody systems}
@@
 common tasks: (1) stability analysis, (2) control system design, (3) control
-system design with output feedback based upon quantities that are dependent
-(i.e., not state variables), (4) determination of measurement equations for use
-in Kalman filter design when measured quantities are at the acceleration level
-(i.e., accelerometers).  The formalism presented here is generic enough to
-cover most examples of time-varying constrained multibody systems with
-arbitrary external inputs and arbitrary specified quantities.  This procedure
-has been created for systems whose equations of motion have been derived with
-Kane's Method; while this is not a strict requirement, systems
-which can be completely described by a set of ODEs is (i.e., no DAEs).
+system or estimator design which involve dependent quanties (i.e, position
+control on a dependent coordinate or measurements at acceleration level).  The
+formalism presented here is generic enough to cover most examples of
+time-varying constrained multibody systems with arbitrary external inputs and
+arbitrary specified quantities.  This procedure has been created for systems
+whose equations of motion have been derived with Kane's Method; while this is
+not a strict requirement, systems which can be completely described by a set of
+ODEs is (i.e., no DAEs).
 

### 2012-03-31 PetersonGedeHubbard2012.tex
Worked on discussion section, added a reference.

@@
 \end{table}
-The first three equations in table \ref{table:assumptions} represent
+The first three equations in Table \ref{table:assumptions} represent
 constraints derived purely from kinematic considerations at the configuration,
@@
 \section{Derivation of linearization algorithm}
-\label{derivations}
+\label{sec:derivations}
 
 We see in Table \ref{table:assumptions} the differential equations which
-describe our system. The goal of this procedure is to create linear equations
+describe our system. The goal of this section is to create linear equations
 of motion with the following structure:
@@
 $\bm{x}_i$ is the state vector of only the independent coordinates and speeds.
-From this equation we can perform numerical integration or stability analysis,
+This equation can be used for simulation, stability analysis,
 or it can be reformulated into the traditional state space equations for
-controller design.
+controller design; these topics are discussed in \ref{sec:discussion}
 
@@
 We now have a linear system of equations which relates the time derivatives of all
-states, dependent or independent, to the independent coordinates, independent
-speeds, and exogenous inputs.
+states (dependent and independent) to the independent states and exogenous inputs.
 
 \section{Discussion}
-
-There are a number of additional steps needed to use these linear equations or
-form them more efficiently. There are also non-singularity requirements
-involved in selection of independent quantities.
-
-In order to use these equations in the state space representation, 2 things
-need to happen. First, matrices $\tilde{A}$ and $\tilde{B}$ need to be
-multiplied by the mass matrix inverse. We can then define
-
+\label{sec:discussion}
+Equation \ref{eq:state_space_constrained} doesn't fit into the standard linear
+system framework because it is non-square.  To obtain a square system (useful
+in stability analysis and control system design) we define
 \begin{align}
@@
 \end{align}
-
-Now we will have matrices $A^\prime$ and $B^\prime$ which are of size $(o +
-n) \times (o - m + n - l)$ and $(o + n) \times p)$. We need to now extract only
-the rows which are associated with the independent quantities. We will call the
-new reduced matrices $A$ and $B$ and compute them with the following
-
+where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
+\bm{R}^{(o + n) \times p)}$.  We can extract the rows corresponding to the
+independent states by defining
 \begin{align}
@@
 \end{align}
+where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
+$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
+state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$.  The standard
+linear systems analyses may now be applied.  It is worth noting that the rows
+of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
+in the output or measurement equations of a linear state space model, as would
+be the case when the control task is to control a dependent state or when
+measurements of dependent state(s) are made.
 
-We will also define the state space vector $x =
-\left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$; we now have the traditional
-state space representation which only contains independent quantities.
-
-This method relies on the nonsingularity of $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$
-and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$.  The implication of this is that
-selection of dependent coordinates and dependent speeds is critical.  The
-permutation matrices effectively select the columns of the constraint gradient
-matrices, and may be selected by examination of the constraint gradient matrix
-through a SVD or other numerical techniques; for small systems it may be
-possible to choose dependent speeds without resorting to numerical techniques
-if the system trajectories stay with the configuration space where these
-matrices are nonsingular.
+The choice on which state variables are taken to be independent may determine
+whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}$ are nonsingular.  These matrices depend upon the
+configuration variables and parameters; for certain configurations or
+parameters, it may be that the choice of independent state variables cannot be
+arbitrary.  While some systems may permit a choice of independent state
+variables which are valid for all configurations of interest, others may not.
+Methods for automatically selecting the ``best'' choice of independent state
+variables are discussed in \cite{Reckdahl1996}; they involve computing
+the SVD of the Jacobian of the constraint matrices to determine a set of
+indepedent speeds which will ensure the nonsingularity of the aforementioned
+matrices.
 

### 2012-04-04 rolling_disc.py
Added rolling disc example py file

@@ -1,0 +1,113 @@
+from sympy import *
+from sympy.physics.mechanics import *
+
+psi, phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z = dynamicsymbols('psi,'
+    'phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z')
+psid, phid, thetad, oxd, oyd, ozd, w_xd, w_yd, w_zd, v_xd, v_yd, v_zd = dynamicsymbols('psi,'
+    'phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z', 1)
+r, m, g, I, J = symbols('r, m, g, I, J')
+t = symbols('t')
+
+N = ReferenceFrame('N')
+Y = N.orientnew('Y', 'Axis', [psi, N.z])
+L = Y.orientnew('L', 'Axis', [phi, Y.x])
+R = L.orientnew('R', 'Axis', [theta, L.y])
+
+O = Point('O')
+C = O.locatenew('C', r * L.z)
+
+R.set_ang_vel(N, w_x * L.x + w_y * L.y + w_z * L.z)
+O.set_vel(N, v_x * L.x + v_y * L.y + v_z * L.z)
+C.v2pt_theory(O, N, R)
+
+ccons = [r * (L.z & Y.z) - oz]
+vcons = [C.vel(N) & Y.x, C.vel(N) & Y.y, C.vel(N) & Y.z]
+
+Disk = RigidBody('Disk', O, R, m, (inertia(L, I, I, J), O))
+
+v_o_exp_n = O.vel(N).express(N)
+
+kd = ([psid - w_z / cos(psid), phid - w_x, thetad - w_y + w_z * tan(phi),
+       (v_o_exp_n & N.x) - oxd, (v_o_exp_n & N.y) - oyd, (v_o_exp_n & N.z) - ozd])
+
+Bodies = [Disk]
+Forces = [(O, m * g * Y.z)]
+
+KM = Kane(N)
+KM.coords([psi, phi, theta, ox, oy], [oz], ccons)
+KM.speeds([w_x, w_y, w_z], [v_x, v_y, v_z], vcons)
+KM.kindiffeq(kd)
+KM.kanes_equations(Forces, Bodies)
+
+mm = KM.mass_matrix_full
+f = KM.forcing_full
+f.simplify()
[…]

### 2012-04-05 PetersonGedeHubbard2012.tex
Fixed typos in eq 11
Made changes suggested by JKM, added reference, work on discussion/conclusion.

@@
 choice of independent coordinates and speeds, and the time-derivatives of all
-coordinates and speeds (both independent and dependent).  This enables several
-common tasks: (1) stability analysis, (2) control system design, (3) control
-system or estimator design which involve dependent quanties (i.e, position
-control on a dependent coordinate or measurements at acceleration level).  The
-formalism presented here is generic enough to cover most examples of
-time-varying constrained multibody systems with arbitrary external inputs and
-arbitrary specified quantities.  This procedure has been created for systems
-whose equations of motion have been derived with Kane's Method; while this is
-not a strict requirement, systems which can be completely described by a set of
-ODEs is (i.e., no DAEs).
+coordinates and speeds (both independent and dependent).  This is necessary for
+tasks which require a linear state space formulation.  Common examples include
+stability analysis, controller and estimator design (including control or
+measurement of a dependent quantity).  The formalism presented here is generic
+enough to cover most examples of time-varying constrained multibody systems
+with arbitrary external inputs and arbitrary specified quantities.  This
+procedure has been created for systems whose equations of motion have been
+derived with Kane's Method.  While this is not a strict requirement, systems
+which can be completely described by a set of ODEs is (i.e., no DAEs).
 
@@
 $\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
-relationships described in table \ref{table:assumptions}.  The system is
-described by $l + m + n + o$ equations which relate $2n + 2o + 2p + 1$
-quantities, which implies that $n - l + o - m + p + 1$ of these quantities are
+relationships described in Table \ref{table:assumptions}.  The system is
+described by $l + m + n + o$ equations which relate $2n + 2o + s + 1$
+quantities, which implies that $n - l + o - m + s + 1$ of these quantities are
 independent.  This paper discusses the first order relationship between these
 independent quantities ($n-l$ independent coordinates, $o-m$ independent
-speeds, $p$ control inputs, and time $t$) and the time derivatives of the $n$
+speeds, $s$ control inputs, and time $t$) and the time derivatives of the $n$
 coordinates and $o$ speeds.
-%It is assumed that there are $l$
-%configuration constraints (), and $m$ velocity constraints , and $m$ acceleration
-%constraints (). Two functions $ relate the coordinate time derivatives to the generalized speeds,
-%and two more functions $\bm{f}_2 : \mathbf{R}^n \times \mathbf{R}^o \times
-%\mathbf{R} \to \mathbf{R}^{o-m}$ and $\bm{f}_3 : \mathbf{R}^n
-%\times \mathbf{R}^n \times \mathbf{R}^o \times \mathbf{R}^p \times \mathbf{R}
-%\to \mathbf{R}^{o-m}$ relate the time derivatives of the speeds $\bm{\dot{u}}$ to
-%the externally applied inputs $\bm{r}\in\bm{R}^q$, coordinates $\bm{q}$ and
-%their time derivatives $\bm{\dot{q}}$,  speeds $\bm{u}$, and time $t$. These
-%assumptions can be summarized by the following equations
-%\begin{align}
-%  \label{eq:configuration}
-%  \bm{f}_{c}(\bm{q}, t) &= \bm{0}\\
-%  \label{eq:velocity}
-%  \bm{f}_{v}(\bm{q}, \bm{u}, t) &= \bm{0}\\
-%  \label{eq:acceleration}
-%  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &= \bm{0}\\
-%  \label{eq:kindiffs}
-%  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t)
-%    + \bm{f}_{1}(\bm{q}, \bm{u}, t) &= \bm{0} \\
-%  \label{eq:dyndiffs}
-%  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t)
-%    + \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) & = \bm{0}
-%\end{align}
 \begin{table}[htbp]
@@
     $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^o$ & Speeds and their time derivatives\\
-    $\bm{r}$ & $\mathbf{R}^p$ & Control inputs \\
+    $\bm{r}$ & $\mathbf{R}^s$ & Exogenous inputs \\
   \end{tabular}
@@
 \end{table}
+
 The first three equations in Table \ref{table:assumptions} represent
@@
 velocity, and acceleration levels, respectively.  The velocity constraints may
-be nonholonomic or differentiated holonomic, while the acceleration constraints
-are assumed to be differentiated versions of the velocity constraints.  The
-fourth equation represents kinematic differential equations which relate
-$\bm{\dot{q}}$ and $\bm{u}$ and will be linear in both of these quantities.
-The fifth equation represents the constraint free nonholonomic dynamic
-differential equations~\cite{Kane1985}, and will necessarily by linear in
-$\bm{\dot{u}}$ and input forces and torques $\bm{r}$, but generally nonlinear
-in all other quantities.  This formalism is exceedingly generic; in practice,
-the structure of the functions in table \ref{table:assumptions} can be
-leveraged to simplify the following calculations.  This will be discussed
-subsequently.
+be nonholonomic or time differentiated holonomic.  The acceleration constraints
+can be time differentiated velocity constraints or kinematic acceleration
+constraints.  The fourth equation represents kinematic differential equations
+which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both of these
+quantities.  The fifth equation represents the constraint-free nonholonomic
+dynamic differential equations~\cite{Kane1985}, and will necessarily be linear
+in $\bm{\dot{u}}$ and the inputs $\bm{r}$, but generally nonlinear in all other
+quantities.  This formalism is exceedingly generic; in practice, the structure
+of the functions in Table \ref{table:assumptions} can be leveraged to simplify
+the following calculations.  This will be discussed subsequently.
 
@@
 (e.g., total mechanical energy) may restrict trajectories to a lower
-dimensional manifold.  In general this cannot be assumed, however, especially
+dimensional manifold.  In general this cannot be assumed, especially
 in the presence of external control inputs $\bm{r}$ which may change the total
@@
 $p=10$ but 5 of these state variables are cyclic coordinates (coordinates of
-rear wheel contact point, heading, and both wheel angles) and are often not of
-much interest in themselves, nor are they necessary for most analyses.
+rear wheel contact point, heading, and both wheel angles) and may only be of
+secondary interest.
 
@@
 or it can be reformulated into the traditional state space equations for
-controller design; these topics are discussed in \ref{sec:discussion}
+a variety of tasks; these topics are discussed in \ref{sec:discussion}
 
@@
   \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
-  \nabla_{\bm{q}}\bm{f}_{c} \cdot \delta \bm{q}\\
+  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
   \label{eq:velocity_expansion}
   \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \cdot \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \cdot \delta \bm{u} \\
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
   \label{eq:acceleration_expansion}
@@
   \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
-t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \cdot \delta \bm{q} +
+t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
 \nabla_{\bm{\dot{q}}}\bm{f}_{a}
-\cdot \delta \bm{\dot{q}} \notag\\
-&+ \nabla_{\bm{u}}\bm{f}_{a} \cdot \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a} \cdot
+ \delta \bm{\dot{q}} \notag\\
+&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
 \delta \bm{\dot{u}}
@@
   \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
-  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \cdot \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \cdot \delta\bm{\dot{q}}\\
+  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
   \label{eq:f1_expansion}
   \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
-  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \cdot \delta\bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{1} \cdot \delta\bm{u}
+  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
 \end{align}
@@
   \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
-  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \cdot \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \cdot \delta\bm{\dot{q}} +
-  \nabla_{\bm{u}}\bm{f}_{1} \cdot \delta\bm{u}
+  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
 \end{align}
@@
       \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
-      \nabla_{\bm{q}}\bm{f}_{2} \cdot \delta\bm{q}
-      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \cdot \delta\bm{\dot{u}}\\
+      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
+      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
   \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
   \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
-  \nabla_{\bm{q}}\bm{f}_{3} \cdot \delta\bm{q}\notag\\
+  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
   \label{eq:f3_expansion}
-  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \cdot \delta\bm{\dot{q}}
-  + \nabla_{\bm{u}}\bm{f}_{3} \cdot \delta \bm{u}
-  + \nabla_{\bm{r}}\bm{f}_{3} \cdot \delta\bm{r}
+  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
+  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
+  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
 \end{align}
@@
   \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3) \cdot
-  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \cdot \delta\bm{\dot{q}}\notag\\
+  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
+  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
   \label{eq:f2_plus_f3_expansion}
-  &+ \nabla_{\bm{u}}\bm{f}_{3} \cdot \delta\bm{u} +
-  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \cdot \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \cdot \delta\bm{r}
+  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
+  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
 \end{align}
@@
 (\ref{eq:acceleration_expansion}),
-and (\ref{eq:f2_plus_f3_expansion}) to zero (as per table
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as per Table
 \ref{table:assumptions}), and introducing the following definitions
@@
 \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
-\tilde{A}_{uqc} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} \bm{f}_a \\
-\tilde{A}_{uuc} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_a & \quad &
-\tilde{A}_{uqd} &\triangleq -\tilde{A}_{qq} - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
-\tilde{A}_{uud} &\triangleq -\tilde{A}_{qu} - \nabla_{\bm{u}} \bm{f}_3 & \quad &
+\tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
+\tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
+\tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
+\tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
 \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
@@
       \begin{array}{c}
-        \bm{0}_{(n + m) \times q} \\
+        \bm{0}_{(n + m) \times s} \\
         \tilde{B}_{u}
@@
 
-Setting the right hand side of equation (\ref{eq:configuration_expansion}) equal to
-zero, and making use of $P_q$, we have
+By assumption, equation
+(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
 \begin{align}
-  \nabla_{\bm{q}}\bm{f}_{c} P_{q} \cdot \delta \bm{\tilde{q}} &=
-  \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \cdot \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \cdot \delta \bm{q_d}\notag\\
-  &= \bm{0}\notag\\
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
+   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
   \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \cdot \delta \bm{q}_i \notag\\
+  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
   \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \cdot \delta
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
     \bm{q}_i
@@
   \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \cdot \delta \bm{q}_i
+  \delta \bm{q} &= C_0 \delta \bm{q}_i
 \end{align}
@@
 \begin{align}
-  \nabla_{\bm{q}}\bm{f}_{v} \cdot \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \cdot \delta \bm{u} &= \nabla_{\bm{q}} \bm{f}_{v} \cdot
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \cdot \delta \bm{u}_i +
-\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \cdot \delta \bm{u}_d \notag\\
-&= \bm{0}\notag\\
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
+  &= \nabla_{\bm{q}} \bm{f}_{v}
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
 \implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \cdot \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \cdot \delta \bm{u}_i \right]\notag\\
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
   \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \cdot \delta \bm{q}\notag\\
+  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
   &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \cdot \delta \bm{u}_i
+    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i
 \end{align}
@@
   \label{eq:delta_u}
-  \delta \bm{u} &= C_1 \cdot \delta \bm{q} + C_2 \cdot \delta \bm{u}_i\notag\\
-  &= C_1 C_0 \cdot \delta \bm{q}_i + C_2 \cdot \delta \bm{u}_i
+  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
+  &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
 \end{align}
@@
       \begin{array}{c}
-        \bm{0}_{(n+m) \times q} \\
+        \bm{0}_{(n+m) \times s} \\
         \tilde{B}_{u}
@@
 \end{align}
-
 We now have a linear system of equations which relates the time derivatives of all
@@
 \begin{align}
+  \label{eq:A_prime}
     A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+  \label{eq:B_prime}
     B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
@@
 where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
-\bm{R}^{(o + n) \times p)}$.  We can extract the rows corresponding to the
+\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
 independent states by defining
 \begin{align}
+  \label{eq:P_prime}
     P^\prime &\triangleq \begin{bmatrix}
@@
     \end{bmatrix} \\
+  \label{eq:A}
     A &\triangleq P^{\prime T} A^\prime \\
+  \label{eq:B}
     B &\triangleq P^{\prime T} B^\prime
@@
 whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}$ are nonsingular.  These matrices depend upon the
-configuration variables and parameters; for certain configurations or
-parameters, it may be that the choice of independent state variables cannot be
-arbitrary.  While some systems may permit a choice of independent state
-variables which are valid for all configurations of interest, others may not.
-Methods for automatically selecting the ``best'' choice of independent state
-variables are discussed in \cite{Reckdahl1996}; they involve computing
-the SVD of the Jacobian of the constraint matrices to determine a set of
-indepedent speeds which will ensure the nonsingularity of the aforementioned
-matrices.
+P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
+and constant parameters; for certain configurations or parameters, it may be
+that the choice of independent state variables cannot be arbitrary.  While some
+systems may permit a choice of independent state variables which are valid for
+all configurations of interest, others may not.  Methods for automatically
+selecting the ``best'' choice of independent state variables are discussed in
+\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
+constraint matrices to determine a set of indepedent speeds which will ensure
+the nonsingularity of the aforementioned matrices.
 
-% Not sure which bibliography style we are supposed to use
-%\bibliographystyle{spphys}       % APS-like style for physics
-%\bibliographystyle{spbasic}      % basic style, author-year citations
+This algorithm as been applied to two well studied systems: the rolling disc
+and an extended version of the Whipple bicycle model\cite{Meijaard2007}.
+Detailed model description, derivation, and benchmarking results are presented
+in the electronic supplementary material.
+
 \section{Conclusion}
-We have presented a procedure for creating linear equations of
-motion for a system with holonomic and nonholonomic constraints. We have also
-described some of the considerations that must be made when following this
-procedure as well validating it with examples.
+We have presented a procedure for creating linear equations of motion for
+systems with holonomic and nonholonomic constraints. We have also described
+some of the considerations that must be made when following this procedure as
+well validating it with examples.
 
@@
         \ref{table:assumptions}.
-    \item Compute the quantities shown in equation (\ref{eq:quant_to_compute}).
-    \item Identify independent coordinates and speeds.
-    \item Form the quantities in equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
+    \item Determine a point of linearization ($\bm{q}^*$, $\dot{\bm{q}}^*$,
+      $\bm{u}^*$, $\dot{\bm{u}}^*$, and $\bm{r}^*$) which satisfies the
+      equations in Table \ref{table:assumptions}.
+    \item Compute the quantities in equation (\ref{eq:quant_to_compute}).
+    \item Identify independent coordinates and speeds; form equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
         (\ref{eq:C_2}).
-    \item From the results of steps 2 and 4, for equation (\ref{eq:state_space_constrained}).
+    \item Form equation (\ref{eq:state_space_constrained}), and if needed,
+      equations (\ref{eq:A_prime})-(\ref{eq:B}).
 \end{enumerate}
@@
   Foundation under award 0928339 and two Google Summer of Code projects (2009,
-  2011).  Aaron Muerer, Brian Granger, Mateuz Paprocki, and Ondrej Certik
-  provided valuable help with SymPy and Python.  Jason Moore, Thomas Johnston,
-  Evan Sperber, and Andrew Kickertz provided valuable feedback during
-  discussions of multibody dynamics and control related issues.
+  2011).  Jason Moore, Thomas Johnston, Evan Sperber, and Andrew Kickertz
+  provided valuable feedback during discussions of multibody dynamics and
+  control related issues.
 \end{acknowledgements}
 
+% Not sure which bibliography style we are supposed to use
+%\bibliographystyle{spphys}       % APS-like style for physics
+%\bibliographystyle{spbasic}      % basic style, author-year citations
 \bibliography{references}   % name your BibTeX data base

### 2012-04-05 PetersonGedeHubbard2012.tex
Minor edits.

@@
   coordinates/speeds (both independent and dependent) and independent
-  generalized coordinates/speeds and exogenous inputs.  These linear ODE's
+  generalized coordinates/speeds and exogenous inputs.  These linear ODEs
   effectively embed the constraints, and enable stability analysis and control
@@
 
-%The system described by equations \ref{eq:configuration}-\ref{eq:dyndiffs} has
-%$o-m$ velocity degrees of freedom.
 Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
@@
 \cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
-constraint matrices to determine a set of indepedent speeds which will ensure
-the nonsingularity of the aforementioned matrices.
+constraint matrices to determine a set of independent speeds which will ensure
+the non-singularity of the aforementioned matrices.
 
@@
   provided valuable feedback during discussions of multibody dynamics and
-  control related issues.
+  control.
 \end{acknowledgements}

### 2012-04-05 PetersonGedeHubbard2012.tex
More minor edits.

@@
 $\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
-state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$.  The standard
-linear systems analyses may now be applied.  It is worth noting that the rows
+state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
+linear systems analyses may applied.  It is worth noting that the rows
 of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
-in the output or measurement equations of a linear state space model, as would
-be the case when the control task is to control a dependent state or when
-measurements of dependent state(s) are made.
+in the output or measurement equations of a linear state space model, as in the
+case when the controlling or measuring dependent state(s).
 

### 2012-04-07 rolling_disc_luke.py
Luke's version of rolling disc derivation.

@@ -1,0 +1,81 @@
+from sympy import symbols, Matrix
+from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point, dot,
+cross, mprint, RigidBody, inertia, Kane)
+
+# Symbols for time and constant parameters
+t, r, m, g, I, J = symbols('t r m g I J')
+
+# Configuration variables and their time derivatives
+# q[0] -- yaw
+# q[1] -- lean
+# q[2] -- spin
+# q[3] -- disc center distance from inertial origin, x direction
+# q[4] -- disc center distance from inertial origin, y direction
+# q[5] -- disc center distance from inertial origin, z direction
+q = dynamicsymbols('q:6')
+qd = [qi.diff(t) for qi in q]
+
+# Generalized speeds and their time derivatives
+# u[0] -- disc angular velocity component, disc fixed x direction
+# u[1] -- disc angular velocity component, disc fixed y direction
+# u[2] -- disc angular velocity component, disc fixed z direction
+# u[3] -- disc velocity component, disc fixed x direction
+# u[4] -- disc velocity component, disc fixed y direction
+# u[5] -- disc velocity component, disc fixed z direction
+u = dynamicsymbols('u:6')
+ud = [ui.diff(t) for ui in u]
+
+# Reference frames
+N = ReferenceFrame('N')                     # Inertial reference frame
+A = N.orientnew('A', 'Axis', [q[0], N.z])   # Yaw intermediate frame
+B = A.orientnew('B', 'Axis', [q[1], A.x])   # Lean intermediate frame
+C = B.orientnew('C', 'Axis', [q[2], B.y])   # Disc fixed frame
+
+# Inertial angular velocity and angular acceleration of disc fixed frame
+C.set_ang_vel(N, u[0]*C.x + u[1]*C.y + u[2]*C.z)
+C.set_ang_acc(N, ud[0]*C.x + ud[1]*C.y + ud[3]*C.z)
+
+# Points
+P = Point('P')                              # Ground disc contact point
+O = P.locatenew('O', -r*B.z)                # Center of disc
+
+# Configuration constraint and its Jacobian w.r.t. q
[…]

### 2012-04-13 rolling_disc.py
Fixed error in inertia and added a print statement.

@@ -23,7 +23,7 @@ C.v2pt_theory(O, N, R)
 ccons = [r * (L.z & Y.z) - oz]
 vcons = [C.vel(N) & Y.x, C.vel(N) & Y.y, C.vel(N) & Y.z]
 
-Disk = RigidBody('Disk', O, R, m, (inertia(L, I, I, J), O))
+Disk = RigidBody('Disk', O, R, m, (inertia(L, I, J, I), O))
 
 v_o_exp_n = O.vel(N).express(N)
 
@@ -40,6 +40,7 @@ KM.kindiffeq(kd)
 KM.kanes_equations(Forces, Bodies)
 
 mm = KM.mass_matrix_full
+mprint(mm)
 f = KM.forcing_full
 f.simplify()
 f = f.subs(KM.kindiffdict())

### 2012-04-13 rolling_disc_luke.py
Added manual derivation.

@@ -1,4 +1,4 @@
-from sympy import symbols, Matrix
+from sympy import symbols, Matrix, eye, zeros
 from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point, dot,
 cross, mprint, RigidBody, inertia, Kane)
 
@@ -33,13 +33,13 @@ C = B.orientnew('C', 'Axis', [q[2], B.y])   # Disc fixed frame
 
 # Inertial angular velocity and angular acceleration of disc fixed frame
 C.set_ang_vel(N, u[0]*C.x + u[1]*C.y + u[2]*C.z)
-C.set_ang_acc(N, ud[0]*C.x + ud[1]*C.y + ud[3]*C.z)
+C.set_ang_acc(N, ud[0]*C.x + ud[1]*C.y + ud[2]*C.z)
 
 # Points
 P = Point('P')                              # Ground disc contact point
 O = P.locatenew('O', -r*B.z)                # Center of disc
 
-# Configuration constraint and its Jacobian w.r.t. q
+# Configuration constraint and its Jacobian w.r.t. q        (Table 1)
 f_c = Matrix([q[5] - dot(O.pos_from(P), N.z)])
 f_c_dq = f_c.jacobian(q)
 
@@ -47,13 +47,13 @@ f_c_dq = f_c.jacobian(q)
 O.set_vel(N, u[3]*C.x + u[4]*C.y + u[5]*C.z)
 O.set_acc(N, O.vel(N).diff(t, C) + cross(C.ang_vel_in(N), O.vel(N)))
 
-# Velocity level constraints
+# Velocity level constraints                                (Table 1)
 v_contact_point = O.vel(N) + cross(C.ang_vel_in(N), P.pos_from(O))
 f_v = Matrix([dot(v_contact_point, uv) for uv in C])
 f_v_dq = f_v.jacobian(q)
 f_v_du = f_v.jacobian(u)
 
-# Acceleration level constraints
+# Acceleration level constraints                            (Table 1)
 f_a = f_v.diff(t)
 
 # Disc angular velocity in N expressed using time derivatives of coordinates
@@ -64,11 +64,89 @@ v_o_n_qd = qd[3]*N.x + qd[4]*N.y + qd[5]*N.z
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in C] + 
[…]

### 2012-04-13 PetersonGedeHubbard2012.tex
Implemented some of Mont's suggested changes.

@@
 \usepackage{bm}
-
-%% Big-O notation.
-\providecommand{\OO}[1]{\ensuremath{\operatorname{O}\bigl(#1\bigr)}}
 
@@
 tasks which require a linear state space formulation.  Common examples include
-stability analysis, controller and estimator design (including control or
+stability analysis and controller and estimator design (including control or
 measurement of a dependent quantity).  The formalism presented here is generic
@@
 
-The complete configuration of the system is assumed to be described by
+The complete configuration of the system is further assumed to be described by
 generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
@@
 constraints.  The fourth equation represents kinematic differential equations
-which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both of these
+which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both these
 quantities.  The fifth equation represents the constraint-free nonholonomic
@@
 (e.g., total mechanical energy) may restrict trajectories to a lower
-dimensional manifold.  In general this cannot be assumed, especially
-in the presence of external control inputs $\bm{r}$ which may change the total
+dimensional manifold.  In general this cannot be assumed, however, especially
+in the presence of exogenous inputs $\bm{r}$ which may change the total
 mechanical energy of the system.  While more than $p$ of the independent state
@@
 possible $n + o$ may be obvious for small systems where there is intuition
-about the system behavior, but in general the optimal choice depends on how the
+about system behavior, but in general the optimal choice depends on how the
 configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
-hand, in systems with cyclic/ignorable coordinates (i.e., coordinates which do
-not appear in the dynamic equations of motion), fewer than $p$ state
+hand, in systems with cyclic (ignorable) coordinates, fewer than $p$ state
 variables may be of interest; one example is the Whipple bicycle model where
-$p=10$ but 5 of these state variables are cyclic coordinates (coordinates of
-rear wheel contact point, heading, and both wheel angles) and may only be of
-secondary interest.
+$p=10$ but 5 of these state variables are cyclic coordinates (Cartesian
+coordinates of rear wheel contact point, heading, and both wheel angles) and
+may only be of secondary interest.
 
@@
 
-We see in Table \ref{table:assumptions} the differential equations which
-describe our system. The goal of this section is to create linear equations
-of motion with the following structure:
+The system, is desribed (in part) by the kinematic and dynamic differential
+equations in Table \ref{table:assumptions} .  The goal of this section is to
+create linear equations of motion with the following structure:
 \begin{align}
@@
 \end{align}
-For convenience, we define
+We define
 \begin{equation}
@@
 \end{align}
-We now have a linear system of equations which relates the time derivatives of all
-states (dependent and independent) to the independent states and exogenous inputs.
+Equation \ref{eq:state_space_constrained} is a linear system of equations which
+relates the time derivatives of all states (dependent and independent) to the
+independent states and exogenous inputs; this is the precisely the form of
+equation \ref{eq:structure}.
 
@@
 
-The choice on which state variables are taken to be independent may determine
+The choice of which state variables are taken to be independent may determine
 whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}

### 2012-04-21 PetersonGedeHubbard2012.tex
Major rearrangment of paper sections.

@@
-\documentclass{svjour3}                     % onecolumn (standard format)
+\documentclass[smallcondensed]{svjour3}                     % onecolumn (standard format)
 \RequirePackage{fix-cm}
@@
 \usepackage{bm}
+%\usepackage[polutonikogreek]{babel}
 
@@
 
+\section{Conclusions}
+\label{sec:conclusions}
+We have provided a procedure to obtain linear state space formulation for
+constrained multibody systems whose nonlinear dynamics are described only by
+ordinary differential equations. With this state space formulation, you can
+perform stability analysis, control system design, and other linear analyses.
+
+In some sense this work is simply applying the chain rule through matrix
+operations in a way that respects system constraints. An advanced C.A.S. system
+which possesses the ability to represent dependent quantities (as functions of
+independent quantities) could perform the chain rule directly, thereby
+eliminating the need for matrix operations. We leave an open question: how does
+computational performance depend on implementation details and problem size?
+
+%In summary, the procedure is:
+%\begin{enumerate}
+%    \item Describe the system in the form of the equations shown in Table
+%        \ref{table:assumptions}.
+%    \item Determine a point of linearization ($\bm{q}^*$, $\dot{\bm{q}}^*$,
+%      $\bm{u}^*$, $\dot{\bm{u}}^*$, and $\bm{r}^*$) which satisfies the
+%      equations in Table \ref{table:assumptions}.
+%    \item Compute the quantities in equation (\ref{eq:quant_to_compute}).
+%    \item Identify independent coordinates and speeds; form equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
+%        (\ref{eq:C_2}).
+%    \item Form equation (\ref{eq:state_space_constrained}), and if needed,
+%      equations (\ref{eq:A_prime})-(\ref{eq:B}).
+%\end{enumerate}
+
 \section{Introduction}
-\label{intro}
+\label{sec:intro}
+Multibody systems, rigid or flexible, and particles often have constraints
+which limit how they can move. In this paper, we will discuss 3 types of
+constraints: configuration, velocity, and acceleration. Configuration
+constraints limit the location or orientation of parts of the system, relative
+to the external world or other parts of the system. Velocity constraints limit
+the speeds in which the configuration can change, either from configuration
+constraints (through time differentiation) or independent application. In this
+paper, acceleration constraints refer to time differentiated velocity
+constraints. Examples of multibody systems with constraints are closed
+kinematic loops or objects rolling without slip.
 
-The goal of the following calculations is to concretely establish the first
-order relationships (about an arbitrary equilibrium point) between an arbitrary
-choice of independent coordinates and speeds, and the time-derivatives of all
-coordinates and speeds (both independent and dependent).  This is necessary for
-tasks which require a linear state space formulation.  Common examples include
-stability analysis and controller and estimator design (including control or
-measurement of a dependent quantity).  The formalism presented here is generic
-enough to cover most examples of time-varying constrained multibody systems
-with arbitrary external inputs and arbitrary specified quantities.  This
-procedure has been created for systems whose equations of motion have been
-derived with Kane's Method.  While this is not a strict requirement, systems
-which can be completely described by a set of ODEs is (i.e., no DAEs).
+For these systems, a linear form of the equations of motion will be necessary
+for many different analyses.  Stability analysis and control system design are
+two possible scenarios.  Linearization of these constrained systems is not
+straightforward and presents significant challenges. Simply computing the
+jacobian matrix of the differential equations will give incorrect results, with
+no immediate indication that the results are wrong.
 
-The complete configuration of the system is further assumed to be described by
-generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
-described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
-All coordinates and speeds, not only the independent ones, are represented by
-$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
-relationships described in Table \ref{table:assumptions}.  The system is
-described by $l + m + n + o$ equations which relate $2n + 2o + s + 1$
-quantities, which implies that $n - l + o - m + s + 1$ of these quantities are
-independent.  This paper discusses the first order relationship between these
-independent quantities ($n-l$ independent coordinates, $o-m$ independent
-speeds, $s$ control inputs, and time $t$) and the time derivatives of the $n$
-coordinates and $o$ speeds.
+Currently, there are only a few approaches to linearizing constrained systems
+which have been presented (citations here). A limitation of these approaches is
+that they have been developed for equations of motion found using Lagrange¿s
+method, which contain Lagrange multipliers. Kane¿s method (citation here)
+allows for formulation of equations of motion for constrained systems in which
+no multipliers are present. The dynamics of these systems can be completely
+described by ordinary differential equations. This significantly different
+formulation requires a different approach to linearization than what the
+existing options provide.
+
+The goal of the following paper is to concretely establish the first order
+relationships (about an arbitrary equilibrium point) between a selection of
+independent coordinates and speeds and the time-derivatives of all of the
+coordinates and speeds (both independent and dependent). The formalism
+presented here is generic enough to cover most examples of time-varying
+constrained multibody systems with arbitrary external inputs and arbitrary
+specified quantities.  This procedure was created for systems whose equations
+of motion have been derived with Kane¿s Method. Although also applicable to
+dynamic system equations formulated using other methods, it is restricted to
+systems which can be completely described by a set of ODEs (i.e., no DAEs are
+present).
+
+We first present our linearization procedure. Next we will offer some
+discussion of the procedure, along with an example. The derivation of the
+procedure is provided in the appendix. Additionally, the electronic
+supplementary material provides more examples (in more detail) and an
+alternative ordering for the independent/dependent quantities.
+
+%The complete configuration of the system is further assumed to be described by
+%generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
+%described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
+%All coordinates and speeds, not only the independent ones, are represented by
+%$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
+%relationships described in Table \ref{table:assumptions}.  The system is
+%described by $l + m + n + o$ equations which relate $2n + 2o + s + 1$
+%quantities, which implies that $n - l + o - m + s + 1$ of these quantities are
+%independent.  This paper discusses the first order relationship between these
+%independent quantities ($n-l$ independent coordinates, $o-m$ independent
+%speeds, $s$ control inputs, and time $t$) and the time derivatives of the $n$
+%coordinates and $o$ speeds.
+%
+%The first three equations in Table \ref{table:assumptions} represent
+%constraints derived purely from kinematic considerations at the configuration,
+%velocity, and acceleration levels, respectively.  The velocity constraints may
+%be nonholonomic or time differentiated holonomic.  The acceleration constraints
+%can be time differentiated velocity constraints or kinematic acceleration
+%constraints.  The fourth equation represents kinematic differential equations
+%which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both these
+%quantities.  The fifth equation represents the constraint-free nonholonomic
+%dynamic differential equations~\cite{Kane1985}, and will necessarily be linear
+%in $\bm{\dot{u}}$ and the inputs $\bm{r}$, but generally nonlinear in all other
+%quantities.  This formalism is exceedingly generic; in practice, the structure
+%of the functions in Table \ref{table:assumptions} can be leveraged to simplify
+%the following calculations.  This will be discussed subsequently.
+%
+%Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
+%manifold embedded in a $n + o$ dimensional space, though conserved quantities
+%(e.g., total mechanical energy) may restrict trajectories to a lower
+%dimensional manifold.  In general this cannot be assumed, however, especially
+%in the presence of exogenous inputs $\bm{r}$ which may change the total
+%mechanical energy of the system.  While more than $p$ of the independent state
+%variables may be of interest (i.e., dependent coordinates or speeds may be of
+%interest in themselves), only $p$ quantities are truly independent.  Which of
+%these $p$ coordinates and speeds should be chosen as independent out of the
+%possible $n + o$ may be obvious for small systems where there is intuition
+%about system behavior, but in general the optimal choice depends on how the
+%configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
+%hand, in systems with cyclic (ignorable) coordinates, fewer than $p$ state
+%variables may be of interest; one example is the Whipple bicycle model where
+%$p=10$ but 5 of these state variables are cyclic coordinates (Cartesian
+%coordinates of rear wheel contact point, heading, and both wheel angles) and
+%may only be of secondary interest.
+
+\section{Results}
+\label{sec:results}
+The linearized equations of motion are
+\begin{align}
+  \label{eq:state_space_constrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q}_i \\
+        \delta \bm{u}_i
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Table \ref{table:assumptions} provides the form in which the nonlinear
+equations of motion and constraints must be arranged.  Equations
+\ref{eq:quant_to_compute} provides the bulk of the terms in equation
+\ref{eq:state_space_constrained}; these terms are Jacobian matrices of the
+differential equations. Equations \ref{eq:C_0}-\ref{eq:C_2} provide the terms
+$C_0$, $C_1$, and $C_2$ which embed the constraints into
+\ref{eq:state_space_constrained}. Finally, the permutation matrices in
+equations \ref{eq:Pq} and \ref{eq:Pu} provide the means to separate the
+independent and dependent speeds.
+
 \begin{table}[htbp]
@@
 
-The first three equations in Table \ref{table:assumptions} represent
-constraints derived purely from kinematic considerations at the configuration,
-velocity, and acceleration levels, respectively.  The velocity constraints may
-be nonholonomic or time differentiated holonomic.  The acceleration constraints
-can be time differentiated velocity constraints or kinematic acceleration
-constraints.  The fourth equation represents kinematic differential equations
-which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both these
-quantities.  The fifth equation represents the constraint-free nonholonomic
-dynamic differential equations~\cite{Kane1985}, and will necessarily be linear
-in $\bm{\dot{u}}$ and the inputs $\bm{r}$, but generally nonlinear in all other
-quantities.  This formalism is exceedingly generic; in practice, the structure
-of the functions in Table \ref{table:assumptions} can be leveraged to simplify
-the following calculations.  This will be discussed subsequently.
+\begin{align}
+\label{eq:quant_to_compute}
+  \begin{array}{llcll}
+    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
+    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
+    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
+    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
+    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
+    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
+    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
+    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
+    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
+    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
+    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
+  \end{array}
+\end{align}
 
-Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
-manifold embedded in a $n + o$ dimensional space, though conserved quantities
-(e.g., total mechanical energy) may restrict trajectories to a lower
-dimensional manifold.  In general this cannot be assumed, however, especially
-in the presence of exogenous inputs $\bm{r}$ which may change the total
-mechanical energy of the system.  While more than $p$ of the independent state
-variables may be of interest (i.e., dependent coordinates or speeds may be of
-interest in themselves), only $p$ quantities are truly independent.  Which of
-these $p$ coordinates and speeds should be chosen as independent out of the
-possible $n + o$ may be obvious for small systems where there is intuition
-about system behavior, but in general the optimal choice depends on how the
-configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
-hand, in systems with cyclic (ignorable) coordinates, fewer than $p$ state
-variables may be of interest; one example is the Whipple bicycle model where
-$p=10$ but 5 of these state variables are cyclic coordinates (Cartesian
-coordinates of rear wheel contact point, heading, and both wheel angles) and
-may only be of secondary interest.
+\begin{align}
+  \label{eq:C_0}
+  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
+  \label{eq:C_1}
+  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \\
+  \label{eq:C_2}
+  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui}\\
+  \label{eq:Pq}
+  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
+  \label{eq:Pu}
+  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
+\end{align}
+
+\section{Discussion}
+\label{sec:discussion}
+Equation \ref{eq:state_space_constrained} doesn't fit into the standard linear
+system framework because it is non-square.  To obtain a square system (useful
+in stability analysis and control system design) we define
+\begin{align}
+  \label{eq:A_prime}
+    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+  \label{eq:B_prime}
+    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
+\end{align}
+where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
+\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
+independent states by defining
+\begin{align}
+  \label{eq:P_prime}
+    P^\prime &\triangleq \begin{bmatrix}
+        P_{qi} & \bm{O}_{n \times (o - m)} \\
+        \bm{O}_{o \times (n - l)} & P_{ui}
+    \end{bmatrix} \\
+  \label{eq:A}
+    A &\triangleq P^{\prime T} A^\prime \\
+  \label{eq:B}
+    B &\triangleq P^{\prime T} B^\prime
+\end{align}
+where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
+$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
+state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
+linear systems analyses may applied.  It is worth noting that the rows
+of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
+in the output or measurement equations of a linear state space model, as in the
+case when the controlling or measuring dependent state(s).
+
+The choice of which state variables are taken to be independent may determine
+whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
+and constant parameters; for certain configurations or parameters, it may be
+that the choice of independent state variables cannot be arbitrary.  While some
+systems may permit a choice of independent state variables which are valid for
+all configurations of interest, others may not.  Methods for automatically
+selecting the ``best'' choice of independent state variables are discussed in
+\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
+constraint matrices to determine a set of independent speeds which will ensure
+the non-singularity of the aforementioned matrices.
+
+This algorithm as been applied to two well studied systems: the rolling disc
+and an extended version of the Whipple bicycle model\cite{Meijaard2007}.
+Detailed model description, derivation, and benchmarking results are presented
+in the electronic supplementary material.
+
+\section{Example}
+\label{example}
+We apply this linearization technique to the well studied rolling disc.  Figure
+XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
+assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
+transverse inertia $I$.  To illustrate the concept of dependent quantities, we
+purposefully choose a nonminimal set of coordinates and speeds.  In this
+particular example, this can be avoided, but in more complicated systems, it is
+difficult (if not impossible) to describe the system with a minimal choice.  The
+orientation of the disc fixed frame $C$ relative to the inertial frame $N$ is given by
+a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw), $q_1$
+(lean), $q_2$ (spin).  The location
+of the center of disc relative to the inertial origin is defined by coordinates
+q3, q4, q5.  The angular velocity of the disc is defined by three speeds
+\[u_0 \triangleq {^{N}}\bm{\omega}^{C} \cdot \bm{c}_{x}\]
 
@@
 \ref{table:assumptions}), and introducing the following definitions
-\begin{align}
-    \label{eq:quant_to_compute}
-    \begin{array}{llcll}
-\tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
-\tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
-\tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
-\tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
-\tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
-\tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-\tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
-\tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
-\tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
-\tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
-\tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
-\tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
-\end{array}
-\end{align}
 enables the unconstrained linear state space equations to be written as
@@
 We define
-\begin{equation}
-  \label{eq:C_0}
-  C_0 \triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}
-\end{equation}
 which enables us to write
@@
 We define
-\begin{align}
-  \label{eq:C_1}
-  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \\
-  \label{eq:C_2}
-  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui}
-\end{align}
 which enables us to write
@@
 rewrite equation (\ref{eq:state_space_unconstrained}) as
-\begin{align}
-  \label{eq:state_space_constrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q}_i \\
-        \delta \bm{u}_i
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
+%\begin{align}
+%  \label{eq:state_space_constrained}
+%  \left[
+%    \begin{array}{cc}
+%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+%      \tilde{M}_{uqd} & \tilde{M}_{uud}
+%    \end{array}
+%    \right]
+%    \left[
+%      \begin{array}{c}
+%        \delta \bm{\dot{q}} \\
+%        \delta \bm{\dot{u}}
+%      \end{array}
+%    \right]
+%   &=
+%   \left[
+%     \begin{array}{cc}
+%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+%     \end{array}
+%   \right]
+%    \left[
+%      \begin{array}{c}
+%        \delta \bm{q}_i \\
+%        \delta \bm{u}_i
+%      \end{array}
+%    \right]
+%    +
+%    \left[
+%      \begin{array}{c}
+%        \bm{0}_{(n+m) \times s} \\
+%        \tilde{B}_{u}
+%      \end{array}
+%    \right]
+%    \delta \bm{r}
+%\end{align}
 Equation \ref{eq:state_space_constrained} is a linear system of equations which
@@
 equation \ref{eq:structure}.
-
-\section{Discussion}
-\label{sec:discussion}
-Equation \ref{eq:state_space_constrained} doesn't fit into the standard linear
-system framework because it is non-square.  To obtain a square system (useful
-in stability analysis and control system design) we define
-\begin{align}
-  \label{eq:A_prime}
-    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
-  \label{eq:B_prime}
-    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
-\end{align}
-where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
-\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
-independent states by defining
-\begin{align}
-  \label{eq:P_prime}
-    P^\prime &\triangleq \begin{bmatrix}
-        P_{qi} & \bm{O}_{n \times (o - m)} \\
-        \bm{O}_{o \times (n - l)} & P_{ui}
-    \end{bmatrix} \\
-  \label{eq:A}
-    A &\triangleq P^{\prime T} A^\prime \\
-  \label{eq:B}
-    B &\triangleq P^{\prime T} B^\prime
-\end{align}
-where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
-$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
-state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
-linear systems analyses may applied.  It is worth noting that the rows
-of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
-in the output or measurement equations of a linear state space model, as in the
-case when the controlling or measuring dependent state(s).
-
-The choice of which state variables are taken to be independent may determine
-whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
-and constant parameters; for certain configurations or parameters, it may be
-that the choice of independent state variables cannot be arbitrary.  While some
-systems may permit a choice of independent state variables which are valid for
-all configurations of interest, others may not.  Methods for automatically
-selecting the ``best'' choice of independent state variables are discussed in
-\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
-constraint matrices to determine a set of independent speeds which will ensure
-the non-singularity of the aforementioned matrices.
-
-This algorithm as been applied to two well studied systems: the rolling disc
-and an extended version of the Whipple bicycle model\cite{Meijaard2007}.
-Detailed model description, derivation, and benchmarking results are presented
-in the electronic supplementary material.
-
-\section{Conclusion}
-We have presented a procedure for creating linear equations of motion for
-systems with holonomic and nonholonomic constraints. We have also described
-some of the considerations that must be made when following this procedure as
-well validating it with examples.
-
-In summary, the procedure is:
-\begin{enumerate}
-    \item Describe the system in the form of the equations shown in Table
-        \ref{table:assumptions}.
-    \item Determine a point of linearization ($\bm{q}^*$, $\dot{\bm{q}}^*$,
-      $\bm{u}^*$, $\dot{\bm{u}}^*$, and $\bm{r}^*$) which satisfies the
-      equations in Table \ref{table:assumptions}.
-    \item Compute the quantities in equation (\ref{eq:quant_to_compute}).
-    \item Identify independent coordinates and speeds; form equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
-        (\ref{eq:C_2}).
-    \item Form equation (\ref{eq:state_space_constrained}), and if needed,
-      equations (\ref{eq:A_prime})-(\ref{eq:B}).
-\end{enumerate}
 

### 2012-04-21 PetersonGedeHubbard2012.tex
Finished putting in rolling disc equations.

@@
 transverse inertia $I$.  To illustrate the concept of dependent quantities, we
-purposefully choose a nonminimal set of coordinates and speeds.  In this
-particular example, this can be avoided, but in more complicated systems, it is
-difficult (if not impossible) to describe the system with a minimal choice.  The
-orientation of the disc fixed frame $C$ relative to the inertial frame $N$ is given by
-a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw), $q_1$
-(lean), $q_2$ (spin).  The location
-of the center of disc relative to the inertial origin is defined by coordinates
-q3, q4, q5.  The angular velocity of the disc is defined by three speeds
-\[u_0 \triangleq {^{N}}\bm{\omega}^{C} \cdot \bm{c}_{x}\]
+purposefully choose to describe the configuration and velocity of the disc with
+a non-minimal set of coordinates and speeds.  This choice is not purely
+academic; a physical disc equipped with a rate gyroscope and accelerometer
+measure these body fixed quantities (or their time derivatives) directly,
+rather than the more traditional set which is chosen for analytic convenience.
+In more complicated systems with constraints, it is difficult (if not
+impossible) to describe the system with a minimal choice of coordinates and
+speeds; the ideas demonstrated by this simply example apply analogously to more
+complicated ones.
 
+The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
+is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
+$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
+of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
+$C^*$ in $N$ are
+\begin{align}
+  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
+  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
+  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
+\end{align}
+If the disc is to make contact with the ground, it should be obvious to the
+reader that there is a configuration constraint involving the height $q_5$ of
+the disc center and the lean $q_1$.  Similarly, if the disc rolls without
+slip, it should be obvious that there are constraints which relate the
+generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
+constraints and the equations of motion are presented in the electronic
+supplementary material and can be found in the literature (reference O'Reilly
+and others?).
+\begin{table}[htbp]
+  \centering
+  \caption{Rolling disc governing definitions and equations}
+  \label{table:rollingdisc}
+  \begin{tabular}[c]{l l l}
+    Quantity & Shape & Description\\
+    \hline
+    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
+    derivatives\\
+    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
+  \end{tabular}
+  \begin{tabular}[c]{r @{ $=$ } l l}
+    \multicolumn{3}{c}{ } \\
+    \multicolumn{2}{c}{Equation} &           \\
+    \hline
+    $rc_1 + q_5$ & $0$ & $\bm{f}_c = \bm{0}$ \\
+    $\left[\begin{array}{c}
+        r u_1 c_2 + u_3 \\
+        -r (u_0 c_2 + u_2 s_2) + u_4 \\
+        r u_1 s_2 + u_5
+    \end{array}\right] $ &
+    $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$
+      & $\bm{f}_v = \bm{0}$\\
+    $\left[\begin{array}{c}
+        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
+        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
+      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
+      \end{array}\right]$ &
+      $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$ & $\bm{f}_a =
+        \bm{0}$\\
+        $\left[\begin{array}{c}
+-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
+\dot{q}_0 s_1 + \dot{q}_2 \\
+\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
+\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
++ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
+- \dot{q}_5 s_2 c_1\\
+- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
+\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
+         \end{array}\right] + \left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
+             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ &
+             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{array}\right]$
+                 & $\bm{f}_0 + \bm{f}_1 = \bm{0}$\\
+                 $\left[\begin{array}{c}
+-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
+-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
+-I \dot{u}_2 - m r s_2 \dot{u}_4
+                   \end{array}\right]
++
+                 \left[\begin{array}{c}
+-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
+                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
+ I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
+                   \end{array}\right]$ &
+             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \end{array}\right]$
+               & $\bm{f}_2 + \bm{f}_3 = \bm{0}$
+  \end{tabular}
+\end{table}
+
+%f_c:
+%[r*cos(q1) + q5]
+%f_v:
+%[                r*u1*cos(q2) + u3]
+%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
+%[                r*u1*sin(q2) + u5]
+%f_a:
+%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
+%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
+%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
+%f_0:
+%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
+%[                                                                                                       sin(q1)*q0' + q2']
+%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
+%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
+%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
+%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
+%f_1:
+%[-u0]
+%[-u1]
+%[-u2]
+%[-u3]
+%[-u4]
+%[-u5]
+%f_2:
+%[                  -I*u0' - m*r*cos(q2)*u4']
+%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
+%[                  -I*u2' - m*r*sin(q2)*u4']
+%f_3:
+%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
+%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
+%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]
 \section{Derivation of linearization algorithm}

### 2012-04-21 rolling_disc_luke.py

@@ -1,6 +1,6 @@
-from sympy import symbols, Matrix, eye, zeros
+from sympy import symbols, Matrix, eye, zeros, latex
 from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point, dot,
-cross, mprint, RigidBody, inertia, Kane)
+cross, mprint, RigidBody, inertia, Kane, mlatex)
 
 # Symbols for time and constant parameters
 t, r, m, g, I, J = symbols('t r m g I J')
@@ -140,12 +140,22 @@ for i in [0, 1, 2]:
     # All other terms that don't have du/dt in them
     f_3[i] = gif_con[i].subs({udi : 0 for udi in ud}) + gaf_con[i]
 
-mprint(f_2)
-mprint(f_3)
-mprint(Muud)
+print("f_c:")
+mprint(f_c)
+print("f_v:")
+mprint(f_v)
+print("f_a:")
 mprint(f_a)
-
+print("f_0:")
+mprint(f_0)
+print("f_1:")
+mprint(f_1)
+print("f_2:")
+mprint(f_2)
+print("f_3:")
+mprint(f_3)
 stop
+
 # Kane's dynamic equations via sympy.physics.mechanics
 Bodies_List = [RigidBody('Disk', O, C, m, (inertia(C, I, J, I), O))]
 Forces_List = [(O, m*g*A.z)]

### 2012-04-21 PetersonGedeHubbard2012.tex
Reworked the abstract.

@@
 \begin{abstract}
-  We present a systematic procedure for linearizing equations of motion
-  governing constrained multibody systems which have been derived with Kane's
-  method.  The procedure avoids differential algebraic equations and yields
-  linear relationships between time derivatives of generalized
-  coordinates/speeds (both independent and dependent) and independent
-  generalized coordinates/speeds and exogenous inputs.  These linear ODEs
-  effectively embed the constraints, and enable stability analysis and control
-  system design without the need to explicitly address the constraints.  We
-  numerically validate the results on a well studied model of a bicycle which
-  includes both configuration and velocity constraints.
+% Context
+Many common mechanical systems have configuration or velocity constraints.
+% Need  (what we have and what we want)
+Analyses of such systems often require linearized forms of the motion
+equations.  The derivation of the nonlinear motion equations is well
+understood, but very little has been presented on the proper approach to
+forming linearized versions which satisfy the constraints.
+% Task
+To address this issue, we formed a procedure of organizing the constraints
+and motion equations, and subsequently linearizing these equations.
+% Object of the document
+In this paper we present this procedure to correctly linearize the
+nonlinear motion equations of constrained multibody systems and is illustrated
+by a familiar example.
+% Findings
+We applied this procedure to a model of a bicycle which includes a closed
+kinematic loop and four nonholonomic constraints and found that the resulting
+symbolic linearized motion equations are identical to previously published
+results.  The technique can be performed symbolically (enabling explict
+analytical forms of the linearized dynamics) or numerically.
+% Conclusion
+Motion equations for constrained multibody systems can now be linearized in a
+systematic fashion, symbolically or numerically, thereby facilitating common
+linear analyses (e.g., stability analysis and control system design).
+% Perspectives
+Efficiency of computer implementations of this procedure will need to be
+considered in future studies.
 \keywords{symbolic dynamics, control, constrained multibody systems}

### 2012-04-23 PetersonGedeHubbard2012.tex
Cleaned up derivation a little
also made a small change to the abstract

@@
 % Object of the document
-In this paper we present this procedure to correctly linearize the
-nonlinear motion equations of constrained multibody systems and is illustrated
-by a familiar example.
+In this paper we present this procedure to correctly linearize the nonlinear
+motion equations of constrained multibody systems, and illustrate it with a
+familiar example.
 % Findings
@@
 
-The system, is desribed (in part) by the kinematic and dynamic differential
-equations in Table \ref{table:assumptions} .  The goal of this section is to
-create linear equations of motion with the following structure:
-\begin{align}
-    \label{eq:structure}
-    \tilde{M} \dot{\bm{x}} = \tilde{A} \bm{x}_i + \tilde{B} \bm{r}
-\end{align}
-where $\bm{x}$ is the state vector of all the coordinates and speeds and
-$\bm{x}_i$ is the state vector of only the independent coordinates and speeds.
-This equation can be used for simulation, stability analysis,
-or it can be reformulated into the traditional state space equations for
-a variety of tasks; these topics are discussed in \ref{sec:discussion}
+This section outlines the derivation of the procedure for determining linear
+equations of motion \ref{eq:state_space_constrained}. This equation can be used
+for simulation, stability analysis, or it can be reformulated into the
+traditional state space equations for a variety of tasks; these topics are
+discussed in \ref{sec:discussion}.
 
@@
 (\ref{eq:acceleration_expansion}),
-and (\ref{eq:f2_plus_f3_expansion}) to zero (as per Table
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
 \ref{table:assumptions}), and introducing the following definitions
@@
     \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
-    \bm{q}_i
-\end{align}
-We define
-which enables us to write
-\begin{align}
+    \bm{q}_i\\
   \label{eq:delta_q}
@@
   &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i
-\end{align}
-We define
-which enables us to write
-\begin{align}
+    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
+  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
   \label{eq:delta_u}
-  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
-  &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
+  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
 \end{align}

### 2012-06-21 PetersonGedeHubbard2012.tex
changed table 2, fixed typos, add O'Reilly reference.

@@
 impossible) to describe the system with a minimal choice of coordinates and
-speeds; the ideas demonstrated by this simply example apply analogously to more
+speeds; the ideas demonstrated by this simple example can be applied to more
 complicated ones.
@@
 constraints and the equations of motion are presented in the electronic
-supplementary material and can be found in the literature (reference O'Reilly
-and others?).
+supplementary material and can be found in the literature \cite{O'Reilly1996}.
 \begin{table}[htbp]
@@
   \end{tabular}
-  \begin{tabular}[c]{r @{ $=$ } l l}
-    \multicolumn{3}{c}{ } \\
-    \multicolumn{2}{c}{Equation} &           \\
+  \begin{tabular}[c]{r @{ $=$ } l}
+%    \multicolumn{3}{c}{ } \\
+%    \multicolumn{2}{c}{Equation} &           \\
     \hline
-    $rc_1 + q_5$ & $0$ & $\bm{f}_c = \bm{0}$ \\
-    $\left[\begin{array}{c}
+    % Configuration constraint
+    $\bm{f}_c$ & $rc_1 + q_5$ \\
+    % Velocity constraints
+    $\bm{f}_v$ & $\left[\begin{array}{c}
         r u_1 c_2 + u_3 \\
@@
         r u_1 s_2 + u_5
-    \end{array}\right] $ &
-    $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$
-      & $\bm{f}_v = \bm{0}$\\
-    $\left[\begin{array}{c}
+    \end{array}\right] $ \\
+    % Acceleration constraints
+    $\bm{f}_a$ & $\left[\begin{array}{c}
         -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
@@
       r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
-      \end{array}\right]$ &
-      $\left[\begin{array}{c}$0$\\$0$\\$0$\end{array}\right]$ & $\bm{f}_a =
-        \bm{0}$\\
+      \end{array}\right]$ \\
+    % Kinematic differential equations
+    $\bm{f}_0$ &
         $\left[\begin{array}{c}
@@
 \dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
-         \end{array}\right] + \left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
-             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ &
-             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \\ 0 \\ 0 \\ 0 \end{array}\right]$
-                 & $\bm{f}_0 + \bm{f}_1 = \bm{0}$\\
-                 $\left[\begin{array}{c}
+         \end{array}\right]$ \\
+    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
+             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
+    % Dynamic differential equations
+  $\bm{f}_2$ & $\left[\begin{array}{c}
 -I \dot{u}_0 - m r c_2 \dot{u}_4 \\
@@
 -I \dot{u}_2 - m r s_2 \dot{u}_4
-                   \end{array}\right]
-+
-                 \left[\begin{array}{c}
+                   \end{array}\right]$ \\
+
+  $\bm{f}_3$ & $\left[\begin{array}{c}
 -I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
@@
  I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
-                   \end{array}\right]$ &
-             $\left[\begin{array}{c} 0 \\ 0 \\ 0 \end{array}\right]$
-               & $\bm{f}_2 + \bm{f}_3 = \bm{0}$
+                   \end{array}\right]$
   \end{tabular}
@@
 By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
-rewrite equation (\ref{eq:state_space_unconstrained}) as
-%\begin{align}
-%  \label{eq:state_space_constrained}
-%  \left[
-%    \begin{array}{cc}
-%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-%      \tilde{M}_{uqd} & \tilde{M}_{uud}
-%    \end{array}
-%    \right]
-%    \left[
-%      \begin{array}{c}
-%        \delta \bm{\dot{q}} \\
-%        \delta \bm{\dot{u}}
-%      \end{array}
-%    \right]
-%   &=
-%   \left[
-%     \begin{array}{cc}
-%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-%     \end{array}
-%   \right]
-%    \left[
-%      \begin{array}{c}
-%        \delta \bm{q}_i \\
-%        \delta \bm{u}_i
-%      \end{array}
-%    \right]
-%    +
-%    \left[
-%      \begin{array}{c}
-%        \bm{0}_{(n+m) \times s} \\
-%        \tilde{B}_{u}
-%      \end{array}
-%    \right]
-%    \delta \bm{r}
-%\end{align}
-Equation \ref{eq:state_space_constrained} is a linear system of equations which
-relates the time derivatives of all states (dependent and independent) to the
-independent states and exogenous inputs; this is the precisely the form of
-equation \ref{eq:structure}.
+rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
+\ref{eq:state_space_constrained}, a linear system of equations which relates
+the time derivatives of all states (dependent and independent) to the
+independent states and exogenous inputs.
 

### 2012-06-22 PetersonGedeHubbard2012.tex
Added rolling disc figure.

@@
 \usepackage{bm}
+\usepackage{pgf}
 %\usepackage[polutonikogreek]{babel}
@@
 Currently, there are only a few approaches to linearizing constrained systems
-which have been presented (citations here). A limitation of these approaches is
-that they have been developed for equations of motion found using Lagrange¿s
-method, which contain Lagrange multipliers. Kane¿s method (citation here)
+which have been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these approaches is
+that they have been developed for equations of motion found using Lagrange's
+method, which contain Lagrange multipliers. Kane's method \cite{Kane1985}
 allows for formulation of equations of motion for constrained systems in which
@@
 
-The goal of the following paper is to concretely establish the first order
+The goal of this paper is to concretely establish the first order
 relationships (about an arbitrary equilibrium point) between a selection of
@@
 specified quantities.  This procedure was created for systems whose equations
-of motion have been derived with Kane¿s Method. Although also applicable to
+of motion have been derived with Kane's Method. Although also applicable to
 dynamic system equations formulated using other methods, it is restricted to
-systems which can be completely described by a set of ODEs (i.e., no DAEs are
-present).
+systems which can be completely described by a set of ODEs (i.e., DAEs needn't
+be solved).
 
@@
 discussion of the procedure, along with an example. The derivation of the
-procedure is provided in the appendix. Additionally, the electronic
+procedure is provided in Section \ref{sec:derivations}. Additionally, the electronic
 supplementary material provides more examples (in more detail) and an
@@
 equations of motion and constraints must be arranged.  Equations
-\ref{eq:quant_to_compute} provides the bulk of the terms in equation
-\ref{eq:state_space_constrained}; these terms are Jacobian matrices of the
-differential equations. Equations \ref{eq:C_0}-\ref{eq:C_2} provide the terms
-$C_0$, $C_1$, and $C_2$ which embed the constraints into
-\ref{eq:state_space_constrained}. Finally, the permutation matrices in
-equations \ref{eq:Pq} and \ref{eq:Pu} provide the means to separate the
-independent and dependent speeds.
+(\ref{eq:quant_to_compute}) provides the bulk of the terms in equation
+(\ref{eq:state_space_constrained}); these terms are Jacobian matrices of the
+differential equations. Equations (\ref{eq:C_0})-(\ref{eq:C_2}) provide the
+terms $C_0$, $C_1$, and $C_2$ which embed the constraints into
+(\ref{eq:state_space_constrained}). Finally, the permutation matrices in
+equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the independent and
+dependent speeds.
 
@@
 \label{sec:discussion}
-Equation \ref{eq:state_space_constrained} doesn't fit into the standard linear
+Equation (\ref{eq:state_space_constrained}) doesn't fit into the standard linear
 system framework because it is non-square.  To obtain a square system (useful
@@
 %[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]
+
+\begin{figure}
+  \centering
+  \input{rollingdisc.pgf}
+  \caption{Rolling disc configuration}
+  \label{fig:disc}
+\end{figure}
 \section{Derivation of linearization algorithm}
@@
 traditional state space equations for a variety of tasks; these topics are
-discussed in \ref{sec:discussion}.
+discussed in Section \ref{sec:discussion}.
 

### 2012-08-24 PetersonGedeHubbard2012.tex
Added example's equations

@@
 \documentclass[smallcondensed]{svjour3}                     % onecolumn (standard format)
+%\documentclass{article}
 \RequirePackage{fix-cm}
@@
 \usepackage{graphicx}
-\usepackage{amsmath}
+\usepackage[fleqn]{amsmath}
 \usepackage{amssymb}
@@
 \end{figure}
+
+The matrices which make up the linear equations were created using the method
+presented in Section \ref{sec:results}. It is very important to note that the
+quantities in the matrices represent the states at the point of linearization.
+This leads to $q$'s, $u$'s, $\dot{q}$'s, and $\dot{u}$'s all showing up. The
+nonlinear equations need to be evaluated at a point to determine all of these
+quantities. The $\dot{u}$ quantities are not actually required to be 0, so
+equilibrium is not a requirement of the linearization point chosen.
+
+The independent state vector is:
+\begin{align}
+\begin{bmatrix}
+\delta \bm{q}_i \\
+\delta \bm{u}_i
+\end{bmatrix} =
+\begin{bmatrix}
+q_0 &
+q_1 &
+q_2 &
+q_3 &
+q_4 &
+u_0 &
+u_1 &
+u_2
+\end{bmatrix}^T
+\end{align}
+
+And the complete state vector is:
+\begin{align}
+\begin{bmatrix}
+\delta \bm{q} \\
+\delta \bm{u}
+\end{bmatrix} =
+\left[
+\begin{array}{cccccccccccc}
+q_0 &
+q_1 &
+q_2 &
+q_3 &
+q_4 &
+q_5 &
+u_0 &
+u_1 &
+u_2 &
+u_3 &
+u_4 &
+u_5
+\end{array}
+\right]^T
+\end{align}
+
+The matrices which make up the linear equations have been partitioned for
+readability.
+
+\begin{equation}
+\begin{aligned}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]=\\
+\end{aligned}
+\end{equation}
+\begin{equation*}
+\begin{aligned}
+\left[\begin{smallmatrix}- c_{1} s_{2} & c_{2} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} c_{2} & s_{2} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & c_{0} c_{2} - s_{0} s_{1} s_{2} & c_{0} s_{1} s_{2} + c_{2} s_{0} & - c_{1} s_{2} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & - c_{1} s_{0} & c_{0} c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & c_{0} s_{2} + c_{2} s_{0} s_{1} & - c_{0} c_{2} s_{1} + s_{0} s_{2} & c_{1} c_{2} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r s_{2} u_{1} & 0 & 0 & 0 & 0 & c_{2} r & 0 & 1 & 0 & 0\\0 & 0 & r \left(- c_{2} u_{2} + s_{2} u_{0}\right) & 0 & 0 & 0 & - c_{2} r & 0 & - r s_{2} & 0 & 1 & 0\\0 & 0 & c_{2} r u_{1} & 0 & 0 & 0 & 0 & r s_{2} & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - c_{2} m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & c_{2} m r & 0 & m r s_{2}\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & - m r s_{2} & 0\end{smallmatrix}\right]
+\end{aligned}
+\end{equation*}
+
+\begin{equation}
+\begin{aligned}
+\left[
+    \begin{array}{c}
+    (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 \\
+    (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0\\
+    (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0
+    \end{array}
+\right] =\\
+\end{aligned}
+\end{equation}
+
+\scriptsize
+\begin{equation*}
+\begin{aligned}
+\left[\begin{smallmatrix}0 & - s_{1} s_{2} \dot{q}_{0} & c_{1} c_{2} \dot{q}_{0} + s_{2} \dot{q}_{1} & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0\\0 & c_{2} s_{1} \dot{q}_{0} & c_{1} s_{2} \dot{q}_{0} - c_{2} \dot{q}_{1} & 0 & 0\\\left(- c_{0} c_{2} + s_{0} s_{1} s_{2}\right) \dot{q}_{4} + \left(c_{0} s_{1} s_{2} + c_{2} s_{0}\right) \dot{q}_{3} & s_{2} \left(- c_{0} c_{1} \dot{q}_{4} + c_{1} s_{0} \dot{q}_{3} - s_{1} \dot{q}_{5}\right) & c_{1} c_{2} \dot{q}_{5} + r s_{2} u_{1} + \left(c_{0} s_{2} + c_{2} s_{0} s_{1}\right) \dot{q}_{3} + \left(- c_{0} c_{2} s_{1} + s_{0} s_{2}\right) \dot{q}_{4} & 0 & 0\\c_{1} \left(c_{0} \dot{q}_{3} + s_{0} \dot{q}_{4}\right) & c_{0} s_{1} \dot{q}_{4} - c_{1} \dot{q}_{5} - s_{0} s_{1} \dot{q}_{3} & r \left(c_{2} u_{2} - s_{2} u_{0}\right) & 0 & 0\\- \left(c_{0} s_{2} + c_{2} s_{0} s_{1}\right) \dot{q}_{4} + \left(- c_{0} c_{2} s_{1} + s_{0} s_{2}\right) \dot{q}_{3} & c_{2} \left(c_{0} c_{1} \dot{q}_{4} - c_{1} s_{0} \dot{q}_{3} + s_{1} \dot{q}_{5}\right) & c_{1} s_{2} \dot{q}_{5} - c_{2} r u_{1} + \left(- c_{0} c_{2} + s_{0} s_{1} s_{2}\right) \dot{q}_{3} - \left(c_{0} s_{1} s_{2} + c_{2} s_{0}\right) \dot{q}_{4} & 0 & 0\\0 & 0 & c_{2} r u_{1} \dot{q}_{2} & 0 & 0\\0 & 0 & r \left(- c_{2} u_{0} - s_{2} u_{2}\right) \dot{q}_{2} & 0 & 0\\0 & 0 & r s_{2} u_{1} \dot{q}_{2} & 0 & 0\\0 & - c_{1} c_{2} g m r & m r \left(c_{2}^{2} r u_{0} u_{1} + c_{2} r s_{2} u_{1} u_{2} + g s_{1} s_{2} + s_{2} u_{0} u_{5} - s_{2} u_{2} u_{3} - s_{2} \dot{u}_{4}\right) & 0 & 0\\0 & 0 & m r \left(c_{2}^{2} r u_{2} - 2 c_{2} r s_{2} u_{0} u_{2} - c_{2} u_{0} u_{4} + c_{2} u_{1} u_{3} - c_{2} \dot{u}_{5} + r s_{2}^{2} u_{0} + r u_{1} + s_{2} u_{1} u_{5} - s_{2} u_{2} u_{4} + s_{2} \dot{u}_{3}\right) & 0 & 0\\0 & - c_{1} g m r s_{2} & m r \left(- c_{2} g s_{1} + c_{2} r s_{2} u_{0} u_{1} - c_{2} u_{0} u_{5} + c_{2} u_{2} u_{3} + c_{2} \dot{u}_{4} + r s_{2}^{2} u_{1} u_{2}\right) & 0 & 0\end{smallmatrix}\right]
+\end{aligned}
+\end{equation*}
+
+\normalsize
+
+\begin{equation}
+\begin{aligned}
+\left[
+\begin{array}{c}
+\tilde{A}_{qu} C_2 \\
+\tilde{A}_{uuc} C_2 \\
+\tilde{A}_{uud} C_2 \\
+    \end{array}
+\right] =\\
+\end{aligned}
+\end{equation}
+\begin{equation*}
+\begin{aligned}
+\left[\begin{smallmatrix}1 & 0 & 0\\0 & 1 & 0\\0 & 0 & 1\\0 & - c_{2} r & 0\\c_{2} r & 0 & r s_{2}\\0 & - r s_{2} & 0\\0 & r s_{2} \dot{q}_{2} & 0\\- r s_{2} \dot{q}_{2} & 0 & c_{2} r \dot{q}_{2}\\0 & - c_{2} r \dot{q}_{2} & 0\\- c_{2} m r u_{5} & \frac{1}{4} m r^{2} \left(4 c_{2} s_{2} u_{0} + 4 s_{2}^{2} u_{2} - 5 u_{2}\right) & \frac{1}{4} m r \left(4 c_{2} u_{3} - r u_{1}\right)\\m r \left(c_{2}^{2} r u_{2} - c_{2} r s_{2} u_{0} - s_{2} u_{4}\right) & m r \left(- c_{2} u_{5} + s_{2} u_{3}\right) & m r \left(c_{2} r s_{2} u_{2} + c_{2} u_{4} - r s_{2}^{2} u_{0}\right)\\\frac{1}{4} m r \left(r u_{1} - 4 s_{2} u_{5}\right) & \frac{1}{4} m r^{2} \left(- 4 c_{2} s_{2} u_{2} + 4 s_{2}^{2} u_{0} + u_{0}\right) & m r s_{2} u_{3}\end{smallmatrix}\right]
+\end{aligned}
+\end{equation*}
+
 \section{Derivation of linearization algorithm}

### 2012-08-28 PetersonGedeHubbard2012.tex
The A matrix is now a reasonable size.

--- 2012-06-22 PetersonGedeHubbard2012.tex
@@
 \end{figure}
+
+The matrices which make up the linear equations were created using the method
+presented in Section \ref{sec:results}. It is very important to note that the
+quantities in the matrices represent the states at the point of linearization.
+This leads to $q$'s, $u$'s, $\dot{q}$'s, and $\dot{u}$'s all showing up. The
+nonlinear equations need to be evaluated at a point to determine all of these
+quantities. The $\dot{u}$ quantities are not actually required to be 0, so
+equilibrium is not a requirement of the linearization point chosen. To reduce
+the complexity of the equations, we chose $q_0$ and $q_2$ to be equal to 0.
+This means that the other quantities in the linearized equations (other $q$'s,
+the $\dot{q}$'s, the $u$'s, and the $\dot{u}$'s need to be solved for with
+these 2 quantities already set to 0).
+
+The independent state vector is:
+\begin{align}
+\begin{bmatrix}
+\delta \bm{q}_i \\
+\delta \bm{u}_i
+\end{bmatrix} =
+\begin{bmatrix}
+q_0 &
+q_1 &
+q_2 &
+q_3 &
+q_4 &
+u_0 &
+u_1 &
+u_2
+\end{bmatrix}^T
+\end{align}
+
+And the complete state vector is:
+\begin{align}
+\begin{bmatrix}
+\delta \bm{q} \\
+\delta \bm{u}
+\end{bmatrix} =
+\left[
+\begin{array}{cccccccccccc}
+q_0 &
+q_1 &
+q_2 &
+q_3 &
+q_4 &
+q_5 &
+u_0 &
+u_1 &
+u_2 &
+u_3 &
+u_4 &
+u_5
+\end{array}
+\right]^T
+\end{align}
+
+The two matrices which make up the linearized equation:
+
+\begin{equation}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]=
+\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
+\end{equation}
+
+\begin{equation*}
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]=
+\end{equation*}
+\begin{equation}
+\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
+\end{equation}
+
 \section{Derivation of linearization algorithm}

### 2012-08-28 PetersonGedeHubbard2012.tex
Added a sentence to results section.

--- 2012-08-24 PetersonGedeHubbard2012.tex
@@
 Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged.  Equations
-(\ref{eq:quant_to_compute}) provides the bulk of the terms in equation
-(\ref{eq:state_space_constrained}); these terms are Jacobian matrices of the
-differential equations. Equations (\ref{eq:C_0})-(\ref{eq:C_2}) provide the
-terms $C_0$, $C_1$, and $C_2$ which embed the constraints into
-(\ref{eq:state_space_constrained}). Finally, the permutation matrices in
-equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the independent and
-dependent speeds.
+equations of motion and constraints must be arranged in order to carry out the
+linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+are Jacobian matrices of the differential equations. Equations
+(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+independent and dependent speeds.
 
@@
 \end{table}
+
+The matrices in Equation (\ref{eq:state_space_constrained}) are evaluated at
+the point of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*$; this point must satisfy the
+equations in Table \ref{table:assumptions}, and may or may not be an
+equilibrium point.
 

### 2012-08-28 PetersonGedeHubbard2012.tex
Revised a couple sentences in example.

@@
 The matrices which make up the linear equations were created using the method
-presented in Section \ref{sec:results}. It is very important to note that the
-quantities in the matrices represent the states at the point of linearization.
-This leads to $q$'s, $u$'s, $\dot{q}$'s, and $\dot{u}$'s all showing up. The
-nonlinear equations need to be evaluated at a point to determine all of these
-quantities. The $\dot{u}$ quantities are not actually required to be 0, so
-equilibrium is not a requirement of the linearization point chosen. To reduce
-the complexity of the equations, we chose $q_0$ and $q_2$ to be equal to 0.
-This means that the other quantities in the linearized equations (other $q$'s,
-the $\dot{q}$'s, the $u$'s, and the $\dot{u}$'s need to be solved for with
-these 2 quantities already set to 0).
+presented in Section \ref{sec:results}. It is important to note that the
+entries in these matrices must be evaluated at the point of linearization
+($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
+linearization must satisfy the equations in Table \ref{table:assumptions}.
+There is no requirement that this point be an equilibrium point, which implies
+that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
+reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
+means that the other quantities in the linearized equations (other $q$'s, the
+$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s need to be solved for with these 2
+quantities already set to 0).
 

### 2012-08-28 PetersonGedeHubbard2012.tex
Fixed a parens

@@
 means that the other quantities in the linearized equations (other $q$'s, the
-$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s need to be solved for with these 2
-quantities already set to 0).
+$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
+quantities already set to 0.
 

### 2012-09-29 PetersonGedeHubbard2012.tex
Revised abstract.

@@
 Analyses of such systems often require linearized forms of the motion
-equations.  The derivation of the nonlinear motion equations is well
-understood, but very little has been presented on the proper approach to
+equations.  Although derivation of the nonlinear motion equations is well
+understood, very little has been presented on the proper approach to
 forming linearized versions which satisfy the constraints.
 % Task
-To address this issue, we formed a procedure of organizing the constraints
+To address this issue, we developed a procedure for organizing the constraints
 and motion equations, and subsequently linearizing these equations.
@@
 motion equations of constrained multibody systems, and illustrate it with a
-familiar example.
+familiar example, the rolling disc.
 % Findings
-We applied this procedure to a model of a bicycle which includes a closed
-kinematic loop and four nonholonomic constraints and found that the resulting
+We apply this procedure to a model of a bicycle that includes a closed
+kinematic loop and four nonholonomic constraints and find that the resulting
 symbolic linearized motion equations are identical to previously published
@@
 % Perspectives
-Efficiency of computer implementations of this procedure will need to be
-considered in future studies.
+Efficiency of computer implementations of this procedure will be considered in
+future studies.
 \keywords{symbolic dynamics, control, constrained multibody systems}

### 2012-09-29 PetersonGedeHubbard2012.tex
Implemented the remainder of Mont's trivial revisions.

@@
 to the external world or other parts of the system. Velocity constraints limit
-the speeds in which the configuration can change, either from configuration
+the speeds at which the configuration can change, either from configuration
 constraints (through time differentiation) or independent application. In this
 paper, acceleration constraints refer to time differentiated velocity
-constraints. Examples of multibody systems with constraints are closed
+constraints. Examples of multibody systems with constraints include closed
 kinematic loops or objects rolling without slip.
 
-For these systems, a linear form of the equations of motion will be necessary
+For these systems, a linear form of the equations of motion is necessary
 for many different analyses.  Stability analysis and control system design are
@@
 straightforward and presents significant challenges. Simply computing the
-jacobian matrix of the differential equations will give incorrect results, with
+Jacobian matrix of the differential equations will give incorrect results, with
 no immediate indication that the results are wrong.
 
-Currently, there are only a few approaches to linearizing constrained systems
-which have been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these approaches is
-that they have been developed for equations of motion found using Lagrange's
-method, which contain Lagrange multipliers. Kane's method \cite{Kane1985}
-allows for formulation of equations of motion for constrained systems in which
-no multipliers are present. The dynamics of these systems can be completely
-described by ordinary differential equations. This significantly different
-formulation requires a different approach to linearization than what the
+Currently, only a few approaches to linearizing constrained systems which have
+been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these
+approaches is that they have been developed for equations of motion found using
+Lagrange's method, which contain Lagrange multipliers. Kane's method
+\cite{Kane1985} allows formulation of equations of motion for constrained
+systems in which no multipliers are present. The dynamics of these systems can
+be completely described by ordinary differential equations. This significantly
+different formulation requires a different approach to linearization than what
 existing options provide.
 
-The goal of this paper is to concretely establish the first order
-relationships (about an arbitrary equilibrium point) between a selection of
-independent coordinates and speeds and the time-derivatives of all of the
-coordinates and speeds (both independent and dependent). The formalism
-presented here is generic enough to cover most examples of time-varying
-constrained multibody systems with arbitrary external inputs and arbitrary
-specified quantities.  This procedure was created for systems whose equations
-of motion have been derived with Kane's Method. Although also applicable to
-dynamic system equations formulated using other methods, it is restricted to
-systems which can be completely described by a set of ODEs (i.e., DAEs needn't
-be solved).
+The goal of this paper is to establish concretely the first order relationships
+(about an arbitrary equilibrium point) between a selection of independent
+coordinates and speeds and the time-derivatives of all coordinates and speeds
+(both independent and dependent). The formalism presented here is generic
+enough to cover most examples of time-varying constrained multibody systems
+with arbitrary external inputs and arbitrary specified quantities.  The
+procedure was created for systems whose equations of motion have been derived
+with Kane's method. Although also applicable to dynamic system equations
+formulated using other methods, it is restricted to systems which can be
+completely described by a set of ODEs (i.e., DAEs needn't be solved).
 
-We first present our linearization procedure. Next we will offer some
-discussion of the procedure, along with an example. The derivation of the
-procedure is provided in Section \ref{sec:derivations}. Additionally, the electronic
+We first present the linearization procedure, followed by some discussion of
+the procedure, together with an example. The derivation of the procedure is
+provided in Section \ref{sec:derivations}. Additionally, the electronic
 supplementary material provides more examples (in more detail) and an
@@
 
-The matrices in Equation (\ref{eq:state_space_constrained}) are evaluated at
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at
 the point of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
@@
 We apply this linearization technique to the well studied rolling disc.  Figure
-XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
-assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
-transverse inertia $I$.  To illustrate the concept of dependent quantities, we
-purposefully choose to describe the configuration and velocity of the disc with
-a non-minimal set of coordinates and speeds.  This choice is not purely
-academic; a physical disc equipped with a rate gyroscope and accelerometer
-measure these body fixed quantities (or their time derivatives) directly,
-rather than the more traditional set which is chosen for analytic convenience.
-In more complicated systems with constraints, it is difficult (if not
-impossible) to describe the system with a minimal choice of coordinates and
-speeds; the ideas demonstrated by this simple example can be applied to more
-complicated ones.
+\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
+The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
+$J$, and transverse inertia $I$.  To illustrate the concept of dependent
+quantities, we purposefully choose to describe the configuration and velocity
+of the disc with a non-minimal set of coordinates and speeds.  This choice is
+not purely academic.  It is convenient for a physical disc equipped with a rate
+gyroscope and accelerometer to measure these body fixed quantities (or their
+time derivatives) directly, rather than the more traditional set which is
+chosen for analytic convenience.  In more complicated systems with constraints,
+it is difficult (if not impossible) to describe the system with a minimal
+choice of coordinates and speeds; the ideas demonstrated by this simple example
+can be applied to more complicated ones.
 
-The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
-is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
-$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
-of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
-$C^*$ in $N$ are
+The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
+is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
+(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
+location of disc center $C^*$ relative to the inertial origin $N^*$, and the
+velocity of $C^*$ in $N$ are
 \begin{align}
@@
 \end{align}
-If the disc is to make contact with the ground, it should be obvious to the
-reader that there is a configuration constraint involving the height $q_5$ of
-the disc center and the lean $q_1$.  Similarly, if the disc rolls without
-slip, it should be obvious that there are constraints which relate the
-generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
-constraints and the equations of motion are presented in the electronic
-supplementary material and can be found in the literature \cite{O'Reilly1996}.
+If the disc is in contact with the ground, a configuration constraint involving
+the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
+rolls without slip, other constraints relate the generalized speeds $u_i, (i =
+0,\dots,5)$.  The derivations of these constraints and the equations of motion
+are presented in the electronic supplementary material and can also be found in
+the literature \cite{O'Reilly1996}.
 \begin{table}[htbp]
@@
 
-And the complete state vector is:
+And the complete (independent and dependent) state vector is:
 \begin{align}
@@
 
-The two matrices which make up the linearized equation:
-
+The coefficient matrices of the linearized equations (Equation
+\ref{eq:state_space_constrained}) are:
 \begin{equation}
@@
 
-\section{Derivation of linearization algorithm}
+\appendix
+
+\section{Derivation of linearization procedure}
 \label{sec:derivations}
 
-This section outlines the derivation of the procedure for determining linear
-equations of motion \ref{eq:state_space_constrained}. This equation can be used
-for simulation, stability analysis, or it can be reformulated into the
-traditional state space equations for a variety of tasks; these topics are
-discussed in Section \ref{sec:discussion}.
+This appendix outlines the derivation of Equation
+(\ref{eq:state_space_constrained}. This equation can be used for simulation,
+stability analysis, or it can be reformulated into the traditional state space
+equations for a variety of tasks; these topics are discussed in Section
+\ref{sec:discussion}.
 
@@
 satisfied by these quantities.  In the interest of brevity, we omit writing
-this equilibrium after each gradient in the calculations below; all are
-evaluated at these equilibrium conditions.  Expansion of the three constraint
-equations yields
+this point of linearization in each gradient in the calculations below; all are
+evaluated at this point.  Expansion of the three constraint
+equations, keeping only first order terms, yields
 \begin{align}
@@
 Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet only $p = n - l + o - m$ of these quantities are independent.
-To address this issue, a particular set of independent
-coordinates and speeds must be selected. To this end, consider the following
-partitioning of the generalized coordinates and generalized speeds:
++ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
+address this issue, a smaller set of independent coordinates and speeds must be
+selected. To this end, consider the following partitioning of the generalized
+coordinates and generalized speeds:
 \begin{equation*}
@@
 independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
-\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities
-($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the
-original ordering of the coordinates and speeds.  We use the notation $P_{qi}$
-and $P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$,
-respectively; similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while
-$P_{ud}$ is the last $m$ columns of $P_u$.
+\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
+($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
+ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
+$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
+similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
+last $m$ columns of $P_u$.
 

### 2012-10-03 PetersonGedeHubbard2012.tex
Ready to send out

@@
 % Object of the document
-In this paper we present this procedure to correctly linearize the nonlinear
-motion equations of constrained multibody systems, and illustrate it with a
-familiar example, the rolling disc.
+In this paper we present this procedure to linearize the nonlinear motion
+equations of constrained multibody systems, and illustrate it with a familiar
+example, the rolling disc.
 % Findings
@@
 \end{abstract}
-
-\section{Conclusions}
-\label{sec:conclusions}
-We have provided a procedure to obtain linear state space formulation for
-constrained multibody systems whose nonlinear dynamics are described only by
-ordinary differential equations. With this state space formulation, you can
-perform stability analysis, control system design, and other linear analyses.
-
-In some sense this work is simply applying the chain rule through matrix
-operations in a way that respects system constraints. An advanced C.A.S. system
-which possesses the ability to represent dependent quantities (as functions of
-independent quantities) could perform the chain rule directly, thereby
-eliminating the need for matrix operations. We leave an open question: how does
-computational performance depend on implementation details and problem size?
 
@@
 
-For these systems, a linear form of the equations of motion is necessary
-for many different analyses.  Stability analysis and control system design are
-two possible scenarios.  Linearization of these constrained systems is not
-straightforward and presents significant challenges. Simply computing the
-Jacobian matrix of the differential equations will give incorrect results, with
-no immediate indication that the results are wrong.
+Linear equations of motion are necessary for many different analyses.
+Stability analysis and control system design are two possible scenarios.
+Linearization of these constrained systems is not straightforward and presents
+significant challenges. Computing the Jacobian matrix of the differential
+equations - without considering the system's constraints - will give incorrect
+results.
 
@@
 The goal of this paper is to establish concretely the first order relationships
-(about an arbitrary equilibrium point) between a selection of independent
+(about an arbitrary point of linearization) between a selection of independent
 coordinates and speeds and the time-derivatives of all coordinates and speeds
@@
 
+With this example, we have used our linearization procedure to find linear
+equations of motion, which respect the system's constraints. Parametric
+stability analysis (eigenvalue analysis) can now be performed on these equations.
+These computations were implemented symbolically with the open source symbolic
+manipulator SymPy \cite{SymPy2012}.
+
 \appendix
-
 \section{Derivation of linearization procedure}
@@
 The first terms are identically zero because of the assumption that the
-equilibrium point satisfies the constraints.  The Taylor series expansion of
-the kinematic differential equations is
+point of linearization satisfies the constraints.  The Taylor series expansion
+of the kinematic differential equations is
 \begin{align}

### 2012-10-03 PetersonGedeHubbard2012.tex
Added conclusion.

@@
 
+\section{Conclusions}
+A procedure for forming linearized equations of motion for constrained
+multi-body systems has been presented.  This procedure can be implemented
+symbolically or numerically, and handles configuration, velocity, and
+acceleration level constraints.  The procedure has been implemented
+symbolically in the \texttt{sympy.physics.mechanics} sub-module of the open
+source symbolic manipulator SymPy\cite{SymPy2012}.
+
+\begin{acknowledgements}
+ This material is based upon work partially supported by the National Science
+ Foundation under award 0928339 and three Google Summer of Code projects (2009,
+ 2011, 2012).  Jason Moore, Thomas Johnston, Evan Sperber, and Andrew Kickertz
+ provided valuable feedback during discussions of multibody dynamics and
+ control.
+\end{acknowledgements}
+
 \appendix
@@
 
-\begin{acknowledgements}
-  This material is based upon work partially supported by the National Science
-  Foundation under award 0928339 and two Google Summer of Code projects (2009,
-  2011).  Jason Moore, Thomas Johnston, Evan Sperber, and Andrew Kickertz
-  provided valuable feedback during discussions of multibody dynamics and
-  control.
-\end{acknowledgements}
-
 % Not sure which bibliography style we are supposed to use

### 2012-10-03 PetersonGedeHubbard2012.tex
Fixed spelling error.
Changed .pgf figure to pdf and updated .tex file accordingly.

@@
 \usepackage{bm}
-\usepackage{pgf}
+%\usepackage{pgf}
 %\usepackage[polutonikogreek]{babel}
@@
 symbolic linearized motion equations are identical to previously published
-results.  The technique can be performed symbolically (enabling explict
+results.  The technique can be performed symbolically (enabling explicit
 analytical forms of the linearized dynamics) or numerically.
@@
 future studies.
-\keywords{symbolic dynamics, control, constrained multibody systems}
+\keywords{linearization, constrained multibody systems, symbolic dynamics, control}
 \end{abstract}
@@
   \centering
-  \input{rollingdisc.pgf}
+  \includegraphics{rollingdisc.pdf}
   \caption{Rolling disc configuration}
@@
 With this example, we have used our linearization procedure to find linear
-equations of motion, which respect the system's constraints. Parametric
-stability analysis (eigenvalue analysis) can now be performed on these equations.
-These computations were implemented symbolically with the open source symbolic
-manipulator SymPy \cite{SymPy2012}.
+equations of motion which respect the system's constraints. Parametric
+stability analysis (eigenvalue analysis) can now be performed on these
+equations.  These computations were implemented symbolically with the open
+source symbolic manipulator SymPy \cite{SymPy2012}.
 

### 2013-02-17 PetersonGedeHubbard2012.tex
Fixed typo in M_{uqd}
Clarified point of linearization just below Table 1.

@@
 
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at
-the point of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*$; this point must satisfy the
-equations in Table \ref{table:assumptions}, and may or may not be an
-equilibrium point.
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+an equilibrium point.
 
@@
     \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
-    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_2 \\
+    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
     \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &

### 2013-02-17 PetersonGedeHubbard2012.tex
Added subscript to identity matrix in definition of C_2
Changed speeds to states in sentence discussing constraint Jacobians.

@@
   \label{eq:C_2}
-  C_2 &\triangleq \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
     \bm{f}_{v} \right] P_{ui}\\
@@
 \cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
-constraint matrices to determine a set of independent speeds which will ensure
+constraint matrices to determine a set of independent states which will ensure
 the non-singularity of the aforementioned matrices.

### 2013-05-25 rolling_disc_with_arm.tex
WIP: writing up rolling disc with arm.

\documentclass{article}

\begin{document}
\section{Rolling disc with skate attached to arm}
Consider a thin disc of mass $m$, radius $r$, radial inertia $mr^2/2$, and,
transverse inertia $mr^2/4$.  At the disc center is an axle attached to one end
of a massless rod of length $l$; to the other end of the rod a skate is
affixed.  The skate permits frictionless sliding along the line of intersection
between the disc plane and the ground plane upon which the disc rolls without
slip.  The skate permits no sliding direction perpendicular to this line.

The inertial frame $N$ has $n_z$ aligned with the local gravitational field,
while $n_x$ and $n_y$ span the ground plane with $n_z = n_x \times n_y$.  Four
generalized coordinates $q_0, q_1, q_2, q_3$ sequentially orient four frames
$A, B, C, D$.  $A$ is oriented relative to $N$ by a $q_0$ rotation about $n_z$;
this is the yaw frame.  $B$ is oriented relative to $A$ by a $q_1$ rotation
about $a_x$; this is the lean frame. $C$ and is oriented relative to $B$ by a
$q_2$ rotations about $b_y$; this is the disc frame ($c_y$ is perpendicular to
the disc plane).  $D$ is oriented relative to $B$ by a $q_3$ rotation about
$b_y$; this is the arm frame ($d_z$ is parallel to the line between the disc
center and the arm-ground contact point).

A torque $\tau$ is applied between the disc and the arm by means of a motor.

\end{document}

### 2013-06-01 PetersonGedeHubbard2012.tex
Fixed dimensions of P' in Discussion section.
Added dynamics overview and rewrote intro
Introduction still needs work to complete it. Dynamics overview probably needs
review, but is "complete".

@@
-\documentclass[smallcondensed]{svjour3}                     % onecolumn (standard format)
+\documentclass[smallcondensed,final]{svjour3}                     % onecolumn (standard format)
 %\documentclass{article}
@@
 
-\title{A linearization procedure for constrained multibody systems}
+\title{Symbolic linearization of constrained multibody systems}
 
-\titlerunning{A linearization procedure for constrained multibody systems}        % if too long for running head
+\titlerunning{Symbolic linearization of constrained multibody systems}        % if too long for running head
 
@@
 Analyses of such systems often require linearized forms of the motion
-equations.  Although derivation of the nonlinear motion equations is well
-understood, very little has been presented on the proper approach to
-forming linearized versions which satisfy the constraints.
+equations.
 % Task
@@
 and motion equations, and subsequently linearizing these equations.
+This procedure was specifically developed for equations of motion generated by
+Kane's method, but is compatible with any method as long as the following are
+true: only ordinary differential equations are required to describe the system,
+and independent and dependent states remain in the final equations (i.e. no
+DAEs, no substitution of variables).
 % Object of the document
-In this paper we present this procedure to linearize the nonlinear motion
-equations of constrained multibody systems, and illustrate it with a familiar
-example, the rolling disc.
-% Findings
-We apply this procedure to a model of a bicycle that includes a closed
-kinematic loop and four nonholonomic constraints and find that the resulting
-symbolic linearized motion equations are identical to previously published
-results.  The technique can be performed symbolically (enabling explicit
-analytical forms of the linearized dynamics) or numerically.
-% Conclusion
-Motion equations for constrained multibody systems can now be linearized in a
-systematic fashion, symbolically or numerically, thereby facilitating common
-linear analyses (e.g., stability analysis and control system design).
-% Perspectives
-Efficiency of computer implementations of this procedure will be considered in
-future studies.
+Following a brief review of Kane's method and the structure of the equations it
+generates, we present the procedure to symbolically linearize the nonlinear
+motion equations of constrained multibody systems, and illustrate it with the
+familiar example of the rolling disc.
+
 \keywords{linearization, constrained multibody systems, symbolic dynamics, control}
@@
 \label{sec:intro}
+Designing a feedback controller or performing stability analysis are two common
+analyses performed on models of multibody systems.
+A linear form of the equations of motion is required to apply the most
+frequently used techniques to such systems.
+Normally, obtaining the equations of motion is as simple as taking the Jacobian
+of the right hand side function of the equations of motion with respect to the
+states.
+However, when the multibody system in questions has constraints, this process
+is more complicated.
+If one has used Kane's method \cite{Kane1985} to generate the equations for a
+constrained system, the resulting equations will have a particular form: a set
+of ordinary differential equations containing both the independent and
+dependent states.
+Simple inspection of the equations does not reveal that they represent a
+constrained system - their form is similar to the equations of an unconstrained
+system.
+It is critical to state that linearizing these equations by computing the
+Jacobian will give incorrect results.
+This paper provides a way to correctly generate the linear equations of motion
+for constrained systems which have been found using Kane's method (or any other
+method which results in equations with the same form).
+
 Multibody systems, rigid or flexible, and particles often have constraints
-which limit how they can move. In this paper, we will discuss 3 types of
-constraints: configuration, velocity, and acceleration. Configuration
-constraints limit the location or orientation of parts of the system, relative
-to the external world or other parts of the system. Velocity constraints limit
-the speeds at which the configuration can change, either from configuration
-constraints (through time differentiation) or independent application. In this
-paper, acceleration constraints refer to time differentiated velocity
-constraints. Examples of multibody systems with constraints include closed
-kinematic loops or objects rolling without slip.
+which limit how they can move or be configured. In this paper, we will discuss
+three types of constraints: configuration, velocity, and acceleration.
+Configuration constraints limit the location or orientation of parts of the
+system, relative to the external world or other parts of the system. Velocity
+constraints limit the speeds at which the configuration can change, either from
+configuration constraints (through time differentiation) or independent
+application. In this paper, acceleration constraints refer to time
+differentiated velocity constraints. Motion constraints most often appear in
+systems where there is rolling without slip or there are closed kinematic
+loops.
 
-Linear equations of motion are necessary for many different analyses.
-Stability analysis and control system design are two possible scenarios.
-Linearization of these constrained systems is not straightforward and presents
-significant challenges. Computing the Jacobian matrix of the differential
-equations - without considering the system's constraints - will give incorrect
-results.
+% TODO: Improve literature review.
+Bottema appears to be the first author to show that special considerations are
+needed for linearizing non-holonomic systems \cite{Bottema1949}.
+While other authors have examined linearization of nonholonomic systems, we
+have found isssues with applying any of them to equations generated by Kane's
+method.
+Kang et al. \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both
+explored linearization, but these methods have been developed for equations of
+motion found using Lagrange's method, which contain Lagrange multipliers (which
+are not present in equations of motion generated from Kane's method).
+Minaker and Rieveley \cite{Minaker2010} and Schwab and Meijaard
+\cite{Schwab2003} both have developed techniques for generating equations of
+motion (and linear forms), however both techniques are more numerical in
+nature - they do not provide symbolic equations of motion or symbolic
+linearized equations.
+These processes, along with others such as Neuman's %TODO: fix reference
+, formulate the linear equations as a step in the equations of motion
+generation process.
+They are not capable of linearizing existing equations of motion.
 
-Currently, only a few approaches to linearizing constrained systems which have
-been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these
-approaches is that they have been developed for equations of motion found using
-Lagrange's method, which contain Lagrange multipliers. Kane's method
-\cite{Kane1985} allows formulation of equations of motion for constrained
-systems in which no multipliers are present. The dynamics of these systems can
-be completely described by ordinary differential equations. This significantly
-different formulation requires a different approach to linearization than what
-existing options provide.
-
+% TODO:reiterate limiations of model
 The goal of this paper is to establish concretely the first order relationships
@@
 
-We first present the linearization procedure, followed by some discussion of
-the procedure, together with an example. The derivation of the procedure is
-provided in Section \ref{sec:derivations}. Additionally, the electronic
-supplementary material provides more examples (in more detail) and an
-alternative ordering for the independent/dependent quantities.
+% TODO:rewrite organization of paper
+The use of Kane's method in generating equations of motion is reviewed first,
+in order to properly orient the reader to the format and some of the qualities
+of the generated equations.
+Kane's method is then briefly used to generate the equations of motion for a
+simple system, a rolling disk with a trailing arm.
+The need for a different linearization technique is also demonstrated with this
+model.
+In response to this need, we present the newly developed linearization
+procedure, followed by its derivation. % maybe flip the order?
+This new linearization procedure is applied to the example system, to provide
+an example of its application.
+Finally, we discuss other details and nuances of the procedure and its use.
 
-%The complete configuration of the system is further assumed to be described by
-%generalized coordinates $\bm{q}\in\bm{R}^n$ and its velocity completely
-%described by generalized speeds $\bm{u}\in\bm{R}^o$, with $n \ne o$ in general.
-%All coordinates and speeds, not only the independent ones, are represented by
-%$\bm{q}$ and $\bm{u}$.  The system is assumed to be governed by the
-%relationships described in Table \ref{table:assumptions}.  The system is
-%described by $l + m + n + o$ equations which relate $2n + 2o + s + 1$
-%quantities, which implies that $n - l + o - m + s + 1$ of these quantities are
-%independent.  This paper discusses the first order relationship between these
-%independent quantities ($n-l$ independent coordinates, $o-m$ independent
-%speeds, $s$ control inputs, and time $t$) and the time derivatives of the $n$
-%coordinates and $o$ speeds.
-%
-%The first three equations in Table \ref{table:assumptions} represent
-%constraints derived purely from kinematic considerations at the configuration,
-%velocity, and acceleration levels, respectively.  The velocity constraints may
-%be nonholonomic or time differentiated holonomic.  The acceleration constraints
-%can be time differentiated velocity constraints or kinematic acceleration
-%constraints.  The fourth equation represents kinematic differential equations
-%which relate $\bm{\dot{q}}$ to $\bm{u}$ and are linear in both these
-%quantities.  The fifth equation represents the constraint-free nonholonomic
-%dynamic differential equations~\cite{Kane1985}, and will necessarily be linear
-%in $\bm{\dot{u}}$ and the inputs $\bm{r}$, but generally nonlinear in all other
-%quantities.  This formalism is exceedingly generic; in practice, the structure
-%of the functions in Table \ref{table:assumptions} can be leveraged to simplify
-%the following calculations.  This will be discussed subsequently.
-%
-%Trajectories of the system exist on a $p \triangleq n - l + o - m$ dimensional
-%manifold embedded in a $n + o$ dimensional space, though conserved quantities
-%(e.g., total mechanical energy) may restrict trajectories to a lower
-%dimensional manifold.  In general this cannot be assumed, however, especially
-%in the presence of exogenous inputs $\bm{r}$ which may change the total
-%mechanical energy of the system.  While more than $p$ of the independent state
-%variables may be of interest (i.e., dependent coordinates or speeds may be of
-%interest in themselves), only $p$ quantities are truly independent.  Which of
-%these $p$ coordinates and speeds should be chosen as independent out of the
-%possible $n + o$ may be obvious for small systems where there is intuition
-%about system behavior, but in general the optimal choice depends on how the
-%configuration $\bm{q}$ affects the gradients of $f_c$ and $f_v$.  On the other
-%hand, in systems with cyclic (ignorable) coordinates, fewer than $p$ state
-%variables may be of interest; one example is the Whipple bicycle model where
-%$p=10$ but 5 of these state variables are cyclic coordinates (Cartesian
-%coordinates of rear wheel contact point, heading, and both wheel angles) and
-%may only be of secondary interest.
+\section{Kane's Method, Briefly}
+\label{sec:kane_method}
+To familiarize the reader with Kane's method, three concepts relating to its
+use will be examined.
+The first is the use of generalized speeds in addition to generalized
+coordinates.
+The second concept is in how these generalized speeds can be used to project
+permissible motions of the system, and how this removes the need to consider
+non-contributing (internal constraint) forces.
+Finally, the mathematical steps required to use Kane's method are shown and
+described.
+
+In using Lagrange's method, generalized coordinates are used to describe the
+configuration of a system while the time derivatives of the gen. coordinates
+describe the velocities of a system.
+Kane's method allows for the velocities of a system to be written in terms of
+generalized speeds, which are not required to be the time derivatives of
+coordinates (although such a definition is permitted).
+
+The benefit of using both generalized speeds and coordinates can be seen in
+the following example.
+If the orientation of a rigid body were to be described using using Euler
+parameters (quaternions), its angular velocity is relatively complicated when
+using the time derivatives of the coordinates.
+Using generalized speeds allows for the angular velocity to be written in a
+much simpler form:
+\begin{align}
+\label{eq:ex1}
+{^N}\bar{\omega}^B = \it{\omega}_1 \hat{b}_1 + \it{\omega}_2 \hat{b}_2 +
+\it{\omega}_3 \hat{b}_3
+\end{align}
+or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
+defined as the sum of three generalized-speed/basis-vector products.
+The derivatives of these generalized speeds will then appear in the equations
+of motion, rather than the twice time differentiated generalized coordinates.
+
+A downside of allowing such definitions of system velocities is that the
+kinematic differential equations, which relate the rate of change of the gen.
+coordinates to the gen. speeds, become more complicated.
+This penalty is usually offset by the vastly simplified equations of motion
+that are generated.
+
+Using and understanding Kane's method requires another concept: partial
+velocities.
+Partial velocities are defined as the partial derivatives of the velocity
+vectors with respect to the generalized speeds.
+They can be defined for both translational and rotational velocities.
+Using the angular velocity of body $B$ in reference frame $N$ as shown in
+(\ref{eq:ex1}), the partial velocities can be written as:
+\[
+{^N}\bar{\omega}^B_{\omega_1} = \hat{b}_1 \quad \quad
+{^N}\bar{\omega}^B_{\omega_2} = \hat{b}_2 \quad \quad
+{^N}\bar{\omega}^B_{\omega_3} = \hat{b}_3
+\]
+The partial velocities represent the direction of motion associated with each
+generalized speed.
+
+By taking the dot product of each partial velocity and both sides of Newton's
+second law, $F=ma$ (or the rotational equivalent), only terms that are
+parallel to the partial velocities are left in the equation.
+One equation is generated for each generalized speed.
+
+An important byproduct of this is that internal constraint forces
+(non-contributing forces) generated by things such as pin or sliding joints no
+longer need to be considered.
+Assuming the generalized speeds have been defined in an intelligent manner, the
+forces involved in the internal constraints cannot affect the rate of change of
+any generalized speed, and thus do not need to be considered.
+
+In summary, Kane's method allows for velocities to be defined in terms of
+generalized speeds, allowing for simpler velocity expressions.
+The dot product of each partial velocity (the direction of motion associated
+with each generalized speed) with Newton's second law (or Euler's equations)
+removes all terms which are not related to each generalized speed.
+This generally results in simple equations of motion, which are relatively easy
+to form.
+
+The mathematical formalism for applying Kane's method to multibody systems
+follows.
+Consider a system composed of rigid bodies $B_1,...,B_g$, particles
+$P_1,...,P_h$, with points of interest $Q_1,...,Q_k$, and reference frames of
+interest $E_1,...,E_c$, all defined relative to the inertial reference frame
+$N$.
+This system has generalized coordinates $q_1,...,q_n$, generalized speeds
+$u_1,...,u_o$.
+Additionally, there are $l$ configuration constraints and $m$ velocity
+constraints.
+
+In order to properly apply Kane's method, the velocities of each component in
+the system need to be written in the following form.
+Translational velocity of particles:
+\begin{align}
+\label{eq:particle_translational}
+{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{u_j} u_j + {^N}\bar{v}^{P_i}_t
+\quad \quad i=1,...,g
+\end{align}
+Translational velocity of rigid body mass centers:
+\begin{align}
+\label{eq:rb_translational}
+{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{u_j} u_j +
+{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
+\end{align}
+Translational velocity of points of interest:
+\begin{align}
+\label{eq:points_translational}
+{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{u_j} u_j + {^N}\bar{v}^{Q_i}_t
+\quad \quad i=1,...,k
+\end{align}
+Rotational velocity of rigid bodies:
+\begin{align}
+\label{eq:rb_rotational}
+{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
+{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
+\end{align}
+Rotational velocity of reference frames of interest:
+\begin{align}
+\label{eq:rb_rotational}
+{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
+{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
+\end{align}
+
+In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents the
+velocity or angular velocity component which is a prescribed function of time,
+and has no dependence on generalized coordinates or speeds (e.g., a crank
+driven at a fixed rate).
+
+For this system, the configuration constraints are defined as:
+\begin{align}
+\label{eq:configuration_constraints}
+\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
+\end{align}
+where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
+The velocity constraints are defined as:
+\begin{align}
+\label{eq:velocity_constraints}
+\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
+\end{align}
+where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$.
+It is important that the velocity constraints include the time-differentiated
+configuration constraints or equivalent constraints which produce the same
+behavior.
+Furthermore, velocity constraints can not be written with $\dot{\mathbf{q}}$
+terms (the time derivatives of the coordinates).
+Acceleration constraints need to be defined as well.
+Typically, they will simply by time-differentiated velocity constraints and
+have the form:
+\begin{align}
+\label{eq:acceleration_constraints}
+\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
+\mathbf{0}
+\end{align}
+
+At this point, it is convienient to define a matrix $\mathbf{B}$, which is the
+gradient of the velocity constraints $\mathbf{f}_v$ with respect to all of the
+generalized speeds $\mathbf{u}$.
+\begin{align}
+\label{eq:constraint_B}
+\mathbf{B} \triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
+t)
+\end{align}
+and from (\ref{eq:velocity_constraints}) and (\ref{eq:constraint_B}) the
+following can be written:
+\begin{align}
+\label{eq:constraint_Bu0}
+\mathbf{0} = \mathbf{B}\mathbf{u}
+\end{align}
+
+With these definitions established, Kane's method can be applied.
+The output of Kane's method is the equation $\mathbf{F} + \mathbf{F}^* =
+\mathbf{0}$ - the sum of the generalized active forces and generalized inertial
+forces is 0.
+If Newton's second law were written as $f - ma = 0$, the $\mathbf{F}$ is the
+multibody generalization of $f$ and $\mathbf{F}^*$ is the multibody
+generalization of $-ma$.
+These terms are created by taking the dot product of the partial
+velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
+- (\ref{eq:rb_rotational}), with the resultant forces/torques, $\bar{R}$/
+$\bar{T}$, and the inertial forces/torques, $\bar{R}^*$/$\bar{T}^*$.
+
+The resultant force at a point is defined as the sum of all force vectors
+acting directly on that point.
+The resultant torque on a reference frame is defined similarly - it is the sum
+of all torque vectors acting on that reference frame.
+Note that distance forces do not need to be transformed into a resultant
+force/torque for any components of the system; they only need to be considered
+at the point of application (e.g., if a force $\bar{f}$ is the only force
+applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$ and
+$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
+These resultant forces and torques are used to write the generalized active
+forces:
+
+\begin{equation}
+\label{eq:definition_F}
+\mathbf{F} =
+\begin{bmatrix}
+\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
+\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
+\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_1 +
+\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 +
+\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_1 \\
+\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
+\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
+\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_2 +
+\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 +
+\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_2 \\
+\displaystyle \vdots \\
+\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
+\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
+\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_o +
+\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o +
+\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_o
+\end{bmatrix}
+\end{equation}
+
+The inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$, will now be
+defined. Note: the reference frame these are defined is not written - at this
+point in the derivation of the equations of motion, they are assumed to be in
+the inertial frame, represented here as $N$.
+
+Particles:
+\begin{align}
+\label{eq:particle_gen_inertia}
+\bar{R}^*_P \triangleq -m_p {^N}\bar{a}^P
+\end{align}
+Rigid Bodies:
+\begin{align}
+\label{eq:rb_translational_gen_inertia}
+\bar{R}^*_{B^*} \triangleq -m_B {^N}\bar{a}^{B^*}
+\end{align}
+\begin{align}
+\label{eq:rb_rotational_gen_inertia}
+\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \times \bar{\bar{I}}^{B/B^*} -
+{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\omega^B
+\end{align}
+Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
+$B$ about its mass center $B^*$ \ref{Kane1985}.
+
+If the basis vectors of the reference frame associated with the rigid body $B$
+are aligned with the principal axes of $B$, then the following definitions can
+be used:
+\begin{align}
+\alpha_i \triangleq {^N}\alpha^B \cdot \hat{b}_i \quad \quad
+\omega_i \triangleq {^N}\omega^B \cdot \hat{b}_i \quad \quad
+I_i \triangleq \hat{b}_i \cdot \bar{\bar{I}}^{B/B^*} \cdot \hat{b}_i
+\quad \quad for i=1,2,3
+\end{align}
+In this case, $I_1$, $I_2$, and $I_3$ will be the principal inertias of rigid
+body $B$.
+This allows the inertia torque to be written as:
+\begin{align}
+\label{eq:simple_rb_rotational_gen_inertia}
+\bar{T}_B^* = -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1
+              -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2
+              -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3
+\end{align}
+
+Now $\mathbf{F}^*$ can be constructed in a manner similar to $\mathbf{F}$.
+
+\begin{equation}
+\label{eq:definition_Fstar}
+\mathbf{F} =
+\begin{bmatrix}
+\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
+\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
+\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 \\
+\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
+\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
+\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 \\
+\displaystyle \vdots \\
+\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
+\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
+\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o
+\end{bmatrix}
+\end{equation}
+
+Now the equations $\mathbf{F} + \mathbf{F}^* = \mathbf{0}$ are defined.
+Bu what has been written is valid only if there are no motion constraints.
+The equations need to be transformed in order to not violate the behaviors
+defined by the constraints.
+
+Kane's Method allows for the equations of motion to be written using both
+independent and dependent speeds, $\mathbf{u}_i$ and $\mathbf{u}_d$, without
+the need for Lagrange multipliers.
+Even though both independent and dependent speeds are in the final equations, a
+choice still needs to be made as to which speeds will be the constrained
+speeds.
+The methods for doing this are outside the scope of this paper, but it is
+important that whatever selection is made avoids singularities.
+
+To simplify notation, it is assumed that the generalized speeds are already
+ordered in a favorable manner:
+\begin{align}
+\label{eq:speeds_independent_dependent}
+\mathbf{u} =
+\begin{bmatrix}
+\mathbf{u}_i \\
+\mathbf{u}_d
+\end{bmatrix}
+\end{align}
+The matrix $\mathbf{B}$ from (\ref{eq:constraint_B}) needs to be rewritten as
+well:
+\begin{align}
+\label{eq:constraint_B_independent_dependent}
+\mathbf{0} &= \mathbf{B} \mathbf{u}\\
+\mathbf{0} &= \begin{bmatrix}
+\mathbf{B}_i & \mathbf{B}_d
+\end{bmatrix} \begin{bmatrix}
+\mathbf{u}_i \\ \mathbf{u}_d
+\end{bmatrix}
+\end{align}
+Where $\mathbf{B}_i$ has the shape $m \times o - m$ and $\mathbf{B}_d$ has the
+shape $m \times m$. This equation can be rewritten as:
+\begin{align}
+\label{eq:constraint_A}
+\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d = \mathbf{0} \\
+\mathbf{B}_d \mathbf{u}_d = - \mathbf{B}_i \mathbf{u}_i \\
+\mathbf{u}_d = -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
+\mathbf{A} \triangle - \mathbf{B}_d^{-1} \mathbf{B}_i \\
+\mathbf{u}_d = \mathbf{A} \mathbf{u}_i
+\end{align}
+Where $\mathbf{A}$ has the shape $m \times m$.
+
+Finally, to constrain the generalized active and inertia forces, they need to
+be written as:
+\begin{align}
+\label{eq:F_Fstar_rewrite}
+\mathbf{F} = \begin{bmatrix} \mathbf{F}_i \\ \mathbf{F}_d \end{bmatrix} \quad
+\quad \mathbf{F}^* = \begin{bmatrix} \mathbf{F}^*_i \\ \mathbf{F}^*_d
+\end{bmatrix}
+\end{align}
+allowing the nonholonomic generalized forces to be written:
+\begin{align}
+\label{eq:F_Fstar_nonholonomic}
+\tilde{\mathbf{F}} = \mathbf{F}_i + \mathbf{A}^T \mathbf{F}_d \quad \quad
+\tilde{\mathbf{F}}^* = \mathbf{F}_i^* + \mathbf{A}^T \mathbf{F}_d^*
+\end{align}
+
+After constraining the equations of motion generated, we have reduced the
+number of equations from $o$ to $o-m$; the equations still contain $o$ unknown
+terms that need to be solved for ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$).
+In order to solve these second order differential equations for the
+$\dot{\mathbf{u}}$ terms, additional equations are needed.
+The acceleration constrain equations fill this role, as there are $m$ of these
+equations, giving:
+\begin{align}
+\label{eq:F_Fstar_f_a}
+\begin{bmatrix}
+\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* \\
+\mathbf{f}_a (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t)
+\end{bmatrix} = \mathbf{0}
+\end{align}
+which gives a total of $o$ equations, enough to be able to solve for the
+$\dot{\mathbf{u}}$'s.
+
+As a final step, the arrangement of the equations as used in this paper is
+shown in Table \ref{table:assumptions}.
+These equations are all drawing from the preceding steps, but have been
+re-organized for convenience (as will be shown later).
+
+\begin{table}[htbp]
+  \centering
+  \caption{Constrained multibody system governing definitions and equations}
+  \label{table:assumptions}
+  \begin{tabular}[c]{l l l}
+    Quantity & Shape & Description\\
+    \hline
+    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^n$ & Coordinates and their time
+    derivatives\\
+    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^o$ & Speeds and their time derivatives\\
+    $\bm{r}$ & $\mathbf{R}^s$ & Exogenous inputs \\
+  \end{tabular}
+  \begin{tabular}[c]{r @{ $=$ } l l}
+    \multicolumn{3}{c}{ } \\
+    \multicolumn{2}{c}{Equation} & Description \\
+    \hline
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration constraints \\
+    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity constraints \\
+    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
+    \times 1}$ & Acceleration constraints \\
+    $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
+    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
+    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential equations
+  \end{tabular}
+\end{table}
+
+The only new terms are $\mathbf{f}_0$, $\mathbf{f}_1$, $\mathbf{f}_2$, and
+$\mathbf{f}_3$.
+The kinematic differential equations are split up into the quantities
+$\mathbf{f}_0$ and $\mathbf{f}_1$, depending on whether each term has a
+$\mathbf{u}$ or $\dot{\mathbf{q}}$ present.
+The same is true for $\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a
+$\dot{\mathbf{u}}$, it is placed into $\mathbf{f}_2$ and if not it is placed
+into $\mathbf{f}_3$.
+
+\section{Dynamics Example}
+TODO: Luke writes here?
 
@@
 \end{align}
-where $P^\prime \in \bm{R}^{(o - m + n - l) \times (o - m + n - l)}$.  Defining
+where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
 $\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square

### 2013-06-02 PetersonGedeHubbard2012.tex
Work on derivation of linearization section.
Linearization derivation section, results section, and parts of the discussion
section need to be merged. Examples still need to be added.

@@
 
+It is critical to note that just taking the gradient of (\ref{eq:F_Fstar_f_a})
+with respect to $\mathbf{u}_i$ does not give the correct linear equations. The
+reasons for this will be covered later.
+
 \section{Dynamics Example}
@@
 
-\section{Results}
-\label{sec:results}
-The linearized equations of motion are
-\begin{align}
-  \label{eq:state_space_constrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
+\section{Derivation of linearization procedure}
+\label{sec:derivations}
+
+In response to the need we have demonstrated, we present a solution to this
+constrained, linearization problem.
+There is a reason that the equations generated by taking the gradient with
+respect to the generalized speeds are incorrect.
+When finding the equations of motion for a constrained system, some generalized
+coordinates and speeds are classified as dependent speeds, meaning their
+behavior is determined by the other coordinates and speeds.
+These terms should in fact be written as:
+\begin{align}
+\label{eq:q_d_redefined}
+\mathbf{q}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t) \\
+\label{eq:u_d_redefined}
+\mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
+\end{align}
+Then, upon taking the derivative of each of the terms with respect to an
+independent speed, the dependent speed would need to be solved for and its
+derivative found.
+No previous authors appear to have made clear that this distinction should be
+made when using Kane's method.
+
+What we present addresses this issue.
+Rather then substituting in the solutions to the dependent speeds though, we
+use knowledge of the structure of the equations and constraints (as defined in
+Table \ref{table:assumptions}) to generate linear equations.
+
+We begin with a first order Taylor series expansion of the equations in Table
+\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
+$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
+assumed that all of the equations in Table \ref{table:assumptions} are
+satisfied by these quantities.  In the interest of brevity, we omit writing
+this point of linearization in each gradient in the calculations below; all are
+evaluated at this point.  Expansion of the three constraint
+equations, keeping only first order terms, yields
+\begin{align}
+  \label{eq:configuration_expansion}
+  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
+  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
+  \label{eq:velocity_expansion}
+  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
+  \label{eq:acceleration_expansion}
+  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
+  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
+t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
+\nabla_{\bm{\dot{q}}}\bm{f}_{a}
+ \delta \bm{\dot{q}} \notag\\
+&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
 \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q}_i \\
-        \delta \bm{u}_i
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged in order to carry out the
-linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
-bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
-are Jacobian matrices of the differential equations. Equations
-(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
-embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
-permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
-independent and dependent speeds.
-
-\begin{table}[htbp]
-  \centering
-  \caption{Constrained multibody system governing definitions and equations}
-  \label{table:assumptions}
-  \begin{tabular}[c]{l l l}
-    Quantity & Shape & Description\\
-    \hline
-    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^n$ & Coordinates and their time
-    derivatives\\
-    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^o$ & Speeds and their time derivatives\\
-    $\bm{r}$ & $\mathbf{R}^s$ & Exogenous inputs \\
-  \end{tabular}
-  \begin{tabular}[c]{r @{ $=$ } l l}
-    \multicolumn{3}{c}{ } \\
-    \multicolumn{2}{c}{Equation} & Description \\
-    \hline
-    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration constraints \\
-    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity constraints \\
-    $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
-    \times 1}$ & Acceleration constraints \\
-    $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
-    $\bm{0}_{n \times 1}$ & Kinematic differential equations \\
-    $\bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-    \bm{u}, \bm{r}, t)$ & $\bm{0}_{(o - m) \times 1}$ & Dynamic differential equations
-  \end{tabular}
-\end{table}
-
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
-of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
-satisfy the equations in Table \ref{table:assumptions}, and may or may not be
-an equilibrium point.
-
+\end{align}
+The first terms are identically zero because of the assumption that the
+point of linearization satisfies the constraints.  The Taylor series expansion
+of the kinematic differential equations is
+\begin{align}
+  \label{eq:f0_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
+  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
+  \label{eq:f1_expansion}
+  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
+  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
+that the sum of the first term on the right hand side of each equation must
+equal zero, we obtain
+\begin{align}
+  \label{eq:f0_plus_f1_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
+  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Similarly, a Taylor series expansion of the dynamic differential equations, we
+obtain
+\begin{align}
+  \label{eq:f2_expansion}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
+      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
+      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
+      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
+  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
+  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
+  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
+  \label{eq:f3_expansion}
+  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
+  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
+  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
+that the sum of the first term on the right hand sides of these equations must
+equal zero, we obtain
+\begin{align}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
+  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
+  \label{eq:f2_plus_f3_expansion}
+  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
+  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+
+Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
+(\ref{eq:acceleration_expansion}),
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
+\ref{table:assumptions}), and introducing the following definitions
 \begin{align}
@@
 \end{align}
+enables the unconstrained linear state space equations to be written as
+\begin{align}
+  \label{eq:state_space_unconstrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       \tilde{A}_{qq} & \tilde{A}_{qu} \\
+       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
+       \tilde{A}_{uqd} & \tilde{A}_{uud}
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q} \\
+        \delta \bm{u}
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n + m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
++ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
+address this issue, a smaller set of independent coordinates and speeds must be
+selected. To this end, consider the following partitioning of the generalized
+coordinates and generalized speeds:
+\begin{equation*}
+  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
+      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
+      \qquad\qquad
+  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
+      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
+\end{equation*}
+where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
+are invertible permutation matrices which map an ordering which has the
+independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
+\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
+($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
+ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
+$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
+similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
+last $m$ columns of $P_u$.
+
+By assumption, equation
+(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
+   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
+  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
+  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
+  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
+    \bm{q}_i\\
+  \label{eq:delta_q}
+  \delta \bm{q} &= C_0 \delta \bm{q}_i
+\end{align}
+
+Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
+  &= \nabla_{\bm{q}} \bm{f}_{v}
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
+\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
+  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
+  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
+  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
+  \label{eq:delta_u}
+  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
+\end{align}
+
+By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
+rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
+\ref{eq:state_space_constrained}, a linear system of equations which relates
+the time derivatives of all states (dependent and independent) to the
+independent states and exogenous inputs.
+
+\section{Results}
+\label{sec:results}
+The linearized equations of motion are
+\begin{align}
+  \label{eq:state_space_constrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q}_i \\
+        \delta \bm{u}_i
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Table \ref{table:assumptions} provides the form in which the nonlinear
+equations of motion and constraints must be arranged in order to carry out the
+linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+are Jacobian matrices of the differential equations. Equations
+(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+independent and dependent speeds.
+
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+an equilibrium point.
 
@@
 
-%f_c:
-%[r*cos(q1) + q5]
-%f_v:
-%[                r*u1*cos(q2) + u3]
-%[-r*(u0*cos(q2) + u2*sin(q2)) + u4]
-%[                r*u1*sin(q2) + u5]
-%f_a:
-%[                                -r*u1*sin(q2)*q2' + r*cos(q2)*u1' + u3']
-%[-r*(-u0*sin(q2)*q2' + u2*cos(q2)*q2' + sin(q2)*u2' + cos(q2)*u0') + u4']
-%[                                 r*u1*cos(q2)*q2' + r*sin(q2)*u1' + u5']
-%f_0:
-%[                                                                                      -sin(q2)*cos(q1)*q0' + cos(q2)*q1']
-%[                                                                                                       sin(q1)*q0' + q2']
-%[                                                                                       sin(q2)*q1' + cos(q1)*cos(q2)*q0']
-%[(sin(q0)*cos(q2) + sin(q1)*sin(q2)*cos(q0))*q4' + (-sin(q0)*sin(q1)*sin(q2) + cos(q0)*cos(q2))*q3' - sin(q2)*cos(q1)*q5']
-%[                                                                -sin(q0)*cos(q1)*q3' + sin(q1)*q5' + cos(q0)*cos(q1)*q4']
-%[ (sin(q0)*sin(q2) - sin(q1)*cos(q0)*cos(q2))*q4' + (sin(q0)*sin(q1)*cos(q2) + sin(q2)*cos(q0))*q3' + cos(q1)*cos(q2)*q5']
-%f_1:
-%[-u0]
-%[-u1]
-%[-u2]
-%[-u3]
-%[-u4]
-%[-u5]
-%f_2:
-%[                  -I*u0' - m*r*cos(q2)*u4']
-%[-J*u1' + m*r*sin(q2)*u5' + m*r*cos(q2)*u3']
-%[                  -I*u2' - m*r*sin(q2)*u4']
-%f_3:
-%[-I*u1*u2 + J*u1*u2 + g*m*r*sin(q1)*cos(q2) - m*r*(-u0*u5 + u2*u3)*cos(q2)]
-%[                m*r*(u0*u4 - u1*u3)*sin(q2) + m*r*(u1*u5 - u2*u4)*cos(q2)]
-%[ I*u0*u1 - J*u0*u1 + g*m*r*sin(q1)*sin(q2) - m*r*(-u0*u5 + u2*u3)*sin(q2)]
-
 \begin{figure}
@@
 \appendix
-\section{Derivation of linearization procedure}
-\label{sec:derivations}
-
-This appendix outlines the derivation of Equation
-(\ref{eq:state_space_constrained}. This equation can be used for simulation,
-stability analysis, or it can be reformulated into the traditional state space
-equations for a variety of tasks; these topics are discussed in Section
-\ref{sec:discussion}.
-
-We begin with a first order Taylor series expansion of the equations in Table
-\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
-$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
-assumed that all of the equations in Table \ref{table:assumptions} are
-satisfied by these quantities.  In the interest of brevity, we omit writing
-this point of linearization in each gradient in the calculations below; all are
-evaluated at this point.  Expansion of the three constraint
-equations, keeping only first order terms, yields
-\begin{align}
-  \label{eq:configuration_expansion}
-  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
-  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
-  \label{eq:velocity_expansion}
-  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
-  \label{eq:acceleration_expansion}
-  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
-  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
-t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
-\nabla_{\bm{\dot{q}}}\bm{f}_{a}
- \delta \bm{\dot{q}} \notag\\
-&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
-\delta \bm{\dot{u}}
-\end{align}
-The first terms are identically zero because of the assumption that the
-point of linearization satisfies the constraints.  The Taylor series expansion
-of the kinematic differential equations is
-\begin{align}
-  \label{eq:f0_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
-  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
-  \label{eq:f1_expansion}
-  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
-  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
-that the sum of the first term on the right hand side of each equation must
-equal zero, we obtain
-\begin{align}
-  \label{eq:f0_plus_f1_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
-  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Similarly, a Taylor series expansion of the dynamic differential equations, we
-obtain
-\begin{align}
-  \label{eq:f2_expansion}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
-      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
-      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
-      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
-  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
-  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
-  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
-  \label{eq:f3_expansion}
-  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
-  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
-  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
-that the sum of the first term on the right hand sides of these equations must
-equal zero, we obtain
-\begin{align}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
-  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
-  \label{eq:f2_plus_f3_expansion}
-  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
-  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-
-Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
-(\ref{eq:acceleration_expansion}),
-and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
-\ref{table:assumptions}), and introducing the following definitions
-enables the unconstrained linear state space equations to be written as
-\begin{align}
-  \label{eq:state_space_unconstrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       \tilde{A}_{qq} & \tilde{A}_{qu} \\
-       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
-       \tilde{A}_{uqd} & \tilde{A}_{uud}
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q} \\
-        \delta \bm{u}
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n + m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
-address this issue, a smaller set of independent coordinates and speeds must be
-selected. To this end, consider the following partitioning of the generalized
-coordinates and generalized speeds:
-\begin{equation*}
-  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
-      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
-      \qquad\qquad
-  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
-      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
-\end{equation*}
-where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
-are invertible permutation matrices which map an ordering which has the
-independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
-\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
-($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
-ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
-$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
-similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
-last $m$ columns of $P_u$.
-
-By assumption, equation
-(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
-   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
-  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
-  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
-    \bm{q}_i\\
-  \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \delta \bm{q}_i
-\end{align}
-
-Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
-  &= \nabla_{\bm{q}} \bm{f}_{v}
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
-\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
-\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
-  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
-  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
-  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
-  \label{eq:delta_u}
-  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
-\end{align}
-
-By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
-rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
-\ref{eq:state_space_constrained}, a linear system of equations which relates
-the time derivatives of all states (dependent and independent) to the
-independent states and exogenous inputs.
-
 % Not sure which bibliography style we are supposed to use

### 2013-06-07 PetersonGedeHubbard2012.tex
Further updates and refinements to doc.
Still needs examples sections.

@@
 
-\title{Symbolic linearization of constrained multibody systems}
-
-\titlerunning{Symbolic linearization of constrained multibody systems}        % if too long for running head
+\title{Symbolic linearization of equations of motion of constrained multibody systems}
+
+\titlerunning{Symbolic linearization of equations of motion of constrained multibody systems}        % if too long for running head
 
@@
 % Task
-To address this issue, we developed a procedure for organizing the constraints
-and motion equations, and subsequently linearizing these equations.
+To address this issue, we developed a procedure for organizing the constraint
+and motion equations and their subsequent linearization.
 This procedure was specifically developed for equations of motion generated by
@@
 motion equations of constrained multibody systems, and illustrate it with the
+% TODO: change name of example
 familiar example of the rolling disc.
@@
 A linear form of the equations of motion is required to apply the most
-frequently used techniques to such systems.
-Normally, obtaining the equations of motion is as simple as taking the Jacobian
-of the right hand side function of the equations of motion with respect to the
-states.
+frequently used techniques for the analyses to such systems.
+Normally, obtaining the linearized equations of motion is as simple as taking
+the Jacobian of the right hand side function of the equations of motion with
+respect to the states.
 However, when the multibody system in questions has constraints, this process
@@
 system.
-It is critical to state that linearizing these equations by computing the
-Jacobian will give incorrect results.
+It is critical to understand linearizing these equations by computing the
+Jacobian gives incorrect results.
 This paper provides a way to correctly generate the linear equations of motion
@@
 
-Multibody systems, rigid or flexible, and particles often have constraints
-which limit how they can move or be configured. In this paper, we will discuss
-three types of constraints: configuration, velocity, and acceleration.
+Multibody systems (rigid or flexible) and particles often have constraints
+which limit how they can move or be configured.
+In this paper, we will discuss three types of constraints: configuration,
+velocity, and acceleration.
 Configuration constraints limit the location or orientation of parts of the
-system, relative to the external world or other parts of the system. Velocity
-constraints limit the speeds at which the configuration can change, either from
-configuration constraints (through time differentiation) or independent
-application. In this paper, acceleration constraints refer to time
-differentiated velocity constraints. Motion constraints most often appear in
-systems where there is rolling without slip or there are closed kinematic
-loops.
+system, relative to the external world or other parts of the system.
+Velocity constraints limit the speeds at which the configuration can change,
+either from configuration constraints (through time differentiation) or
+independent application.
+Velocity constraints most often appear in systems where there is rolling
+without slip or there are closed kinematic loops.
+Within this paper, acceleration constraints refer to time differentiated
+velocity constraints.
 
@@
 Bottema appears to be the first author to show that special considerations are
-needed for linearizing non-holonomic systems \cite{Bottema1949}.
+needed for linearizing nonholonomic systems \cite{Bottema1949}.
 While other authors have examined linearization of nonholonomic systems, we
@@
 \cite{Schwab2003} both have developed techniques for generating equations of
-motion (and linear forms), however both techniques are more numerical in
-nature - they do not provide symbolic equations of motion or symbolic
+motion (and linear forms thereof), however both techniques are more numerical
+in nature - they do not provide symbolic equations of motion or symbolic
 linearized equations.
@@
 % TODO:reiterate limiations of model
-The goal of this paper is to establish concretely the first order relationships
+The goal of this paper is to correctly establish the first order relationships
 (about an arbitrary point of linearization) between a selection of independent
@@
 enough to cover most examples of time-varying constrained multibody systems
-with arbitrary external inputs and arbitrary specified quantities.  The
+with arbitrary external inputs and arbitrary specified quantities.
+% TODO: electronic supplementary material with examples for other methods
+% TODO: rewrite following sentences after e.s.m.
+The
 procedure was created for systems whose equations of motion have been derived
@@
 of the generated equations.
-Kane's method is then briefly used to generate the equations of motion for a
-simple system, a rolling disk with a trailing arm.
+Kane's method is then used to generate the equations of motion for a simple
+system, 
+%TODO: update example name 
+a rolling disk with a trailing arm.
 The need for a different linearization technique is also demonstrated with this
@@
 In response to this need, we present the newly developed linearization
-procedure, followed by its derivation. % maybe flip the order?
-This new linearization procedure is applied to the example system, to provide
-an example of its application.
+procedure and its derivation.
+This new linearization procedure is then performed on the example system to
+provide an example of its application.
 Finally, we discuss other details and nuances of the procedure and its use.
@@
 \label{sec:kane_method}
-To familiarize the reader with Kane's method, three concepts relating to its
-use will be examined.
+To familiarize the reader with Kane's method, some concepts relating to its
+use will be described.
 The first is the use of generalized speeds in addition to generalized
@@
 non-contributing (internal constraint) forces.
-Finally, the mathematical steps required to use Kane's method are shown and
+Then, the mathematical steps required to use Kane's method are shown and
 described.
 
-In using Lagrange's method, generalized coordinates are used to describe the
+When using Lagrange's method, generalized coordinates are used to describe the
 configuration of a system while the time derivatives of the gen. coordinates
@@
 using the time derivatives of the coordinates.
-Using generalized speeds allows for the angular velocity to be written in a
-much simpler form:
+Using generalized speeds allows for the angular velocity to be written in the
+much simpler form
 \begin{align}
@@
 The derivatives of these generalized speeds will then appear in the equations
-of motion, rather than the twice time differentiated generalized coordinates.
-
-A downside of allowing such definitions of system velocities is that the
-kinematic differential equations, which relate the rate of change of the gen.
-coordinates to the gen. speeds, become more complicated.
-This penalty is usually offset by the vastly simplified equations of motion
-that are generated.
-
-Using and understanding Kane's method requires another concept: partial
+of motion, rather than the twice time differentiated generalized coordinates
+($\ddot{\mathbf{q}}$'s).
+
+A downside of allowing such definitions is clear when formulating the kinematic
+differential equations, which relate the rate of change of the generalized
+coordinates to the generalized speeds.
+These equations become much more complicated, but this is usually offset by the
+significantly simpler equations of motion.
+
+Use and understanding of Kane's method requires another concept, partial
 velocities.
@@
 vectors with respect to the generalized speeds.
-They can be defined for both translational and rotational velocities.
+They are defined for both translational and rotational velocities.
 Using the angular velocity of body $B$ in reference frame $N$ as shown in
-(\ref{eq:ex1}), the partial velocities can be written as:
+(\ref{eq:ex1}), the partial velocities can be written
 \[
@@
 \]
-The partial velocities represent the direction of motion associated with each
-generalized speed.
+and it can be seen that the partial velocities represent the direction of
+motion associated with each generalized speed.
 
 By taking the dot product of each partial velocity and both sides of Newton's
-second law, $F=ma$ (or the rotational equivalent), only terms that are
-parallel to the partial velocities are left in the equation.
+second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
+are parallel to the partial velocities are left in the equation.
 One equation is generated for each generalized speed.
@@
 An important byproduct of this is that internal constraint forces
-(non-contributing forces) generated by things such as pin or sliding joints no
-longer need to be considered.
-Assuming the generalized speeds have been defined in an intelligent manner, the
-forces involved in the internal constraints cannot affect the rate of change of
-any generalized speed, and thus do not need to be considered.
+(non-contributing forces) imposed by constructs such as pin or sliding joints
+no longer need to be considered.
 
@@
 removes all terms which are not related to each generalized speed.
-This generally results in simple equations of motion, which are relatively easy
-to form.
+This generally results in simple equations of motion, which are easier to form.
 
@@
 Consider a system composed of rigid bodies $B_1,...,B_g$, particles
-$P_1,...,P_h$, with points of interest $Q_1,...,Q_k$, and reference frames of
-interest $E_1,...,E_c$, all defined relative to the inertial reference frame
-$N$.
-This system has generalized coordinates $q_1,...,q_n$, generalized speeds
+$P_1,...,P_h$, with points of force application $Q_1,...,Q_k$, and reference
+frames of torque application $E_1,...,E_c$, all defined relative to the
+inertial reference frame $N$.
+This system has generalized coordinates $q_1,...,q_n$ and generalized speeds
 $u_1,...,u_o$.
@@
 the system need to be written in the following form.
+
 Translational velocity of particles:
@@
 velocity or angular velocity component which is a prescribed function of time,
-and has no dependence on generalized coordinates or speeds (e.g., a crank
+and has no dependence on generalized speeds (e.g., a crank
 driven at a fixed rate).
@@
 With these definitions established, Kane's method can be applied.
-The output of Kane's method is the equation $\mathbf{F} + \mathbf{F}^* =
-\mathbf{0}$ - the sum of the generalized active forces and generalized inertial
-forces is 0.
+The output of Kane's method is the equation
+\begin{align}
+\label{eq:kanes_eq}
+\mathbf{F} + \mathbf{F}^* = \mathbf{0}
+\end{align}
+or, the sum of the generalized active forces and generalized inertial forces is
+0.
 If Newton's second law were written as $f - ma = 0$, the $\mathbf{F}$ is the
@@
 velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
-- (\ref{eq:rb_rotational}), with the resultant forces/torques, $\bar{R}$/
-$\bar{T}$, and the inertial forces/torques, $\bar{R}^*$/$\bar{T}^*$.
+- (\ref{eq:rb_rotational}), with the resultant forces/torques ($\bar{R}$/
+$\bar{T}$) and the inertial forces/torques ($\bar{R}^*$/$\bar{T}^*$).
 
@@
 Note that distance forces do not need to be transformed into a resultant
-force/torque for any components of the system; they only need to be considered
-at the point of application (e.g., if a force $\bar{f}$ is the only force
-applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$ and
-$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
+force/torque pair for any components of the system; they only need to be
+considered at the point of application (e.g., if a force $\bar{f}$ is the only
+force applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$
+and $\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
 These resultant forces and torques are used to write the generalized active
@@
 point in the derivation of the equations of motion, they are assumed to be in
-the inertial frame, represented here as $N$.
+the inertial frame which is $N$ for this system.
 
@@
 \label{eq:rb_rotational_gen_inertia}
-\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \times \bar{\bar{I}}^{B/B^*} -
-{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\omega^B
+\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \cdot \bar{\bar{I}}^{B/B^*} -
+{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\bar{\omega}^B
 \end{align}
 Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
-$B$ about its mass center $B^*$ \ref{Kane1985}.
+$B$ about its mass center $B^*$ \cite{Kane1985}.
 
@@
 are aligned with the principal axes of $B$, then the following definitions can
-be used:
-\begin{align}
-\alpha_i \triangleq {^N}\alpha^B \cdot \hat{b}_i \quad \quad
-\omega_i \triangleq {^N}\omega^B \cdot \hat{b}_i \quad \quad
+be used
+\begin{align}
+\alpha_i \triangleq {^N}\bar{\alpha}^B \cdot \hat{b}_i \quad \quad
+\omega_i \triangleq {^N}\bar{\omega}^B \cdot \hat{b}_i \quad \quad
 I_i \triangleq \hat{b}_i \cdot \bar{\bar{I}}^{B/B^*} \cdot \hat{b}_i
-\quad \quad for i=1,2,3
+\quad \quad \text{for }i=1,2,3
 \end{align}
@@
 body $B$.
-This allows the inertia torque to be written as:
+This allows the inertia torque to be written
 \begin{align}
 \label{eq:simple_rb_rotational_gen_inertia}
-\bar{T}_B^* = -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1
-              -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2
-              -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3
+\bar{T}_B^* =& -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1\\
+             & -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2\\
+             & -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3\\
 \end{align}
@@
 
-Now the equations $\mathbf{F} + \mathbf{F}^* = \mathbf{0}$ are defined.
-Bu what has been written is valid only if there are no motion constraints.
+Now all the terms in (\ref{eq:kanes_eq}) are defined.
+But this is valid only if there are no motion constraints.
 The equations need to be transformed in order to not violate the behaviors
@@
 
-Kane's Method allows for the equations of motion to be written using both
-independent and dependent speeds, $\mathbf{u}_i$ and $\mathbf{u}_d$, without
-the need for Lagrange multipliers.
+Kane's Method allows for the constrained equations of motion to be written
+using both independent and dependent speeds, $\mathbf{u}_i$ and $\mathbf{u}_d$,
+without the need for Lagrange multipliers.
 Even though both independent and dependent speeds are in the final equations, a
-choice still needs to be made as to which speeds will be the constrained
-speeds.
+choice still needs to be made as to which speeds will be dependent.
 The methods for doing this are outside the scope of this paper, but it is
@@
 To simplify notation, it is assumed that the generalized speeds are already
-ordered in a favorable manner:
+ordered in a favorable manner.
 \begin{align}
@@
 \end{align}
-The matrix $\mathbf{B}$ from (\ref{eq:constraint_B}) needs to be rewritten as
-well:
+The matrix $\mathbf{B}$ from (\ref{eq:constraint_B}) also needs to be
+partitioned.
 \begin{align}
@@
 \end{align}
-Where $\mathbf{B}_i$ has the shape $m \times o - m$ and $\mathbf{B}_d$ has the
-shape $m \times m$. This equation can be rewritten as:
+where $\mathbf{B}_i$ has the shape $m \times o - m$ and $\mathbf{B}_d$ has the
+shape $m \times m$.
+This equation can then be rewritten as
 \begin{align}
@@
 \mathbf{u}_d = -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
-\mathbf{A} \triangle - \mathbf{B}_d^{-1} \mathbf{B}_i \\
+\mathbf{A} \triangleq - \mathbf{B}_d^{-1} \mathbf{B}_i \\
 \mathbf{u}_d = \mathbf{A} \mathbf{u}_i
 \end{align}
-Where $\mathbf{A}$ has the shape $m \times m$.
+where $\mathbf{A}$ has the shape $m \times m$.
 
 Finally, to constrain the generalized active and inertia forces, they need to
-be written as:
+be partitioned in a similar manner.
 \begin{align}
@@
 
-After constraining the equations of motion generated, we have reduced the
-number of equations from $o$ to $o-m$; the equations still contain $o$ unknown
-terms that need to be solved for ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$).
+After constraining the equations of motion, we have reduced the number of
+equations from $o$ to $o-m$; but the equations still contain $o$ unknown terms
+that need to be solved for ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$).
 In order to solve these second order differential equations for the
@@
 
+%TODO: Expand on r (exogenous inputs)
 \begin{table}[htbp]
@@
 The same is true for $\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a
-$\dot{\mathbf{u}}$, it is placed into $\mathbf{f}_2$ and if not it is placed
+$\dot{\mathbf{u}}$ it is placed into $\mathbf{f}_2$ and if not it is placed
 into $\mathbf{f}_3$.
@@
 \label{eq:q_d_redefined}
-\mathbf{q}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t) \\
+\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
 \label{eq:u_d_redefined}
-\mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
-\end{align}
-Then, upon taking the derivative of each of the terms with respect to an
+\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
+\end{align}
+Then, upon taking the derivative of each of these terms with respect to an
 independent speed, the dependent speed would need to be solved for and its
@@
 assumed that all of the equations in Table \ref{table:assumptions} are
-satisfied by these quantities.  In the interest of brevity, we omit writing
-this point of linearization in each gradient in the calculations below; all are
-evaluated at this point.  Expansion of the three constraint
-equations, keeping only first order terms, yields
+satisfied by these quantities (i.e., the system is in a valid state).
+In the interest of brevity, we omit writing this point of linearization in each
+gradient in the calculations below; all are evaluated at this point.
+Expansion of the three constraint equations, keeping only first order terms,
+yields
 \begin{align}
@@
 
-\section{Results}
-\label{sec:results}
 The linearized equations of motion are
@@
 \end{align}
-Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged in order to carry out the
-linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
-bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
-are Jacobian matrices of the differential equations. Equations
-(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
-embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
-permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
-independent and dependent speeds.
-
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
-of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
-satisfy the equations in Table \ref{table:assumptions}, and may or may not be
-an equilibrium point.
+
+where
 
@@
 
+Table \ref{table:assumptions} provides the form in which the nonlinear
+equations of motion and constraints must be arranged in order to carry out the
+linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+are Jacobian matrices of the differential equations. Equations
+(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+independent and dependent speeds.
+
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+an equilibrium point.
+
 \section{Discussion}
 \label{sec:discussion}
-Equation (\ref{eq:state_space_constrained}) doesn't fit into the standard linear
-system framework because it is non-square.  To obtain a square system (useful
-in stability analysis and control system design) we define
+Most modern control techniques assume a system can be written as
+\begin{align}
+\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B}\mathbf{u}
+\end{align}
+but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
+linear system framework because it is non-square.
+To obtain a square system we define
+\begin{align}
+\tilde{A} &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]\\
+\tilde{B} &= 
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+\end{align}
+from (\ref{eq:state_space_constrained}) and
 \begin{align}
@@
 
-The choice of which state variables are taken to be independent may determine
-whether $\nabla_{\bm{q}}\bm{f}_{c} P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}$ are nonsingular.  These matrices depend upon the configuration $\bm{q}$
-and constant parameters; for certain configurations or parameters, it may be
-that the choice of independent state variables cannot be arbitrary.  While some
-systems may permit a choice of independent state variables which are valid for
-all configurations of interest, others may not.  Methods for automatically
-selecting the ``best'' choice of independent state variables are discussed in
-\cite{Reckdahl1996}; they involve computing the SVD of the Jacobian of the
-constraint matrices to determine a set of independent states which will ensure
-the non-singularity of the aforementioned matrices.
-
-This algorithm as been applied to two well studied systems: the rolling disc
-and an extended version of the Whipple bicycle model\cite{Meijaard2007}.
-Detailed model description, derivation, and benchmarking results are presented
-in the electronic supplementary material.
-
-\section{Example}
-\label{example}
-We apply this linearization technique to the well studied rolling disc.  Figure
-\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
-The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
-$J$, and transverse inertia $I$.  To illustrate the concept of dependent
-quantities, we purposefully choose to describe the configuration and velocity
-of the disc with a non-minimal set of coordinates and speeds.  This choice is
-not purely academic.  It is convenient for a physical disc equipped with a rate
-gyroscope and accelerometer to measure these body fixed quantities (or their
-time derivatives) directly, rather than the more traditional set which is
-chosen for analytic convenience.  In more complicated systems with constraints,
-it is difficult (if not impossible) to describe the system with a minimal
-choice of coordinates and speeds; the ideas demonstrated by this simple example
-can be applied to more complicated ones.
-
-The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
-is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
-(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
-location of disc center $C^*$ relative to the inertial origin $N^*$, and the
-velocity of $C^*$ in $N$ are
-\begin{align}
-  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
-  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
-  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
-\end{align}
-If the disc is in contact with the ground, a configuration constraint involving
-the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
-rolls without slip, other constraints relate the generalized speeds $u_i, (i =
-0,\dots,5)$.  The derivations of these constraints and the equations of motion
-are presented in the electronic supplementary material and can also be found in
-the literature \cite{O'Reilly1996}.
-\begin{table}[htbp]
-  \centering
-  \caption{Rolling disc governing definitions and equations}
-  \label{table:rollingdisc}
-  \begin{tabular}[c]{l l l}
-    Quantity & Shape & Description\\
-    \hline
-    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
-    derivatives\\
-    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
-  \end{tabular}
-  \begin{tabular}[c]{r @{ $=$ } l}
-%    \multicolumn{3}{c}{ } \\
-%    \multicolumn{2}{c}{Equation} &           \\
-    \hline
-    % Configuration constraint
-    $\bm{f}_c$ & $rc_1 + q_5$ \\
-    % Velocity constraints
-    $\bm{f}_v$ & $\left[\begin{array}{c}
-        r u_1 c_2 + u_3 \\
-        -r (u_0 c_2 + u_2 s_2) + u_4 \\
-        r u_1 s_2 + u_5
-    \end{array}\right] $ \\
-    % Acceleration constraints
-    $\bm{f}_a$ & $\left[\begin{array}{c}
-        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
-        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
-      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
-      \end{array}\right]$ \\
-    % Kinematic differential equations
-    $\bm{f}_0$ &
-        $\left[\begin{array}{c}
--\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
-\dot{q}_0 s_1 + \dot{q}_2 \\
-\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
-\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
-+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
-- \dot{q}_5 s_2 c_1\\
-- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
-\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
-         \end{array}\right]$ \\
-    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
-             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
-    % Dynamic differential equations
-  $\bm{f}_2$ & $\left[\begin{array}{c}
--I \dot{u}_0 - m r c_2 \dot{u}_4 \\
--J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
--I \dot{u}_2 - m r s_2 \dot{u}_4
-                   \end{array}\right]$ \\
-
-  $\bm{f}_3$ & $\left[\begin{array}{c}
--I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
-                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
- I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
-                   \end{array}\right]$
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \centering
-  \includegraphics{rollingdisc.pdf}
-  \caption{Rolling disc configuration}
-  \label{fig:disc}
-\end{figure}
-
-The matrices which make up the linear equations were created using the method
-presented in Section \ref{sec:results}. It is important to note that the
-entries in these matrices must be evaluated at the point of linearization
-($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
-linearization must satisfy the equations in Table \ref{table:assumptions}.
-There is no requirement that this point be an equilibrium point, which implies
-that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
-reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
-means that the other quantities in the linearized equations (other $q$'s, the
-$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
-quantities already set to 0.
-
-The independent state vector is:
-\begin{align}
-\begin{bmatrix}
-\delta \bm{q}_i \\
-\delta \bm{u}_i
-\end{bmatrix} =
-\begin{bmatrix}
-q_0 &
-q_1 &
-q_2 &
-q_3 &
-q_4 &
-u_0 &
-u_1 &
-u_2
-\end{bmatrix}^T
-\end{align}
-
-And the complete (independent and dependent) state vector is:
-\begin{align}
-\begin{bmatrix}
-\delta \bm{q} \\
-\delta \bm{u}
-\end{bmatrix} =
-\left[
-\begin{array}{cccccccccccc}
-q_0 &
-q_1 &
-q_2 &
-q_3 &
-q_4 &
-q_5 &
-u_0 &
-u_1 &
-u_2 &
-u_3 &
-u_4 &
-u_5
-\end{array}
-\right]^T
-\end{align}
-
-The coefficient matrices of the linearized equations (Equation
-\ref{eq:state_space_constrained}) are:
-\begin{equation}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]=
-\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
-\end{equation}
-
-\begin{equation*}
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]=
-\end{equation*}
-\begin{equation}
-\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
-\end{equation}
-
-With this example, we have used our linearization procedure to find linear
-equations of motion which respect the system's constraints. Parametric
-stability analysis (eigenvalue analysis) can now be performed on these
-equations.  These computations were implemented symbolically with the open
-source symbolic manipulator SymPy \cite{SymPy2012}.
-
+It is notable that there is a second opportunity (after constraining the
+nonlinear equations of motion) to determine which speeds are dependent.
+Now that the equations are linear, the choice of which state variables are
+taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
+P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
+These matrices depend upon the configuration $\bm{q}$ and constant parameters;
+for certain configurations or parameters, it may be that the choice of
+independent state variables cannot be arbitrary.
+While some systems may permit a choice of independent state variables which are
+valid for all configurations of interest, others may not.
+Methods for automatically selecting the ``best'' choice of independent state
+variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
+of the Jacobian of the constraint matrices to determine a set of independent
+states which will ensure the non-singularity of the aforementioned matrices.
+
+%\section{Example}
+%\label{example}
+%We apply this linearization technique to the well studied rolling disc.  Figure
+%\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
+%The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
+%$J$, and transverse inertia $I$.  To illustrate the concept of dependent
+%quantities, we purposefully choose to describe the configuration and velocity
+%of the disc with a non-minimal set of coordinates and speeds.  This choice is
+%not purely academic.  It is convenient for a physical disc equipped with a rate
+%gyroscope and accelerometer to measure these body fixed quantities (or their
+%time derivatives) directly, rather than the more traditional set which is
+%chosen for analytic convenience.  In more complicated systems with constraints,
+%it is difficult (if not impossible) to describe the system with a minimal
+%choice of coordinates and speeds; the ideas demonstrated by this simple example
+%can be applied to more complicated ones.
+%
+%The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
+%is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
+%(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
+%location of disc center $C^*$ relative to the inertial origin $N^*$, and the
+%velocity of $C^*$ in $N$ are
+%\begin{align}
+%  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
+%  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
+%  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
+%\end{align}
+%If the disc is in contact with the ground, a configuration constraint involving
+%the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
+%rolls without slip, other constraints relate the generalized speeds $u_i, (i =
+%0,\dots,5)$.  The derivations of these constraints and the equations of motion
+%are presented in the electronic supplementary material and can also be found in
+%the literature \cite{O'Reilly1996}.
+%\begin{table}[htbp]
+%  \centering
+%  \caption{Rolling disc governing definitions and equations}
+%  \label{table:rollingdisc}
+%  \begin{tabular}[c]{l l l}
+%    Quantity & Shape & Description\\
+%    \hline
+%    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
+%    derivatives\\
+%    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
+%  \end{tabular}
+%  \begin{tabular}[c]{r @{ $=$ } l}
+%%    \multicolumn{3}{c}{ } \\
+%%    \multicolumn{2}{c}{Equation} &           \\
+%    \hline
+%    % Configuration constraint
+%    $\bm{f}_c$ & $rc_1 + q_5$ \\
+%    % Velocity constraints
+%    $\bm{f}_v$ & $\left[\begin{array}{c}
+%        r u_1 c_2 + u_3 \\
+%        -r (u_0 c_2 + u_2 s_2) + u_4 \\
+%        r u_1 s_2 + u_5
+%    \end{array}\right] $ \\
+%    % Acceleration constraints
+%    $\bm{f}_a$ & $\left[\begin{array}{c}
+%        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
+%        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
+%      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
+%      \end{array}\right]$ \\
+%    % Kinematic differential equations
+%    $\bm{f}_0$ &
+%        $\left[\begin{array}{c}
+%-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
+%\dot{q}_0 s_1 + \dot{q}_2 \\
+%\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
+%\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+%+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
+%- \dot{q}_5 s_2 c_1\\
+%- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
+%\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
+%         \end{array}\right]$ \\
+%    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
+%             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
+%    % Dynamic differential equations
+%  $\bm{f}_2$ & $\left[\begin{array}{c}
+%-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
+%-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
+%-I \dot{u}_2 - m r s_2 \dot{u}_4
+%                   \end{array}\right]$ \\
+%
+%  $\bm{f}_3$ & $\left[\begin{array}{c}
+%-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
+%                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
+% I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
+%                   \end{array}\right]$
+%  \end{tabular}
+%\end{table}
+%
+%
+%\begin{figure}
+%  \centering
+%  \includegraphics{rollingdisc.pdf}
+%  \caption{Rolling disc configuration}
+%  \label{fig:disc}
+%\end{figure}
+%
+%The matrices which make up the linear equations were created using the method
+%presented in Section \ref{sec:results}. It is important to note that the
+%entries in these matrices must be evaluated at the point of linearization
+%($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
+%linearization must satisfy the equations in Table \ref{table:assumptions}.
+%There is no requirement that this point be an equilibrium point, which implies
+%that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
+%reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
+%means that the other quantities in the linearized equations (other $q$'s, the
+%$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
+%quantities already set to 0.
+%
+%The independent state vector is:
+%\begin{align}
+%\begin{bmatrix}
+%\delta \bm{q}_i \\
+%\delta \bm{u}_i
+%\end{bmatrix} =
+%\begin{bmatrix}
+%q_0 &
+%q_1 &
+%q_2 &
+%q_3 &
+%q_4 &
+%u_0 &
+%u_1 &
+%u_2
+%\end{bmatrix}^T
+%\end{align}
+%
+%And the complete (independent and dependent) state vector is:
+%\begin{align}
+%\begin{bmatrix}
+%\delta \bm{q} \\
+%\delta \bm{u}
+%\end{bmatrix} =
+%\left[
+%\begin{array}{cccccccccccc}
+%q_0 &
+%q_1 &
+%q_2 &
+%q_3 &
+%q_4 &
+%q_5 &
+%u_0 &
+%u_1 &
+%u_2 &
+%u_3 &
+%u_4 &
+%u_5
+%\end{array}
+%\right]^T
+%\end{align}
+%
+%The coefficient matrices of the linearized equations (Equation
+%\ref{eq:state_space_constrained}) are:
+%\begin{equation}
+%  \left[
+%    \begin{array}{cc}
+%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+%      \tilde{M}_{uqd} & \tilde{M}_{uud}
+%    \end{array}
+%    \right]=
+%\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
+%\end{equation}
+%
+%
+%\begin{equation*}
+%   \left[
+%     \begin{array}{cc}
+%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+%     \end{array}
+%   \right]=
+%\end{equation*}
+%\begin{equation}
+%\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
+%\end{equation}
+%
+%With this example, we have used our linearization procedure to find linear
+%equations of motion which respect the system's constraints. Parametric
+%stability analysis (eigenvalue analysis) can now be performed on these
+%equations.  These computations were implemented symbolically with the open
+%source symbolic manipulator SymPy \cite{SymPy2012}.
+%
 \section{Conclusions}
@@
 symbolically or numerically, and handles configuration, velocity, and
-acceleration level constraints.  The procedure has been implemented
+acceleration level constraints.
+
+%TODO: mention if it is compatible with other procedures.
+
+The procedure has been implemented
 symbolically in the \texttt{sympy.physics.mechanics} sub-module of the open

### 2013-07-04 PetersonGedeHubbard2012.tex
Incorporated Mont's changes.

@@
 % Context
-Many common mechanical systems have configuration or velocity constraints.
+Many common dynamic mechanical systems have configuration or velocity constraints.
 % Need  (what we have and what we want)
@@
 frequently used techniques for the analyses to such systems.
-Normally, obtaining the linearized equations of motion is as simple as taking
-the Jacobian of the right hand side function of the equations of motion with
-respect to the states.
+Normally, obtaining the linearized equations of motion is as simple as
+calculating the Jacobian of the right hand side function of the equations of
+motion with respect to the states.
 However, when the multibody system in questions has constraints, this process
@@
 % TODO: Improve literature review.
-Bottema appears to be the first author to show that special considerations are
-needed for linearizing nonholonomic systems \cite{Bottema1949}.
+Bottema appears to be the first author to have shown that special
+considerations are needed for linearizing nonholonomic systems
+\cite{Bottema1949}.
 While other authors have examined linearization of nonholonomic systems, we
@@
 The need for a different linearization technique is also demonstrated with this
-model.
+example.
 In response to this need, we present the newly developed linearization
 procedure and its derivation.
-This new linearization procedure is then performed on the example system to
-provide an example of its application.
+This new linearization procedure is then performed on the system to provide an
+example of its application.
 Finally, we discuss other details and nuances of the procedure and its use.
@@
 When using Lagrange's method, generalized coordinates are used to describe the
-configuration of a system while the time derivatives of the gen. coordinates
-describe the velocities of a system.
+configuration of a system while the time derivatives of the generalized
+coordinates describe the velocities of a system.
 Kane's method allows for the velocities of a system to be written in terms of
@@
 
-The benefit of using both generalized speeds and coordinates can be seen in
-the following example.
+The benefit of choosing generalized speeds to be other than simply the
+derivatives of the corresponding generalized coordinates can be seen in the
+following example.
 If the orientation of a rigid body were to be described using using Euler
-parameters (quaternions), its angular velocity is relatively complicated when
-using the time derivatives of the coordinates.
+parameters as coordinates, its angular velocity is relatively complicated when
+using the time derivatives of the coordinates and requires a velocity
+constraint.
 Using generalized speeds allows for the angular velocity to be written in the
@@
 
-A downside of allowing such definitions is clear when formulating the kinematic
+A drawback of allowing such definitions is clear when formulating the kinematic
 differential equations, which relate the rate of change of the generalized
@@
 Partial velocities are defined as the partial derivatives of the velocity
-vectors with respect to the generalized speeds.
-They are defined for both translational and rotational velocities.
+vectors with respect to the generalized speeds and are defined for both
+translational and rotational velocities.
 Using the angular velocity of body $B$ in reference frame $N$ as shown in
-(\ref{eq:ex1}), the partial velocities can be written
+(\ref{eq:ex1}), the three partial velocities can be written
 \[
@@
 second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
-are parallel to the partial velocities are left in the equation.
+are parallel to the partial velocities remain in the equation.
 One equation is generated for each generalized speed.
 
-An important byproduct of this is that internal constraint forces
-(non-contributing forces) imposed by constructs such as pin or sliding joints
+An important byproduct of this choice is that internal constraint forces
+(non-contributing forces) imposed by objects such as pin or sliding joints
 no longer need to be considered.
@@
 with each generalized speed) with Newton's second law (or Euler's equations)
-removes all terms which are not related to each generalized speed.
+removes all terms which are not related to the rate of change of each
+generalized speed.
 This generally results in simple equations of motion, which are easier to form.
@@
 In order to properly apply Kane's method, the velocities of each component in
-the system need to be written in the following form.
+the system need to be written as linear combinations of the generalized speeds.
 
@@
 
-In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents the
+In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents a
 velocity or angular velocity component which is a prescribed function of time,
 and has no dependence on generalized speeds (e.g., a crank
-driven at a fixed rate).
+driven at a specified rate).
 
@@
 behavior.
-Furthermore, velocity constraints can not be written with $\dot{\mathbf{q}}$
+Furthermore, velocity constraints should not be written with $\dot{\mathbf{q}}$
 terms (the time derivatives of the coordinates).
@@
 \label{eq:constraint_A}
-\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d = \mathbf{0} \\
-\mathbf{B}_d \mathbf{u}_d = - \mathbf{B}_i \mathbf{u}_i \\
-\mathbf{u}_d = -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
-\mathbf{A} \triangleq - \mathbf{B}_d^{-1} \mathbf{B}_i \\
-\mathbf{u}_d = \mathbf{A} \mathbf{u}_i
+\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d &= \mathbf{0} \\
+\mathbf{B}_d \mathbf{u}_d &= - \mathbf{B}_i \mathbf{u}_i \\
+\mathbf{u}_d &= -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
+\mathbf{A} &\triangleq - \mathbf{B}_d^{-1} \mathbf{B}_i \\
+\mathbf{u}_d &= \mathbf{A} \mathbf{u}_i
 \end{align}
@@
 $\dot{\mathbf{u}}$ terms, additional equations are needed.
-The acceleration constrain equations fill this role, as there are $m$ of these
+The acceleration constraint equations fill this role, as there are $m$ of these
 equations, giving:

### 2013-07-04 PetersonGedeHubbard2012.tex
Fix multiply defined references.

--- 2013-06-07 PetersonGedeHubbard2012.tex
@@
 \begin{align}
-\label{eq:rb_rotational}
+\label{eq:rb_rotational_bodies}
 {^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
@@
 \begin{align}
-\label{eq:rb_rotational}
+\label{eq:rb_rotational_frames}
 {^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
@@
 velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
-- (\ref{eq:rb_rotational}), with the resultant forces/torques ($\bar{R}$/
+- (\ref{eq:rb_rotational_frames}), with the resultant forces/torques ($\bar{R}$/
 $\bar{T}$) and the inertial forces/torques ($\bar{R}^*$/$\bar{T}^*$).

### 2013-07-29 PetersonGedeHubbard2012.tex
WIP: work on rolling disc example.

@@
 \section{Dynamics Example}
-TODO: Luke writes here?
-
-\section{Derivation of linearization procedure}
-\label{sec:derivations}
-
-In response to the need we have demonstrated, we present a solution to this
-constrained, linearization problem.
-There is a reason that the equations generated by taking the gradient with
-respect to the generalized speeds are incorrect.
-When finding the equations of motion for a constrained system, some generalized
-coordinates and speeds are classified as dependent speeds, meaning their
-behavior is determined by the other coordinates and speeds.
-These terms should in fact be written as:
-\begin{align}
-\label{eq:q_d_redefined}
-\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
-\label{eq:u_d_redefined}
-\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
-\end{align}
-Then, upon taking the derivative of each of these terms with respect to an
-independent speed, the dependent speed would need to be solved for and its
-derivative found.
-No previous authors appear to have made clear that this distinction should be
-made when using Kane's method.
-
-What we present addresses this issue.
-Rather then substituting in the solutions to the dependent speeds though, we
-use knowledge of the structure of the equations and constraints (as defined in
-Table \ref{table:assumptions}) to generate linear equations.
-
-We begin with a first order Taylor series expansion of the equations in Table
-\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
-$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
-assumed that all of the equations in Table \ref{table:assumptions} are
-satisfied by these quantities (i.e., the system is in a valid state).
-In the interest of brevity, we omit writing this point of linearization in each
-gradient in the calculations below; all are evaluated at this point.
-Expansion of the three constraint equations, keeping only first order terms,
-yields
-\begin{align}
-  \label{eq:configuration_expansion}
-  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
-  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
-  \label{eq:velocity_expansion}
-  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
-  \label{eq:acceleration_expansion}
-  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
-  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
-t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
-\nabla_{\bm{\dot{q}}}\bm{f}_{a}
- \delta \bm{\dot{q}} \notag\\
-&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
-\delta \bm{\dot{u}}
-\end{align}
-The first terms are identically zero because of the assumption that the
-point of linearization satisfies the constraints.  The Taylor series expansion
-of the kinematic differential equations is
-\begin{align}
-  \label{eq:f0_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
-  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
-  \label{eq:f1_expansion}
-  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
-  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
-that the sum of the first term on the right hand side of each equation must
-equal zero, we obtain
-\begin{align}
-  \label{eq:f0_plus_f1_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
-  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Similarly, a Taylor series expansion of the dynamic differential equations, we
-obtain
-\begin{align}
-  \label{eq:f2_expansion}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
-      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
-      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
-      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
-  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
-  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
-  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
-  \label{eq:f3_expansion}
-  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
-  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
-  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
-that the sum of the first term on the right hand sides of these equations must
-equal zero, we obtain
-\begin{align}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
-  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
-  \label{eq:f2_plus_f3_expansion}
-  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
-  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-
-Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
-(\ref{eq:acceleration_expansion}),
-and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
-\ref{table:assumptions}), and introducing the following definitions
-\begin{align}
-\label{eq:quant_to_compute}
-  \begin{array}{llcll}
-    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
-    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
-    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
-    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
-    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
-    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
-    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
-    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
-    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
-    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
-    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
-  \end{array}
-\end{align}
-enables the unconstrained linear state space equations to be written as
-\begin{align}
-  \label{eq:state_space_unconstrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       \tilde{A}_{qq} & \tilde{A}_{qu} \\
-       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
-       \tilde{A}_{uqd} & \tilde{A}_{uud}
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q} \\
-        \delta \bm{u}
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n + m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
-address this issue, a smaller set of independent coordinates and speeds must be
-selected. To this end, consider the following partitioning of the generalized
-coordinates and generalized speeds:
+To illustrate Kane's method in the presence of configuration and velocity
+constraints, we consider a thin disc $C$ rolling without slip on a horizontal
+plane. The dextral unit vectors $\bm{c}_x, \bm{c}_y, \bm{c}_z$ are fixed to $C$
+with $\bm{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ to be the
+disc radius and mass, respectively, and assume the mass distribution of a thin
+disc so that $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2$. We take the inertial
+(Newtonian) frame $N$ to be similarly equipped with dextral unit vectors
+$\bm{n}_x, \bm{n}_y, \bm{n}_z$, with $\bm{n}_z$ perpendicular to the ground
+plane and aligned with local gravitational field ($\bm{n}_z$ points downwards).
+To orient the disc, we first align the $C$ with $N$, then apply successive body
+fixed ZYX rotations: yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$
+and $B$ the two frames associated with the first two intermediate rotations in
+the sequence; we refer to these as the instantaneous yaw and lean frames,
+respectively. The unit vector $\bm{b}_z$ fixed in the $B$ frame is parallel to
+the disc plane and aligned with the position vector from the disc center $C^*$
+to the lowest point of the disc $P^*$. Traditionally, to locate the center of
+the disc $C^*$ relative to the inertial origin $N^*$, two coordinates are used
+to locate the contact point in the ground plane; those two coordinates plus the
+lean angle $q_2$ determine the location of the center of the disc. The
+advantage of this approach is that the choice of coordinates guarantees the
+disc remain in contact with the ground plane. To illustrate how dependent
+coordinates must be accounted for when linearizing the equations of motion, we
+purposefully locate the center of the disc $C^*$ relative to the inertial
+origin $N^*$ with a non-minimal choice of coordinates
 \begin{equation*}
-  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
-      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
-      \qquad\qquad
-  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
-      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
+  \bm{r}^{C^*} = q_4 \bm{n}_x + q_5 \bm{n}_y + q_6 \bm{n}_z
 \end{equation*}
-where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
-are invertible permutation matrices which map an ordering which has the
-independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
-\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
-($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
-ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
-$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
-similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
-last $m$ columns of $P_u$.
-
-By assumption, equation
-(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
-   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
-  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
-  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
-    \bm{q}_i\\
-  \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \delta \bm{q}_i
-\end{align}
-
-Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
-  &= \nabla_{\bm{q}} \bm{f}_{v}
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
-\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
-\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
-  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
-  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
-  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
-  \label{eq:delta_u}
-  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
-\end{align}
-
-By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
-rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
-\ref{eq:state_space_constrained}, a linear system of equations which relates
-the time derivatives of all states (dependent and independent) to the
-independent states and exogenous inputs.
-
-The linearized equations of motion are
-\begin{align}
-  \label{eq:state_space_constrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q}_i \\
-        \delta \bm{u}_i
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-
-where
-
-\begin{align}
-  \label{eq:C_0}
-  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
-  \label{eq:C_1}
-  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \\
-  \label{eq:C_2}
-  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui}\\
-  \label{eq:Pq}
-  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
-  \label{eq:Pu}
-  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
-\end{align}
-
-Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged in order to carry out the
-linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
-bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
-are Jacobian matrices of the differential equations. Equations
-(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
-embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
-permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
-independent and dependent speeds.
-
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
-of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
-satisfy the equations in Table \ref{table:assumptions}, and may or may not be
-an equilibrium point.
-
-\section{Discussion}
-\label{sec:discussion}
-Most modern control techniques assume a system can be written as
-\begin{align}
-\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B}\mathbf{u}
-\end{align}
-but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
-linear system framework because it is non-square.
-To obtain a square system we define
-\begin{align}
-\tilde{A} &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]\\
-\tilde{B} &= 
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-\end{align}
-from (\ref{eq:state_space_constrained}) and
-\begin{align}
-  \label{eq:A_prime}
-    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
-  \label{eq:B_prime}
-    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
-\end{align}
-where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
-\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
-independent states by defining
-\begin{align}
-  \label{eq:P_prime}
-    P^\prime &\triangleq \begin{bmatrix}
-        P_{qi} & \bm{O}_{n \times (o - m)} \\
-        \bm{O}_{o \times (n - l)} & P_{ui}
-    \end{bmatrix} \\
-  \label{eq:A}
-    A &\triangleq P^{\prime T} A^\prime \\
-  \label{eq:B}
-    B &\triangleq P^{\prime T} B^\prime
-\end{align}
-where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
-$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
-state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
-linear systems analyses may applied.  It is worth noting that the rows
-of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
-in the output or measurement equations of a linear state space model, as in the
-case when the controlling or measuring dependent state(s).
-
-It is notable that there is a second opportunity (after constraining the
-nonlinear equations of motion) to determine which speeds are dependent.
-Now that the equations are linear, the choice of which state variables are
-taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
-P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
-These matrices depend upon the configuration $\bm{q}$ and constant parameters;
-for certain configurations or parameters, it may be that the choice of
-independent state variables cannot be arbitrary.
-While some systems may permit a choice of independent state variables which are
-valid for all configurations of interest, others may not.
-Methods for automatically selecting the ``best'' choice of independent state
-variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
-of the Jacobian of the constraint matrices to determine a set of independent
-states which will ensure the non-singularity of the aforementioned matrices.
-
-%\section{Example}
-%\label{example}
+which results in the configuration constraint
+\begin{equation}
+ r\cos{q_2} + q_6 = 0
+\end{equation}
+which must be satisfied for the disc to remain in contact with the ground.
+While this constraint is easily avoided by appropriate choice of coordinates,
+it serves to illustrate considerations which must be made when analyzing
+systems where there is no obvious way to eliminate configuration constraints.
+
+Again for the purpose of illustrating how velocity constraints must be
+accounted for when obtaining linearized equations of motion, we define
+the angular velocity of $C$ and the velocity of $C^*$ relative to $N$ using a
+non-minimal choice of generalized speeds
+\begin{align}
+  \label{w_u}
+  \bm{\omega}^C &= u_1 \bm{c}_x + u_2 \bm{c}_y + u_3 \bm{c}_z \\
+  \label{v_u}
+  \bm{v}^{C^*} &= u_4 \bm{c}_x + u_5 \bm{c}_y + u_6 \bm{c}_z
+\end{align}
+Given these definitions, the velocity of the lowest point of the disc $P^*$ is
+\begin{align*}
+  \bm{v}^{P^*} &= \bm{v}^{C^*} + \bm{\omega}^{C} \times r^{P^*/C^*} \\
+  &= (r u_2 \cos q_3  + u_4) \bm{c}_x
+   + (-r (u_1 \cos q_3  + u_3 \sin q_3) + u_5) \bm{c}_y \\
+   &+ (r u_2 \sin q_3  + u_6) \bm{c}_z
+\end{align*}
+where we have made use of the fact that $\bm{r}^{P^*/C^*} = r\bm{b}_z$.
+Under the assumption of pure rolling this immediately yields three velocity
+constraints
+\begin{align}
+  r u_2 \cos q_3  + u_4 &= 0\\
+  -r (u_1 \cos q_3  + u_3 \sin q_3) + u_5 &= 0\\
+  r u_2 \sin q_3  + u_6 &= 0
+\end{align}
+These equations are linear in the $u_i$'s, nonlinear in the $q_i$'s, involve
+geometric system parameters (but not mass or inertial parameters), and do not
+explicitly involve $\dot{q}_i$'s. This structure will be taken advantage of
+subsequently.
+
+At this point, the reader familiar with the rolling disc problem is certainly
+wondering why we have chosen such a diabolical set of coordinates and speeds.
+In this simple example, there is an \textit{obvious and minimal} choice of
+coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this choice
+will be valid for all parameters and configurations; further, the derivation of
+the equations of motion can be done without ever introducing these quantities),
+for some systems this is not the case (e.g., Stewart's platform, bicycle). The
+numeric values of system parameters and the configurations in which the system
+will operate will determine which coordinates and speeds can be taken to be
+dependent (and hence potentially eliminated by algebraic manipulations).
+However, it may not be possible \textit{apriori} to easily determine which
+coordinates and speeds should be taken as dependent; further, some systems may
+operate in distinct enough regions of the configuration space that one
+\textit{must} use different choices of dependent coordinates and/or speeds as
+the system moves from one region to another. The purpose of our diabolical
+choice of coordinates and speeds is to illustrate, in the context of a familiar
+example, how constraints must be taken into account when linearizing equations
+of motion. It is our hope that by doing this for a well known and relatively
+simple example, readers can apply the same techniques to more complicated
+systems where its use is more appropriate.
+
+Time differentiating the velocity constraints yields
+\begin{align}
+-r u_2 \dot{q}_3 \sin q_3 + r \dot{u}_2 \cos q_3 + \dot{u}_4 &= 0 \\
+r(u_1 \dot{q}_3 \sin q_3 - u_3 \dot{q}_3 \cos q_3) - r \dot{u}_1 \cos q_3 - r \dot{u}_3 \sin q_3 + \dot{u}_5 &= 0 \\
+r u_2 \dot{q}_3 \cos q_3 + r \dot{u}_2 \sin q_3 + \dot{u}_6 &= 0
+\end{align}
+which must also be satisfied during general motions. There are two types of
+terms in these equations; 1) terms involving $\dot{u}_i$'s and $q_i$'s, and 2)
+terms involving $\dot{q}_i$'s, $u_i$'s, and $q_i$'s. This structure will be
+taken advantage of subsequently.
+
+The kinematic differential equations relate the time derivatives of the
+coordinates and the generalized speeds. From the addition theorem for angular
+velocity, we may write the angular velocity of $C$ as
+\begin{equation}
+  \label{w_qdot}
+  \bm{\omega}^C = \dot{q}_1 \bm{n}_z + \dot{q}_2 \bm{a}_x + \dot{q}_3 \bm{b}_y
+\end{equation}
+while time differentiating $r^{C^*}$ in $N$ yields
+\begin{equation}
+  \label{v_qdot}
+  \bm{v}^{C^*} = \dot{q}_4 \bm{n}_x + \dot{q}_5 \bm{n}_y + \dot{q}_6 \bm{n}_z
+\end{equation}
+Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot}),
+and resolving these vector equations into the $B$ frame, and rearranging so
+that all terms appear to the left of the equality sign, we obtain the kinematic
+differential equations
+\begin{align}
+  \dot{q}_2 - u_1 \cos q_3  - u_3 \sin q_3 &= 0 \\
+  \dot{q}_1 \sin q_2 + \dot{q}_3 - u_2 &= 0 \\
+  \dot{q}_1 \cos q_2 + u_1 \sin q_3  - u_3 \cos q_3  &= 0 \\
+  \dot{q}_4 \cos q_1 + \dot{q}_5 \sin q_1 -u_4 \cos q_3 - u_6 \sin q_3 &= 0 \\
+ -\dot{q}_4 \sin q_1 \cos q_2 + \dot{q}_5 \cos q_1 \cos q_2 + \dot{q}_6 \sin q_2 - u_5 &= 0 \\
+  \dot{q}_4 \sin q_1 \sin q_2 - \dot{q}_5 \sin q_2 \cos q_1 + \dot{q}_6 \cos q_2 + u_4 \sin q_3 - u_6 \cos q_3 &= 0
+\end{align}
+Before formulating the dynamic equations, we
+
 %We apply this linearization technique to the well studied rolling disc.  Figure
@@
 %source symbolic manipulator SymPy \cite{SymPy2012}.
+
+\section{Derivation of linearization procedure}
+\label{sec:derivations}
+
+In response to the need we have demonstrated, we present a solution to this
+constrained, linearization problem.
+There is a reason that the equations generated by taking the gradient with
+respect to the generalized speeds are incorrect.
+When finding the equations of motion for a constrained system, some generalized
+coordinates and speeds are classified as dependent speeds, meaning their
+behavior is determined by the other coordinates and speeds.
+These terms should in fact be written as:
+\begin{align}
+\label{eq:q_d_redefined}
+\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
+\label{eq:u_d_redefined}
+\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
+\end{align}
+Then, upon taking the derivative of each of these terms with respect to an
+independent speed, the dependent speed would need to be solved for and its
+derivative found.
+No previous authors appear to have made clear that this distinction should be
+made when using Kane's method.
+
+What we present addresses this issue.
+Rather then substituting in the solutions to the dependent speeds though, we
+use knowledge of the structure of the equations and constraints (as defined in
+Table \ref{table:assumptions}) to generate linear equations.
+
+We begin with a first order Taylor series expansion of the equations in Table
+\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
+$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
+assumed that all of the equations in Table \ref{table:assumptions} are
+satisfied by these quantities (i.e., the system is in a valid state).
+In the interest of brevity, we omit writing this point of linearization in each
+gradient in the calculations below; all are evaluated at this point.
+Expansion of the three constraint equations, keeping only first order terms,
+yields
+\begin{align}
+  \label{eq:configuration_expansion}
+  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
+  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
+  \label{eq:velocity_expansion}
+  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
+  \label{eq:acceleration_expansion}
+  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
+  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
+t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
+\nabla_{\bm{\dot{q}}}\bm{f}_{a}
+ \delta \bm{\dot{q}} \notag\\
+&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
+\delta \bm{\dot{u}}
+\end{align}
+The first terms are identically zero because of the assumption that the
+point of linearization satisfies the constraints.  The Taylor series expansion
+of the kinematic differential equations is
+\begin{align}
+  \label{eq:f0_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
+  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
+  \label{eq:f1_expansion}
+  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
+  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
+that the sum of the first term on the right hand side of each equation must
+equal zero, we obtain
+\begin{align}
+  \label{eq:f0_plus_f1_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
+  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Similarly, a Taylor series expansion of the dynamic differential equations, we
+obtain
+\begin{align}
+  \label{eq:f2_expansion}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
+      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
+      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
+      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
+  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
+  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
+  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
+  \label{eq:f3_expansion}
+  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
+  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
+  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
+that the sum of the first term on the right hand sides of these equations must
+equal zero, we obtain
+\begin{align}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
+  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
+  \label{eq:f2_plus_f3_expansion}
+  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
+  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+
+Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
+(\ref{eq:acceleration_expansion}),
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
+\ref{table:assumptions}), and introducing the following definitions
+\begin{align}
+\label{eq:quant_to_compute}
+  \begin{array}{llcll}
+    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
+    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
+    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
+    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
+    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
+    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
+    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
+    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
+    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
+    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
+    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
+  \end{array}
+\end{align}
+enables the unconstrained linear state space equations to be written as
+\begin{align}
+  \label{eq:state_space_unconstrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       \tilde{A}_{qq} & \tilde{A}_{qu} \\
+       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
+       \tilde{A}_{uqd} & \tilde{A}_{uud}
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q} \\
+        \delta \bm{u}
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n + m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
++ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
+address this issue, a smaller set of independent coordinates and speeds must be
+selected. To this end, consider the following partitioning of the generalized
+coordinates and generalized speeds:
+\begin{equation*}
+  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
+      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
+      \qquad\qquad
+  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
+      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
+\end{equation*}
+where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
+are invertible permutation matrices which map an ordering which has the
+independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
+\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
+($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
+ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
+$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
+similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
+last $m$ columns of $P_u$.
+
+By assumption, equation
+(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
+   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
+  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
+  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
+  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
+    \bm{q}_i\\
+  \label{eq:delta_q}
+  \delta \bm{q} &= C_0 \delta \bm{q}_i
+\end{align}
+
+Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
+  &= \nabla_{\bm{q}} \bm{f}_{v}
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
+\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
+  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
+  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
+  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
+  \label{eq:delta_u}
+  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
+\end{align}
+
+By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
+rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
+\ref{eq:state_space_constrained}, a linear system of equations which relates
+the time derivatives of all states (dependent and independent) to the
+independent states and exogenous inputs.
+
+The linearized equations of motion are
+\begin{align}
+  \label{eq:state_space_constrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q}_i \\
+        \delta \bm{u}_i
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+
+where
+
+\begin{align}
+  \label{eq:C_0}
+  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
+  \label{eq:C_1}
+  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \\
+  \label{eq:C_2}
+  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui}\\
+  \label{eq:Pq}
+  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
+  \label{eq:Pu}
+  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
+\end{align}
+
+Table \ref{table:assumptions} provides the form in which the nonlinear
+equations of motion and constraints must be arranged in order to carry out the
+linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+are Jacobian matrices of the differential equations. Equations
+(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+independent and dependent speeds.
+
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+an equilibrium point.
+
+\section{Discussion}
+\label{sec:discussion}
+Most modern control techniques assume a system can be written as
+\begin{align}
+\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B}\mathbf{u}
+\end{align}
+but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
+linear system framework because it is non-square.
+To obtain a square system we define
+\begin{align}
+\tilde{A} &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]\\
+\tilde{B} &= 
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+\end{align}
+from (\ref{eq:state_space_constrained}) and
+\begin{align}
+  \label{eq:A_prime}
+    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+  \label{eq:B_prime}
+    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
+\end{align}
+where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
+\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
+independent states by defining
+\begin{align}
+  \label{eq:P_prime}
+    P^\prime &\triangleq \begin{bmatrix}
+        P_{qi} & \bm{O}_{n \times (o - m)} \\
+        \bm{O}_{o \times (n - l)} & P_{ui}
+    \end{bmatrix} \\
+  \label{eq:A}
+    A &\triangleq P^{\prime T} A^\prime \\
+  \label{eq:B}
+    B &\triangleq P^{\prime T} B^\prime
+\end{align}
+where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
+$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
+state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
+linear systems analyses may applied.  It is worth noting that the rows
+of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
+in the output or measurement equations of a linear state space model, as in the
+case when the controlling or measuring dependent state(s).
+
+It is notable that there is a second opportunity (after constraining the
+nonlinear equations of motion) to determine which speeds are dependent.
+Now that the equations are linear, the choice of which state variables are
+taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
+P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
+These matrices depend upon the configuration $\bm{q}$ and constant parameters;
+for certain configurations or parameters, it may be that the choice of
+independent state variables cannot be arbitrary.
+While some systems may permit a choice of independent state variables which are
+valid for all configurations of interest, others may not.
+Methods for automatically selecting the ``best'' choice of independent state
+variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
+of the Jacobian of the constraint matrices to determine a set of independent
+states which will ensure the non-singularity of the aforementioned matrices.
+
+%\section{Example}
+%\label{example}
+%We apply this linearization technique to the well studied rolling disc.  Figure
+%\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
+%The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
+%$J$, and transverse inertia $I$.  To illustrate the concept of dependent
+%quantities, we purposefully choose to describe the configuration and velocity
+%of the disc with a non-minimal set of coordinates and speeds.  This choice is
+%not purely academic.  It is convenient for a physical disc equipped with a rate
+%gyroscope and accelerometer to measure these body fixed quantities (or their
+%time derivatives) directly, rather than the more traditional set which is
+%chosen for analytic convenience.  In more complicated systems with constraints,
+%it is difficult (if not impossible) to describe the system with a minimal
+%choice of coordinates and speeds; the ideas demonstrated by this simple example
+%can be applied to more complicated ones.
+%
+%The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
+%is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
+%(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
+%location of disc center $C^*$ relative to the inertial origin $N^*$, and the
+%velocity of $C^*$ in $N$ are
+%\begin{align}
+%  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
+%  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
+%  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
+%\end{align}
+%If the disc is in contact with the ground, a configuration constraint involving
+%the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
+%rolls without slip, other constraints relate the generalized speeds $u_i, (i =
+%0,\dots,5)$.  The derivations of these constraints and the equations of motion
+%are presented in the electronic supplementary material and can also be found in
+%the literature \cite{O'Reilly1996}.
+%\begin{table}[htbp]
+%  \centering
+%  \caption{Rolling disc governing definitions and equations}
+%  \label{table:rollingdisc}
+%  \begin{tabular}[c]{l l l}
+%    Quantity & Shape & Description\\
+%    \hline
+%    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
+%    derivatives\\
+%    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
+%  \end{tabular}
+%  \begin{tabular}[c]{r @{ $=$ } l}
+%%    \multicolumn{3}{c}{ } \\
+%%    \multicolumn{2}{c}{Equation} &           \\
+%    \hline
+%    % Configuration constraint
+%    $\bm{f}_c$ & $rc_1 + q_5$ \\
+%    % Velocity constraints
+%    $\bm{f}_v$ & $\left[\begin{array}{c}
+%        r u_1 c_2 + u_3 \\
+%        -r (u_0 c_2 + u_2 s_2) + u_4 \\
+%        r u_1 s_2 + u_5
+%    \end{array}\right] $ \\
+%    % Acceleration constraints
+%    $\bm{f}_a$ & $\left[\begin{array}{c}
+%        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
+%        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
+%      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
+%      \end{array}\right]$ \\
+%    % Kinematic differential equations
+%    $\bm{f}_0$ &
+%        $\left[\begin{array}{c}
+%-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
+%\dot{q}_0 s_1 + \dot{q}_2 \\
+%\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
+%\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+%+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
+%- \dot{q}_5 s_2 c_1\\
+%- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
+%\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
+%         \end{array}\right]$ \\
+%    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
+%             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
+%    % Dynamic differential equations
+%  $\bm{f}_2$ & $\left[\begin{array}{c}
+%-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
+%-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
+%-I \dot{u}_2 - m r s_2 \dot{u}_4
+%                   \end{array}\right]$ \\
+%
+%  $\bm{f}_3$ & $\left[\begin{array}{c}
+%-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
+%                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
+% I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
+%                   \end{array}\right]$
+%  \end{tabular}
+%\end{table}
+%
+%
+%\begin{figure}
+%  \centering
+%  \includegraphics{rollingdisc.pdf}
+%  \caption{Rolling disc configuration}
+%  \label{fig:disc}
+%\end{figure}
+%
+%The matrices which make up the linear equations were created using the method
+%presented in Section \ref{sec:results}. It is important to note that the
+%entries in these matrices must be evaluated at the point of linearization
+%($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
+%linearization must satisfy the equations in Table \ref{table:assumptions}.
+%There is no requirement that this point be an equilibrium point, which implies
+%that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
+%reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
+%means that the other quantities in the linearized equations (other $q$'s, the
+%$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
+%quantities already set to 0.
+%
+%The independent state vector is:
+%\begin{align}
+%\begin{bmatrix}
+%\delta \bm{q}_i \\
+%\delta \bm{u}_i
+%\end{bmatrix} =
+%\begin{bmatrix}
+%q_0 &
+%q_1 &
+%q_2 &
+%q_3 &
+%q_4 &
+%u_0 &
+%u_1 &
+%u_2
+%\end{bmatrix}^T
+%\end{align}
+%
+%And the complete (independent and dependent) state vector is:
+%\begin{align}
+%\begin{bmatrix}
+%\delta \bm{q} \\
+%\delta \bm{u}
+%\end{bmatrix} =
+%\left[
+%\begin{array}{cccccccccccc}
+%q_0 &
+%q_1 &
+%q_2 &
+%q_3 &
+%q_4 &
+%q_5 &
+%u_0 &
+%u_1 &
+%u_2 &
+%u_3 &
+%u_4 &
+%u_5
+%\end{array}
+%\right]^T
+%\end{align}
+%
+%The coefficient matrices of the linearized equations (Equation
+%\ref{eq:state_space_constrained}) are:
+%\begin{equation}
+%  \left[
+%    \begin{array}{cc}
+%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+%      \tilde{M}_{uqd} & \tilde{M}_{uud}
+%    \end{array}
+%    \right]=
+%\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
+%\end{equation}
+%
+%
+%\begin{equation*}
+%   \left[
+%     \begin{array}{cc}
+%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+%     \end{array}
+%   \right]=
+%\end{equation*}
+%\begin{equation}
+%\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
+%\end{equation}
+%
+%With this example, we have used our linearization procedure to find linear
+%equations of motion which respect the system's constraints. Parametric
+%stability analysis (eigenvalue analysis) can now be performed on these
+%equations.  These computations were implemented symbolically with the open
+%source symbolic manipulator SymPy \cite{SymPy2012}.
 %

### 2013-07-29 rolling_disc_luke.py

@@ -1,5 +1,5 @@
 from sympy import (symbols, Matrix, eye, zeros, pi, trigsimp,
-solve_linear_system_LU, solve)
+        solve_linear_system_LU, solve)
 from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point, dot,
 cross, mprint, RigidBody, inertia, KanesMethod, mlatex)
 from sympy import *
@@ -9,88 +9,94 @@ from sympy.physics.mechanics import *
 t, r, m, g, I, J = symbols('t r m g I J')
 
 # Configuration variables and their time derivatives
-# q[0] -- yaw
-# q[1] -- lean
-# q[2] -- spin
-# q[3] -- disc center distance from inertial origin, N.x direction
-# q[4] -- disc center distance from inertial origin, N.y direction
-# q[5] -- disc center distance from inertial origin, N.z direction
-q = dynamicsymbols('q:6')
-qd = [qi.diff(t) for qi in q]
+# q1 -- yaw
+# q2 -- lean
+# q3 -- spin
+# q4 -- disc center distance from inertial origin, N.x direction
+# q5 -- disc center distance from inertial origin, N.y direction
+# q6 -- disc center distance from inertial origin, N.z direction
+q1, q2, q3, q4, q5, q6 = q = dynamicsymbols('q1:7')
+q1d, q2d, q3d, q4d, q5d, q6d = qd = [qi.diff(t) for qi in q]
 q_zero = {qi : 0 for qi in q}
 qd_zero = {qdi : 0 for qdi in qd}
 
 # Generalized speeds and their time derivatives
-# u[0] -- disc angular velocity component, disc fixed x direction
-# u[1] -- disc angular velocity component, disc fixed y direction
-# u[2] -- disc angular velocity component, disc fixed z direction
-# u[3] -- disc velocity component, disc fixed x direction
-# u[4] -- disc velocity component, disc fixed y direction
-# u[5] -- disc velocity component, disc fixed z direction
+# u1 -- disc angular velocity component, disc fixed x direction
+# u2 -- disc angular velocity component, disc fixed y direction
[…]

### 2013-07-29 PetersonGedeHubbard2012.tex
Work on example.

@@
 With these definitions established, Kane's method can be applied.
-The output of Kane's method is the equation
+The output of Kane's method, in vector form, is the equation
 \begin{align}
@@
 constraints, we consider a thin disc $C$ rolling without slip on a horizontal
-plane. The dextral unit vectors $\bm{c}_x, \bm{c}_y, \bm{c}_z$ are fixed to $C$
-with $\bm{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ to be the
+plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
+with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ to be the
 disc radius and mass, respectively, and assume the mass distribution of a thin
-disc so that $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2$. We take the inertial
+disc so that $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial
 (Newtonian) frame $N$ to be similarly equipped with dextral unit vectors
-$\bm{n}_x, \bm{n}_y, \bm{n}_z$, with $\bm{n}_z$ perpendicular to the ground
-plane and aligned with local gravitational field ($\bm{n}_z$ points downwards).
+$\hat{n}_x, \hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground
+plane and aligned with local gravitational field ($\hat{n}_z$ points downwards).
 To orient the disc, we first align the $C$ with $N$, then apply successive body
@@
 the sequence; we refer to these as the instantaneous yaw and lean frames,
-respectively. The unit vector $\bm{b}_z$ fixed in the $B$ frame is parallel to
+respectively. The unit vector $\hat{b}_z$ fixed in the $B$ frame is parallel to
 the disc plane and aligned with the position vector from the disc center $C^*$
@@
 \begin{equation*}
-  \bm{r}^{C^*} = q_4 \bm{n}_x + q_5 \bm{n}_y + q_6 \bm{n}_z
+  \bar{r}^{C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
 \end{equation*}
@@
   \label{w_u}
-  \bm{\omega}^C &= u_1 \bm{c}_x + u_2 \bm{c}_y + u_3 \bm{c}_z \\
+  {^N}\bar{\omega}^C &= u_1 \hat{c}_x + u_2 \hat{c}_y + u_3 \hat{c}_z \\
   \label{v_u}
-  \bm{v}^{C^*} &= u_4 \bm{c}_x + u_5 \bm{c}_y + u_6 \bm{c}_z
-\end{align}
-Given these definitions, the velocity of the lowest point of the disc $P^*$ is
+  {^N}\bar{v}^{C^*} &= u_4 \hat{c}_x + u_5 \hat{c}_y + u_6 \hat{c}_z
+\end{align}
+It should be immediately obvious from these defintions that the partial angular
+velocity and partial velocity of $C^*$, respectively, are
+\begin{align}
+  {^N}\bar{\omega}^C_u &= [\hat{c}_x, \hat{c}_y, \hat{c}_z, 0, 0, 0] \\
+  {^N}\bar{v}^{C^*}_u &= [0, 0, 0, \hat{c}_x, \hat{c}_y, \hat{c}_z]
+\end{align}
+where we present the partial velocities as a $1\times6$ matrix whose entries
+are populated with standard 3-vectors. Given these definitions, the velocity of
+the lowest point of the disc $P^*$ is
 \begin{align*}
@@
   \label{w_qdot}
-  \bm{\omega}^C = \dot{q}_1 \bm{n}_z + \dot{q}_2 \bm{a}_x + \dot{q}_3 \bm{b}_y
+  {^N}\bar{\omega}^C = \dot{q}_1 \hat{n}_z + \dot{q}_2 \hat{a}_x + \dot{q}_3 \hat{b}_y
 \end{equation}
@@
   \label{v_qdot}
-  \bm{v}^{C^*} = \dot{q}_4 \bm{n}_x + \dot{q}_5 \bm{n}_y + \dot{q}_6 \bm{n}_z
+  \bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
 \end{equation}
@@
 and resolving these vector equations into the $B$ frame, and rearranging so
-that all terms appear to the left of the equality sign, we obtain the kinematic
-differential equations
+that all terms appear to the left of the equality sign, we obtain
 \begin{align}
@@
 \end{align}
-Before formulating the dynamic equations, we
+The angular acceleration of $C$ and the translation acceleration of $C^*$,
+relative to $N$, is
+\begin{align}
+  {^N}\bar{\alpha}^{C} &= \dot{u}_1 \hat{c}_x + \dot{u}_2 \hat{c}_y + \dot{u}_3 \hat{c}_z  \\
+  {^N}\bar{a}^{C^*} &= (u_2 u_6 - u_3 u_5 + \dot{u}_4) \hat{c}_x + (u_3 u_4 - u_1 u_6 +
++ \dot{u}_5) \hat{c}_y \notag\\&+ (u_1 u_5 - u_2 u_4 + \dot{u}_6) \hat{c}_z
+\end{align}
 

### 2013-07-29 rolling_disc_luke.py
More work on example

@@ -6,7 +6,7 @@ from sympy import *
 from sympy.physics.mechanics import *
 
 # Symbols for time and constant parameters
-t, r, m, g, I, J = symbols('t r m g I J')
+t, r, m, g, v = symbols('t r m g v')
 
 # Configuration variables and their time derivatives
 # q1 -- yaw
@@ -44,12 +44,16 @@ C = B.orientnew('C', 'Axis', [q3, B.y])   # Disc fixed frame
 # Inertial angular velocity and angular acceleration of disc fixed frame
 C.set_ang_vel(N, u1*C.x + u2*C.y + u3*C.z)
 C.set_ang_acc(N, u1d*C.x + u2d*C.y + u3d*C.z)
+print("Angular acceleration")
+mprint(C.ang_acc_in(N))
 
 # Points
 NO = Point('NO')                                       # Inertial origin
 CO = Point('CO')                                       # Disc center
 CO.set_vel(N, u4*C.x + u5*C.y + u6*C.z)
-CO.set_acc(N, CO.vel(N).dt(N) + cross(C.ang_vel_in(N), CO.vel(N)))
+CO.set_acc(N, CO.vel(N).dt(C) + cross(C.ang_vel_in(N), CO.vel(N)))
+print("Translational acceleration")
+mprint(CO.acc(N))
 
 P = CO.locatenew('P', r*B.z)                           # Disc-ground contact
 P.v2pt_theory(CO, N, C)
@@ -85,7 +89,6 @@ for kd_i in kindiffs:
 f_0 = kindiffs.subs(u_zero)
 f_1 = kindiffs.subs(qd_zero)
 
-
 # f_0 == f_0_coef_matrix * qd, used to solve for qdots
 f_0_coef_matrix = zeros((6,6))
 for i in range(6):
@@ -96,45 +99,50 @@ for i in range(6):
 f_a = f_v.diff(t)
 print("Acceleration level constraints")
 mprint(f_a)
-stop
+
 
 # Kane's dynamic equations via elbow grease
 # Partial angular velocities and velocities
 partial_w_C = [C.ang_vel_in(N).diff(ui, N) for ui in u]
-partial_v_O = [O.vel(N).diff(ui, N) for ui in u]
+partial_v_CO = [CO.vel(N).diff(ui, N) for ui in u]
 
 # Active forces
 F_O = m*g*A.z
 # Generalized active forces (unconstrained)
[…]

### 2013-07-30 rolling_disc_luke.py
Obtained equilibrium conditions and evaluated M and A matrices.

@@ -19,6 +19,8 @@ q1, q2, q3, q4, q5, q6 = q = dynamicsymbols('q1:7')
 q1d, q2d, q3d, q4d, q5d, q6d = qd = [qi.diff(t) for qi in q]
 q_zero = {qi : 0 for qi in q}
 qd_zero = {qdi : 0 for qdi in qd}
+qd_sym = dict(zip(qd, symbols('qd1:7')))
+qd_sym_inv = {v: k for (k, v) in qd_sym.items()}
 
 # Generalized speeds and their time derivatives
 # u1 -- disc angular velocity component, disc fixed x direction
@@ -74,6 +76,9 @@ mprint(f_v)
 
 # Disc angular velocity in N expressed using time derivatives of coordinates
 w_c_n_qd = q1d*A.z + q2d*B.x + q3d*B.y
+# Disc angular acceleration in N expressed using time derivatives of coordinates
+alpha_c_n_qd = w_c_n_qd.dt(N)
+
 # Disc center velocity in N expressed using time derivatives of coordinates
 v_co_n_qd = q4d*N.x + q5d*N.y + q6d*N.z
 
@@ -117,13 +122,13 @@ mprint(gaf)
 # Inertia force
 R_star_O = -m*CO.acc(N)
 
-JJ = m * r**2 / 2
-II = m * r**2 / 4
+I = m * r**2 / 4
+J = m * r**2 / 2
 
 # Inertia torque
-I_C_CO = inertia(C, II, JJ, II)     # Inertia of disc C about point CO
-T_star_C = (-dot(I_C_CO, C.ang_acc_in(N))
-            - cross(C.ang_vel_in(N), dot(I_C_CO, C.ang_vel_in(N))))
+I_C_CO = inertia(C, I, J, I)     # Inertia of disc C about point CO
+T_star_C = -(dot(I_C_CO, C.ang_acc_in(N))
+             + cross(C.ang_vel_in(N), dot(I_C_CO, C.ang_vel_in(N))))
 
 # Generalized inertia forces (unconstrained)
 gif = [dot(R_star_O, pv) + dot(T_star_C, pav) for pv, pav in
@@ -214,43 +219,122 @@ o = len(u)
 #m = len(udep)
 
 # Point of linearization
-eq_point = {q1: 0,
[…]

### 2013-07-30 rolling_disc_luke.py
Eigenvalues obtained.

@@ -357,3 +357,17 @@ Amat.simplify()
 print("A:")
 mprint(Amat)
 
+upright_nominal = {q1d: 0, q2: 0, m: 1, r: 1, g: 1}
+M_upright = M.subs(upright_nominal)
+A_upright = Amat.subs(upright_nominal)
+A_ss = M_upright.inv() * A_upright
+pprint(A_ss)
+perm = Pqi.row_join(zeros(6, 3)).col_join(zeros(6, 5).row_join(Pui))
+A_ss_red = perm.T * A_ss
+pprint(A_ss_red)
+
+evals = A_ss_red.eigenvals()
+print(evals.keys())
+q3d_check = 1/sqrt(3)
+print([evalf.N(i.subs({q3d: q3d_check})) for i in evals.keys()])
+

### 2013-07-30 rolling_disc_luke.py
WIP: trying to find equilibrium conditions.

@@ -42,6 +42,8 @@ N = ReferenceFrame('N')                   # Inertial frame
 A = N.orientnew('A', 'Axis', [q1, N.z])   # Yaw intermediate frame
 B = A.orientnew('B', 'Axis', [q2, A.x])   # Lean intermediate frame
 C = B.orientnew('C', 'Axis', [q3, B.y])   # Disc fixed frame
+# Disc angular velocity in N expressed using time derivatives of coordinates
+w_c_n_qd = C.ang_vel_in(N)
 
 # Inertial angular velocity and angular acceleration of disc fixed frame
 C.set_ang_vel(N, u1*C.x + u2*C.y + u3*C.z)
@@ -51,7 +53,10 @@ mprint(C.ang_acc_in(N))
 
 # Points
 NO = Point('NO')                                       # Inertial origin
-CO = Point('CO')                                       # Disc center
+CO = NO.locatenew('CO', q4*N.x + q5*N.y + q6*N.z)      # Disc center
+# Disc center velocity in N expressed using time derivatives of coordinates
+v_co_n_qd = CO.pos_from(NO).dt(N)
+
 CO.set_vel(N, u4*C.x + u5*C.y + u6*C.z)
 CO.set_acc(N, CO.vel(N).dt(C) + cross(C.ang_vel_in(N), CO.vel(N)))
 print("Translational acceleration")
@@ -74,13 +79,8 @@ f_v_du = f_v.jacobian(u)
 print("Velocity constraints")
 mprint(f_v)
 
-# Disc angular velocity in N expressed using time derivatives of coordinates
-w_c_n_qd = q1d*A.z + q2d*B.x + q3d*B.y
 # Disc angular acceleration in N expressed using time derivatives of coordinates
 alpha_c_n_qd = w_c_n_qd.dt(N)
-
-# Disc center velocity in N expressed using time derivatives of coordinates
-v_co_n_qd = q4d*N.x + q5d*N.y + q6d*N.z
 
 # Kinematic differential equations
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in B] +
[…]

### 2013-07-30 rolling_disc_luke.py
WIP: more work to find equilibrium.

@@ -58,7 +58,6 @@ CO = NO.locatenew('CO', q4*N.x + q5*N.y + q6*N.z)      # Disc center
 v_co_n_qd = CO.pos_from(NO).dt(N)
 
 CO.set_vel(N, u4*C.x + u5*C.y + u6*C.z)
-CO.set_acc(N, CO.vel(N).dt(C) + cross(C.ang_vel_in(N), CO.vel(N)))
 print("Translational acceleration")
 mprint(CO.acc(N))
 
@@ -79,9 +78,6 @@ f_v_du = f_v.jacobian(u)
 print("Velocity constraints")
 mprint(f_v)
 
-# Disc angular acceleration in N expressed using time derivatives of coordinates
-alpha_c_n_qd = w_c_n_qd.dt(N)
-
 # Kinematic differential equations
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in B] +
                   [dot(v_co_n_qd - CO.vel(N), uv) for uv in B])
@@ -95,10 +91,7 @@ f_0 = kindiffs.subs(u_zero)
 f_1 = kindiffs.subs(qd_zero)
 
 # f_0 == f_0_coef_matrix * qd, used to solve for qdots
-f_0_coef_matrix = zeros((6,6))
-for i in range(6):
-    for j in range(6):
-        f_0_coef_matrix[i, j] = f_0[i].diff(qd[j])
+f_0_coef_matrix = f_0.jacobian(qd)
 
 # Acceleration level constraints                            (Table 1)
 #v_co_n = cross(C.ang_vel_in(N), CO.pos_from(P))
@@ -114,17 +107,17 @@ partial_w_C = [C.ang_vel_in(N).diff(ui, N) for ui in u]
 partial_v_CO = [CO.vel(N).diff(ui, N) for ui in u]
 
 # Active forces
-F_O = m*g*A.z
+F_CO = m*g*A.z
 # Generalized active forces (unconstrained)
-gaf = [dot(F_O, pv) for pv in partial_v_CO]
+gaf = [dot(F_CO, pv) for pv in partial_v_CO]
 print("Generalized active forces (unconstrained)")
 mprint(gaf)
 
 # Inertia force
-R_star_O = -m*CO.acc(N)
+R_star_CO = -m*CO.acc(N)
 
-I = m * r**2 / 4
-J = m * r**2 / 2
+I = (m * r**2) / 4
+J = (m * r**2) / 2
 
 # Inertia torque
[…]

### 2013-07-31 rolling_disc_luke.py
Nearly finished example.

@@ -44,12 +44,10 @@ B = A.orientnew('B', 'Axis', [q2, A.x])   # Lean intermediate frame
 C = B.orientnew('C', 'Axis', [q3, B.y])   # Disc fixed frame
 # Disc angular velocity in N expressed using time derivatives of coordinates
 w_c_n_qd = C.ang_vel_in(N)
+w_b_n_qd = B.ang_vel_in(N)
 
 # Inertial angular velocity and angular acceleration of disc fixed frame
-C.set_ang_vel(N, u1*C.x + u2*C.y + u3*C.z)
-C.set_ang_acc(N, u1d*C.x + u2d*C.y + u3d*C.z)
-print("Angular acceleration")
-mprint(C.ang_acc_in(N))
+C.set_ang_vel(N, u1*B.x + u2*B.y + u3*B.z)
 
 # Points
 NO = Point('NO')                                       # Inertial origin
@@ -58,48 +56,35 @@ CO = NO.locatenew('CO', q4*N.x + q5*N.y + q6*N.z)      # Disc center
 v_co_n_qd = CO.pos_from(NO).dt(N)
 
 CO.set_vel(N, u4*C.x + u5*C.y + u6*C.z)
-print("Translational acceleration")
-mprint(CO.acc(N))
 
 P = CO.locatenew('P', r*B.z)                           # Disc-ground contact
 P.v2pt_theory(CO, N, C)
-P.a2pt_theory(CO, N, C)
 
 # Configuration constraint and its Jacobian w.r.t. q        (Table 1)
 f_c = Matrix([q6 - dot(CO.pos_from(P), N.z)])
 f_c_dq = f_c.jacobian(q)
-print("Configuration constraint")
-mprint(f_c)
 
 # Velocity level constraints                                (Table 1)
 f_v = Matrix([dot(P.vel(N), uv) for uv in C])
 f_v_dq = f_v.jacobian(q)
 f_v_du = f_v.jacobian(u)
-print("Velocity constraints")
-mprint(f_v)
 
 # Kinematic differential equations
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in B] +
-                  [dot(v_co_n_qd - CO.vel(N), uv) for uv in B])
[…]

### 2013-08-01 rolling_disc_luke.py
Fix naive eigenvalue calculations.

@@ -271,8 +271,8 @@ print([evalf.N(i.subs({q3d: q3d_check})) for i in evals.keys()])
 
 zero = (f_0+f_1).col_join(f_2+f_3).col_join(f_a)
 rhs = solve(zero, qd+ud)
-rhs = Matrix([rhs[i].subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv) for i in qd+ud])
-A_ss_bad_sym = rhs.jacobian(q+u)
-A_ss_bad_num = A_ss_bad_sym.subs({q1d: 0, q2: 0, m: 1, r: 1})
+A_ss_bad_sym = Matrix([rhs[i] for i in qd+ud]).jacobian(q+u)
+A_ss_bad_num = A_ss_bad_sym.subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv).subs({q1d: 0, q2: 0, m: 1, r: 1, g: 1})
 nevals = A_ss_bad_num.eigenvals()
 print([evalf.N(i.subs({q3d: q3d_check})) for i in nevals.keys()])
+

### 2013-08-01 PetersonGedeHubbard2012.tex
Rewrote example for new speeds.
Now uses the lean-frame fixed speeds for the rolling disc example. Might need a
little cleanup, and probably the figure.

@@
 \label{eq:definition_Fstar}
-\mathbf{F} =
+\mathbf{F}^* =
 \begin{bmatrix}
@@
 plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
-with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ to be the
-disc radius and mass, respectively, and assume the mass distribution of a thin
-disc so that $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial
-(Newtonian) frame $N$ to be similarly equipped with dextral unit vectors
-$\hat{n}_x, \hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground
-plane and aligned with local gravitational field ($\hat{n}_z$ points downwards).
-To orient the disc, we first align the $C$ with $N$, then apply successive body
-fixed ZYX rotations: yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$
-and $B$ the two frames associated with the first two intermediate rotations in
-the sequence; we refer to these as the instantaneous yaw and lean frames,
-respectively. The unit vector $\hat{b}_z$ fixed in the $B$ frame is parallel to
-the disc plane and aligned with the position vector from the disc center $C^*$
-to the lowest point of the disc $P^*$. Traditionally, to locate the center of
-the disc $C^*$ relative to the inertial origin $N^*$, two coordinates are used
-to locate the contact point in the ground plane; those two coordinates plus the
-lean angle $q_2$ determine the location of the center of the disc. The
-advantage of this approach is that the choice of coordinates guarantees the
-disc remain in contact with the ground plane. To illustrate how dependent
-coordinates must be accounted for when linearizing the equations of motion, we
-purposefully locate the center of the disc $C^*$ relative to the inertial
-origin $N^*$ with a non-minimal choice of coordinates
+with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ as the
+disc radius and mass, and assume a thin disc mass distribution so that
+$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial (Newtonian)
+frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
+\hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
+(downwards) and aligned with local gravitational field. To orient the disc, we
+first align the $C$ with $N$, then apply successive body fixed ZXY rotations:
+yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$ and $B$ the two frames
+associated with the first two intermediate rotations in the sequence; we refer
+to these as the instantaneous yaw and lean frames, respectively. The unit
+vector $\hat{b}_z$ fixed in the $B$ frame is parallel to the disc plane and
+aligned with the position vector from the disc center $C^*$ to the lowest point
+of the disc $P^*$. Traditionally, to locate the center of the disc $C^*$
+relative to the inertial origin $N^*$, two coordinates are used to locate the
+contact point in the ground plane; those two coordinates plus the lean angle
+$q_2$ determine the location of the center of the disc. The advantage of this
+approach is that the choice of coordinates guarantees the disc remain in
+contact with the ground plane. To illustrate how dependent coordinates must be
+accounted for when linearizing the equations of motion, we purposefully locate
+the center of the disc $C^*$ relative to the inertial origin $N^*$ with a
+non-minimal choice of coordinates
 \begin{equation*}
-  \bar{r}^{C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
+  \bar{r}^{N^*C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
 \end{equation*}
-which results in the configuration constraint
+which results in the configuration constraint $\mathbf{f}_c$
 \begin{equation}
+  \label{rd:f_c}
   r\cos{q_2} + q_6 = 0
@@
 
-Again for the purpose of illustrating how velocity constraints must be
-accounted for when obtaining linearized equations of motion, we define
-the angular velocity of $C$ and the velocity of $C^*$ relative to $N$ using a
-non-minimal choice of generalized speeds
-\begin{align}
+Again to demonstrate how velocity constraints must be accounted for when
+obtaining linearized equations of motion, we define the velocity of $C^*$ in
+$N$ with a non-minimal choice of speeds, defined in the body-fixed ($C$) frame.
+The angular velocity of $C$ in $N$, however, is defined in the intermediate
+frame $B$, which is the instantaneous lean frame.
+\begin{align}
+  \label{v_u}
+  {^N}\bar{v}^{C^*} &= u_4 \hat{c}_x + u_5 \hat{c}_y + u_6 \hat{c}_z \\
   \label{w_u}
-  {^N}\bar{\omega}^C &= u_1 \hat{c}_x + u_2 \hat{c}_y + u_3 \hat{c}_z \\
-  \label{v_u}
-  {^N}\bar{v}^{C^*} &= u_4 \hat{c}_x + u_5 \hat{c}_y + u_6 \hat{c}_z
+  {^N}\bar{\omega}^C &= u_1 \hat{b}_x + u_2 \hat{b}_y + u_3 \hat{b}_z
 \end{align}
@@
 \begin{align}
-  {^N}\bar{\omega}^C_u &= [\hat{c}_x, \hat{c}_y, \hat{c}_z, 0, 0, 0] \\
-  {^N}\bar{v}^{C^*}_u &= [0, 0, 0, \hat{c}_x, \hat{c}_y, \hat{c}_z]
+  {^N}\bar{v}^{C^*}_u &= [0, 0, 0, \hat{c}_x, \hat{c}_y, \hat{c}_z] \\
+  {^N}\bar{\omega}^C_u &= [\hat{b}_x, \hat{b}_y, \hat{b}_z, 0, 0, 0] 
 \end{align}
@@
 \begin{align*}
-  \bm{v}^{P^*} &= \bm{v}^{C^*} + \bm{\omega}^{C} \times r^{P^*/C^*} \\
-  &= (r u_2 \cos q_3  + u_4) \bm{c}_x
-   + (-r (u_1 \cos q_3  + u_3 \sin q_3) + u_5) \bm{c}_y \\
-   &+ (r u_2 \sin q_3  + u_6) \bm{c}_z
+    {^N}\bm{v}^{P^*} &= \bm{v}^{C^*} + \bm{\omega}^{C} \times r^{P^*/C^*} \\
+                     &= u_4\hat{c}_x + u_5\hat{c}_y + u_6\hat{c}_z + r u_2
+                        \hat{b}_x - r u_1 \hat{b}_y \\
+                     &= (r u_2 \cos(q_3) + u_4) \hat{c}_x + (-r
+                        u_1 + u_5)\hat{c}_y + r u_2 \sin(q_3) +
+                        u6) \hat{c}_z
 \end{align*}
@@
 Under the assumption of pure rolling this immediately yields three velocity
-constraints
-\begin{align}
-  r u_2 \cos q_3  + u_4 &= 0\\
-  -r (u_1 \cos q_3  + u_3 \sin q_3) + u_5 &= 0\\
-  r u_2 \sin q_3  + u_6 &= 0
-\end{align}
-These equations are linear in the $u_i$'s, nonlinear in the $q_i$'s, involve
-geometric system parameters (but not mass or inertial parameters), and do not
-explicitly involve $\dot{q}_i$'s. This structure will be taken advantage of
-subsequently.
+constraint equations which make up $\mathbf{f}_v = \mathbf{0}$
+\begin{subequations}
+\label{rd:f_v}
+\begin{align}
+    r u_2 \cos(q_3) + u_4 &= 0\\
+            -r u_1 + u_5 &= 0\\
+    r u_2 \sin(q_3) + u_6 &= 0
+\end{align}
+\end{subequations}
+These equations are linear in the $u_i$ terms, nonlinear in the $q_i$ terms,
+involve geometric system parameters (but not mass or inertial parameters), and
+do not explicitly involve $\dot{q}_i$ terms. This structure will be taken
+advantage of subsequently.
 
@@
 wondering why we have chosen such a diabolical set of coordinates and speeds.
-In this simple example, there is an \textit{obvious and minimal} choice of
+In this simple example, there is an {\bf \textit{obvious and minimal}} choice of
 coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this choice
@@
 simple example, readers can apply the same techniques to more complicated
-systems where its use is more appropriate.
-
-Time differentiating the velocity constraints yields
-\begin{align}
--r u_2 \dot{q}_3 \sin q_3 + r \dot{u}_2 \cos q_3 + \dot{u}_4 &= 0 \\
-r(u_1 \dot{q}_3 \sin q_3 - u_3 \dot{q}_3 \cos q_3) - r \dot{u}_1 \cos q_3 - r \dot{u}_3 \sin q_3 + \dot{u}_5 &= 0 \\
-r u_2 \dot{q}_3 \cos q_3 + r \dot{u}_2 \sin q_3 + \dot{u}_6 &= 0
-\end{align}
+systems where its use is more appropriate or necessary.
+
+Time differentiating the velocity constraint equations yields the acceleration
+constraint equations, $\mathbf{f}_a = \mathbf{0}$, written out as:
+\begin{subequations}
+\label{rd:f_a}
+\begin{align}
+    -r u_{2} \sin(q_{3}) \dot{q}_{3} + r \cos(q_{3}) \dot{u}_{2} + \dot{u}_{4} &=
+    0\\
+    - r \dot{u}_{1} + \dot{u}_{5} &= 0\\
+    r u_{2} \cos(q_{3}) \dot{q}_{3} + r \sin(q_{3}) \dot{u}_{2} + \dot{u}_{6} &=
+    0
+\end{align}
+\end{subequations}
 which must also be satisfied during general motions. There are two types of
@@
 The kinematic differential equations relate the time derivatives of the
-coordinates and the generalized speeds. From the addition theorem for angular
-velocity, we may write the angular velocity of $C$ as
+coordinates and the generalized speeds. Here they are solved by equating the
+velocities written with time-differentiated generalized coordinates and the
+velocities written with generalized speeds.  From the addition theorem for
+angular velocity, we may write the angular velocity of $C$ as
 \begin{equation}
@@
 \begin{align}
-  \dot{q}_2 - u_1 \cos q_3  - u_3 \sin q_3 &= 0 \\
-  \dot{q}_1 \sin q_2 + \dot{q}_3 - u_2 &= 0 \\
-  \dot{q}_1 \cos q_2 + u_1 \sin q_3  - u_3 \cos q_3  &= 0 \\
-  \dot{q}_4 \cos q_1 + \dot{q}_5 \sin q_1 -u_4 \cos q_3 - u_6 \sin q_3 &= 0 \\
- -\dot{q}_4 \sin q_1 \cos q_2 + \dot{q}_5 \cos q_1 \cos q_2 + \dot{q}_6 \sin q_2 - u_5 &= 0 \\
-  \dot{q}_4 \sin q_1 \sin q_2 - \dot{q}_5 \sin q_2 \cos q_1 + \dot{q}_6 \cos q_2 + u_4 \sin q_3 - u_6 \cos q_3 &= 0
-\end{align}
-The angular acceleration of $C$ and the translation acceleration of $C^*$,
-relative to $N$, is
-\begin{align}
-  {^N}\bar{\alpha}^{C} &= \dot{u}_1 \hat{c}_x + \dot{u}_2 \hat{c}_y + \dot{u}_3 \hat{c}_z  \\
-  {^N}\bar{a}^{C^*} &= (u_2 u_6 - u_3 u_5 + \dot{u}_4) \hat{c}_x + (u_3 u_4 - u_1 u_6 +
-+ \dot{u}_5) \hat{c}_y \notag\\&+ (u_1 u_5 - u_2 u_4 + \dot{u}_6) \hat{c}_z
-\end{align}
+    \label{rd:f_0_f_1}
+\underbrace{\left[\begin{matrix}\dot{q}_{2}\\\sin\left(q_{2}\right)
+    \dot{q}_{1} + \dot{q}_{3}\\\cos\left(q_{2}\right)
+    \dot{q}_{1}\\\dot{q}_{4}\\\dot{q}_{5}\\\dot{q}_{6}\end{matrix}\right]}_{\mathbf{f}_0}
+    + 
+\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\r u_{1}
+    \sin\left(q_{1}\right) \cos\left(q_{2}\right) +
+    r u_{2} \cos\left(q_{1}\right)\\- r u_{1}
+    \cos\left(q_{1}\right) \cos\left(q_{2}\right) +
+    r u_{2} \sin\left(q_{1}\right)\\- r u_{1}
+    \sin\left(q_{2}\right)\end{matrix}\right]}_{\mathbf{f}_1}
+    = \left[\begin{matrix} 0\\ 0\\ 0\\ 0\\ 0\\ 0\end{matrix}\right]
+\end{align}
+The translation acceleration of $C^*$ and the angular acceleration of $C$,
+relative to $N$, are
+\begin{align}
+    {^N}\bar{a}^{C^*} =& (-(u_1 \sin(q_3) + u_3 \cos(q_3)) u_5 + u_2 u_6 +
+                         \dot{u}_4) \hat{c}_x \notag \\
+                      & + ((u_1 \sin(q_3) + u_3 \cos(q_3)) u_4 - (u_1
+                         \cos(q_3) - u_3 \sin(q_3)) u_6 + \dot{u}_5) \hat{c}_y
+                         \notag\\
+                      & + ((u_1 \cos(q_3) - u_3 \sin(q_3)) u_5 - u_2 u_4 +
+                         \dot{u}_6) \hat{c}_z \\
+    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 \tan(q_2) + \dot{u}_1) \hat{b}_x
+                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 \tan(q2) +
+                         \dot{u}_3) \hat{b}_z
+\end{align}
+
+As stated, the mass of the disc is $m$, and the inertia dyadic of the disc can
+be written as
+\begin{align}
+    \bar{\bar{\mathbf{I}}}^{C/C^*} = \frac{m r^2}{4} \hat{c}_x\hat{c}_x +
+    \frac{m r^2}{2} \hat{c}_y\hat{c}_y + \frac{m r^2}{4} \hat{c}_z\hat{c}_z
+\end{align}
+Using equations (\ref{eq:rb_translational_gen_inertia}) and
+(\ref{eq:rb_rotational_gen_inertia}), $\bar{R}^*_{C^*}$ and $\bar{T}^*_C$ can be
+written.
+\begin{align}
+    \bar{R}^*_{C^*} =& - m (- (u_{1} \sin(q_{3}) + u_{3}
+                       \cos(q_{3})) u_{5} + u_{2} u_{6} +
+                       \dot{u}_{4}) \hat{c}_x \notag \\
+                     & - m ((u_{1} \sin(q_{3}) + u_{3}
+                       \cos(q_{3})) u_{4} - (u_{1}
+                       \cos(q_{3}) - u_{3}
+                       \sin(q_{3})) u_{6} + \dot{u}_{5})
+                       \hat{c}_y \notag \\
+                     & - m ((u_{1} \cos(q_{3}) - u_{3}
+                       \sin(q_{3})) u_{5} - u_{2} u_{4} +
+                   \dot{u}_{6}) \hat{c}_z \\
+    \bar{T}^*_C =& \frac{m r^{2}}{4} (2 u_{1} u_{2} \sin(q_{3})
+                   - u_{1} u_{3} \sin(q_{3}) \tan(q_{2})
+                   + 2 u_{2} u_{3} \cos(q_{3}) \notag\\&\quad\quad- u_{3}^2
+                   \cos(q_{3}) \tan(q_{2}) +
+                   \sin(q_{3}) \dot{u}_{3} - \cos(q_{3})
+                   \dot{u}_{1}) \hat{c}_x \notag \\
+                 & - \frac{m r^{2}}{2} \dot{u}_{2} \hat{c}_y \notag \\
+                 & + \frac{m r^{2}}{4} (- 2 u_{1} u_{2}
+                   \cos(q_{3}) + u_{1} u_{3} \cos(q_{3})
+                   \tan(q_{2}) + 2 u_{2} u_{3}
+                   \sin(q_{3}) \notag\\&\quad\quad - u_{3}^2 \sin(q_{3})
+                   \tan(q_{2}) - \sin(q_{3}) \dot{u}_{1}
+                   - \cos(q_{3}) \dot{u}_{3}) \hat{c}_z
+\end{align}
+
+The generalized active forces for this example are
+\begin{align}
+    \bar{R}_{C^*} &= m g \hat{n}_z \\
+    \bar{T}_C &= 0
+\end{align}
+
+Equations (\ref{eq:definition_F}), (\ref{eq:definition_Fstar}), and
+(\ref{eq:kanes_eq}) can be used to construct the equations of motion.
+For brevity, we omit the results of each step involved, and only write the
+results, after they have been partitioned into the $\mathbf{f}_2$ and
+$\mathbf{f}_3$ vectors.
+\begin{align}
+    \label{rd:f_2_f_3}
+    & \left[\begin{matrix} 0\\ 0\\ 0\end{matrix}\right] =
+    \underbrace{\left[\begin{matrix}- \frac{m r}{4} \left(r \dot{u}_{1} + 4
+                \dot{u}_{5}\right)\\\frac{m r}{2} \left(- r \dot{u}_{2} + 2
+                \sin\left(q_{3}\right) \dot{u}_{6} + 2 \cos\left(q_{3}\right)
+                \dot{u}_{4}\right)\\- \frac{m r^{2}}{4}
+                \dot{u}_{3}\end{matrix}\right]}_{\mathbf{f}_2} + \notag\\
+    &\underbrace{\left[\begin{matrix}
+                m r \left( g \sin(q_2) - u_1 u_4 \sin(q_3) + u_1 u_6 \cos(q_3)
+                    + (\frac{r u_2}{2} - \frac{r u_3}{4} \tan(q_2) - u_4
+                \cos(q_3) - u_6 \sin(q_3)) u_3 \right)
+                \\m r \left(- u_{2} u_{4}
+                \sin\left(q_{3}\right) + u_{2} u_{6}
+                \cos\left(q_{3}\right) - u_{3} u_{5}\right)\\\frac{m r^{2}}{4}
+                \left(- 2 u_{2} + u_{3} \tan\left(q_{2}\right)\right)
+                u_{1}\end{matrix}\right]}_{\mathbf{f}_3} 
+\end{align}
+
+As there are 12 unknown quantities, $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$, we
+need 3 additional equations so solve for the time-derivative terms.
+The acceleration constraints $\mathbf{f}_a$ provide the final 3 equations.
+Combining (\ref{rd:f_0_f_1}), (\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the
+complete second order differential equations can be written.
+\begin{align}
+\label{rd:ode}
+\begin{bmatrix} \mathbf{f}_0 + \mathbf{f}_1\\
+                \mathbf{f}_2 + \mathbf{f}_3\\
+                \mathbf{f}_a \end{bmatrix} = \mathbf{0}
+\end{align}
+
+The naive approach to linearizing these equations of motion would be to solve
+for the ($\dot{q}_i$, $\dot{u}_i$) terms to construct the right-hand side of
+\begin{align}
+\begin{bmatrix}\dot{\mathbf{q}} \\ \dot{\mathbf{u}}\end{bmatrix} =
+    \mathbf{f}(\mathbf{q}, \mathbf{u}, t)
+\end{align}
+and then take the jacobian of $\mathbf{f}$ with respect to $[\mathbf{q}
+\quad \mathbf{u}]^T$. However, this will give the incorrect results. As
+discussed in Schwab, this system has a critical speed where all the eigenvalues
+are 0 under the following conditions: the system is linearized about the
+upright rolling steady-state configuration, and all parameters ($r$, $m$, and
+$g$) are unity. When these conditions are met, the critical speed is
+$v=\frac{1}{\sqrt{3}}$.
+
+When following the naive approach, 8 of the eigenvalues are 0 and 4 are
+non-zero under the same conditions:
+\begin{align}
+\sqrt{-v^2},\quad -\sqrt{-v^2},\quad \sqrt{-\frac{12}{5} v^2 +
+\frac{4}{5}},\quad -\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
+\end{align}
+At the critical speed, two of these non-zero eigenvalues go to zero. But at any
+forward speed, there is a nonzero imaginary eigenvalue pair. Having illustrated
+the failings of the simple approach to linearization of a constrained multibody
+system defined with Kane's method, we will now outline a linearization
+procedure which addresses this issue and then revisit this example applying the
+proper linearization procedure.
 

### 2013-08-01 PetersonGedeHubbard2012.tex
Incorporated Mont's sect 3 changes.

@@
 plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
-with $\hat{c}_y$ perpendicular to the disc plane. We take $r$ and $m$ as the
+with $\hat{c}_y$ perpendicular to the disc plane. Let $r$ and $m$ be the
 disc radius and mass, and assume a thin disc mass distribution so that
-$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. We take the inertial (Newtonian)
+$I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. Take the inertial (Newtonian)
 frame $N$ to be similarly equipped with dextral unit vectors $\hat{n}_x,
 \hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
-(downwards) and aligned with local gravitational field. To orient the disc, we
-first align the $C$ with $N$, then apply successive body fixed ZXY rotations:
-yaw $q_1$, roll $q_2$, and spin $q_3$.  We denote by $A$ and $B$ the two frames
+(downwards) and aligned with the local gravitational field. To orient the disc,
+first align $C$ with $N$, then apply successive body fixed ZXY rotations:
+yaw $q_1$, roll $q_2$, and spin $q_3$.  Denote by $A$ and $B$ the two frames
 associated with the first two intermediate rotations in the sequence; we refer
 to these as the instantaneous yaw and lean frames, respectively. The unit
-vector $\hat{b}_z$ fixed in the $B$ frame is parallel to the disc plane and
+vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disc plane and
 aligned with the position vector from the disc center $C^*$ to the lowest point
-of the disc $P^*$. Traditionally, to locate the center of the disc $C^*$
+of the disc $P^*$.
+
+Traditionally, to locate the center of the disc $C^*$
 relative to the inertial origin $N^*$, two coordinates are used to locate the
-contact point in the ground plane; those two coordinates plus the lean angle
-$q_2$ determine the location of the center of the disc. The advantage of this
-approach is that the choice of coordinates guarantees the disc remain in
-contact with the ground plane. To illustrate how dependent coordinates must be
+contact point in the ground plane; these two coordinates and the lean angle
+and yaw angles determine the location of the center of the disc. The advantage of this
+approach is that the choice of coordinates guarantees that the disc remain in
+contact with the ground plane.
+
+To illustrate how dependent coordinates must be
 accounted for when linearizing the equations of motion, we purposefully locate
@@
 it serves to illustrate considerations which must be made when analyzing
-systems where there is no obvious way to eliminate configuration constraints.
+systems in which there is no obvious way to eliminate configuration constraints.
 
@@
 \end{align}
-It should be immediately obvious from these defintions that the partial angular
-velocity and partial velocity of $C^*$, respectively, are
+The unconstrained partial angular velocity and partial velocity of $C^*$,
+respectively, are then
 \begin{align}
@@
 dependent (and hence potentially eliminated by algebraic manipulations).
-However, it may not be possible \textit{apriori} to easily determine which
+However, it may not be possible \textit{apriori} to determine easily which
 coordinates and speeds should be taken as dependent; further, some systems may
@@
 \textit{must} use different choices of dependent coordinates and/or speeds as
-the system moves from one region to another. The purpose of our diabolical
+the system moves from one region to another. The purpose of our complex
 choice of coordinates and speeds is to illustrate, in the context of a familiar

### 2013-08-02 PetersonGedeHubbard2012.tex
Wordsmithed introduction.

@@
 To address this issue, we developed a procedure for organizing the constraint
-and motion equations and their subsequent linearization.
-This procedure was specifically developed for equations of motion generated by
-Kane's method, but is compatible with any method as long as the following are
-true: only ordinary differential equations are required to describe the system,
-and independent and dependent states remain in the final equations (i.e. no
-DAEs, no substitution of variables).
+and motion equations and their subsequent linearization.  This procedure was
+specifically developed for equations of motion generated by Kane's method, but
+is compatible with any method as long as the following are true: only ordinary
+differential equations are required to describe the system, and independent and
+dependent states remain in the final equations (i.e. no DAEs, no substitution
+of variables).
 % Object of the document
@@
 motion equations of constrained multibody systems, and illustrate it with the
-% TODO: change name of example
 familiar example of the rolling disc.
@@
 \label{sec:intro}
-Designing a feedback controller or performing stability analysis are two common
-analyses performed on models of multibody systems.
-A linear form of the equations of motion is required to apply the most
-frequently used techniques for the analyses to such systems.
-Normally, obtaining the linearized equations of motion is as simple as
-calculating the Jacobian of the right hand side function of the equations of
-motion with respect to the states.
-However, when the multibody system in questions has constraints, this process
-is more complicated.
-If one has used Kane's method \cite{Kane1985} to generate the equations for a
-constrained system, the resulting equations will have a particular form: a set
-of ordinary differential equations containing both the independent and
-dependent states.
-Simple inspection of the equations does not reveal that they represent a
-constrained system - their form is similar to the equations of an unconstrained
-system.
-It is critical to understand linearizing these equations by computing the
-Jacobian gives incorrect results.
-This paper provides a way to correctly generate the linear equations of motion
-for constrained systems which have been found using Kane's method (or any other
-method which results in equations with the same form).
-
-Multibody systems (rigid or flexible) and particles often have constraints
-which limit how they can move or be configured.
-In this paper, we will discuss three types of constraints: configuration,
-velocity, and acceleration.
-Configuration constraints limit the location or orientation of parts of the
-system, relative to the external world or other parts of the system.
-Velocity constraints limit the speeds at which the configuration can change,
-either from configuration constraints (through time differentiation) or
-independent application.
-Velocity constraints most often appear in systems where there is rolling
-without slip or there are closed kinematic loops.
-Within this paper, acceleration constraints refer to time differentiated
-velocity constraints.
+Design of linear state feedback controllers and first order stability analysis
+of equilibria are two of the most common uses of multibody system dynamics
+models. Both of these uses require linearized equations of motion. For
+unconstrained systems, obtaining the linearized equations of motion is as
+simple as arranging the equations of motion in first order form ($\dot{x} =
+f(x, t)$) and evaluating the Jacobian of the right hand side function, with
+respect to the states, evaluated at the equilibria point of interest.  However,
+when the system in questions has constraints, i.e., not all the states are
+independent, this process is not as straightforward.
+
+For a system system with constraints, Kane's dynamical equations\cite{Kane195}
+are first ordinary differential equations (in the generalized speeds) which
+include both the independent and dependent states (assuming the analyst has not
+taken steps to algebraically eliminate states deemed to be dependent). The
+dynamical equations and the kinematical differential equations can be solved
+for the time derivatives of the generalized coordinates and generalized speeds,
+respectively. Taken together, it is tempting to think that this system of first
+order differential equations could be linearized in the same fashion as an
+unconstrained system. Unfortunately, this is not the case. Linearizing these
+equations in the same manner as one would for an unconstrained system (by
+directly computing the Jacobian) is incorrect and will yield incorrect
+linearized equations (and subsequently incorrect eigenvalues, in the case of a
+stability analysis). We illustrate this by means of a familiar example, the
+rolling disc, formulated with a non-traditional set of coordinates and
+generalized speeds which are not all independent. We then provide a systematic
+method to correctly generate the linear equations of motion for constrained
+systems which have been found using Kane's method (or any other method which
+results in equations with the same form). We then apply this method to the
+rolling disc example to illustrate that it yields eigenvalues that match
+previously publish results.
+
+We consider systems subject to three types of constraints: configuration,
+velocity, and acceleration.  Configuration constraints limit the location or
+orientation of parts of the system, relative to the external world or other
+parts of the system.  Velocity constraints limit the speeds at which the
+configuration can change, either from configuration constraints (through time
+differentiation) or independent application.  Velocity constraints most often
+appear in systems where there is rolling without slip or there are closed
+kinematic loops.  In the context of this paper, acceleration constraints arise
+either from time differentiated velocity constraints or similar kinematic
+considerations.
 
@@
 considerations are needed for linearizing nonholonomic systems
-\cite{Bottema1949}.
-While other authors have examined linearization of nonholonomic systems, we
-have found isssues with applying any of them to equations generated by Kane's
-method.
-Kang et al. \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both
-explored linearization, but these methods have been developed for equations of
-motion found using Lagrange's method, which contain Lagrange multipliers (which
-are not present in equations of motion generated from Kane's method).
-Minaker and Rieveley \cite{Minaker2010} and Schwab and Meijaard
-\cite{Schwab2003} both have developed techniques for generating equations of
-motion (and linear forms thereof), however both techniques are more numerical
-in nature - they do not provide symbolic equations of motion or symbolic
-linearized equations.
-These processes, along with others such as Neuman's %TODO: fix reference
-, formulate the linear equations as a step in the equations of motion
-generation process.
-They are not capable of linearizing existing equations of motion.
-
-% TODO:reiterate limiations of model
-The goal of this paper is to correctly establish the first order relationships
-(about an arbitrary point of linearization) between a selection of independent
-coordinates and speeds and the time-derivatives of all coordinates and speeds
-(both independent and dependent). The formalism presented here is generic
-enough to cover most examples of time-varying constrained multibody systems
-with arbitrary external inputs and arbitrary specified quantities.
+\cite{Bottema1949}. While other authors have examined linearization of
+nonholonomic systems, we have found that most are not directly applicable to
+equations formed using Kane's method, and while we do not doubt that they
+\textit{could} be applied correctly, the details of how to do so, as far as the
+authors are aware, have not been explicitly presented. Kang et al.
+\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
+linearization, but these methods have been developed for equations of motion
+found using Lagrange's method, which contain Lagrange multipliers (which are
+not present in equations of motion generated from Kane's method). Further, both
+of these papers lacked a thorough discussion of basic concepts such as how many
+quantities are independent, and in the case of eigenvalue computation, how many
+eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
+Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
+generating equations of motion (and linear forms thereof), however both
+techniques are numerical in nature - they do not provide symbolic equations of
+motion or symbolic linearized equations. While Neuman presents a symbolic (as
+opposed to numeric) formulation for linearization of the Q-matrix formulation
+of the Lagrangian\cite{Neuman1984}, his work makes no mention of constrained
+systems and attempts to contact the author regarding the software (algebraic
+robot modeler) as well as internet searches for the software have been
+fruitless.
+
+% TODO:reiterate limitations of model
+The goal of this paper is to definitively establish how the first time
+derivative of all coordinates and speeds (independent and dependent) depend, to
+first order, upon a selection of independent coordinates and independent
+speeds, for an arbitrary point of linearization. We demonstrate how a reduced
+portion of these linearized equations can be used to perform standard stability
+analysis (i.e., for control system design or linear stability analysis). The
+formalism presented here is generic enough to cover most examples of
+time-varying constrained multibody systems with arbitrary external inputs and
+arbitrary specified quantities.
 % TODO: electronic supplementary material with examples for other methods
 % TODO: rewrite following sentences after e.s.m.
-The
-procedure was created for systems whose equations of motion have been derived
-with Kane's method. Although also applicable to dynamic system equations
-formulated using other methods, it is restricted to systems which can be
-completely described by a set of ODEs (i.e., DAEs needn't be solved).
-
-% TODO:rewrite organization of paper
-The use of Kane's method in generating equations of motion is reviewed first,
-in order to properly orient the reader to the format and some of the qualities
-of the generated equations.
-Kane's method is then used to generate the equations of motion for a simple
-system,
-%TODO: update example name
-a rolling disk with a trailing arm.
-The need for a different linearization technique is also demonstrated with this
-example.
-In response to this need, we present the newly developed linearization
-procedure and its derivation.
-This new linearization procedure is then performed on the system to provide an
-example of its application.
-Finally, we discuss other details and nuances of the procedure and its use.
+The procedure was created for systems whose equations of motion have been
+derived with Kane's method. Although also applicable to dynamic system
+equations formulated using other methods, it is restricted to systems which can
+be completely described by a set of ODEs (i.e., DAEs needn't be solved).
+
+We begin with a review of Kane's method for generating equations of motion is
+reviewed first, in order to properly orient the reader to the format and some
+of the qualities of the generated equations. Next, we apply Kane's method to
+generate the equations of motion for the rolling disk, but we purposefully
+introduce a dependent coordinate and three dependent speeds in the formulation.
+The need for a different linearization technique is demonstrated with this
+example. In response to this need, we present the newly developed linearization
+procedure and its derivation.  This new linearization procedure is then applied
+the same rolling disk example to illustrate its validity and an example of its
+application.  Finally, we discuss other details and nuances of the procedure
+and its use, as well as directions for future work.
 

### 2013-08-02 PetersonGedeHubbard2012.tex
Wordsmith section on Kane's method.

@@
 differential equations are required to describe the system, and independent and
-dependent states remain in the final equations (i.e. no DAEs, no substitution
+dependent states remain in the final equations (i.e. no DAE's, no substitution
 of variables).
@@
 motion equations of constrained multibody systems, and illustrate it with the
-familiar example of the rolling disc.
+familiar example of the rolling disk.
 
@@
 \end{abstract}
-
-%In summary, the procedure is:
-%\begin{enumerate}
-%    \item Describe the system in the form of the equations shown in Table
-%        \ref{table:assumptions}.
-%    \item Determine a point of linearization ($\bm{q}^*$, $\dot{\bm{q}}^*$,
-%      $\bm{u}^*$, $\dot{\bm{u}}^*$, and $\bm{r}^*$) which satisfies the
-%      equations in Table \ref{table:assumptions}.
-%    \item Compute the quantities in equation (\ref{eq:quant_to_compute}).
-%    \item Identify independent coordinates and speeds; form equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
-%        (\ref{eq:C_2}).
-%    \item Form equation (\ref{eq:state_space_constrained}), and if needed,
-%      equations (\ref{eq:A_prime})-(\ref{eq:B}).
-%\end{enumerate}
 
@@
 
-For a system system with constraints, Kane's dynamical equations\cite{Kane195}
-are first ordinary differential equations (in the generalized speeds) which
-include both the independent and dependent states (assuming the analyst has not
-taken steps to algebraically eliminate states deemed to be dependent). The
-dynamical equations and the kinematical differential equations can be solved
-for the time derivatives of the generalized coordinates and generalized speeds,
-respectively. Taken together, it is tempting to think that this system of first
-order differential equations could be linearized in the same fashion as an
-unconstrained system. Unfortunately, this is not the case. Linearizing these
-equations in the same manner as one would for an unconstrained system (by
-directly computing the Jacobian) is incorrect and will yield incorrect
-linearized equations (and subsequently incorrect eigenvalues, in the case of a
-stability analysis). We illustrate this by means of a familiar example, the
-rolling disc, formulated with a non-traditional set of coordinates and
-generalized speeds which are not all independent. We then provide a systematic
-method to correctly generate the linear equations of motion for constrained
-systems which have been found using Kane's method (or any other method which
-results in equations with the same form). We then apply this method to the
-rolling disc example to illustrate that it yields eigenvalues that match
-previously publish results.
-
-We consider systems subject to three types of constraints: configuration,
-velocity, and acceleration.  Configuration constraints limit the location or
-orientation of parts of the system, relative to the external world or other
-parts of the system.  Velocity constraints limit the speeds at which the
-configuration can change, either from configuration constraints (through time
-differentiation) or independent application.  Velocity constraints most often
-appear in systems where there is rolling without slip or there are closed
-kinematic loops.  In the context of this paper, acceleration constraints arise
-either from time differentiated velocity constraints or similar kinematic
-considerations.
+For a system system with constraints, Kane's dynamical equations\cite{Kane1985}
+are first order ODE's (in the generalized speeds) which include both the
+independent and dependent states (assuming the dependents states have not been
+algebraically eliminated). The dynamical equations and the kinematical
+differential equations can be solved for the time derivatives of the
+generalized coordinates and generalized speeds, respectively. Taken together,
+it is tempting to think that this system of first order differential equations
+could be linearized in the same fashion as an unconstrained system.
+Unfortunately, this is not the case. Linearizing these equations in the same
+manner as one would for an unconstrained system (by directly computing the
+Jacobian) is incorrect and will yield incorrect linearized equations (and
+subsequently incorrect eigenvalues, in the case of a stability analysis). We
+illustrate this by means of a familiar example, the rolling disk, formulated
+with a non-traditional set of coordinates and generalized speeds which are not
+all independent. We then provide a systematic method to correctly generate the
+linear equations of motion for constrained systems which have been found using
+Kane's method (or any other method which results in equations with the same
+form). We then apply this method to the rolling disk example to illustrate that
+it yields eigenvalues that match previously published results.
+
+We consider three types of constraints: configuration, velocity, and
+acceleration. Configuration constraints limit the location or orientation of
+parts of the system, relative to the external world or other parts of the
+system. Velocity constraints limit the speeds at which the configuration can
+change, either from configuration constraints (through time differentiation) or
+independent application. Velocity constraints most often appear in systems
+where there is rolling without slip or there are closed kinematic loops. In the
+context of this paper, acceleration constraints arise either from time
+differentiated velocity constraints or similar kinematic considerations.
 
@@
 \cite{Bottema1949}. While other authors have examined linearization of
-nonholonomic systems, we have found that most are not directly applicable to
-equations formed using Kane's method, and while we do not doubt that they
-\textit{could} be applied correctly, the details of how to do so, as far as the
-authors are aware, have not been explicitly presented. Kang et al.
+nonholonomic systems, we have found that most techniques are not directly
+applicable to equations formed using Kane's method, and while we do not doubt
+that they \textit{could} be applied correctly, the details of how to do so, as
+far as the authors are aware, have not been explicitly presented. Kang et al.
 \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
@@
 not present in equations of motion generated from Kane's method). Further, both
-of these papers lacked a thorough discussion of basic concepts such as how many
-quantities are independent, and in the case of eigenvalue computation, how many
+of these papers lack a discussion of basic concepts such as how many quantities
+are independent, and in the case of eigenvalue computation, how many
 eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
@@
 generating equations of motion (and linear forms thereof), however both
-techniques are numerical in nature - they do not provide symbolic equations of
+techniques are numerical in nature -- they do not provide symbolic equations of
 motion or symbolic linearized equations. While Neuman presents a symbolic (as
@@
 systems and attempts to contact the author regarding the software (algebraic
-robot modeler) as well as internet searches for the software have been
+robot modeler -- Arm) as well as internet searches for the software have been
 fruitless.
@@
 The goal of this paper is to definitively establish how the first time
-derivative of all coordinates and speeds (independent and dependent) depend, to
-first order, upon a selection of independent coordinates and independent
-speeds, for an arbitrary point of linearization. We demonstrate how a reduced
-portion of these linearized equations can be used to perform standard stability
-analysis (i.e., for control system design or linear stability analysis). The
-formalism presented here is generic enough to cover most examples of
-time-varying constrained multibody systems with arbitrary external inputs and
-arbitrary specified quantities.
+derivative of all coordinates and speeds (independent \textit{and} dependent)
+depend, to first order, upon a selection of independent coordinates and
+independent speeds, for an arbitrary point of linearization. We demonstrate how
+a reduced portion of these linearized equations can be used to perform standard
+stability analysis (i.e., for control system design or linear stability
+analysis). The formalism presented here is generic enough to cover most
+examples of time-varying constrained multibody systems with arbitrary external
+inputs and arbitrary specified quantities.
 % TODO: electronic supplementary material with examples for other methods
@@
 equations formulated using other methods, it is restricted to systems which can
-be completely described by a set of ODEs (i.e., DAEs needn't be solved).
+be completely described by a set of ODE's (i.e., DAE's needn't be solved).
 
@@
 \label{sec:kane_method}
-To familiarize the reader with Kane's method, some concepts relating to its
-use will be described.
-The first is the use of generalized speeds in addition to generalized
-coordinates.
-The second concept is in how these generalized speeds can be used to project
-permissible motions of the system, and how this removes the need to consider
-non-contributing (internal constraint) forces.
-Then, the mathematical steps required to use Kane's method are shown and
-described.
+To familiarize the reader with Kane's method, some concepts relating to its use
+will be described.  The first is the use of generalized speeds in addition to
+generalized coordinates.  The second concept is in how these generalized speeds
+can be used to project permissible motions of the system, and how this removes
+the need to consider non-contributing (internal constraint) forces.  Then, the
+mathematical steps required to use Kane's method are shown and described.
 
@@
 configuration of a system while the time derivatives of the generalized
-coordinates describe the velocities of a system.
-Kane's method allows for the velocities of a system to be written in terms of
-generalized speeds, which are not required to be the time derivatives of
-coordinates (although such a definition is permitted).
+coordinates describe the velocities of a system.  Kane's method allows for the
+velocities of a system to be written in terms of generalized speeds, which are
+not required to be the time derivatives of coordinates (although such a
+definition is permitted).
 
@@
 derivatives of the corresponding generalized coordinates can be seen in the
-following example.
-If the orientation of a rigid body were to be described using using Euler
-parameters as coordinates, its angular velocity is relatively complicated when
-using the time derivatives of the coordinates and requires a velocity
-constraint.
-Using generalized speeds allows for the angular velocity to be written in the
-much simpler form
+following example. If the orientation of a rigid body were to be described
+using using Euler parameters or quaternions as coordinates, its angular
+velocity in terms of time derivatives of those coordinates is relatively
+complicated; it also requires a velocity constraint. Using generalized speeds
+allows for the angular velocity to be expressed more simply as
 \begin{align}
@@
 
-A drawback of allowing such definitions is clear when formulating the kinematic
-differential equations, which relate the rate of change of the generalized
-coordinates to the generalized speeds.
-These equations become much more complicated, but this is usually offset by the
-significantly simpler equations of motion.
-
-Use and understanding of Kane's method requires another concept, partial
-velocities.
-Partial velocities are defined as the partial derivatives of the velocity
-vectors with respect to the generalized speeds and are defined for both
-translational and rotational velocities.
-Using the angular velocity of body $B$ in reference frame $N$ as shown in
-(\ref{eq:ex1}), the three partial velocities can be written
+A drawback of allowing such definitions can be seen when formulating the
+kinematic differential equations, which relate the rate of change of the
+generalized coordinates to the generalized speeds. These equations become more
+complicated (in comparison to $\dot{q}_i = u_i$), but this is usually offset by
+the significantly simpler dynamical equations of motion\cite{Mitiguy1996}
+
+Use and understanding of Kane's method requires the use and understanding of
+partial velocities and partial angular velocities. These are simply the the
+partial derivatives of the (angular) velocity vectors with respect to the
+generalized speeds. For example, for the angular velocity in equation
+(\ref{eq:ex1}), the partial angular velocities of body $B$ in reference frame
+$N$ are
 \[
@@
 and it can be seen that the partial velocities represent the direction of
-motion associated with each generalized speed.
-
-By taking the dot product of each partial velocity and both sides of Newton's
+motion associated with each generalized speed. The geometric interpretations of
+these partial (angular) velocities are discussed at length by
+Lesser\cite{Lesser1992}.
+
+Taking the dot product of each partial velocity and both sides of Newton's
 second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
-are parallel to the partial velocities remain in the equation.
-One equation is generated for each generalized speed.
-
-An important byproduct of this choice is that internal constraint forces
-(non-contributing forces) imposed by objects such as pin or sliding joints
-no longer need to be considered.
+are parallel to the partial velocities remain; one scalar equation is generated
+for each generalized speed. An important byproduct is that internal constraint forces
+(non-contributing forces) imposed by objects such as pin or sliding joints no
+longer need to be considered. The reason for this is that partial velocities
+are \textit{by construction} perpendicular to forces of constraint. Constraint
+forces can be included in the formulation, but will not appear in the final
+equations of motion because of this very desirable property.
 
 In summary, Kane's method allows for velocities to be defined in terms of
-generalized speeds, allowing for simpler velocity expressions.
-The dot product of each partial velocity (the direction of motion associated
-with each generalized speed) with Newton's second law (or Euler's equations)
-removes all terms which are not related to the rate of change of each
-generalized speed.
+generalized speeds, allowing for simpler velocity expressions.  The dot product
+of each partial velocity (the direction of motion associated with each
+generalized speed) with Newton's second law (or Euler's equations) removes all
+terms which are not related to the rate of change of each generalized speed.
 This generally results in simple equations of motion, which are easier to form.
@@
 The mathematical formalism for applying Kane's method to multibody systems
-follows.
-Consider a system composed of rigid bodies $B_1,...,B_g$, particles
-$P_1,...,P_h$, with points of force application $Q_1,...,Q_k$, and reference
+follows. Consider a system composed of particles $P_1,...,P_h$, rigid bodies
+$B_1,...,B_g$, with points of force application $Q_1,...,Q_k$, and reference
 frames of torque application $E_1,...,E_c$, all defined relative to the
-inertial reference frame $N$.
-This system has generalized coordinates $q_1,...,q_n$ and generalized speeds
-$u_1,...,u_o$.
-Additionally, there are $l$ configuration constraints and $m$ velocity
-constraints.
+inertial reference frame $N$. This system has generalized coordinates
+$q_1,...,q_n$ and generalized speeds $u_1,...,u_o$. Additionally, there are $l$
+configuration constraints and $m$ velocity constraints.
 
@@
 the system need to be written as linear combinations of the generalized speeds.
-
-Translational velocity of particles:
+The translational velocity of the particles are
 \begin{align}
@@
 \end{align}
-Translational velocity of rigid body mass centers:
+while the translational velocity of rigid body mass centers are
 \begin{align}
@@
 \end{align}
-Translational velocity of points of interest:
+and finally the translational velocity of points which have applied forces are
 \begin{align}
@@
 \end{align}
-Rotational velocity of rigid bodies:
+Next, the angular velocity of rigid bodies are
 \begin{align}
@@
 \end{align}
-Rotational velocity of reference frames of interest:
+and the angular velocity of reference frames which have applied torques are
 \begin{align}
@@
 velocity or angular velocity component which is a prescribed function of time,
-and has no dependence on generalized speeds (e.g., a crank
-driven at a specified rate).
-
-For this system, the configuration constraints are defined as:
+and has no dependence on generalized speeds (e.g., a motor or crank driven at a
+specified rate).
+
+For this system, we consider configuration constraints of the form
 \begin{align}
@@
 where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
-The velocity constraints are defined as:
+The velocity constraints are taken to have the form
 \begin{align}
@@
 \end{align}
-where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$.
-It is important that the velocity constraints include the time-differentiated
-configuration constraints or equivalent constraints which produce the same
-behavior.
-Furthermore, velocity constraints should not be written with $\dot{\mathbf{q}}$
-terms (the time derivatives of the coordinates).
-Acceleration constraints need to be defined as well.
-Typically, they will simply by time-differentiated velocity constraints and
-have the form:
+where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$,
+and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
+velocity constraints include the time-differentiated configuration constraints
+or equivalent constraints which produce the same behavior. Furthermore,
+velocity constraints should be written such that generalized speeds and
+coordinates appear, but $\dot{\mathbf{q}}$ terms do not.  Finally, acceleration
+constraints are typically time-differentiated velocity constraints (or
+similarly formed from kinematic considerations) and are assumed to have the
+form
 \begin{align}
@@
 \end{align}
-
-At this point, it is convienient to define a matrix $\mathbf{B}$, which is the
-gradient of the velocity constraints $\mathbf{f}_v$ with respect to all of the
-generalized speeds $\mathbf{u}$.
+In contrast to the velocity constraints, our formulation permits the
+acceleration constraints to involve $\dot{\mathbf{q}}$ terms (althernatively
+they can be eliminated by appealing to the kinematic differential equations).
+
+By assumption, the velocity constraints are linear with respect to
+$\mathbf{u}$, so we introduce
 \begin{align}
 \label{eq:constraint_B}
-\mathbf{B} \triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
+\mathbf{B} &\triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
 t)
 \end{align}
-and from (\ref{eq:velocity_constraints}) and (\ref{eq:constraint_B}) the
-following can be written:
+which implies
 \begin{align}
 \label{eq:constraint_Bu0}
-\mathbf{0} = \mathbf{B}\mathbf{u}
-\end{align}
-
-With these definitions established, Kane's method can be applied.
-The output of Kane's method, in vector form, is the equation
+\mathbf{B}\mathbf{u} + \mathbf{f}_{vt}(\mathbf{q}, t) &= \mathbf{0}
+\end{align}
+where $\mathbf{f}_{vt}$ arises from explicit time varying terms that occur in
+the velocity constraints. These typically arise from specified quantites whose
+dynamics are being ignored (e.g., a motor spinning at constant speed).
+
+With these definitions established, Kane's method can be applied. In vector
+form, Kane's dynamical equations are
 \begin{align}
@@
 \end{align}
-or, the sum of the generalized active forces and generalized inertial forces is
-0.
-If Newton's second law were written as $f - ma = 0$, the $\mathbf{F}$ is the
-multibody generalization of $f$ and $\mathbf{F}^*$ is the multibody
-generalization of $-ma$.
-These terms are created by taking the dot product of the partial
-velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
-- (\ref{eq:rb_rotational_frames}), with the resultant forces/torques ($\bar{R}$/
-$\bar{T}$) and the inertial forces/torques ($\bar{R}^*$/$\bar{T}^*$).
+where $\mathbf{F}$ are the generalized active forces and $\mathbf{F}^*$ are the
+generalized inertia forces. If Newton's second law is written as $f - ma = 0$,
+$\mathbf{F}$ is the multibody generalization of $f$ and $\mathbf{F}^*$ is the
+multibody generalization of $-ma$. As will be shown, these forces are obtained
+by taking the dot product of the partial velocities/angular-velocities, as
+defined in (\ref{eq:particle_translational}) - (\ref{eq:rb_rotational_frames}),
+with the resultant forces/torques ($\bar{R}$/ $\bar{T}$) and with inertial
+forces/torques ($\bar{R}^*$/$\bar{T}^*$).
 
 The resultant force at a point is defined as the sum of all force vectors
-acting directly on that point.
-The resultant torque on a reference frame is defined similarly - it is the sum
-of all torque vectors acting on that reference frame.
-Note that distance forces do not need to be transformed into a resultant
+acting directly on that point. The resultant torque on a reference frame is
+defined similarly -- it is the sum of all torque vectors acting on that
+reference frame. Distance forces do not need to be transformed into a resultant
 force/torque pair for any components of the system; they only need to be
 considered at the point of application (e.g., if a force $\bar{f}$ is the only
-force applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$
-and $\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
-These resultant forces and torques are used to write the generalized active
-forces:
-
+force applied to body $B$ at point $Q$, then $\bar{R}_{B^*}=0$ and
+$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The generalized active forces are
 \begin{equation}
@@
 
-The inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$, will now be
-defined. Note: the reference frame these are defined is not written - at this
-point in the derivation of the equations of motion, they are assumed to be in
-the inertial frame which is $N$ for this system.
-
-Particles:
+We now define the inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$. We
+leave out the inertial frame $N$ from the expressions; it is assumed to be $N$.
+The inertia force for particle $P$ is
 \begin{align}
 \label{eq:particle_gen_inertia}
-\bar{R}^*_P \triangleq -m_p {^N}\bar{a}^P
-\end{align}
-Rigid Bodies:
+\bar{R}^*_P \triangleq -m_P {^N}\bar{a}^P
+\end{align}
+while the inertia force and inertia torque for rigid body $B$ is
 \begin{align}
 \label{eq:rb_translational_gen_inertia}
-\bar{R}^*_{B^*} \triangleq -m_B {^N}\bar{a}^{B^*}
-\end{align}
-\begin{align}
+\bar{R}^*_{B^*} &\triangleq -m_B {^N}\bar{a}^{B^*} \\
 \label{eq:rb_rotational_gen_inertia}
-\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \cdot \bar{\bar{I}}^{B/B^*} -
+\bar{T}^*_B &\triangleq -{^N}\bar{\alpha}^B \cdot \bar{\bar{I}}^{B/B^*} -
 {^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\bar{\omega}^B
@@
 Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
-$B$ about its mass center $B^*$ \cite{Kane1985}.
-
-If the basis vectors of the reference frame associated with the rigid body $B$
-are aligned with the principal axes of $B$, then the following definitions can
-be used
-\begin{align}
-\alpha_i \triangleq {^N}\bar{\alpha}^B \cdot \hat{b}_i \quad \quad
-\omega_i \triangleq {^N}\bar{\omega}^B \cdot \hat{b}_i \quad \quad
-I_i \triangleq \hat{b}_i \cdot \bar{\bar{I}}^{B/B^*} \cdot \hat{b}_i
-\quad \quad \text{for }i=1,2,3
-\end{align}
-In this case, $I_1$, $I_2$, and $I_3$ will be the principal inertias of rigid
-body $B$.
-This allows the inertia torque to be written
-\begin{align}
-\label{eq:simple_rb_rotational_gen_inertia}
-\bar{T}_B^* =& -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1\\
-             & -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2\\
-             & -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3\\
-\end{align}
-
-Now $\mathbf{F}^*$ can be constructed in a manner similar to $\mathbf{F}$.
-
+$B$ about its mass center $B^*$ \cite{Kane1985}. The generalized inertia forces
+are
 \begin{equation}
@@
 
-Now all the terms in (\ref{eq:kanes_eq}) are defined.
-But this is valid only if there are no motion constraints.
-The equations need to be transformed in order to not violate the behaviors
-defined by the constraints.
-
-Kane's Method allows for the constrained equations of motion to be written
-using both independent and dependent speeds, $\mathbf{u}_i$ and $\mathbf{u}_d$,
-without the need for Lagrange multipliers.
-Even though both independent and dependent speeds are in the final equations, a
-choice still needs to be made as to which speeds will be dependent.
+All the terms in (\ref{eq:kanes_eq}) are now defined, but this equation is
+valid only if there are no velocity constraints. If there are velocity
+constraints, the equations must be transformed. As we present them, Kane's
+nonholonomic dynamical equations will be of dimension $o - m$, where $o$ is the
+total number of generalized speeds and $m$ is the number of velocity
+constraints. However, these $o-m$ equations generally involve all $o$ speeds
+(we assume the dependent speeds and their time derivatives have \textit{not}
+been algebraically eliminated). In order to obtain the nonholonomic dynamical
+equations, however, a choice must be made as to which speeds will be dependent.
 The methods for doing this are outside the scope of this paper, but it is
-important that whatever selection is made avoids singularities.
-
-To simplify notation, it is assumed that the generalized speeds are already
-ordered in a favorable manner.
-\begin{align}
-\label{eq:speeds_independent_dependent}
-\mathbf{u} =
-\begin{bmatrix}
-\mathbf{u}_i \\
-\mathbf{u}_d
-\end{bmatrix}
-\end{align}
-The matrix $\mathbf{B}$ from (\ref{eq:constraint_B}) also needs to be
-partitioned.
-\begin{align}
-\label{eq:constraint_B_independent_dependent}
-\mathbf{0} &= \mathbf{B} \mathbf{u}\\
-\mathbf{0} &= \begin{bmatrix}
-\mathbf{B}_i & \mathbf{B}_d
-\end{bmatrix} \begin{bmatrix}
-\mathbf{u}_i \\ \mathbf{u}_d
-\end{bmatrix}
-\end{align}
-where $\mathbf{B}_i$ has the shape $m \times o - m$ and $\mathbf{B}_d$ has the
-shape $m \times m$.
-This equation can then be rewritten as
+important that whatever selection is made avoids singularities which can occur
+in certain configurations or sets of system parameters. For a careful treatment
+of this, we encourage the reader to examine the work of Reckdahl
+\cite{Reckdahl1996}.
+
+We assume the $o$ generalized speeds are ordered $\mathbf{u} = [\mathbf{u}_i,
+\mathbf{u}_d]^T$, where $\mathbf{u}_i\in\mathbf{R}^{o-m}$ and
+$\mathbf{u}_d\in\mathbf{R}^{m}$ are the independent and dependent speeds,
+respectively. The velocity constraint Jacobian matrix $B$ can be similarly
+rearranged as $\mathbf{B} = [\mathbf{B}_i, \mathbf{B}_d]$, where
+$\mathbf{B}_i\in \mathbf{R}^{m \times o-m}$ and $\mathbf{B}_d \in \mathbf{R}^{m
+\times m}$. With this in mind, equation (\ref{eq:constraint_Bu0}) can be
+written as
+\begin{align}
+\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d + \mathbf{f}_{vt}(\mathbf{q}, t) &= \mathbf{0} \\
+\implies \mathbf{u}_d &= -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i - \mathbf{f}_{vt}(\mathbf{q}, t)
+\end{align}
+For convenience, we define
 \begin{align}
 \label{eq:constraint_A}
-\mathbf{B}_i \mathbf{u}_i + \mathbf{B}_d \mathbf{u}_d &= \mathbf{0} \\
-\mathbf{B}_d \mathbf{u}_d &= - \mathbf{B}_i \mathbf{u}_i \\
-\mathbf{u}_d &= -\mathbf{B}_d^{-1} \mathbf{B}_i \mathbf{u}_i \\
-\mathbf{A} &\triangleq - \mathbf{B}_d^{-1} \mathbf{B}_i \\
-\mathbf{u}_d &= \mathbf{A} \mathbf{u}_i
-\end{align}
-where $\mathbf{A}$ has the shape $m \times m$.
-
-Finally, to constrain the generalized active and inertia forces, they need to
-be partitioned in a similar manner.
+\mathbf{A} \triangleq - \mathbf{B}_{d}^{-1} \mathbf{B}_i
+\end{align}
+and note that $\mathbf{A}\in\mathbf{R}^{m \times (o - m)}$. To constrain the
+generalized active and inertia forces, we partitioned them in a similar
+manner
 \begin{align}
@@
 \end{align}
-
-After constraining the equations of motion, we have reduced the number of
-equations from $o$ to $o-m$; but the equations still contain $o$ unknown terms
-that need to be solved for ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$).
-In order to solve these second order differential equations for the
-$\dot{\mathbf{u}}$ terms, additional equations are needed.
-The acceleration constraint equations fill this role, as there are $m$ of these
-equations, giving:
+which allows Kane's nonholonomic dynamical equations to be written as
+\begin{align}
+\label{eq:Kanes_nonholonomic}
+\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
+\end{align}
+This process of constraining the equations of motion reduces the number of
+equations from $o$ to $o-m$. However, the (\ref{eq:Kanes_nonholonomic}) still
+contain $o$ unknowns ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$). By
+augmenting (\ref{eq:Kanes_nonholonomic}) with the acceleration constraint
+equations (\ref{eq:acceleration_constraints}), we end up with $o$ equations and
+$o$ unknowns. Note that equations (\ref{eq:Kanes_nonholonomic}) and
+(\ref{eq:acceleration_constraints}) are linear in the $\dot{u}_i$'s. The
+augmented system of equations is
 \begin{align}
@@
 \end{align}
-which gives a total of $o$ equations, enough to be able to solve for the
-$\dot{\mathbf{u}}$'s.
-
-As a final step, the arrangement of the equations as used in this paper is
-shown in Table \ref{table:assumptions}.
-These equations are all drawing from the preceding steps, but have been
-re-organized for convenience (as will be shown later).
+
+For convenience, we summarize the critical equations in Table
+\ref{table:assumptions}.
+
+% This is incomplete unless we mention the kinematic equations so I am leaving
+% it out for now. I think we have stated this earlier.
+%With regards to obtaining linearized equations, it is
+%critical to note that solving equations (\ref{eq:F_Fstar_f_a}) for the
+%$\dot{\mathbf{u}}$'s, 
+%with respect to $\mathbf{u}_i$ does not give the correct linear equations. The
+%reasons for this will be covered later.
 
@@
 The only new terms are $\mathbf{f}_0$, $\mathbf{f}_1$, $\mathbf{f}_2$, and
-$\mathbf{f}_3$.
-The kinematic differential equations are split up into the quantities
-$\mathbf{f}_0$ and $\mathbf{f}_1$, depending on whether each term has a
-$\mathbf{u}$ or $\dot{\mathbf{q}}$ present.
-The same is true for $\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a
-$\dot{\mathbf{u}}$ it is placed into $\mathbf{f}_2$ and if not it is placed
-into $\mathbf{f}_3$.
-
-It is critical to note that just taking the gradient of (\ref{eq:F_Fstar_f_a})
-with respect to $\mathbf{u}_i$ does not give the correct linear equations. The
-reasons for this will be covered later.
+$\mathbf{f}_3$.  The kinematic differential equations are split up into the
+quantities $\mathbf{f}_0$ and $\mathbf{f}_1$, depending on whether each term
+has a $\mathbf{u}$ or $\dot{\mathbf{q}}$ present.  The same is true for
+$\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a $\dot{\mathbf{u}}$ it
+is placed into $\mathbf{f}_2$ and if not it is placed into $\mathbf{f}_3$. The
+motivation for this organizational scheme will be made apparent in Section
+\ref{sec:derivations}.
 
@@
 To illustrate Kane's method in the presence of configuration and velocity
-constraints, we consider a thin disc $C$ rolling without slip on a horizontal
+constraints, we consider a thin disk $C$ rolling without slip on a horizontal
 plane. The dextral unit vectors $\hat{c}_x, \hat{c}_y, \hat{c}_z$ are fixed to $C$
-with $\hat{c}_y$ perpendicular to the disc plane. Let $r$ and $m$ be the
-disc radius and mass, and assume a thin disc mass distribution so that
+with $\hat{c}_y$ perpendicular to the disk plane. Let $r$ and $m$ be the
+disk radius and mass, and assume a thin disk mass distribution so that
 $I_{xx}=I_{zz}=mr^2/4$ and $I_{yy} = mr^2/2$. Take the inertial (Newtonian)
@@
 \hat{n}_y, \hat{n}_z$, with $\hat{n}_z$ perpendicular to the ground plane
-(downwards) and aligned with the local gravitational field. To orient the disc,
+(downwards) and aligned with the local gravitational field. To orient the disk,
 first align $C$ with $N$, then apply successive body fixed ZXY rotations:
@@
 to these as the instantaneous yaw and lean frames, respectively. The unit
-vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disc plane and
-aligned with the position vector from the disc center $C^*$ to the lowest point
-of the disc $P^*$.
-
-Traditionally, to locate the center of the disc $C^*$
+vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disk plane and
+aligned with the position vector from the disk center $C^*$ to the lowest point
+of the disk $P^*$.
+
+Traditionally, to locate the center of the disk $C^*$
 relative to the inertial origin $N^*$, two coordinates are used to locate the
 contact point in the ground plane; these two coordinates and the lean angle
-and yaw angles determine the location of the center of the disc. The advantage of this
-approach is that the choice of coordinates guarantees that the disc remain in
+and yaw angles determine the location of the center of the disk. The advantage of this
+approach is that the choice of coordinates guarantees that the disk remain in
 contact with the ground plane.
@@
 accounted for when linearizing the equations of motion, we purposefully locate
-the center of the disc $C^*$ relative to the inertial origin $N^*$ with a
+the center of the disk $C^*$ relative to the inertial origin $N^*$ with a
 non-minimal choice of coordinates
 \begin{equation*}
-  \bar{r}^{N^*C^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
+  \bar{r}^{C^*/N^*} = q_4 \hat{n}_x + q_5 \hat{n}_y + q_6 \hat{n}_z
 \end{equation*}
@@
 \end{equation}
-which must be satisfied for the disc to remain in contact with the ground.
+which must be satisfied for the disk to remain in contact with the ground.
 While this constraint is easily avoided by appropriate choice of coordinates,
@@
 The angular velocity of $C$ in $N$, however, is defined in the intermediate
-frame $B$, which is the instantaneous lean frame.
+lean frame $B$.
 \begin{align}
@@
 are populated with standard 3-vectors. Given these definitions, the velocity of
-the lowest point of the disc $P^*$ is
+the lowest point of the disk $P^*$ is
 \begin{align*}
@@
 
-At this point, the reader familiar with the rolling disc problem is certainly
+At this point, the reader familiar with the rolling disk problem is certainly
 wondering why we have chosen such a diabolical set of coordinates and speeds.
@@
 
-As stated, the mass of the disc is $m$, and the inertia dyadic of the disc can
+As stated, the mass of the disk is $m$, and the inertia dyadic of the disk can
 be written as
@@
 
-As there are 12 unknown quantities, $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$, we
-need 3 additional equations so solve for the time-derivative terms.
-The acceleration constraints $\mathbf{f}_a$ provide the final 3 equations.
-Combining (\ref{rd:f_0_f_1}), (\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the
-complete second order differential equations can be written.
+The twelve unknown quantities appearing in the nine kinematic and dynamic
+differential equations are $\dot{q}_{1-6}$ and $\dot{u}_{1-6}$. The three
+acceleration constraint equations provide the final three equations which allow
+for all twelve unkowns to be solved. Combining (\ref{rd:f_0_f_1}),
+(\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete second order differential
+equations can be written.
 \begin{align}
@@
 \end{align}
-and then take the jacobian of $\mathbf{f}$ with respect to $[\mathbf{q}
-\quad \mathbf{u}]^T$. However, this will give the incorrect results. As
-discussed in Schwab, this system has a critical speed where all the eigenvalues
-are 0 under the following conditions: the system is linearized about the
-upright rolling steady-state configuration, and all parameters ($r$, $m$, and
-$g$) are unity. When these conditions are met, the critical speed is
-$v=\frac{1}{\sqrt{3}}$.
-
-When following the naive approach, 8 of the eigenvalues are 0 and 4 are
-non-zero under the same conditions:
-\begin{align}
-\sqrt{-v^2},\quad -\sqrt{-v^2},\quad \sqrt{-\frac{12}{5} v^2 +
-\frac{4}{5}},\quad -\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
-\end{align}
-At the critical speed, two of these non-zero eigenvalues go to zero. But at any
-forward speed, there is a nonzero imaginary eigenvalue pair. Having illustrated
-the failings of the simple approach to linearization of a constrained multibody
-system defined with Kane's method, we will now outline a linearization
-procedure which addresses this issue and then revisit this example applying the
-proper linearization procedure.
-
-%We apply this linearization technique to the well studied rolling disc.  Figure
-%\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
-%The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
+and then take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
+\mathbf{u}]^T$. However, this will give the incorrect results. As discussed in
+\cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity, and the system
+is linearized about the upright steady rolling condition, the critical speed is
+$v^*=\pm\frac{1}{\sqrt{3}}$. When following the naive approach, and
+evaluating the Jacobian at the same operating conditions and parameters, eight
+of the twelve eigenvalues are identically zero while the remaining four are
+\begin{align}
+  \lambda_{1,2}=\pm\sqrt{-v^2},\quad \lambda_{3,4} = \pm\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
+\end{align}
+which demonstrates the incorrectness of this naive approach (since
+$v=\pm\frac{1}{\sqrt{3}}$ is not a critical point of these eigenvalues).
+Further, the fact that twelve eigenvalues are obtained should be an alert that
+something is incorrect since the number of independent quantities is only eight
+(five coordinates and three speeds). We now outline a linearization procedure
+which addresses this issue and then revisit this example applying the proper
+linearization procedure.
+
+\section{Derivation of linearization procedure}
+\label{sec:derivations}
+
+In response to the need we have demonstrated, we present a solution to this
+constrained, linearization problem.
+There is a reason that the equations generated by taking the gradient with
+respect to the generalized speeds are incorrect.
+When finding the equations of motion for a constrained system, some generalized
+coordinates and speeds are classified as dependent speeds, meaning their
+behavior is determined by the other coordinates and speeds.
+These terms should in fact be written as:
+\begin{align}
+\label{eq:q_d_redefined}
+\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
+\label{eq:u_d_redefined}
+\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
+\end{align}
+Then, upon taking the derivative of each of these terms with respect to an
+independent speed, the dependent speed would need to be solved for and its
+derivative found.
+No previous authors appear to have made clear that this distinction should be
+made when using Kane's method.
+
+What we present addresses this issue.
+Rather then substituting in the solutions to the dependent speeds though, we
+use knowledge of the structure of the equations and constraints (as defined in
+Table \ref{table:assumptions}) to generate linear equations.
+
+We begin with a first order Taylor series expansion of the equations in Table
+\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
+$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
+assumed that all of the equations in Table \ref{table:assumptions} are
+satisfied by these quantities (i.e., the system is in a valid state).
+In the interest of brevity, we omit writing this point of linearization in each
+gradient in the calculations below; all are evaluated at this point.
+Expansion of the three constraint equations, keeping only first order terms,
+yields
+\begin{align}
+  \label{eq:configuration_expansion}
+  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
+  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
+  \label{eq:velocity_expansion}
+  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
+  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
+  \label{eq:acceleration_expansion}
+  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
+  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
+t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
+\nabla_{\bm{\dot{q}}}\bm{f}_{a}
+ \delta \bm{\dot{q}} \notag\\
+&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
+\delta \bm{\dot{u}}
+\end{align}
+The first terms are identically zero because of the assumption that the
+point of linearization satisfies the constraints.  The Taylor series expansion
+of the kinematic differential equations is
+\begin{align}
+  \label{eq:f0_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
+  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
+  \label{eq:f1_expansion}
+  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
+  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
+that the sum of the first term on the right hand side of each equation must
+equal zero, we obtain
+\begin{align}
+  \label{eq:f0_plus_f1_expansion}
+  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
+  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
+  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
+  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
+\end{align}
+Similarly, a Taylor series expansion of the dynamic differential equations, we
+obtain
+\begin{align}
+  \label{eq:f2_expansion}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
+      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
+      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
+      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
+  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
+  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
+  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
+  \label{eq:f3_expansion}
+  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
+  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
+  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
+that the sum of the first term on the right hand sides of these equations must
+equal zero, we obtain
+\begin{align}
+  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
+  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
+  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
+  \label{eq:f2_plus_f3_expansion}
+  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
+  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
+\end{align}
+
+Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
+(\ref{eq:acceleration_expansion}),
+and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
+\ref{table:assumptions}), and introducing the following definitions
+\begin{align}
+\label{eq:quant_to_compute}
+  \begin{array}{llcll}
+    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
+    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
+    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
+    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
+    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
+    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
+    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
+    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
+    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
+    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
+    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
+    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
+  \end{array}
+\end{align}
+enables the unconstrained linear state space equations to be written as
+\begin{align}
+  \label{eq:state_space_unconstrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       \tilde{A}_{qq} & \tilde{A}_{qu} \\
+       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
+       \tilde{A}_{uqd} & \tilde{A}_{uud}
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q} \\
+        \delta \bm{u}
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n + m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
++ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
+address this issue, a smaller set of independent coordinates and speeds must be
+selected. To this end, consider the following partitioning of the generalized
+coordinates and generalized speeds:
+\begin{equation*}
+  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
+      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
+      \qquad\qquad
+  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
+      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
+\end{equation*}
+where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
+are invertible permutation matrices which map an ordering which has the
+independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
+\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
+($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
+ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
+$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
+similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
+last $m$ columns of $P_u$.
+
+By assumption, equation
+(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
+   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
+  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
+  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
+  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
+  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
+    \bm{q}_i\\
+  \label{eq:delta_q}
+  \delta \bm{q} &= C_0 \delta \bm{q}_i
+\end{align}
+
+Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
+\begin{align}
+  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
+  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
+  &= \nabla_{\bm{q}} \bm{f}_{v}
+\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
+\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
+\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
+  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
+  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
+  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
+  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
+  \label{eq:delta_u}
+  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
+\end{align}
+
+By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
+rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
+\ref{eq:state_space_constrained}, a linear system of equations which relates
+the time derivatives of all states (dependent and independent) to the
+independent states and exogenous inputs.
+
+The linearized equations of motion are
+\begin{align}
+  \label{eq:state_space_constrained}
+  \left[
+    \begin{array}{cc}
+      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+      \tilde{M}_{uqd} & \tilde{M}_{uud}
+    \end{array}
+    \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{\dot{q}} \\
+        \delta \bm{\dot{u}}
+      \end{array}
+    \right]
+   &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]
+    \left[
+      \begin{array}{c}
+        \delta \bm{q}_i \\
+        \delta \bm{u}_i
+      \end{array}
+    \right]
+    +
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+    \delta \bm{r}
+\end{align}
+
+where
+
+\begin{align}
+  \label{eq:C_0}
+  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
+  \label{eq:C_1}
+  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v} \\
+  \label{eq:C_2}
+  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+    \bm{f}_{v} \right] P_{ui}\\
+  \label{eq:Pq}
+  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
+  \label{eq:Pu}
+  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
+\end{align}
+
+Table \ref{table:assumptions} provides the form in which the nonlinear
+equations of motion and constraints must be arranged in order to carry out the
+linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+are Jacobian matrices of the differential equations. Equations
+(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+independent and dependent speeds.
+
+The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+an equilibrium point.
+
+\section{Dynamics Example, Revisted}
+
+\section{Discussion}
+\label{sec:discussion}
+Most modern control techniques assume a system can be written as
+\begin{align}
+\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B}\mathbf{u}
+\end{align}
+but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
+linear system framework because it is non-square.
+To obtain a square system we define
+\begin{align}
+\tilde{A} &=
+   \left[
+     \begin{array}{cc}
+       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
+       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
+       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
+     \end{array}
+   \right]\\
+\tilde{B} &= 
+    \left[
+      \begin{array}{c}
+        \bm{0}_{(n+m) \times s} \\
+        \tilde{B}_{u}
+      \end{array}
+    \right]
+\end{align}
+from (\ref{eq:state_space_constrained}) and
+\begin{align}
+  \label{eq:A_prime}
+    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
+  \label{eq:B_prime}
+    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
+\end{align}
+where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
+\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
+independent states by defining
+\begin{align}
+  \label{eq:P_prime}
+    P^\prime &\triangleq \begin{bmatrix}
+        P_{qi} & \bm{O}_{n \times (o - m)} \\
+        \bm{O}_{o \times (n - l)} & P_{ui}
+    \end{bmatrix} \\
+  \label{eq:A}
+    A &\triangleq P^{\prime T} A^\prime \\
+  \label{eq:B}
+    B &\triangleq P^{\prime T} B^\prime
+\end{align}
+where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
+$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
+state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
+linear systems analyses may applied.  It is worth noting that the rows
+of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
+in the output or measurement equations of a linear state space model, as in the
+case when the controlling or measuring dependent state(s).
+
+It is notable that there is a second opportunity (after constraining the
+nonlinear equations of motion) to determine which speeds are dependent.
+Now that the equations are linear, the choice of which state variables are
+taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
+P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
+These matrices depend upon the configuration $\bm{q}$ and constant parameters;
+for certain configurations or parameters, it may be that the choice of
+independent state variables cannot be arbitrary.
+While some systems may permit a choice of independent state variables which are
+valid for all configurations of interest, others may not.
+Methods for automatically selecting the ``best'' choice of independent state
+variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
+of the Jacobian of the constraint matrices to determine a set of independent
+states which will ensure the non-singularity of the aforementioned matrices.
+
+%\section{Example}
+%\label{example}
+%We apply this linearization technique to the well studied rolling disk.  Figure
+%\ref{fig:disk} illustrates the choice of coordinates, speeds, and parameters.
+%The disk is assumed to be knife edged of radius $r$, mass $m$, spin inertia
 %$J$, and transverse inertia $I$.  To illustrate the concept of dependent
 %quantities, we purposefully choose to describe the configuration and velocity
-%of the disc with a non-minimal set of coordinates and speeds.  This choice is
-%not purely academic.  It is convenient for a physical disc equipped with a rate
+%of the disk with a non-minimal set of coordinates and speeds.  This choice is
+%not purely academic.  It is convenient for a physical disk equipped with a rate
 %gyroscope and accelerometer to measure these body fixed quantities (or their
@@
 %
-%The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
+%The orientation of the disk-fixed frame $C$ relative to the inertial frame $N$
 %is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
 %(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
-%location of disc center $C^*$ relative to the inertial origin $N^*$, and the
+%location of disk center $C^*$ relative to the inertial origin $N^*$, and the
 %velocity of $C^*$ in $N$ are
@@
 %\end{align}
-%If the disc is in contact with the ground, a configuration constraint involving
-%the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
+%If the disk is in contact with the ground, a configuration constraint involving
+%the height $q_5$ of the disk center and the lean $q_1$.  Similarly, if the disk
 %rolls without slip, other constraints relate the generalized speeds $u_i, (i =
@@
 %  \centering
-%  \caption{Rolling disc governing definitions and equations}
-%  \label{table:rollingdisc}
+%  \caption{Rolling disk governing definitions and equations}
+%  \label{table:rollingdisk}
 %  \begin{tabular}[c]{l l l}
@@
 %  \centering
-%  \includegraphics{rollingdisc.pdf}
-%  \caption{Rolling disc configuration}
-%  \label{fig:disc}
-%\end{figure}
-%
-%The matrices which make up the linear equations were created using the method
-%presented in Section \ref{sec:results}. It is important to note that the
-%entries in these matrices must be evaluated at the point of linearization
-%($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
-%linearization must satisfy the equations in Table \ref{table:assumptions}.
-%There is no requirement that this point be an equilibrium point, which implies
-%that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
-%reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
-%means that the other quantities in the linearized equations (other $q$'s, the
-%$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
-%quantities already set to 0.
-%
-%The independent state vector is:
-%\begin{align}
-%\begin{bmatrix}
-%\delta \bm{q}_i \\
-%\delta \bm{u}_i
-%\end{bmatrix} =
-%\begin{bmatrix}
-%q_0 &
-%q_1 &
-%q_2 &
-%q_3 &
-%q_4 &
-%u_0 &
-%u_1 &
-%u_2
-%\end{bmatrix}^T
-%\end{align}
-%
-%And the complete (independent and dependent) state vector is:
-%\begin{align}
-%\begin{bmatrix}
-%\delta \bm{q} \\
-%\delta \bm{u}
-%\end{bmatrix} =
-%\left[
-%\begin{array}{cccccccccccc}
-%q_0 &
-%q_1 &
-%q_2 &
-%q_3 &
-%q_4 &
-%q_5 &
-%u_0 &
-%u_1 &
-%u_2 &
-%u_3 &
-%u_4 &
-%u_5
-%\end{array}
-%\right]^T
-%\end{align}
-%
-%The coefficient matrices of the linearized equations (Equation
-%\ref{eq:state_space_constrained}) are:
-%\begin{equation}
-%  \left[
-%    \begin{array}{cc}
-%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-%      \tilde{M}_{uqd} & \tilde{M}_{uud}
-%    \end{array}
-%    \right]=
-%\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
-%\end{equation}
-%
-%
-%\begin{equation*}
-%   \left[
-%     \begin{array}{cc}
-%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-%     \end{array}
-%   \right]=
-%\end{equation*}
-%\begin{equation}
-%\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
-%\end{equation}
-%
-%With this example, we have used our linearization procedure to find linear
-%equations of motion which respect the system's constraints. Parametric
-%stability analysis (eigenvalue analysis) can now be performed on these
-%equations.  These computations were implemented symbolically with the open
-%source symbolic manipulator SymPy \cite{SymPy2012}.
-
-\section{Derivation of linearization procedure}
-\label{sec:derivations}
-
-In response to the need we have demonstrated, we present a solution to this
-constrained, linearization problem.
-There is a reason that the equations generated by taking the gradient with
-respect to the generalized speeds are incorrect.
-When finding the equations of motion for a constrained system, some generalized
-coordinates and speeds are classified as dependent speeds, meaning their
-behavior is determined by the other coordinates and speeds.
-These terms should in fact be written as:
-\begin{align}
-\label{eq:q_d_redefined}
-\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
-\label{eq:u_d_redefined}
-\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
-\end{align}
-Then, upon taking the derivative of each of these terms with respect to an
-independent speed, the dependent speed would need to be solved for and its
-derivative found.
-No previous authors appear to have made clear that this distinction should be
-made when using Kane's method.
-
-What we present addresses this issue.
-Rather then substituting in the solutions to the dependent speeds though, we
-use knowledge of the structure of the equations and constraints (as defined in
-Table \ref{table:assumptions}) to generate linear equations.
-
-We begin with a first order Taylor series expansion of the equations in Table
-\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
-$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
-assumed that all of the equations in Table \ref{table:assumptions} are
-satisfied by these quantities (i.e., the system is in a valid state).
-In the interest of brevity, we omit writing this point of linearization in each
-gradient in the calculations below; all are evaluated at this point.
-Expansion of the three constraint equations, keeping only first order terms,
-yields
-\begin{align}
-  \label{eq:configuration_expansion}
-  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
-  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
-  \label{eq:velocity_expansion}
-  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
-  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
-  \label{eq:acceleration_expansion}
-  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
-  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
-t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
-\nabla_{\bm{\dot{q}}}\bm{f}_{a}
- \delta \bm{\dot{q}} \notag\\
-&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
-\delta \bm{\dot{u}}
-\end{align}
-The first terms are identically zero because of the assumption that the
-point of linearization satisfies the constraints.  The Taylor series expansion
-of the kinematic differential equations is
-\begin{align}
-  \label{eq:f0_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
-  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
-  \label{eq:f1_expansion}
-  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
-  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
-that the sum of the first term on the right hand side of each equation must
-equal zero, we obtain
-\begin{align}
-  \label{eq:f0_plus_f1_expansion}
-  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
-  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
-  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
-  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
-\end{align}
-Similarly, a Taylor series expansion of the dynamic differential equations, we
-obtain
-\begin{align}
-  \label{eq:f2_expansion}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
-      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
-      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
-      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
-  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
-  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
-  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
-  \label{eq:f3_expansion}
-  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
-  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
-  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
-that the sum of the first term on the right hand sides of these equations must
-equal zero, we obtain
-\begin{align}
-  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
-  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
-  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
-  \label{eq:f2_plus_f3_expansion}
-  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
-  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
-\end{align}
-
-Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
-(\ref{eq:acceleration_expansion}),
-and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
-\ref{table:assumptions}), and introducing the following definitions
-\begin{align}
-\label{eq:quant_to_compute}
-  \begin{array}{llcll}
-    \tilde{M}_{qq}  &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_0 & \quad &
-    \tilde{M}_{uqc} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_a \\
-    \tilde{M}_{uuc} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_a & \quad &
-    \tilde{M}_{uqd} &\triangleq \nabla_{\bm{\dot{q}}}\bm{f}_3 \\
-    \tilde{M}_{uud} &\triangleq \nabla_{\bm{\dot{u}}}\bm{f}_2 & \quad &
-    \tilde{A}_{qq}  &\triangleq -\nabla_{\bm{q}}(\bm{f}_0 + \bm{f}_1) \\
-    \tilde{A}_{qu}  &\triangleq -\nabla_{\bm{u}}\bm{f}_1 & \quad &
-    \tilde{A}_{uqc} &\triangleq - \nabla_{\bm{q}} \bm{f}_a \\
-    \tilde{A}_{uuc} &\triangleq - \nabla_{\bm{u}} \bm{f}_a & \quad &
-    \tilde{A}_{uqd} &\triangleq - \nabla_{\bm{q}} (\bm{f}_2 + \bm{f}_3) \\
-    \tilde{A}_{uud} &\triangleq - \nabla_{\bm{u}} \bm{f}_3 & \quad &
-    \tilde{B}_{u}   &\triangleq -\nabla_{\bm{r}}\bm{f}_{3}
-  \end{array}
-\end{align}
-enables the unconstrained linear state space equations to be written as
-\begin{align}
-  \label{eq:state_space_unconstrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       \tilde{A}_{qq} & \tilde{A}_{qu} \\
-       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
-       \tilde{A}_{uqd} & \tilde{A}_{uud}
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q} \\
-        \delta \bm{u}
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n + m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
-+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
-address this issue, a smaller set of independent coordinates and speeds must be
-selected. To this end, consider the following partitioning of the generalized
-coordinates and generalized speeds:
-\begin{equation*}
-  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
-      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
-      \qquad\qquad
-  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
-      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
-\end{equation*}
-where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
-are invertible permutation matrices which map an ordering which has the
-independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
-\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
-($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
-ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
-$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
-similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
-last $m$ columns of $P_u$.
-
-By assumption, equation
-(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
-   &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
-  \nabla_{\bm{q}}\bm{f}_{c} P_{qd} \delta \bm{q_d}\notag\\
-  \implies \delta \bm{q}_d &= -(\nabla_{\bm{q}}\bm{f}_{c} P_{qd})^{-1}
-  (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
-  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
-    \bm{q}_i\\
-  \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \delta \bm{q}_i
-\end{align}
-
-Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
-  &= \nabla_{\bm{q}} \bm{f}_{v}
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
-\nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
-\implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
-  \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
-  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
-  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
-  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
-  \label{eq:delta_u}
-  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
-\end{align}
-
-By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
-rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
-\ref{eq:state_space_constrained}, a linear system of equations which relates
-the time derivatives of all states (dependent and independent) to the
-independent states and exogenous inputs.
-
-The linearized equations of motion are
-\begin{align}
-  \label{eq:state_space_constrained}
-  \left[
-    \begin{array}{cc}
-      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-      \tilde{M}_{uqd} & \tilde{M}_{uud}
-    \end{array}
-    \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{\dot{q}} \\
-        \delta \bm{\dot{u}}
-      \end{array}
-    \right]
-   &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]
-    \left[
-      \begin{array}{c}
-        \delta \bm{q}_i \\
-        \delta \bm{u}_i
-      \end{array}
-    \right]
-    +
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-    \delta \bm{r}
-\end{align}
-
-where
-
-\begin{align}
-  \label{eq:C_0}
-  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
-  \label{eq:C_1}
-  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \\
-  \label{eq:C_2}
-  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui}\\
-  \label{eq:Pq}
-  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
-  \label{eq:Pu}
-  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
-\end{align}
-
-Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged in order to carry out the
-linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
-bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
-are Jacobian matrices of the differential equations. Equations
-(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
-embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
-permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
-independent and dependent speeds.
-
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
-of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
-satisfy the equations in Table \ref{table:assumptions}, and may or may not be
-an equilibrium point.
-
-\section{Discussion}
-\label{sec:discussion}
-Most modern control techniques assume a system can be written as
-\begin{align}
-\dot{\mathbf{x}} = \mathbf{A}\mathbf{x} + \mathbf{B}\mathbf{u}
-\end{align}
-but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
-linear system framework because it is non-square.
-To obtain a square system we define
-\begin{align}
-\tilde{A} &=
-   \left[
-     \begin{array}{cc}
-       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-     \end{array}
-   \right]\\
-\tilde{B} &= 
-    \left[
-      \begin{array}{c}
-        \bm{0}_{(n+m) \times s} \\
-        \tilde{B}_{u}
-      \end{array}
-    \right]
-\end{align}
-from (\ref{eq:state_space_constrained}) and
-\begin{align}
-  \label{eq:A_prime}
-    A^\prime &\triangleq \tilde{M}^{-1} \tilde{A} \\
-  \label{eq:B_prime}
-    B^\prime &\triangleq \tilde{M}^{-1} \tilde{B}
-\end{align}
-where  $A^\prime \in \bm{R}^{(o + n) \times (o - m + n -l)}$, $B^\prime \in
-\bm{R}^{(o + n) \times s}$.  We can extract the rows corresponding to the
-independent states by defining
-\begin{align}
-  \label{eq:P_prime}
-    P^\prime &\triangleq \begin{bmatrix}
-        P_{qi} & \bm{O}_{n \times (o - m)} \\
-        \bm{O}_{o \times (n - l)} & P_{ui}
-    \end{bmatrix} \\
-  \label{eq:A}
-    A &\triangleq P^{\prime T} A^\prime \\
-  \label{eq:B}
-    B &\triangleq P^{\prime T} B^\prime
-\end{align}
-where $P^\prime \in \bm{R}^{(o + n) \times (o - m + n - l)}$.  Defining
-$\bm{x}_i = \left[\delta\bm{q}_i,\,\delta\bm{u}_i\right]^{T}$ yields the square
-state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
-linear systems analyses may applied.  It is worth noting that the rows
-of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
-in the output or measurement equations of a linear state space model, as in the
-case when the controlling or measuring dependent state(s).
-
-It is notable that there is a second opportunity (after constraining the
-nonlinear equations of motion) to determine which speeds are dependent.
-Now that the equations are linear, the choice of which state variables are
-taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
-P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
-These matrices depend upon the configuration $\bm{q}$ and constant parameters;
-for certain configurations or parameters, it may be that the choice of
-independent state variables cannot be arbitrary.
-While some systems may permit a choice of independent state variables which are
-valid for all configurations of interest, others may not.
-Methods for automatically selecting the ``best'' choice of independent state
-variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
-of the Jacobian of the constraint matrices to determine a set of independent
-states which will ensure the non-singularity of the aforementioned matrices.
-
-%\section{Example}
-%\label{example}
-%We apply this linearization technique to the well studied rolling disc.  Figure
-%\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
-%The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
-%$J$, and transverse inertia $I$.  To illustrate the concept of dependent
-%quantities, we purposefully choose to describe the configuration and velocity
-%of the disc with a non-minimal set of coordinates and speeds.  This choice is
-%not purely academic.  It is convenient for a physical disc equipped with a rate
-%gyroscope and accelerometer to measure these body fixed quantities (or their
-%time derivatives) directly, rather than the more traditional set which is
-%chosen for analytic convenience.  In more complicated systems with constraints,
-%it is difficult (if not impossible) to describe the system with a minimal
-%choice of coordinates and speeds; the ideas demonstrated by this simple example
-%can be applied to more complicated ones.
-%
-%The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
-%is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
-%(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
-%location of disc center $C^*$ relative to the inertial origin $N^*$, and the
-%velocity of $C^*$ in $N$ are
-%\begin{align}
-%  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
-%  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
-%  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
-%\end{align}
-%If the disc is in contact with the ground, a configuration constraint involving
-%the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
-%rolls without slip, other constraints relate the generalized speeds $u_i, (i =
-%0,\dots,5)$.  The derivations of these constraints and the equations of motion
-%are presented in the electronic supplementary material and can also be found in
-%the literature \cite{O'Reilly1996}.
-%\begin{table}[htbp]
-%  \centering
-%  \caption{Rolling disc governing definitions and equations}
-%  \label{table:rollingdisc}
-%  \begin{tabular}[c]{l l l}
-%    Quantity & Shape & Description\\
-%    \hline
-%    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
-%    derivatives\\
-%    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
-%  \end{tabular}
-%  \begin{tabular}[c]{r @{ $=$ } l}
-%%    \multicolumn{3}{c}{ } \\
-%%    \multicolumn{2}{c}{Equation} &           \\
-%    \hline
-%    % Configuration constraint
-%    $\bm{f}_c$ & $rc_1 + q_5$ \\
-%    % Velocity constraints
-%    $\bm{f}_v$ & $\left[\begin{array}{c}
-%        r u_1 c_2 + u_3 \\
-%        -r (u_0 c_2 + u_2 s_2) + u_4 \\
-%        r u_1 s_2 + u_5
-%    \end{array}\right] $ \\
-%    % Acceleration constraints
-%    $\bm{f}_a$ & $\left[\begin{array}{c}
-%        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
-%        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
-%      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
-%      \end{array}\right]$ \\
-%    % Kinematic differential equations
-%    $\bm{f}_0$ &
-%        $\left[\begin{array}{c}
-%-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
-%\dot{q}_0 s_1 + \dot{q}_2 \\
-%\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
-%\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
-%+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
-%- \dot{q}_5 s_2 c_1\\
-%- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
-%\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
-%         \end{array}\right]$ \\
-%    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
-%             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
-%    % Dynamic differential equations
-%  $\bm{f}_2$ & $\left[\begin{array}{c}
-%-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-%-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-%-I \dot{u}_2 - m r s_2 \dot{u}_4
-%                   \end{array}\right]$ \\
-%
-%  $\bm{f}_3$ & $\left[\begin{array}{c}
-%-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
-%                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
-% I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
-%                   \end{array}\right]$
-%  \end{tabular}
-%\end{table}
-%
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics{rollingdisc.pdf}
-%  \caption{Rolling disc configuration}
-%  \label{fig:disc}
+%  \includegraphics{rollingdisk.pdf}
+%  \caption{Rolling disk configuration}
+%  \label{fig:disk}
 %\end{figure}

### 2013-08-03 nonholonomic_constraint_differentiation.tex
This note shows equivalence of two ways of differentiating motion constraints.
It is obvious to me now, but when I wrote this (>1 year ago) it apparently
wasn't obvious and I felt the need to write it up.

\documentclass[letterpaper,11pt]{article}
\usepackage[round]{natbib}
\usepackage[margin=1in,centering]{geometry}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[pdftex]{hyperref}
\hypersetup{
    pdftitle={Nonholonomic constraint time differentation},
    pdfauthor={Dale Lukas Peterson},
    pdfsubject={Subject},
    pdfkeywords={Keywords}}

\pagestyle{fancy}
\fancyhead[L]{}
\fancyhead[R]{}  % page number on the right
\fancyfoot[L,C,R]{}  %  No footer on left, center or right, on even or odd pages

\begin{document}
\begin{abstract}
  When linear velocity constraints are encountered in the modelling of a
  mechanical system, it is generally possible to find a set of independent
  speeds which allow for these constraints to be solved for the remaining
  dependent speeds in terms of the independent speeds.  Under these conditions
  we investigate whether time differentation of the motion constraints, and
  subsequently solving these time differentiated equations for the time
  derivatives of the dependent speeds yields the same results as if the
  dependent speeds are differentiated directly.  We find that they are equivalent, a finding that has practical implications for symbolic multibody
  dynamic codes.
\end{abstract}

\section{Problem statement}
Assume the constraints are of the form
\begin{equation}
  \label{eq:velocity_constraints}
  B u = 0
\end{equation}
where $u\in\mathbf{R}^o$, and $B \in \mathbf{R}^{m \times o}$.  Assume that $u
= \left[\begin{array}{cc}u_i & u_d\end{array}\right]^{T}$, i.e., that $o-m$
  independent speeds come first, followed by the $m$ dependent speeds.  We can
  then rewrite equation \ref{eq:velocity_constraints} as
\begin{align}
  \label{eq:BiuiBdud}
  B_i u_i + B_d u_d &= 0 \\
  \label{eq:ud}
  \implies u_d &= -B_d^{-1} B_i u_i
\end{align}

Direct differentiation of equation \ref{eq:BiuiBdud} yields
\begin{align}
  \frac{d}{dt} \left(B_i u_i + B_d u_d \right) &= \frac{d}{dt} 0 \\
  \dot{B}_i u_i + B_i \dot{u}_i + \dot{B}_d u_d + B_d \dot{u}_d &= 0 \\
  \implies \dot{u}_d &= -B_d^{-1} \left(\dot{B}_i u_i + B_i \dot{u}_i + \dot{B}_d u_d \right)\\
  &= -B_d^{-1} \left(\dot{B}_i u_i + B_i \dot{u}_i+ \dot{B}_d \left(-B_d^{-1} B_i u_i \right) 
  \right) \\
  &= -B_d^{-1} \left( \dot{B}_i - \dot{B}_d B_d^{-1} B_i \right) u_i - B_d^{-1}
  B_i \dot{u}_i
\end{align}
In contrast, direct differentiation of equation \ref{eq:ud} yields
\begin{align}
  \frac{d}{dt} u_d &= -\frac{d}{dt}\left(B_d^{-1} B_i u_i\right)\\
\implies  \dot{u}_d &= -\frac{d}{dt}\left(B_d^{-1} B_i \right) u_i -
  \left(B_d^{-1} B_i \right) \frac{d}{dt}\left(u_i\right)\\
  &= -\dot{B}_d^{-1} B_i u_i - B_d^{-1} \dot{B}_i u_i
                    - B_d^{-1} B_i \dot{u}_i \\
  &= -\left(\dot{B}_d^{-1} B_i + B_d^{-1} \dot{B}_i\right) u_i - B_d^{-1} B_i \dot{u}_i
\end{align}

The latter terms of both versions of $\dot{u}_d$ are clearly equal.  For the
first term to be equal, we must then have

\begin{align}
-B_d^{-1} \left( \dot{B}_i - \dot{B}_d B_d^{-1} B_i \right) &=
-\left(\dot{B}_d^{-1} B_i + B_d^{-1} \dot{B}_i\right) \\
\implies B_d^{-1} \dot{B}_d B_d^{-1} B_i &= -\dot{B}_d^{-1} B_i\\
\implies B_d^{-1} \dot{B}_d B_d^{-1} &= -\dot{B}_d^{-1} \\
\implies \dot{B}_d B_d^{-1} + B_d \dot{B}_d^{-1} &= 0 \\
\implies \frac{d}{dt} \left(B_d B_d^{-1} \right) &= 0 \\
\implies \frac{d}{dt} I &= 0
\end{align}

\end{document}

### 2013-08-05 PetersonGedeHubbard2012.tex
Finish example. Add references. Add steady turn equilibrium.

@@
 \section{Dynamics Example}
+\label{sec:example}
 To illustrate Kane's method in the presence of configuration and velocity
@@
 do not explicitly involve $\dot{q}_i$ terms. This structure will be taken
-advantage of subsequently.
+advantage of in Section \ref{sec:derivations}.
 
@@
 is linearized about the upright steady rolling condition, the critical speed is
-$v^*=\pm\frac{1}{\sqrt{3}}$. When following the naive approach, and
+$v \triangleq -r\dot{q}_3 =\pm\frac{1}{\sqrt{3}}$. When following the naive approach, and
 evaluating the Jacobian at the same operating conditions and parameters, eight
@@
 (five coordinates and three speeds). We now outline a linearization procedure
-which addresses this issue and then revisit this example applying the proper
-linearization procedure.
+which addresses this issue; we revisit this example in
+\ref{sec:example_revisited}.
 
@@
 
-\section{Dynamics Example, Revisted}
+\section{Dynamics Example, Revisited}
+\label{sec:example_revisited}
+As shown at the end of Section \ref{sec:example}, obtaining correct linearized
+equations in the presence of constraints requires a technique other than
+straightforward calculation of the Jacobian. In this section, we demonstrate
+that our linearization procedure yields eigenvalues which match published
+results~\cite{Schwab2003,Kane1985,Neimark1972}. The first step in the procedure
+is to form the matrices in Equation (\ref{eq:quant_to_compute}). Once obtained,
+these matrices can be evaluated at the equilibrium conditions and parameters of
+interest.  The standard approach to finding the equilibrium conditions is to
+\textit{choose} the independent quantities and solve the constraint equations
+for the dependent quantities.
+
+We first consider the case of upright steady cruise and begin by choosing the
+independent coordinates to be zero ($q_i^* = 0$, $i = 1,\dots,5$), which
+implies (by appealing to the configuration constraint) that $q_6^* = -r
+\cos{q_2} = -r$.  This corresponds to the disk upright, the lowest point of the
+disc in contact with the ground, and the disk heading aligned with the
+$\hat{n}_x$ unit vector.  Next, we solve the velocity constraint equations for
+the dependent speeds in terms of the independent ones, and substitute these
+expression into the kinematic differential equations. This yields six equations
+with nine unknowns: $u_i (i=1,2,3)$ and $\dot{q}_i (i=1,\dots6)$. We choose yaw
+rate $\dot{q}_1^* = 0$ and lean rate $\dot{q}_2^* = 0$, and let spin rate
+$\dot{q}_3^*$ be a free variable, and solve for the remaining six unknowns.
+This yields $u_1^* = u_3^* = 0$, $u_2^* = \dot{q}_3^*$, $\dot{q}_4^* =
+-r\dot{q}_3^*$, $\dot{q}_5^* = \dot{q}_6^* = 0$. Back substituting the first
+three of these results into the velocity constraint equations yields $u_4^* =
+-r\dot{q}_3^*$, $u_5^* = u_6^* = 0$. Finally, by evaluating the acceleration
+constraints $f_a$ and the dynamic equations $f_2 + f_3$ at these conditions, we
+can solve for $\dot{u}_i^*$, $i = 1,\dots6$. We obtain $\dot{u}_1^* =
+\dot{u}_2^* = \dot{u}_3^* = \dot{u}_4^* = \dot{u}_5^* = 0$, and $\dot{u}_6^* =
+-r \dot{q}_3^{*2}$. The upright equilibrium conditions are thus parameterized
+in terms of the disc spin rate $\dot{q}_3$; for this upright configuration
+($q_2 = 0$) it is convenient to introduce the forward speed $v \triangleq
+-r\dot{q}_3^*$.
+
+Evaluating the matrices in Equations (\ref{eq:quant_to_compute}) and
+(\ref{eq:C_0}-\ref{eq:C_2}) at these equilibrium conditions permits the
+matrices in the linearized equations of motion
+(\ref{eq:state_space_constrained}) to be computed. Solving these equations for
+$\delta\dot{\mathbf{q}}$ and $\delta\dot{\mathbf{u}}$ yields the linearized
+relationship between the time derivatives of all state variables and the
+independent state variables. By taking only the rows associated with the
+independent states, the linear relationship between the independent states and
+their time derivatives are formed. The eigenvalues, parameterized by forward
+speed $v$, of this coefficient matrix may be computed symbolically. Six of the
+eight eigenvalues of this matrix are zero, and the remaining two are
+\begin{align}
+\label{eq:upright_evals}
+  \lambda_{1,2} &= \pm \frac{2\sqrt{1 - 3v^2}}{\sqrt{5}}
+\end{align}
+which has critical points at $v^*=\pm\frac{1}{\sqrt{3}}$ and matches previously
+published results~\cite{Schwab2003,Kane1985,Neimark1972}. For $|v| < v^*$,
+the disk is unstable, but for $|v| \geq v^*$, the eigenvalues are purely
+imaginary and are hence marginally stable. We omit the details of these
+calculation and direct the reader to the electronic supplementary material.
+
+A more general analysis of the stability of the rolling disk equilibria
+considers the case where the disk lean $q_2$ is constant but not necessarily
+zero. To satisfy the dynamics, the yaw rate $\dot{q}_1$ and spin rate
+$\dot{q}_3$ must be chosen to satisfy
+\begin{align}
+  \frac{g}{r}\sin{(q_2)}  + \frac{3}{2}\cos{(q_2)} \dot{q}_3 \dot{q}_1 +
+  \frac{5}{4}\cos{(q_2)}\sin{(q_2)} \dot{q}_1^2
+ &= 0
+\end{align}
+which is a quadratic equation in the yaw rate $\dot{q}_1$ with roots
+\begin{align}
+\dot{q}_1 &= \frac{r}{2 g \sin\left(q_{2}\right)} \left( - \frac{3}{2} \dot{q}_{3}\cos\left(q_{2}\right) \pm \sqrt{- \frac{5 g}{r} \sin^{2}\left(q_{2}\right) \cos\left(q_{2}\right) + \frac{9}{4} \dot{q}_{3}^{2} \cos^{2}\left(q_{2}\right)} \right)
+\end{align}
+To be physically meaningful, these roots must be real, which gives rise to the following additional requirement
+\begin{align}
+  \dot{q}_3^2 &\geq \frac{20g}{9r}\sin{(q_2)} \tan{(q_2)}
+\end{align}
+For a given set of parameters the eigenvalues of the linearized dynamics can be
+parameterized by three quantities lean $q_2$, yaw rate $\dot{q}_1$, and spin
+rate $\dot{q}_3$ (though only two can be independent because of the above
+restrictions). For this more general equilibrium condition, there are still six
+zero eigenvalues, while the two non-zero eigenvalues are
+\begin{align}
+\lambda_{1,2} = \pm\sqrt{\frac{4}{5} \cos\left(q_{2}\right) - \dot{q}_{1}^{2} -\frac{14}{5} \sin\left(q_{2}\right) \dot{q}_{1} \dot{q}_{3} - \frac{12}{5}\dot{q}_{3}^{2}}
+\end{align}
+which can be easily shown to reduce to Equation (\ref{eq:upright_evals}) when
+$q_2 = \dot{q}_1 = 0$. For a more detailed stability of steadily rolling disks,
+see~\cite{O'Reilly1996,Neimark1972,Kuleshov2001}.
 
@@
 state space system $\dot{\bm{x}}_i = A \bm{x}_i + B \bm{r}$ to which standard
-linear systems analyses may applied.  It is worth noting that the rows
-of $A^\prime$ and $B^\prime$ which correspond to dependent states can be used
-in the output or measurement equations of a linear state space model, as in the
-case when the controlling or measuring dependent state(s).
+linear systems analyses may applied.  It is worth noting that the rows of
+$A^\prime$ and $B^\prime$ which correspond to dependent states can be used in
+output or measurement equations of a linear state space model (e.g.,
+measurements from an accelerometer, or measurements of dependent speeds).
 
@@
 
-%\section{Example}
-%\label{example}
-%We apply this linearization technique to the well studied rolling disk.  Figure
-%\ref{fig:disk} illustrates the choice of coordinates, speeds, and parameters.
-%The disk is assumed to be knife edged of radius $r$, mass $m$, spin inertia
-%$J$, and transverse inertia $I$.  To illustrate the concept of dependent
-%quantities, we purposefully choose to describe the configuration and velocity
-%of the disk with a non-minimal set of coordinates and speeds.  This choice is
-%not purely academic.  It is convenient for a physical disk equipped with a rate
-%gyroscope and accelerometer to measure these body fixed quantities (or their
-%time derivatives) directly, rather than the more traditional set which is
-%chosen for analytic convenience.  In more complicated systems with constraints,
-%it is difficult (if not impossible) to describe the system with a minimal
-%choice of coordinates and speeds; the ideas demonstrated by this simple example
-%can be applied to more complicated ones.
-%
-%The orientation of the disk-fixed frame $C$ relative to the inertial frame $N$
-%is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
-%(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
-%location of disk center $C^*$ relative to the inertial origin $N^*$, and the
-%velocity of $C^*$ in $N$ are
-%\begin{align}
-%  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
-%  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
-%  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
-%\end{align}
-%If the disk is in contact with the ground, a configuration constraint involving
-%the height $q_5$ of the disk center and the lean $q_1$.  Similarly, if the disk
-%rolls without slip, other constraints relate the generalized speeds $u_i, (i =
-%0,\dots,5)$.  The derivations of these constraints and the equations of motion
-%are presented in the electronic supplementary material and can also be found in
-%the literature \cite{O'Reilly1996}.
-%\begin{table}[htbp]
-%  \centering
-%  \caption{Rolling disk governing definitions and equations}
-%  \label{table:rollingdisk}
-%  \begin{tabular}[c]{l l l}
-%    Quantity & Shape & Description\\
-%    \hline
-%    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
-%    derivatives\\
-%    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
-%  \end{tabular}
-%  \begin{tabular}[c]{r @{ $=$ } l}
-%%    \multicolumn{3}{c}{ } \\
-%%    \multicolumn{2}{c}{Equation} &           \\
-%    \hline
-%    % Configuration constraint
-%    $\bm{f}_c$ & $rc_1 + q_5$ \\
-%    % Velocity constraints
-%    $\bm{f}_v$ & $\left[\begin{array}{c}
-%        r u_1 c_2 + u_3 \\
-%        -r (u_0 c_2 + u_2 s_2) + u_4 \\
-%        r u_1 s_2 + u_5
-%    \end{array}\right] $ \\
-%    % Acceleration constraints
-%    $\bm{f}_a$ & $\left[\begin{array}{c}
-%        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
-%        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
-%      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
-%      \end{array}\right]$ \\
-%    % Kinematic differential equations
-%    $\bm{f}_0$ &
-%        $\left[\begin{array}{c}
-%-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
-%\dot{q}_0 s_1 + \dot{q}_2 \\
-%\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
-%\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
-%+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
-%- \dot{q}_5 s_2 c_1\\
-%- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
-%\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
-%         \end{array}\right]$ \\
-%    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
-%             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
-%    % Dynamic differential equations
-%  $\bm{f}_2$ & $\left[\begin{array}{c}
-%-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-%-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-%-I \dot{u}_2 - m r s_2 \dot{u}_4
-%                   \end{array}\right]$ \\
-%
-%  $\bm{f}_3$ & $\left[\begin{array}{c}
-%-I u_1 u_2 + J u_1 u_2 + g m r s_1 c_2 - m r (-u_0 u_5 + u_2 u_3) c_2 \\
-%                m r (u_0 u_4 - u_1 u_3) s_2 + m r (u_1 u_5 - u_2 u_4) c_2 \\
-% I u_0 u_1 - J u_0 u_1 + g m r s_1 s_2 - m r (-u_0 u_5 + u_2 u_3) s_2
-%                   \end{array}\right]$
-%  \end{tabular}
-%\end{table}
-%
-%
-%\begin{figure}
-%  \centering
-%  \includegraphics{rollingdisk.pdf}
-%  \caption{Rolling disk configuration}
-%  \label{fig:disk}
-%\end{figure}
-%
-%The matrices which make up the linear equations were created using the method
-%presented in Section \ref{sec:results}. It is important to note that the
-%entries in these matrices must be evaluated at the point of linearization
-%($\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*$), and that this point of
-%linearization must satisfy the equations in Table \ref{table:assumptions}.
-%There is no requirement that this point be an equilibrium point, which implies
-%that in general both $\bm{\dot{q}}$ and $\bm{\dot{u}}$ may be non-zero. To
-%reduce the complexity of the equations, we chose $q_0^* = q_2^* = 0$.  This
-%means that the other quantities in the linearized equations (other $q$'s, the
-%$\dot{q}$'s, the $u$'s, and the $\dot{u}$'s) need to be solved for with these 2
-%quantities already set to 0.
-%
-%The independent state vector is:
-%\begin{align}
-%\begin{bmatrix}
-%\delta \bm{q}_i \\
-%\delta \bm{u}_i
-%\end{bmatrix} =
-%\begin{bmatrix}
-%q_0 &
-%q_1 &
-%q_2 &
-%q_3 &
-%q_4 &
-%u_0 &
-%u_1 &
-%u_2
-%\end{bmatrix}^T
-%\end{align}
-%
-%And the complete (independent and dependent) state vector is:
-%\begin{align}
-%\begin{bmatrix}
-%\delta \bm{q} \\
-%\delta \bm{u}
-%\end{bmatrix} =
-%\left[
-%\begin{array}{cccccccccccc}
-%q_0 &
-%q_1 &
-%q_2 &
-%q_3 &
-%q_4 &
-%q_5 &
-%u_0 &
-%u_1 &
-%u_2 &
-%u_3 &
-%u_4 &
-%u_5
-%\end{array}
-%\right]^T
-%\end{align}
-%
-%The coefficient matrices of the linearized equations (Equation
-%\ref{eq:state_space_constrained}) are:
-%\begin{equation}
-%  \left[
-%    \begin{array}{cc}
-%      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
-%      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
-%      \tilde{M}_{uqd} & \tilde{M}_{uud}
-%    \end{array}
-%    \right]=
-%\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
-%\end{equation}
-%
-%
-%\begin{equation*}
-%   \left[
-%     \begin{array}{cc}
-%       (\tilde{A}_{qq} + \tilde{A}_{qu} C_1 ) C_0 & \tilde{A}_{qu} C_2 \\
-%       (\tilde{A}_{uqc} + \tilde{A}_{uuc} C_1 ) C_0 & \tilde{A}_{uuc} C_2\\
-%       (\tilde{A}_{uqd} + \tilde{A}_{uud} C_1 ) C_0 & \tilde{A}_{uud} C_2
-%     \end{array}
-%   \right]=
-%\end{equation*}
-%\begin{equation}
-%\left[\begin{smallmatrix}0 & 0 & c_{1} \dot{q}_{0} & 0 & 0 & 1 & 0 & 0\\0 & - c_{1} \dot{q}_{0} & 0 & 0 & 0 & 0 & 1 & 0\\0 & s_{1} \dot{q}_{0} & - \dot{q}_{1} & 0 & 0 & 0 & 0 & 1\\- \dot{q}_{4} & 0 & c_{1} \dot{q}_{5} - s_{1} \dot{q}_{4} & 0 & 0 & 0 & - r & 0\\c_{1} \dot{q}_{3} & - c_{1} \dot{q}_{5} + s_{1} \dot{q}_{4} & r u_{2} & 0 & 0 & r & 0 & 0\\- s_{1} \dot{q}_{3} & c_{1} \dot{q}_{4} + s_{1} \dot{q}_{5} & - r u_{1} - \dot{q}_{3} & 0 & 0 & 0 & 0 & 0\\0 & 0 & r u_{1} \dot{q}_{2} & 0 & 0 & 0 & 0 & 0\\0 & 0 & - r u_{0} \dot{q}_{2} & 0 & 0 & 0 & 0 & r \dot{q}_{2}\\0 & 0 & 0 & 0 & 0 & 0 & - r \dot{q}_{2} & 0\\0 & - c_{1} g m r & m r^{2} u_{0} u_{1} & 0 & 0 & - m r u_{5} & - \frac{5}{4} m r^{2} u_{2} & \frac{1}{4} m r \left(- r u_{1} + 4 u_{3}\right)\\0 & 0 & m r \left(r u_{1} + r u_{2} - u_{0} u_{4} + u_{1} u_{3} - \dot{u}_{5}\right) & 0 & 0 & m r^{2} u_{2} & - m r u_{5} & m r u_{4}\\0 & 0 & m r \left(- g s_{1} - u_{0} u_{5} + u_{2} u_{3} + \dot{u}_{4}\right) & 0 & 0 & \frac{1}{4} m r^{2} u_{1} & \frac{1}{4} m r^{2} u_{0} & 0\end{smallmatrix}\right]
-%\end{equation}
-%
-%With this example, we have used our linearization procedure to find linear
-%equations of motion which respect the system's constraints. Parametric
-%stability analysis (eigenvalue analysis) can now be performed on these
-%equations.  These computations were implemented symbolically with the open
-%source symbolic manipulator SymPy \cite{SymPy2012}.
-%
 \section{Conclusions}
@@
 
-The procedure has been implemented
-symbolically in the \texttt{sympy.physics.mechanics} sub-module of the open
-source symbolic manipulator SymPy\cite{SymPy2012}.
+The procedure has been implemented symbolically in the
+\texttt{physics.mechanics} sub-module of the open source symbolic
+manipulator SymPy\cite{SymPy2012}. In addition to applying it on the rolling
+disk example, we have also applied it to the Whipple bicycle model and found
+that it matches eigenvalues published for a benchmark set of
+parameters~\cite{Meijaard2007}.
 

### 2013-08-05 rolling_disc_luke.py

@@ -222,45 +222,46 @@ assert(f_a_eq == zeros((3,1)))
 assert(f_0_eq_plus_f_1_eq == zeros((6,1)))
 assert(f_2_eq_plus_f_3_eq == zeros((3,1)))
 
-M_qq = f_0.jacobian(qd).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-M_uqc = f_a.jacobian(qd).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-M_uuc = f_a.jacobian(ud).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-M_uqd = f_2.jacobian(qd).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-M_uud = f_2.jacobian(ud).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-
-A_qq = -(f_0 + f_1).jacobian(q).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-
-A_qu = -f_1.jacobian(u).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-A_uqc = -f_a.jacobian(q).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-A_uuc = -f_a.jacobian(u).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-A_uqd = -(f_2 + f_3).jacobian(q).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-A_uud = -f_3.jacobian(u).subs(eq_ud).subs(eq_u).subs(eq_qd).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
-
[…]

### 2013-08-06 real_roots_steady.py
Add plot to show regions where equilibrium are feasible.

@@ -1,0 +1,29 @@
+import matplotlib.pyplot as plt
+import numpy as np
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+
+g = 1.0
+r = 1.0
+q2_limit = np.pi/2*.999
+q2 = np.linspace(-q2_limit, q2_limit, 500)
+q3dot_1 = -np.sqrt(20.0*g/9/r*np.sin(q2)*np.tan(q2))
+q3dot_2 = np.sqrt(20.0*g/9/r*np.sin(q2)*np.tan(q2))
+fig, ax = plt.subplots()
+
+ax.fill_between(q2, q3dot_1, q3dot_2, facecolor='grey')
+ax.set_xlabel(r'Lean $q_2$ (rad)')
+ax.set_ylabel(r'Spin rate $\dot{q}_3$ (rad/s)')
+ax.xaxis.set_ticks([-q2_limit, -np.pi/4, 0, np.pi/4, q2_limit])
+ax.xaxis.set_ticklabels([r"$-\frac{\pi}{2}$", r"$-\frac{\pi}{4}$", r"$0$", r"$\frac{\pi}{4}$", r"$\frac{\pi}{2}$"])
+ax.yaxis.set_ticks([-2*q2_limit, -2*np.pi/4, 0, 2*np.pi/4, 2*q2_limit])
+ax.yaxis.set_ticklabels([r"$-\pi$", r"$-\frac{\pi}{2}$", r"$0$", r"$\frac{\pi}{2}$", r"$\pi$"])
+ax.set_xbound(-q2_limit, q2_limit)
+ax.set_ybound(-2*q2_limit, 2*q2_limit)
+plt.text(-np.pi/8, np.pi/2, 'Equilibrium feasible')#, fontdict=font)
+plt.text(-np.pi/8, -np.pi/2, 'Equilibrium feasible')#, fontdict=font)
+plt.text(-3*np.pi/8, 0, 'Equilibrium infeasible')#, fontdict=font)
+plt.text(np.pi/8, 0, 'Equilibrium infeasible')#, fontdict=font)
+#plt.title("Feasibility regions for rolling disk, $m = r = g = 1$")
+plt.show()
+

### 2013-08-06 rolling_disc.py
Remove outdated example.

@@ -1,114 +1,0 @@
-from sympy import *
-from sympy.physics.mechanics import *
-
-psi, phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z = dynamicsymbols('psi,'
-    'phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z')
-psid, phid, thetad, oxd, oyd, ozd, w_xd, w_yd, w_zd, v_xd, v_yd, v_zd = dynamicsymbols('psi,'
-    'phi, theta, ox, oy, oz, w_x, w_y, w_z, v_x, v_y, v_z', 1)
-r, m, g, I, J = symbols('r, m, g, I, J')
-t = symbols('t')
-
-N = ReferenceFrame('N')
-Y = N.orientnew('Y', 'Axis', [psi, N.z])
-L = Y.orientnew('L', 'Axis', [phi, Y.x])
-R = L.orientnew('R', 'Axis', [theta, L.y])
-
-O = Point('O')
-C = O.locatenew('C', r * L.z)
-
-R.set_ang_vel(N, w_x * L.x + w_y * L.y + w_z * L.z)
-O.set_vel(N, v_x * L.x + v_y * L.y + v_z * L.z)
-C.v2pt_theory(O, N, R)
-
-ccons = [r * (L.z & Y.z) - oz]
-vcons = [C.vel(N) & Y.x, C.vel(N) & Y.y, C.vel(N) & Y.z]
-
-Disk = RigidBody('Disk', O, R, m, (inertia(L, I, J, I), O))
-
-v_o_exp_n = O.vel(N).express(N)
-
-kd = ([psid - w_z / cos(psid), phid - w_x, thetad - w_y + w_z * tan(phi),
-       (v_o_exp_n & N.x) - oxd, (v_o_exp_n & N.y) - oyd, (v_o_exp_n & N.z) - ozd])
-
-Bodies = [Disk]
-Forces = [(O, m * g * Y.z)]
-
-KM = Kane(N)
-KM.coords([psi, phi, theta, ox, oy], [oz], ccons)
-KM.speeds([w_x, w_y, w_z], [v_x, v_y, v_z], vcons)
-KM.kindiffeq(kd)
-KM.kanes_equations(Forces, Bodies)
-
-mm = KM.mass_matrix_full
-mprint(mm)
-f = KM.forcing_full
[…]

### 2013-08-06 PetersonGedeHubbard2012.tex
Final edits, add references on LCE computation.

@@
 with a non-traditional set of coordinates and generalized speeds which are not
-all independent. We then provide a systematic method to correctly generate the
+all independent. We provide a systematic method to correctly generate the
 linear equations of motion for constrained systems which have been found using
 Kane's method (or any other method which results in equations with the same
-form). We then apply this method to the rolling disk example to illustrate that
+form). We apply this method to the rolling disk example to illustrate that
 it yields eigenvalues that match previously published results.
@@
 example. In response to this need, we present the newly developed linearization
-procedure and its derivation.  This new linearization procedure is then applied
-the same rolling disk example to illustrate its validity and an example of its
-application.  Finally, we discuss other details and nuances of the procedure
-and its use, as well as directions for future work.
-
-\section{Kane's Method, Briefly}
+procedure and its derivation. This new linearization procedure is applied the
+same rolling disk example to illustrate its validity and an example of its
+application. Finally, we discuss other details and nuances of the procedure and
+its use, as well as directions for future work.
+
+\section{Kane's Method, briefly}
 \label{sec:kane_method}
@@
 or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
-defined as the sum of three generalized-speed/basis-vector products.
-The derivatives of these generalized speeds will then appear in the equations
+defined as the sum of three generalized--speed basis--vector products.
+The derivatives of these generalized speeds will appear in the equations
 of motion, rather than the twice time differentiated generalized coordinates
@@
 complicated (in comparison to $\dot{q}_i = u_i$), but this is usually offset by
-the significantly simpler dynamical equations of motion\cite{Mitiguy1996}
+the significantly simpler dynamical equations of motion~\cite{Mitiguy1996}.
 
 Use and understanding of Kane's method requires the use and understanding of
-partial velocities and partial angular velocities. These are simply the the
+partial velocities and partial angular velocities. These are simply the
 partial derivatives of the (angular) velocity vectors with respect to the
@@
 In contrast to the velocity constraints, our formulation permits the
-acceleration constraints to involve $\dot{\mathbf{q}}$ terms (althernatively
+acceleration constraints to involve $\dot{\mathbf{q}}$ terms (alternatively
 they can be eliminated by appealing to the kinematic differential equations).
@@
 where $\mathbf{f}_{vt}$ arises from explicit time varying terms that occur in
-the velocity constraints. These typically arise from specified quantites whose
+the velocity constraints. These typically arise from specified quantities whose
 dynamics are being ignored (e.g., a motor spinning at constant speed).
@@
 \begin{align}
-\label{eq:Kanes_nonholonomic}
+\label{eq:kanes_eq_nonholonomic}
 \tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
@@
 This process of constraining the equations of motion reduces the number of
-equations from $o$ to $o-m$. However, the (\ref{eq:Kanes_nonholonomic}) still
+equations from $o$ to $o-m$. However, the (\ref{eq:kanes_eq_nonholonomic}) still
 contain $o$ unknowns ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$). By
-augmenting (\ref{eq:Kanes_nonholonomic}) with the acceleration constraint
-equations (\ref{eq:acceleration_constraints}), we end up with $o$ equations and
-$o$ unknowns. Note that equations (\ref{eq:Kanes_nonholonomic}) and
+augmenting (\ref{eq:kanes_eq_nonholonomic}) with the acceleration constraint
+equations (\ref{eq:acceleration_constraints}), we obtain $o$ equations in
+$o$ unknowns. Note that equations (\ref{eq:kanes_eq_nonholonomic}) and
 (\ref{eq:acceleration_constraints}) are linear in the $\dot{u}_i$'s. The
@@
 The unconstrained partial angular velocity and partial velocity of $C^*$,
-respectively, are then
+respectively, are
 \begin{align}
@@
 The kinematic differential equations relate the time derivatives of the
-coordinates and the generalized speeds. Here they are solved by equating the
-velocities written with time-differentiated generalized coordinates and the
-velocities written with generalized speeds.  From the addition theorem for
-angular velocity, we may write the angular velocity of $C$ as
+coordinates and the generalized speeds. They are obtained by equating the
+velocities written using time-differentiated generalized coordinates and the
+velocities written using generalized speeds. From the addition theorem for
+angular velocity, the angular velocity of $C$ is
 \begin{equation}
@@
 \end{align}
-and then take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
+and take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
 \mathbf{u}]^T$. However, this will give the incorrect results. As discussed in
@@
 \begin{align}
+  \label{eq:evals_incorrect}
   \lambda_{1,2}=\pm\sqrt{-v^2},\quad \lambda_{3,4} = \pm\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
@@
 (five coordinates and three speeds). We now outline a linearization procedure
-which addresses this issue; we revisit this example in
+which addresses this issue; we revisit this example in Section
 \ref{sec:example_revisited}.
@@
 \label{sec:derivations}
-
-In response to the need we have demonstrated, we present a solution to this
-constrained, linearization problem.
-There is a reason that the equations generated by taking the gradient with
-respect to the generalized speeds are incorrect.
-When finding the equations of motion for a constrained system, some generalized
-coordinates and speeds are classified as dependent speeds, meaning their
-behavior is determined by the other coordinates and speeds.
-These terms should in fact be written as:
+In response to the need we have demonstrated, we present a linearization
+procedure that properly accounts for system constraints. Taking the Jacobian
+of the right hand side of the ODE's as we did at the end of the previous
+section is incorrect because it fails to apply the chain rule and thereby
+properly account for the relationships imposed by configuration, velocity, and
+acceleration constraints. That it isn't immediately obvious that the chain rule
+needs to be applied is a byproduct of the commonly use notation which doesn't
+make it explicitly clear that dependent coordinates and dependent speeds are not
+only functions of time, but also functions of the independent coordinates and
+independent speeds. These dependent terms should in fact be written as
 \begin{align}
 \label{eq:q_d_redefined}
-\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}, t) \\
+\mathbf{q}_d (t) \to \mathbf{q}_d (\mathbf{q}_i, t) \\
 \label{eq:u_d_redefined}
-\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, t)
-\end{align}
-Then, upon taking the derivative of each of these terms with respect to an
-independent speed, the dependent speed would need to be solved for and its
-derivative found.
-No previous authors appear to have made clear that this distinction should be
-made when using Kane's method.
-
-What we present addresses this issue.
-Rather then substituting in the solutions to the dependent speeds though, we
-use knowledge of the structure of the equations and constraints (as defined in
-Table \ref{table:assumptions}) to generate linear equations.
+\mathbf{u}_d (t) \to \mathbf{u}_d (\mathbf{q}_i, \mathbf{u}_i, t)
+\end{align}
+While this might seem obvious, no previous author appears to make this
+explicitly clear when presenting techniques for symbolically linearizing
+equations of motion which are subject to constraints. While the concept is
+simple in principle, correctly accounting for all quantities is tedious and
+error prone. Having a high level, systematic procedure that can be implemented
+reliably in software is therefore a strong argument for the detailed and
+explicit procedure we present.
 
@@
 assumed that all of the equations in Table \ref{table:assumptions} are
-satisfied by these quantities (i.e., the system is in a valid state).
-In the interest of brevity, we omit writing this point of linearization in each
-gradient in the calculations below; all are evaluated at this point.
-Expansion of the three constraint equations, keeping only first order terms,
-yields
+satisfied by these quantities (i.e., the system is in a valid state).  In the
+interest of brevity, we omit writing this point of linearization in each
+gradient in the calculations below; all are evaluated at this point. Expansion
+of the three constraint equations, keeping only first order terms, yields
 \begin{align}
@@
 address this issue, a smaller set of independent coordinates and speeds must be
-selected. To this end, consider the following partitioning of the generalized
-coordinates and generalized speeds:
+selected. To this end, we partition the generalized coordinates and generalized
+speeds as
 \begin{equation*}
@@
 
-By assumption, equation
-(\ref{eq:configuration_expansion}) is zero; making use of $P_q$, we have
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
+Making use of $P_q$ and the assumption that equation
+(\ref{eq:configuration_expansion}) is zero gives
+\begin{align}
+\nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q} &= \nabla_{\bm{q}}\bm{f}_{c} P_{q} \delta \bm{\tilde{q}} \notag \\
    &= \nabla_{\bm{q}}\bm{f}_{c} P_{qi} \delta \bm{q_i} +
@@
   (\nabla_{\bm{q}}\bm{f}_{c} P_{qi}) \delta \bm{q}_i \notag\\
-  \implies \delta \bm{q} &= \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi} \delta
-    \bm{q}_i\\
   \label{eq:delta_q}
-  \delta \bm{q} &= C_0 \delta \bm{q}_i
-\end{align}
-
-Applying the same approach to (\ref{eq:velocity_expansion}), we obtain
-\begin{align}
-  \bm{0} &= \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
-  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}\notag\\
-  &= \nabla_{\bm{q}} \bm{f}_{v}
-\delta \bm{q} + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
+  \implies \delta \bm{q} &= \underbrace{\left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+  \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right]
+  P_{qi}}_{\triangleq C_0} \delta \bm{q}_i
+\end{align}
+Making use of $P_q$, $P_u$, the assumption that equation
+(\ref{eq:velocity_expansion}) is zero, and taking equation (\ref{eq:delta_q})
+into account gives
+\begin{align}
+\nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} + \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u}
+  &= \nabla_{\bm{q}} \bm{f}_{v} C_0 \delta \bm{q}_i +\nabla_{\bm{u}} \bm{f}_{v}
+  \delta P_u \bm{\tilde{u}} \notag \\
+  &=\nabla_{\bm{q}} \bm{f}_{v} C_0 \delta \bm{q}_i + \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i +
 \nabla_{\bm{u}} \bm{f}_{v} P_{ud} \delta \bm{u}_d \notag\\
+%
 \implies \delta \bm{u}_d &= -\left(\nabla_{\bm{u}} \bm{f}_{v}
-P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} \delta\bm{q} +
+P_{ud}\right)^{-1}\left[\nabla_{\bm{q}}\bm{f}_{v} C_0 \delta \bm{q}_i +
   \nabla_{\bm{u}} \bm{f}_{v} P_{ui} \delta \bm{u}_i \right]\notag\\
-  \implies \delta \bm{u} &= -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \delta \bm{q}\notag\\
-  &+ \left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui} \delta \bm{u}_i\\
-  \delta \bm{u} &= C_1 \delta \bm{q} + C_2 \delta \bm{u}_i\notag\\
+  \implies \delta \bm{u} &= \underbrace{-P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+  \nabla_{\bm{q}} \bm{f}_{v}}_{\triangleq C_1} C_0 \delta \bm{q}_i \notag\\
   \label{eq:delta_u}
-  \delta \bm{u} &= C_1 C_0 \delta \bm{q}_i + C_2 \delta \bm{u}_i
-\end{align}
-
-By making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
-rewrite equation (\ref{eq:state_space_unconstrained}).  This gives Equation
-\ref{eq:state_space_constrained}, a linear system of equations which relates
-the time derivatives of all states (dependent and independent) to the
-independent states and exogenous inputs.
-
-The linearized equations of motion are
+  &+ \underbrace{\left[I - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+  \bm{f}_{v} \right] P_{ui}}_{\triangleq C_2} \delta \bm{u}_i
+\end{align}
+
+Making use of equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}), we can
+rewrite equation (\ref{eq:state_space_unconstrained}) as
 \begin{align}
@@
     \right]
-    +
-    \left[
+    \notag \\
+    &+\left[
       \begin{array}{c}
@@
 \end{align}
-
-where
-
-\begin{align}
-  \label{eq:C_0}
-  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
-    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
-  \label{eq:C_1}
-  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
-  \nabla_{\bm{q}} \bm{f}_{v} \\
-  \label{eq:C_2}
-  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
-    \bm{f}_{v} \right] P_{ui}\\
-  \label{eq:Pq}
-  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
-  \label{eq:Pu}
-  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
-\end{align}
-
-Table \ref{table:assumptions} provides the form in which the nonlinear
-equations of motion and constraints must be arranged in order to carry out the
-linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
-bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
-are Jacobian matrices of the differential equations. Equations
-(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
-embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
-permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
-independent and dependent speeds.
-
-The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
-of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
-\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
-satisfy the equations in Table \ref{table:assumptions}, and may or may not be
-an equilibrium point.
+This definitively establishes how the first time derivatives of coordinates and
+speeds (independent \textit{and} dependent) depend, to first order, upon a
+selection of independent coordinates and independent speeds, for an arbitrary
+point of linearization. Note that the only requirement on this point of
+linearization is that it satisfy all the equations in Table
+\ref{table:assumptions}; it may or may not be an equilibrium point.
+
+%Table \ref{table:assumptions} provides the form in which the nonlinear
+%equations of motion and constraints must be arranged in order to carry out the
+%linearization procedure.  Equations (\ref{eq:quant_to_compute}) provides the
+%bulk of the terms in equation (\ref{eq:state_space_constrained}); these terms
+%are Jacobian matrices of the differential equations.
+%
+%The linearized equations of motion are
+%
+%where
+%
+%\begin{align}
+%  \label{eq:C_0}
+%  C_0 &\triangleq \left[ I_{n \times n} - P_{qd}(\nabla_{\bm{q}}
+%    \bm{f}_{c} P_{qd})^{-1} \nabla_{\bm{q}} \bm{f}_{c} \right] P_{qi}\\
+%  \label{eq:C_1}
+%  C_1 &\triangleq -P_{ud}(\nabla_{\bm{u}} \bm{f}_{v} P_{ud})^{-1}
+%  \nabla_{\bm{q}} \bm{f}_{v} \\
+%  \label{eq:C_2}
+%  C_2 &\triangleq \left[I_{o \times o} - P_{ud} (\nabla_{\bm{u}}\bm{f}_{v} P_{ud})^{-1} \nabla_{\bm{u}}
+%    \bm{f}_{v} \right] P_{ui}\\
+%  \label{eq:Pq}
+%  \bm{q} &= P_{qi} \bm{q}_{i} + P_{qd} \bm{q}_{d} \\
+%  \label{eq:Pu}
+%  \bm{u} &= P_{ui} \bm{u}_{i} + P_{ud} \bm{u}_{d}
+%\end{align}
+
+%Equations
+%(\ref{eq:C_0})-(\ref{eq:C_2}) provide the terms $C_0$, $C_1$, and $C_2$ which
+%embed the constraints into (\ref{eq:state_space_constrained}). Finally, the
+%permutation matrices in equations (\ref{eq:Pq}) and (\ref{eq:Pu}) separate the
+%independent and dependent speeds.
+%
+%The matrices in (\ref{eq:state_space_constrained}) are evaluated at the point
+%of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
+%\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*, \bm{r}=\bm{r}^*$; this point must
+%satisfy the equations in Table \ref{table:assumptions}, and may or may not be
+%an equilibrium point.
 
@@
 these matrices can be evaluated at the equilibrium conditions and parameters of
-interest.  The standard approach to finding the equilibrium conditions is to
-\textit{choose} the independent quantities and solve the constraint equations
-for the dependent quantities.
+interest. We follow the standard approach to finding the equilibrium conditions
+by \textit{choosing} the independent quantities and solving the constraint
+equations for the dependent quantities.
 
@@
 the dependent speeds in terms of the independent ones, and substitute these
-expression into the kinematic differential equations. This yields six equations
-with nine unknowns: $u_i (i=1,2,3)$ and $\dot{q}_i (i=1,\dots6)$. We choose yaw
-rate $\dot{q}_1^* = 0$ and lean rate $\dot{q}_2^* = 0$, and let spin rate
+into the kinematic differential equations. This yields six equations with nine
+unknowns: $u_i (i=1,2,3)$ and $\dot{q}_i (i=1,\dots6)$. We \textit{choose} yaw
+rate $\dot{q}_1^* = 0$, lean rate $\dot{q}_2^* = 0$, let spin rate
 $\dot{q}_3^*$ be a free variable, and solve for the remaining six unknowns.
@@
 
-Evaluating the matrices in Equations (\ref{eq:quant_to_compute}) and
-(\ref{eq:C_0}-\ref{eq:C_2}) at these equilibrium conditions permits the
-matrices in the linearized equations of motion
-(\ref{eq:state_space_constrained}) to be computed. Solving these equations for
+Evaluating the matrices in equations (\ref{eq:state_space_constrained}) at
+these equilibrium conditions and subsequently solving these equations for
 $\delta\dot{\mathbf{q}}$ and $\delta\dot{\mathbf{u}}$ yields the linearized
@@
 independent states, the linear relationship between the independent states and
-their time derivatives are formed. The eigenvalues, parameterized by forward
-speed $v$, of this coefficient matrix may be computed symbolically. Six of the
-eight eigenvalues of this matrix are zero, and the remaining two are
+their time derivatives are formed. The eigenvalues of this coefficient matrix
+may be computed symbolically. Six of the eight eigenvalues of this matrix are
+zero, and the remaining two are
 \begin{align}
@@
 imaginary and are hence marginally stable. We omit the details of these
-calculation and direct the reader to the electronic supplementary material.
+calculations and direct the reader to the electronic supplementary material.
 
@@
 \end{align}
-For a given set of parameters the eigenvalues of the linearized dynamics can be
-parameterized by three quantities lean $q_2$, yaw rate $\dot{q}_1$, and spin
+For a given set of parameters, the eigenvalues of the linearized dynamics can be
+parameterized by three quantities: lean $q_2$, yaw rate $\dot{q}_1$, and spin
 rate $\dot{q}_3$ (though only two can be independent because of the above
@@
 
-It is notable that there is a second opportunity (after constraining the
-nonlinear equations of motion) to determine which speeds are dependent.
-Now that the equations are linear, the choice of which state variables are
-taken to be independent may determine whether $\nabla_{\bm{q}}\bm{f}_{c}
-P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_{v} P_{ud}$ are nonsingular.
-These matrices depend upon the configuration $\bm{q}$ and constant parameters;
-for certain configurations or parameters, it may be that the choice of
-independent state variables cannot be arbitrary.
-While some systems may permit a choice of independent state variables which are
-valid for all configurations of interest, others may not.
-Methods for automatically selecting the ``best'' choice of independent state
-variables are discussed in \cite{Reckdahl1996}; they involve computing the SVD
-of the Jacobian of the constraint matrices to determine a set of independent
+Kane's dynamical equations are typically formulated for one particular choice
+of dependent speeds, and computer code is generated for that single choice of
+dependent speeds. It is possible, however, to output computer code for the
+unconstrained equations of motion (equation (\ref{eq:kanes_eq})), along with
+the velocity constraint coefficient matrix (equation (\ref{eq:constraint_B})),
+such that the independent choice of speeds and the constrained equations of
+motion (equation (\ref{eq:kanes_eq_nonholonomic})) are formed at the time of
+numerical evaluation (as opposed to choosing a particular set of dependent
+speeds during symbolic derivation). There are several reasons why this can be
+desirable: 1) the configuration of the system in question changes significantly
+enough during simulation that no single choice of independent speeds will work
+for all regions of the configuration space; 2) the choice of parameters greatly
+affects which states should be selected to be dependent; and 3) to choose the
+dependent states which minimize the effect of numerical round off. This
+approach can also be applied to computation of the linearized dynamics. A task
+where the ability to switch the choice of dependent states ``on the fly'' is
+extremely helpful is in the computation of Lyapunov characteristic exponents
+(LCE's). When computing LCE's, time simulation of the nonlinear dynamics
+\textit{and} linearization at each point along the trajectory a
+required~\cite{Benettin1980a,Benettin1980b,Udwadia2001};
+having a computer code that makes it easy to switch choice of dependent states
+at the time of numerical evaluation is of great benefit in this case. The
+matrices which must be used to determine the best choice of coordinates and
+speeds are $\nabla_{\bm{q}} \bm{f}_c P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_v P_{ud}$;
+these matrices appear in equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}).
+If for some choice of dependent coordinates or dependent speeds either of these
+matrices are singular or nearly singular, a different set of dependent
+coordinates and dependent speeds should be chosen. These matrices generally
+depend only upon parameters and coordinates (not speeds). While some systems
+may permit a choice of independent state variables which are valid for all
+configurations of interest, others may not. Methods for automatically selecting
+the ``best'' choice of independent state variables are discussed in
+\cite{Reckdahl1996}; they involve computing the singular value decomposition
+$\nabla_q \bm{f}_c$ and $\nabla_u \bm{f}_v$  to determine a set of independent
 states which will ensure the non-singularity of the aforementioned matrices.
@@
 symbolically or numerically, and handles configuration, velocity, and
-acceleration level constraints.
-
+acceleration level constraints. The coefficient matrices in equation
+(\ref{eq:state_space_constrained}) can generally be computed symbolically and
+output as efficient C/C++/Fortran routines which may be compiled into highly
+efficient machine code. This permits library routines which are both highly
+efficient (no finite differences) and very general (arbitrary system parameters
+and linearization point).
 %TODO: mention if it is compatible with other procedures.
@@
 The procedure has been implemented symbolically in the
-\texttt{physics.mechanics} sub-module of the open source symbolic
-manipulator SymPy\cite{SymPy2012}. In addition to applying it on the rolling
-disk example, we have also applied it to the Whipple bicycle model and found
-that it matches eigenvalues published for a benchmark set of
-parameters~\cite{Meijaard2007}.
+\texttt{physics.mechanics} sub-module of the open source symbolic manipulator
+SymPy\cite{SymPy2012}. In addition to the rolling disk example, we have also
+applied it to the Whipple bicycle model (and extensions of this model) and
+found that it matches (to at least 14 digits) the eigenvalues published for a
+benchmark set of parameters~\cite{Meijaard2007}. This latter example provides a
+challenging and rigorous test of the procedure which strongly indicates both
+its utility and its correctness.
 

### 2013-08-06 rolling_disk.py
Finalize rolling disk example for electronic supplementary material.

@@ -1,9 +1,6 @@
-from sympy import (symbols, Matrix, eye, zeros, pi, trigsimp,
-        solve_linear_system_LU, solve)
-from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point, dot,
-cross, mprint, RigidBody, inertia, KanesMethod, mlatex)
-from sympy import *
-from sympy.physics.mechanics import *
+from sympy import symbols, Matrix, eye, zeros, solve, cos, sqrt, Poly
+from sympy.physics.mechanics import (dynamicsymbols, ReferenceFrame, Point,
+        dot, cross, mprint, inertia)
 
 # Symbols for time and constant parameters
 t, r, m, g, v = symbols('t r m g v')
@@ -55,6 +52,7 @@ CO = NO.locatenew('CO', q4*N.x + q5*N.y + q6*N.z)      # Disc center
 # Disc center velocity in N expressed using time derivatives of coordinates
 v_co_n_qd = CO.pos_from(NO).dt(N)
 
+# Disc center velocity in N expressed using generalized speeds
 CO.set_vel(N, u4*C.x + u5*C.y + u6*C.z)
 
 P = CO.locatenew('P', r*B.z)                           # Disc-ground contact
@@ -71,20 +69,24 @@ f_v_du = f_v.jacobian(u)
 
 # Kinematic differential equations
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in B] +
-                  [dot(v_co_n_qd - cross(C.ang_vel_in(N), CO.pos_from(P)), uv) for uv in N])
+                  [dot(v_co_n_qd - cross(C.ang_vel_in(N), CO.pos_from(P)),
+                       uv) for uv in N])
 qdots = solve(kindiffs, qd)
 
 B.set_ang_vel(N, w_b_n_qd.subs(qdots))
-C.set_ang_acc(N, C.ang_vel_in(N).dt(B) + cross(B.ang_vel_in(N), C.ang_vel_in(N)))
+C.set_ang_acc(N, C.ang_vel_in(N).dt(B) + cross(B.ang_vel_in(N),
[…]

### 2013-08-06 real_roots_steady.py
Update scripts to generate output.

@@ -24,6 +24,5 @@ plt.text(-np.pi/8, np.pi/2, 'Equilibrium feasible')#, fontdict=font)
 plt.text(-np.pi/8, -np.pi/2, 'Equilibrium feasible')#, fontdict=font)
 plt.text(-3*np.pi/8, 0, 'Equilibrium infeasible')#, fontdict=font)
 plt.text(np.pi/8, 0, 'Equilibrium infeasible')#, fontdict=font)
-#plt.title("Feasibility regions for rolling disk, $m = r = g = 1$")
-plt.show()
-
+plt.title("Feasibility regions for rolling disk, $m = r = g = 1$")
+plt.savefig("rolling_disk_feasibility_regions.pdf")

### 2013-08-06 rolling_disk.py

@@ -97,6 +97,8 @@ partial_v_CO = [CO.vel(N).diff(ui, N) for ui in u]
 F_CO = m*g*A.z
 # Generalized active forces (unconstrained)
 gaf = [dot(F_CO, pv) for pv in partial_v_CO]
+print("Generalized active forces (unconstrained)")
+mprint(gaf)
 
 # Inertia force
 R_star_CO = -m*CO.acc(N)
@@ -113,6 +115,7 @@ T_star_C = -(dot(I_C_CO, C.ang_acc_in(N))
 gif = [dot(R_star_CO, pv) + dot(T_star_C, pav) for pv, pav in
         zip(partial_v_CO, partial_w_C)]
 print("Generalized inertia forces (unconstrained)")
+mprint(gif)
 
 # Constrained dynamic equations
 # Coordinates to be independent: q1, q2, q3, q4, q5
@@ -147,6 +150,12 @@ gaf_dep = Matrix([gaf[i] for i in dep_indices])
 
 gif_con = gif_indep + Bd_inv_Bi.T * gif_dep
 gaf_con = gaf_indep + Bd_inv_Bi.T * gaf_dep
+
+
+print("Generalized inertia forces (constrained)")
+mprint(gaf_con)
+print("Generalized inertia forces (constrained)")
+mprint(gif_con)
 
 # Build the part of f_2 and f_3 that come from Kane's equations, the first three
 # rows of each
@@ -277,6 +286,8 @@ A_ss_reduced.simplify()
 
 evals = A_ss_reduced.eigenvals()
 upright_check = {q3d: 1/sqrt(3)}
+print("Turning eigenvalues, symbolic:")
+print(evals)
 print("Upright steady eigenvalues at d/dt (q_3) = 1/sqrt(3):")
 print([i.subs(upright_check).n() for i in evals.keys()])
 
@@ -296,6 +307,8 @@ A_ss_reduced.simplify()
 
 evals = A_ss_reduced.eigenvals()
 leaned_check = {q1d: 0, q3d: 1/sqrt(3), q2: 0}
+print("Turning eigenvalues, symbolic:")
+print(evals)
 print("Turning eigenvalues, evaluated for upright steady motion at " +
       "d/dt (q_3) = 1/sqrt(3), should match previous.")
 print([i.subs(leaned_check).n() for i in evals.keys()])

### 2014-01-18 PetersonGedeHubbard2012.tex
Fix minor error in naive linearization eigenvalues.

@@
 independent and dependent states (assuming the dependents states have not been
-algebraically eliminated). The dynamical equations and the kinematical
-differential equations can be solved for the time derivatives of the
-generalized coordinates and generalized speeds, respectively. Taken together,
-it is tempting to think that this system of first order differential equations
-could be linearized in the same fashion as an unconstrained system.
-Unfortunately, this is not the case. Linearizing these equations in the same
-manner as one would for an unconstrained system (by directly computing the
-Jacobian) is incorrect and will yield incorrect linearized equations (and
-subsequently incorrect eigenvalues, in the case of a stability analysis). We
-illustrate this by means of a familiar example, the rolling disk, formulated
-with a non-traditional set of coordinates and generalized speeds which are not
-all independent. We provide a systematic method to correctly generate the
-linear equations of motion for constrained systems which have been found using
-Kane's method (or any other method which results in equations with the same
-form). We apply this method to the rolling disk example to illustrate that
-it yields eigenvalues that match previously published results.
+algebraically eliminated). The kinematical and dynamical differential equations
+can be solved for the time derivatives of the generalized coordinates and
+generalized speeds, respectively. Taken together, it is tempting to think that
+this system of first order differential equations could be linearized in the
+same fashion as an unconstrained system.  Unfortunately, this is not the case.
+Linearizing these equations in the same manner as one would for an
+unconstrained system (by directly computing the Jacobian) is incorrect and will
+yield incorrect linearized equations (and subsequently incorrect eigenvalues,
+in the case of a stability analysis). We illustrate this by means of a familiar
+example, the rolling disk, formulated with a non-traditional set of coordinates
+and generalized speeds which are not all independent. We provide a systematic
+method to correctly generate the linear equations of motion for constrained
+systems which have been found using Kane's method (or any other method which
+results in equations with the same form). We apply this method to the rolling
+disk example to illustrate that it yields eigenvalues that match previously
+published results.
 
@@
 Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
-generating equations of motion (and linear forms thereof), however both
-techniques are numerical in nature -- they do not provide symbolic equations of
-motion or symbolic linearized equations. While Neuman presents a symbolic (as
-opposed to numeric) formulation for linearization of the Q-matrix formulation
-of the Lagrangian\cite{Neuman1984}, his work makes no mention of constrained
-systems and attempts to contact the author regarding the software (algebraic
-robot modeler -- Arm) as well as internet searches for the software have been
-fruitless.
+generating equations of motion (and linear forms thereof) using Lagrange's
+method; the work of Schwab and Meijaard can be applied symbolically while
+Minaker and Reiveley's approach is numeric in nature. Neither discuss the
+choice of generalized speeds as in Kane's method: both form the equations of
+motion in terms of second time derivatives of the coordinates.  Further,
+neither of these works address constraints which cannot be eliminated (i.e.,
+nonlinear configuration constraints), nor do they address what approach should
+be taken when a single choice of dependent states may not suffice. While
+Neuman presents a symbolic (as opposed to numeric) formulation for
+linearization of the Q-matrix formulation of the Lagrangian\cite{Neuman1984},
+his work makes no mention of constrained systems and attempts to contact the
+author regarding the software (algebraic robot modeler -- Arm) as well as
+internet searches for the software have been fruitless.
+
+% What does it mean for the equations to be symbolic/analytic in nature?
 
@@
 and take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
-\mathbf{u}]^T$. However, this will give the incorrect results. As discussed in
+\mathbf{u}]^T$. However, this will give incorrect results. As discussed in
 \cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity, and the system
@@
   \label{eq:evals_incorrect}
-  \lambda_{1,2}=\pm\sqrt{-v^2},\quad \lambda_{3,4} = \pm\sqrt{-\frac{12}{5} v^2 + \frac{4}{5}}
+  \lambda_{1,2}=\pm\frac{\sqrt{6}}{3}\sqrt{-v^2},
+  \quad \lambda_{3,4} = \pm\sqrt{\frac{4}{5} - \frac{12}{5} v^2}
 \end{align}

### 2014-01-19 PetersonGedeHubbard2012.tex
Updated drawing, minor changes in text

@@
 % Context
-Many common dynamic mechanical systems have configuration or velocity constraints.
+Many common dynamic mechanical systems are subject to configuration or velocity
+constraints.
 % Need  (what we have and what we want)
@@
 differential equations are required to describe the system, and independent and
-dependent states remain in the final equations (i.e. no DAE's, no substitution
-of variables).
+dependent states remain in the final equations (i.e. no DAE's, dependent
+variables are not substituted away).
 % Object of the document
@@
 generates, we present the procedure to symbolically linearize the nonlinear
-motion equations of constrained multibody systems, and illustrate it with the
-familiar example of the rolling disk.
+motion equations of constrained multibody systems, and illustrate it with an
+example of the rolling disk.
 
@@
 
-For a system system with constraints, Kane's dynamical equations\cite{Kane1985}
-are first order ODE's (in the generalized speeds) which include both the
-independent and dependent states (assuming the dependents states have not been
+For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
+first order ODE's (in the generalized speeds) which include both the
+independent and dependent states (assuming the dependent states have not been
 algebraically eliminated). The kinematical and dynamical differential equations
@@
 
-We consider three types of constraints: configuration, velocity, and
-acceleration. Configuration constraints limit the location or orientation of
-parts of the system, relative to the external world or other parts of the
-system. Velocity constraints limit the speeds at which the configuration can
-change, either from configuration constraints (through time differentiation) or
-independent application. Velocity constraints most often appear in systems
-where there is rolling without slip or there are closed kinematic loops. In the
-context of this paper, acceleration constraints arise either from time
-differentiated velocity constraints or similar kinematic considerations.
+We consider three levels of constraints: configuration, velocity, and
+acceleration. Configuration (holonomic) constraints limit the location or
+orientation of parts of the system, relative to the external world or other
+parts of the system. Velocity (nonholonomic of time differentiated holonomic)
+constraints limit the speeds at which the configuration can change.
+Velocity constraints most often appear in systems where there is rolling
+without slip or there are closed kinematic loops. For this procedure,
+acceleration level constraints are assumed to be time differentiated velocity
+constraints.
 
@@
 % TODO:reiterate limitations of model
-The goal of this paper is to definitively establish how the first time
-derivative of all coordinates and speeds (independent \textit{and} dependent)
-depend, to first order, upon a selection of independent coordinates and
-independent speeds, for an arbitrary point of linearization. We demonstrate how
-a reduced portion of these linearized equations can be used to perform standard
-stability analysis (i.e., for control system design or linear stability
-analysis). The formalism presented here is generic enough to cover most
+This paper presents linear, first order differential equations relating the
+time derivatives of \textit{all} speeds and coordinates to the independent
+speeds and coordinates, for any point of linearization.
+A provided example demonstrates the necessity of these equations.
+We demonstrate how a reduced portion of these linearized equations can be used
+to perform standard stability analysis (i.e., for control system design or
+linear stability analysis).
+The formalism presented here is generic enough to cover most
 examples of time-varying constrained multibody systems with arbitrary external

### 2014-01-19 PetersonGedeHubbard2012.tex
Address reviewer #3 and #1 remarks.
Reviewer #3 remarks 1-3, 5, 7, 8, 9 addressed.
Reviewer #1 remark 5, 7, 8, 11 addressed, part of 12.

Fix bug in rolling_disk.py which swapped a and c terms of the steady
turning equilibrium condition.

@@
 motion in terms of second time derivatives of the coordinates.  Further,
-neither of these works address constraints which cannot be eliminated (i.e.,
+neither of these works address constraints which cannot be eliminated (e.g.,
 nonlinear configuration constraints), nor do they address what approach should
@@
 time derivatives of \textit{all} speeds and coordinates to the independent
-speeds and coordinates, for any point of linearization.
-A provided example demonstrates the necessity of these equations.
-We demonstrate how a reduced portion of these linearized equations can be used
-to perform standard stability analysis (i.e., for control system design or
-linear stability analysis).
-The formalism presented here is generic enough to cover most
-examples of time-varying constrained multibody systems with arbitrary external
-inputs and arbitrary specified quantities.
+speeds and coordinates, for any point of linearization.  A provided example
+demonstrates the necessity of these equations.  We demonstrate how a reduced
+portion of these linearized equations can be used to perform standard stability
+analysis (i.e., for control system design or linear stability analysis).  The
+formalism presented here is generic enough to cover most examples of
+time-varying constrained multibody systems with arbitrary external inputs and
+arbitrary specified quantities.
 % TODO: electronic supplementary material with examples for other methods
@@
 \label{eq:ex1}
-{^N}\bar{\omega}^B = \it{\omega}_1 \hat{b}_1 + \it{\omega}_2 \hat{b}_2 +
-\it{\omega}_3 \hat{b}_3
+{^N}\bar{\omega}^B = u_1 \hat{b}_1 + u_2 \hat{b}_2 + u_3 \hat{b}_3
 \end{align}
@@
 \[
-{^N}\bar{\omega}^B_{\omega_1} = \hat{b}_1 \quad \quad
-{^N}\bar{\omega}^B_{\omega_2} = \hat{b}_2 \quad \quad
-{^N}\bar{\omega}^B_{\omega_3} = \hat{b}_3
+{^N}\bar{\omega}^B_{1} = \hat{b}_1 \quad \quad
+{^N}\bar{\omega}^B_{2} = \hat{b}_2 \quad \quad
+{^N}\bar{\omega}^B_{3} = \hat{b}_3
 \]
@@
 inertial reference frame $N$. This system has generalized coordinates
-$q_1,...,q_n$ and generalized speeds $u_1,...,u_o$. Additionally, there are $l$
-configuration constraints and $m$ velocity constraints.
+$q_1,...,q_n$ and generalized speeds $u_1,...,u_o$.  While it common that $n =
+o$, this is not a requirement when using Kane's method. Additionally, there are
+$l$ configuration constraints and $m$ velocity constraints.
 
@@
 \label{eq:particle_translational}
-{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{u_j} u_j + {^N}\bar{v}^{P_i}_t
+{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{j} u_j + {^N}\bar{v}^{P_i}_t
 \quad \quad i=1,...,g
@@
 \label{eq:rb_translational}
-{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{u_j} u_j +
+{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{j} u_j +
 {^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
@@
 \label{eq:points_translational}
-{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{u_j} u_j + {^N}\bar{v}^{Q_i}_t
+{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{j} u_j + {^N}\bar{v}^{Q_i}_t
 \quad \quad i=1,...,k
@@
 \label{eq:rb_rotational_bodies}
-{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
+{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{j} u_j +
 {^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
@@
 \label{eq:rb_rotational_frames}
-{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
+{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{j} u_j +
 {^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
@@
     \hline
-    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration constraints \\
-    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity constraints \\
+    $\bm{f}_{c}(\bm{q}, t)$ & $\bm{0}_{l \times 1}$ & Configuration level constraints \\
+    $\bm{f}_{v}(\bm{q}, \bm{u}, t)$ & $\bm{0}_{m \times 1}$ & Velocity level constraints \\
     $\bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t)$ & $\bm{0}_{m
-    \times 1}$ & Acceleration constraints \\
+    \times 1}$ & Acceleration level constraints \\
     $\bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t)$ &
@@
 
-The only new terms are $\mathbf{f}_0$, $\mathbf{f}_1$, $\mathbf{f}_2$, and
-$\mathbf{f}_3$.  The kinematic differential equations are split up into the
-quantities $\mathbf{f}_0$ and $\mathbf{f}_1$, depending on whether each term
-has a $\mathbf{u}$ or $\dot{\mathbf{q}}$ present.  The same is true for
-$\mathbf{f}_2$ and $\mathbf{f}_3$; if a term contains a $\dot{\mathbf{u}}$ it
-is placed into $\mathbf{f}_2$ and if not it is placed into $\mathbf{f}_3$. The
-motivation for this organizational scheme will be made apparent in Section
-\ref{sec:derivations}.
+Not all of the $n$ generalized coordinates are independent due to the $l$
+configuration level constraints (e.g., holonomic constraints resulting from
+closed kinematic loops). Similarly, not all of the $o$ generalized speeds are
+independent due to the $m$ velocity level constraints.  Velocity level
+constraints result from time differentiated configuration constraints or
+nonholonomic constraints (e.g., no-slip rolling).  The $m$ acceleration level
+constraints are assumed to be obtained by time differentiating the $m$ velocity
+level constraints.  Just as Kane's method does not require $n = o$,
+neither does our linearization procedure. The $s$ exogenous inputs include
+control system inputs and/or external disturbance forces or torques applied to
+the system.
+
+The kinematic differential equations are organized into the quantities
+$\mathbf{f}_0$ and $\mathbf{f}_1$; terms depending upon $\dot{\mathbf{q}}$ are
+placed into $\mathbf{f}_0$, all other terms are placed into $\mathbf{f}_1$.
+Similarly, the dynamic differential equations are organized into the
+quantities $\mathbf{f}_2$ and $\mathbf{f}_3$; terms depending upon
+$\dot{\mathbf{u}}$ are placed into $\mathbf{f}_2$, all other terms are placed
+into $\mathbf{f}_3$. The motivation for this organizational scheme will be made
+apparent in Section \ref{sec:derivations}.
 
@@
 
-Traditionally, to locate the center of the disk $C^*$
-relative to the inertial origin $N^*$, two coordinates are used to locate the
-contact point in the ground plane; these two coordinates and the lean angle
-and yaw angles determine the location of the center of the disk. The advantage of this
-approach is that the choice of coordinates guarantees that the disk remain in
-contact with the ground plane.
+Traditionally, to locate the center of the disk $C^*$ relative to the inertial
+origin $N^*$, two coordinates are used to locate the contact point in the
+ground plane; these two coordinates and the lean angle and yaw angles determine
+the location of the center of the disk. The advantage of this approach is that
+by using a minimal set of coordinates, the holonomic ground contact constraint
+is always satisfied and hence no holonomic constraint equation is needed.
 
@@
     +
-\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\r u_{1}
-    \sin\left(q_{1}\right) \cos\left(q_{2}\right) +
-    r u_{2} \cos\left(q_{1}\right)\\- r u_{1}
-    \cos\left(q_{1}\right) \cos\left(q_{2}\right) +
-    r u_{2} \sin\left(q_{1}\right)\\- r u_{1}
-    \sin\left(q_{2}\right)\end{matrix}\right]}_{\mathbf{f}_1}
+\underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\
+%
+%[ sin(q1)*sin(q2)*sin(q3) - cos(q1)*cos(q3),  sin(q1)*cos(q2), -sin(q1)*sin(q2)*cos(q3) - sin(q3)*cos(q1)],
+%[-sin(q1)*cos(q3) - sin(q2)*sin(q3)*cos(q1), -cos(q1)*cos(q2), -sin(q1)*sin(q3) + sin(q2)*cos(q1)*cos(q3)],
+%[                           sin(q3)*cos(q2),         -sin(q2),                           -cos(q2)*cos(q3)]]
+\left(\sin(q_1)\sin(q_2)\sin(q_3) - \cos(q_1)\cos(q_3)\right)u_4 +
+\left(\sin(q_1)\cos(q_2)\right)u_5 + \left(-\sin(q_1)\sin(q_2)\cos(q_3) -
+\sin(q_3)\cos(q_1)\right)u_6 \\
+%
+\left(-\sin(q_1)\cos(q_3) - \sin(q_2)\sin(q_3)\cos(q_1)\right)u_4 +
+\left(-\cos(q_1)\cos(q_2)\right)u_5 +
+\left(-\sin(q_1)\sin(q_3) + \sin(q_2)\cos(q_1)\cos(q_3)\right)u_6
+\\
+%
+\sin(q_3)\cos(q_2)u_4  - \sin(q_2)u_5 - \cos(q_2)\cos(q_3)u_6
+\end{matrix}\right]}_{\mathbf{f}_1}
     = \left[\begin{matrix} 0\\ 0\\ 0\\ 0\\ 0\\ 0\end{matrix}\right]
@@
 for all twelve unkowns to be solved. Combining (\ref{rd:f_0_f_1}),
-(\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete second order differential
-equations can be written.
+(\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete set of motion equations
+can be written.
 \begin{align}
@@
   \lambda_{1,2}=\pm\frac{\sqrt{6}}{3}\sqrt{-v^2},
-  \quad \lambda_{3,4} = \pm\sqrt{\frac{4}{5} - \frac{12}{5} v^2}
+%  \quad \lambda_{3,4} = \pm\sqrt{\frac{4}{5} - \frac{12}{5} v^2}
+  \quad \lambda_{3,4} &= \pm \frac{2\sqrt{1 - 3v^2}}{\sqrt{5}}
 \end{align}
@@
 \begin{align}
-\dot{q}_1 &= \frac{r}{2 g \sin\left(q_{2}\right)} \left( - \frac{3}{2} \dot{q}_{3}\cos\left(q_{2}\right) \pm \sqrt{- \frac{5 g}{r} \sin^{2}\left(q_{2}\right) \cos\left(q_{2}\right) + \frac{9}{4} \dot{q}_{3}^{2} \cos^{2}\left(q_{2}\right)} \right)
+\dot{q}_1 &= \frac{2}{5\cos\left(q_{2}\right) \sin\left(q_{2}\right)} \left( - \frac{3}{2} \dot{q}_{3}\cos\left(q_{2}\right) \pm \sqrt{- \frac{5 g}{r} \sin^{2}\left(q_{2}\right) \cos\left(q_{2}\right) + \frac{9}{4} \dot{q}_{3}^{2} \cos^{2}\left(q_{2}\right)} \right)
 \end{align}
@@
 \end{align}
-but equation (\ref{eq:state_space_constrained}) doesn't fit into the standard
-linear system framework because it is non-square.
-To obtain a square system we define
-\begin{align}
+where $\mathbf{x}$ are the system states and $\mathbf{u}$ are the control
+inputs (the $\mathbf{u}$ used in control literature is analagous to the
+$\mathbf{r}$ we use in this paper).  To transform equation
+(\ref{eq:state_space_constrained}) into this standard form, we define
+\begin{align}
+\tilde{M} &=
+\left[
+\begin{array}{cc}
+  \tilde{M}_{qq} & \bm{0}_{n \times o} \\
+  \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
+  \tilde{M}_{uqd} & \tilde{M}_{uud}
+\end{array}
+\right]\\
 \tilde{A} &=

### 2014-01-19 electronic_supplementary_material/rolling_disk.py

@@ -69,8 +69,7 @@ f_v_du = f_v.jacobian(u)
 
 # Kinematic differential equations
 kindiffs = Matrix([dot(w_c_n_qd - C.ang_vel_in(N), uv) for uv in B] +
-                  [dot(v_co_n_qd - cross(C.ang_vel_in(N), CO.pos_from(P)),
-                       uv) for uv in N])
+                  [dot(v_co_n_qd - CO.vel(N), uv) for uv in N])
 qdots = solve(kindiffs, qd)
 
 B.set_ang_vel(N, w_b_n_qd.subs(qdots))
@@ -221,6 +220,7 @@ f_v_eq = f_v.subs(eq_u).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
 f_a_eq = f_a.subs(eq_ud).subs(eq_u).subs(qd_sym).subs(eq_q).subs(qd_sym_inv).expand()
 f_a_eq.simplify()
 f_0_eq_plus_f_1_eq = (f_0 + f_1).subs(eq_qd).subs(eq_u).subs(qd_sym).subs(eq_q).subs(qd_sym_inv)
+f_0_eq_plus_f_1_eq.simplify()
 f_2_eq_plus_f_3_eq = (f_2 + f_3).subs(eq_ud).subs(eq_u).subs(qd_sym).subs(eq_q).subs(qd_sym_inv).expand()
 f_2_eq_plus_f_3_eq.simplify()
 
@@ -326,8 +326,16 @@ print([i.subs(upright_check).n() for i in evals_bad.keys()])
 
 # Steady equilibrium
 steady = f_3[0].subs(eq_u).subs(qd_sym).subs({q1:0,q3:0}).subs(qd_sym_inv)/(m*r*r)
+print("steady equations of motion:")
+mprint(steady.expand().simplify())
 p = Poly(steady, q1d)
-c, b, a = p.coeffs()
+a, b, c = p.coeffs()
+print("a:")
+mprint(a)
+print("b:")
+mprint(b)
+print("c:")
+mprint(c)
 discriminant = b*b - 4*a*c
 root1 = (-b + sqrt(discriminant))/(2*a)
 root2 = (-b - sqrt(discriminant))/(2*a)

### 2014-01-19 PetersonGedeHubbard2012.tex
Shorten trigonometric functions to make equations fit.
sin(q_2) -> s_2, cos(q_2) -> c_2, tan(q_2) -> t_2, and so on.
Updated drawing, addressed r3c6

@@
 evaluating the Jacobian at the same operating conditions and parameters, eight
-of the twelve eigenvalues are identically zero while the remaining four are
+of the twelve eigenvalues are identically zero while the remaining four are:
 \begin{align}
@@
 \end{align}
-which demonstrates the incorrectness of this naive approach (since
-$v=\pm\frac{1}{\sqrt{3}}$ is not a critical point of these eigenvalues).
+
+At the critical speed, all eigenvalues must be zero, and that is clearly not
+the case here; one pair of non-zero eigenvalues exist at what should be the
+critical speed. This indicates the presence of oscillatory modes that are not
+present in the correct derivation.
+
 Further, the fact that twelve eigenvalues are obtained should be an alert that
-something is incorrect since the number of independent quantities is only eight
-(five coordinates and three speeds). We now outline a linearization procedure
-which addresses this issue; we revisit this example in Section
+something is incorrect, since the number of independent quantities is only
+eight (five coordinates and three speeds). We now outline a linearization
+procedure which addresses this issue; we revisit this example in Section
 \ref{sec:example_revisited}.

### 2014-01-20 PetersonGedeHubbard2012.tex
Add figure and fix math mode errors.
Addressed most of Mont's suggestions
About 1 page worth of corrections remains, as well as a minor change to the
figure and a few others.

@@
 and motion equations and their subsequent linearization.  This procedure was
-specifically developed for equations of motion generated by Kane's method, but
-is compatible with any method as long as the following are true: only ordinary
-differential equations are required to describe the system, and independent and
-dependent states remain in the final equations (i.e. no DAE's, dependent
-variables are not substituted away).
+developed specifically for equations of motion generated by Kane's method, but
+is compatible with any method as long as only ordinary differential equations
+are required to describe the system, and independent and dependent states
+remain in the final equations (i.e. no DAE's, dependent variables are not
+substituted away).
 % Object of the document
@@
 of equilibria are two of the most common uses of multibody system dynamics
-models. Both of these uses require linearized equations of motion. For
-unconstrained systems, obtaining the linearized equations of motion is as
-simple as arranging the equations of motion in first order form ($\dot{x} =
-f(x, t)$) and evaluating the Jacobian of the right hand side function, with
-respect to the states, evaluated at the equilibria point of interest.  However,
-when the system in questions has constraints, i.e., not all the states are
-independent, this process is not as straightforward.
+models. Both require linearized equations of motion. For unconstrained systems,
+obtaining the linearized equations of motion is as simple as arranging the
+equations of motion in first order form ($\dot{x} = f(x, t)$) and calculating
+the Jacobian of the right hand side function, with respect to the states,
+evaluated at the equilibria point of interest.  However, when the system in
+questions has constraints, i.e., not all the states are independent, this
+process is not as straightforward.
 
@@
 example, the rolling disk, formulated with a non-traditional set of coordinates
-and generalized speeds which are not all independent. We provide a systematic
-method to correctly generate the linear equations of motion for constrained
-systems which have been found using Kane's method (or any other method which
-results in equations with the same form). We apply this method to the rolling
-disk example to illustrate that it yields eigenvalues that match previously
-published results.
+and generalized speeds which are not all independent. We then provide a
+systematic method to correctly generate the linear equations of motion for
+constrained systems which have been found using Kane's method (or any other
+method that results in equations with the same form). We apply this method to
+the rolling disk example to illustrate that it yields eigenvalues that match
+previously published results.
 
@@
 orientation of parts of the system, relative to the external world or other
-parts of the system. Velocity (nonholonomic of time differentiated holonomic)
+parts of the system. Velocity (nonholonomic or time differentiated holonomic)
 constraints limit the speeds at which the configuration can change.
-Velocity constraints most often appear in systems where there is rolling
-without slip or there are closed kinematic loops. For this procedure,
-acceleration level constraints are assumed to be time differentiated velocity
-constraints.
+Velocity constraints often appear where there is rolling without slip or there
+are closed kinematic loops. For this procedure, acceleration level constraints
+are assumed to be time differentiated velocity constraints.
 
@@
 nonholonomic systems, we have found that most techniques are not directly
-applicable to equations formed using Kane's method, and while we do not doubt
-that they \textit{could} be applied correctly, the details of how to do so, as
-far as the authors are aware, have not been explicitly presented. Kang et al.
+applicable to equations formed using Kane's method. While we do not doubt that
+they \textit{could} be applied correctly, the details of how to do so, as far
+as the authors are aware, have not been explicitly presented. Kang et al.
 \cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
-linearization, but these methods have been developed for equations of motion
-found using Lagrange's method, which contain Lagrange multipliers (which are
-not present in equations of motion generated from Kane's method). Further, both
-of these papers lack a discussion of basic concepts such as how many quantities
-are independent, and in the case of eigenvalue computation, how many
-eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
-Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
-generating equations of motion (and linear forms thereof) using Lagrange's
-method; the work of Schwab and Meijaard can be applied symbolically while
-Minaker and Reiveley's approach is numeric in nature. Neither discuss the
-choice of generalized speeds as in Kane's method: both form the equations of
-motion in terms of second time derivatives of the coordinates.  Further,
-neither of these works address constraints which cannot be eliminated (e.g.,
-nonlinear configuration constraints), nor do they address what approach should
-be taken when a single choice of dependent states may not suffice. While
-Neuman presents a symbolic (as opposed to numeric) formulation for
-linearization of the Q-matrix formulation of the Lagrangian\cite{Neuman1984},
-his work makes no mention of constrained systems and attempts to contact the
-author regarding the software (algebraic robot modeler -- Arm) as well as
-internet searches for the software have been fruitless.
+linearization, but their methods have been developed for equations of motion
+found using Lagrange's method, which contain Lagrange multipliers ( not present
+in equations of motion generated using Kane's method). Further, both of these
+papers lack a discussion of basic concepts such as how many quantities are
+independent, and in the case of eigenvalue computation, how many eigenvalues
+should be present. Minaker and Rieveley \cite{Minaker2010} and Schwab and
+Meijaard \cite{Schwab2003} have developed techniques for generating equations
+of motion (and linear forms thereof) using Lagrange's method; the work of
+Schwab and Meijaard can be applied symbolically while Minaker and Reiveley's
+approach is numeric in nature. Neither discusses the choice of generalized
+speeds as in Kane's method: both form the equations of motion in terms of
+second time derivatives of the coordinates.  Further, neither of these works
+addresses constraints which cannot be eliminated (e.g., nonlinear configuration
+constraints), nor do they address what approach should be taken when a single
+choice of dependent states may not suffice.  Neuman \cite{Neuman1984} presents
+a symbolic (as opposed to numeric) formulation for linearization of the
+Q-matrix formulation of the Lagrangian, but makes no mention of constrained
+systems and attempts to contact the author regarding the software (algebraic
+robot modeler -- Arm) as well as internet searches for the software have been
+fruitless.
 
@@
 speeds and coordinates, for any point of linearization.  A provided example
-demonstrates the necessity of these equations.  We demonstrate how a reduced
-portion of these linearized equations can be used to perform standard stability
+demonstrates the need for these equations.  We demonstrate how a reduced
+set of these linearized equations can be used to perform standard stability
 analysis (i.e., for control system design or linear stability analysis).  The
@@
 
-We begin with a review of Kane's method for generating equations of motion is
-reviewed first, in order to properly orient the reader to the format and some
-of the qualities of the generated equations. Next, we apply Kane's method to
-generate the equations of motion for the rolling disk, but we purposefully
-introduce a dependent coordinate and three dependent speeds in the formulation.
-The need for a different linearization technique is demonstrated with this
-example. In response to this need, we present the newly developed linearization
-procedure and its derivation. This new linearization procedure is applied the
-same rolling disk example to illustrate its validity and an example of its
+We begin with a review of Kane's method for generating equations of motion, in
+order to properly orient the reader to the format and some of the qualities of
+the generated equations. Next, we apply Kane's method to generate the equations
+of motion for the rolling disk, but we purposefully introduce a dependent
+coordinate and three dependent speeds in the formulation.  The need for a
+different linearization technique is demonstrated with this example. In
+response to this need, we present the newly developed linearization procedure
+and its derivation. This new linearization procedure is applied to the same
+rolling disk example to illustrate its validity and an example of its
 application. Finally, we discuss other details and nuances of the procedure and
@@
 \label{sec:kane_method}
-To familiarize the reader with Kane's method, some concepts relating to its use
-will be described.  The first is the use of generalized speeds in addition to
-generalized coordinates.  The second concept is in how these generalized speeds
-can be used to project permissible motions of the system, and how this removes
-the need to consider non-contributing (internal constraint) forces.  Then, the
-mathematical steps required to use Kane's method are shown and described.
-
-When using Lagrange's method, generalized coordinates are used to describe the
-configuration of a system while the time derivatives of the generalized
-coordinates describe the velocities of a system.  Kane's method allows for the
-velocities of a system to be written in terms of generalized speeds, which are
-not required to be the time derivatives of coordinates (although such a
-definition is permitted).
-
-The benefit of choosing generalized speeds to be other than simply the
-derivatives of the corresponding generalized coordinates can be seen in the
-following example. If the orientation of a rigid body were to be described
-using using Euler parameters or quaternions as coordinates, its angular
-velocity in terms of time derivatives of those coordinates is relatively
-complicated; it also requires a velocity constraint. Using generalized speeds
-allows for the angular velocity to be expressed more simply as
+To familiarize the reader with Kane's method, several concepts relating to its
+use are discussed: the use of generalized speeds in addition to generalized
+coordinates; how these generalized speeds can be used to project permissible
+motions of the system, and how this removes the need to consider
+non-contributing (internal constraint) forces.  Then, the mathematical steps
+required to use Kane's method are illustrated.
+
+In Lagrange's method, generalized coordinates describe the configuration of a
+system while time derivatives of the generalized coordinates characterize the
+velocities of a system.  Kane's method allows velocities to be written in terms
+of generalized speeds, which are not required to be the time derivatives of
+coordinates (although such a definition is permitted).
+
+The benefit of choosing generalized speeds other than simply derivatives of
+corresponding generalized coordinates can be seen in the following example. If
+the orientation of a rigid body were described using Euler parameters or
+quaternions as coordinates, its angular velocity in terms of time derivatives
+of those coordinates is relatively complicated; it also requires a velocity
+constraint. Using generalized speeds allows for the angular velocity to be
+expressed more simply as
 \begin{align}
@@
 \end{align}
-or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
+or, the angular velocity of rigid body $B$ in the reference frame $N$ is
 defined as the sum of three generalized--speed basis--vector products.
@@
 of motion, rather than the twice time differentiated generalized coordinates
-($\ddot{\mathbf{q}}$'s).
+($\ddot{q}$'s).
 
@@
 Use and understanding of Kane's method requires the use and understanding of
-partial velocities and partial angular velocities. These are simply the
-partial derivatives of the (angular) velocity vectors with respect to the
-generalized speeds. For example, for the angular velocity in equation
-(\ref{eq:ex1}), the partial angular velocities of body $B$ in reference frame
-$N$ are
+\textit{partial velocities} and \textit{partial angular velocities}. These are
+simply the partial derivatives of the (angular) velocity vectors with respect
+to the generalized speeds. For example, for the angular velocity in equation
+(\ref{eq:ex1}), the three partial angular velocities of body $B$ in reference
+frame $N$ are
 \[
@@
 
-Taking the dot product of each partial velocity and both sides of Newton's
+Taking the dot product of each partial velocity with both sides of Newton's
 second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
@@
 the system need to be written as linear combinations of the generalized speeds.
-The translational velocity of the particles are
+The translational velocities of the particles are
 \begin{align}
@@
 \end{align}
-while the translational velocity of rigid body mass centers are
+while the translational velocities of rigid body mass centers are
 \begin{align}
@@
 \end{align}
-and finally the translational velocity of points which have applied forces are
+and finally the translational velocities of points which have applied forces are
 \begin{align}
@@
 \end{align}
-Next, the angular velocity of rigid bodies are
+Next, the angular velocities of rigid bodies are
 \begin{align}
@@
 \end{align}
-and the angular velocity of reference frames which have applied torques are
+and the angular velocities of reference frames where a torque is being applied
+are
 \begin{align}
@@
 \end{align}
-where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
-The velocity constraints are taken to have the form
+where $\mathbf{f}_c$ and $\mathbf{0}$ have shape $l \times 1$.  The velocity
+constraints have the form
 \begin{align}
@@
 \end{align}
-where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$,
+where $\mathbf{f}_v$ and $\mathbf{0}$ have shape $m \times 1$,
 and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
@@
 (we assume the dependent speeds and their time derivatives have \textit{not}
-been algebraically eliminated). In order to obtain the nonholonomic dynamical
-equations, however, a choice must be made as to which speeds will be dependent.
-The methods for doing this are outside the scope of this paper, but it is
-important that whatever selection is made avoids singularities which can occur
-in certain configurations or sets of system parameters. For a careful treatment
-of this, we encourage the reader to examine the work of Reckdahl
-\cite{Reckdahl1996}.
+been algebraically eliminated). To obtain the nonholonomic dynamical equations,
+however, a choice must be made as to which speeds will be dependent.  The
+methods for doing this are beyond the scope of this paper, but it is important
+that whatever selection is made avoids singularities which can occur in certain
+configurations or sets of system parameters. For a careful treatment , we
+encourage the reader to examine the work of Reckdahl \cite{Reckdahl1996}.
 
@@
 This process of constraining the equations of motion reduces the number of
-equations from $o$ to $o-m$. However, the (\ref{eq:kanes_eq_nonholonomic}) still
+equations from $o$ to $o-m$. However, eqns. (\ref{eq:kanes_eq_nonholonomic}) still
 contain $o$ unknowns ($\dot{\mathbf{u}}_1,...,\dot{\mathbf{u}}_o$). By
@@
 
-Not all of the $n$ generalized coordinates are independent due to the $l$
+Not all $n$ generalized coordinates are independent due to the $l$
 configuration level constraints (e.g., holonomic constraints resulting from
-closed kinematic loops). Similarly, not all of the $o$ generalized speeds are
+closed kinematic loops). Similarly, not all $o$ generalized speeds are
 independent due to the $m$ velocity level constraints.  Velocity level
@@
 constraints are assumed to be obtained by time differentiating the $m$ velocity
-level constraints.  Just as Kane's method does not require $n = o$,
-neither does our linearization procedure. The $s$ exogenous inputs include
-control system inputs and/or external disturbance forces or torques applied to
-the system.
-
-The kinematic differential equations are organized into the quantities
+level constraints.  Just as Kane's method does not require $n = o$, neither
+does our linearization procedure. The $s$ exogenous inputs include control
+system inputs and/or external disturbance forces or torques applied to the
+system.
+
+The kinematic differential equations are separated into the quantities
 $\mathbf{f}_0$ and $\mathbf{f}_1$; terms depending upon $\dot{\mathbf{q}}$ are
 placed into $\mathbf{f}_0$, all other terms are placed into $\mathbf{f}_1$.
-Similarly, the dynamic differential equations are organized into the
+Similarly, the dynamic differential equations are separated into the
 quantities $\mathbf{f}_2$ and $\mathbf{f}_3$; terms depending upon
@@
 yaw $q_1$, roll $q_2$, and spin $q_3$.  Denote by $A$ and $B$ the two frames
-associated with the first two intermediate rotations in the sequence; we refer
-to these as the instantaneous yaw and lean frames, respectively. The unit
+which persist after the first two intermediate rotations in the sequence; we
+refer to these as the instantaneous yaw and lean frames, respectively. The unit
 vector, $\hat{b}_z$ fixed in the $B$ frame, is parallel to the disk plane and
 aligned with the position vector from the disk center $C^*$ to the lowest point
-of the disk $P^*$.
+of the disk $P^*$. Figure \ref{fig:rolling_disc} shows how the generalized
+coordinates are used to configure the disc.
+
+\begin{figure}[!htb]
+    \centering
+    \includegraphics[width=0.8\textwidth]{disk_drawing_cropped.pdf}
+    \caption{Configuration of rolling disc using non-minimal coordinates.}
+    \label{fig:rolling_disc}
+\end{figure}
 
 Traditionally, to locate the center of the disk $C^*$ relative to the inertial
-origin $N^*$, two coordinates are used to locate the contact point in the
-ground plane; these two coordinates and the lean angle and yaw angles determine
-the location of the center of the disk. The advantage of this approach is that
-by using a minimal set of coordinates, the holonomic ground contact constraint
-is always satisfied and hence no holonomic constraint equation is needed.
+origin $N^*$, two coordinates position the contact point in the ground plane;
+these two coordinates and the lean angle and yaw angles determine the location
+of the center of the disk. The advantage of this approach is that by using a
+minimal set of coordinates, the holonomic ground contact constraint is always
+satisfied and hence no holonomic constraint equation is needed.
 
@@
   \label{rd:f_c}
-  r\cos{q_2} + q_6 = 0
+  rc_2 + q_6 = 0
 \end{equation}
@@
 While this constraint is easily avoided by appropriate choice of coordinates,
-it serves to illustrate considerations which must be made when analyzing
-systems in which there is no obvious way to eliminate configuration constraints.
+it serves to illustrate considerations which arise when analyzing systems in
+which there is no obvious way to eliminate configuration constraints.
 
@@
 $N$ with a non-minimal choice of speeds, defined in the body-fixed ($C$) frame.
-The angular velocity of $C$ in $N$, however, is defined in the intermediate
+The angular velocity of $C$ in $N$, however, is expressed in the intermediate
 lean frame $B$.
@@
                         \hat{b}_x - r u_1 \hat{b}_y \\
-                     &= (r u_2 \cos(q_3) + u_4) \hat{c}_x + (-r
-                        u_1 + u_5)\hat{c}_y + r u_2 \sin(q_3) +
+                     &= (r u_2 c_3 + u_4) \hat{c}_x + (-r
+                        u_1 + u_5)\hat{c}_y + r u_2 s_3 +
                         u6) \hat{c}_z
@@
 \begin{align}
-    r u_2 \cos(q_3) + u_4 &= 0\\
+    r u_2 c_3 + u_4 &= 0\\
             -r u_1 + u_5 &= 0\\
-    r u_2 \sin(q_3) + u_6 &= 0
+    r u_2 s_3 + u_6 &= 0
 \end{align}
@@
 \begin{align}
-    -r u_{2} \sin(q_{3}) \dot{q}_{3} + r \cos(q_{3}) \dot{u}_{2} + \dot{u}_{4} &=
+    -r u_{2} s_3 \dot{q}_{3} + r c_3 \dot{u}_{2} + \dot{u}_{4} &=
     0\\
     - r \dot{u}_{1} + \dot{u}_{5} &= 0\\
-    r u_{2} \cos(q_{3}) \dot{q}_{3} + r \sin(q_{3}) \dot{u}_{2} + \dot{u}_{6} &=
+    r u_{2} c_3 \dot{q}_{3} + r s_3 \dot{u}_{2} + \dot{u}_{6} &=
     0
@@
 \end{equation}
-while time differentiating $r^{C^*}$ in $N$ yields
+while time differentiating $\bar{r}^{C^*}$ in $N$ yields
 \begin{equation}
@@
     \label{rd:f_0_f_1}
-\underbrace{\left[\begin{matrix}\dot{q}_{2}\\\sin\left(q_{2}\right)
-    \dot{q}_{1} + \dot{q}_{3}\\\cos\left(q_{2}\right)
+\underbrace{\left[\begin{matrix}\dot{q}_{2}\\s_2
+    \dot{q}_{1} + \dot{q}_{3}\\ c_2
     \dot{q}_{1}\\\dot{q}_{4}\\\dot{q}_{5}\\\dot{q}_{6}\end{matrix}\right]}_{\mathbf{f}_0}
@@
 \underbrace{\left[\begin{matrix}- u_{1}\\- u_{2}\\- u_{3}\\
+\left(s_1 s_2 s_3 - c_1c_3\right)u_4 +
+s_1 c_2 u_5 - \left(s_1 s_2c_3 +
+s_3c_1\right)u_6 \\
 %
-%[ sin(q1)*sin(q2)*sin(q3) - cos(q1)*cos(q3),  sin(q1)*cos(q2), -sin(q1)*sin(q2)*cos(q3) - sin(q3)*cos(q1)],
-%[-sin(q1)*cos(q3) - sin(q2)*sin(q3)*cos(q1), -cos(q1)*cos(q2), -sin(q1)*sin(q3) + sin(q2)*cos(q1)*cos(q3)],
-%[                           sin(q3)*cos(q2),         -sin(q2),                           -cos(q2)*cos(q3)]]
-\left(\sin(q_1)\sin(q_2)\sin(q_3) - \cos(q_1)\cos(q_3)\right)u_4 +
-\left(\sin(q_1)\cos(q_2)\right)u_5 + \left(-\sin(q_1)\sin(q_2)\cos(q_3) -
-\sin(q_3)\cos(q_1)\right)u_6 \\
-%
-\left(-\sin(q_1)\cos(q_3) - \sin(q_2)\sin(q_3)\cos(q_1)\right)u_4 +
-\left(-\cos(q_1)\cos(q_2)\right)u_5 +
-\left(-\sin(q_1)\sin(q_3) + \sin(q_2)\cos(q_1)\cos(q_3)\right)u_6
+\left(-s_1c_3 - s_2 s_3c_1\right)u_4 -
+c_1c_2 u_5 +
+\left(-s_1 s_3 + s_2c_1c_3\right)u_6
 \\
 %
-\sin(q_3)\cos(q_2)u_4  - \sin(q_2)u_5 - \cos(q_2)\cos(q_3)u_6
+s_3c_2u_4  - s_2u_5 - c_2c_3u_6
 \end{matrix}\right]}_{\mathbf{f}_1}
@@
 \end{align}
-The translation acceleration of $C^*$ and the angular acceleration of $C$,
+The translational acceleration of $C^*$ and the angular acceleration of $C$,
 relative to $N$, are
 \begin{align}
-    {^N}\bar{a}^{C^*} =& (-(u_1 \sin(q_3) + u_3 \cos(q_3)) u_5 + u_2 u_6 +
+    {^N}\bar{a}^{C^*} =& (-(u_1 s_3 + u_3 c_3) u_5 + u_2 u_6 +
                          \dot{u}_4) \hat{c}_x \notag \\
-                      & + ((u_1 \sin(q_3) + u_3 \cos(q_3)) u_4 - (u_1
-                         \cos(q_3) - u_3 \sin(q_3)) u_6 + \dot{u}_5) \hat{c}_y
+                      & + ((u_1 s_3 + u_3 c_3) u_4 - (u_1
+                         c_3 - u_3 s_3) u_6 + \dot{u}_5) \hat{c}_y
                          \notag\\
-                      & + ((u_1 \cos(q_3) - u_3 \sin(q_3)) u_5 - u_2 u_4 +
+                      & + ((u_1 c_3 - u_3 s_3) u_5 - u_2 u_4 +
                          \dot{u}_6) \hat{c}_z \\
-    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 \tan(q_2) + \dot{u}_1) \hat{b}_x
-                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 \tan(q2) +
+    {^N}\bar{\alpha}^{C} =& (-u_2 u_3 + u_3^2 t_2 + \dot{u}_1) \hat{b}_x
+                         + \dot{u}_2 \hat{b}_y + (u_1 u_2 - u_1 u_3 t_2 +
                          \dot{u}_3) \hat{b}_z
@@
 \begin{align}
-    \bar{R}^*_{C^*} =& - m (- (u_{1} \sin(q_{3}) + u_{3}
-                       \cos(q_{3})) u_{5} + u_{2} u_{6} +
+    \bar{R}^*_{C^*} =& - m (- (u_{1} s_3 + u_{3}
+                       c_3) u_{5} + u_{2} u_{6} +
                        \dot{u}_{4}) \hat{c}_x \notag \\
-                     & - m ((u_{1} \sin(q_{3}) + u_{3}
-                       \cos(q_{3})) u_{4} - (u_{1}
-                       \cos(q_{3}) - u_{3}
-                       \sin(q_{3})) u_{6} + \dot{u}_{5})
+                     & - m ((u_{1} s_3 + u_{3}
+                       c_3) u_{4} - (u_{1}
+                       c_3 - u_{3}
+                       s_3) u_{6} + \dot{u}_{5})
                        \hat{c}_y \notag \\
-                     & - m ((u_{1} \cos(q_{3}) - u_{3}
-                       \sin(q_{3})) u_{5} - u_{2} u_{4} +
+                     & - m ((u_{1} c_3 - u_{3}
+                       s_3) u_{5} - u_{2} u_{4} +
                    \dot{u}_{6}) \hat{c}_z \\
-    \bar{T}^*_C =& \frac{m r^{2}}{4} (2 u_{1} u_{2} \sin(q_{3})
-                   - u_{1} u_{3} \sin(q_{3}) \tan(q_{2})
-                   + 2 u_{2} u_{3} \cos(q_{3}) \notag\\&\quad\quad- u_{3}^2
-                   \cos(q_{3}) \tan(q_{2}) +
-                   \sin(q_{3}) \dot{u}_{3} - \cos(q_{3})
+    \bar{T}^*_C =& \frac{m r^{2}}{4} (2 u_{1} u_{2} s_3
+                   - u_{1} u_{3} s_3 t_2
+                   + 2 u_{2} u_{3} c_3 \notag\\&\quad\quad- u_{3}^2
+                   c_3 t_2 +
+                   s_3 \dot{u}_{3} - c_3
                    \dot{u}_{1}) \hat{c}_x \notag \\
@@
                  & + \frac{m r^{2}}{4} (- 2 u_{1} u_{2}
-                   \cos(q_{3}) + u_{1} u_{3} \cos(q_{3})
-                   \tan(q_{2}) + 2 u_{2} u_{3}
-                   \sin(q_{3}) \notag\\&\quad\quad - u_{3}^2 \sin(q_{3})
-                   \tan(q_{2}) - \sin(q_{3}) \dot{u}_{1}
-                   - \cos(q_{3}) \dot{u}_{3}) \hat{c}_z
+                   c_3 + u_{1} u_{3} c_3
+                   t_2 + 2 u_{2} u_{3}
+                   s_3 \notag\\&\quad\quad - u_{3}^2 s_3
+                   t_2 - s_3 \dot{u}_{1}
+                   - c_3 \dot{u}_{3}) \hat{c}_z
 \end{align}
@@
                 \dot{u}_{5}\right)\\\frac{m r}{2} \left(- r \dot{u}_{2} + 2
-                \sin\left(q_{3}\right) \dot{u}_{6} + 2 \cos\left(q_{3}\right)
-                \dot{u}_{4}\right)\\- \frac{m r^{2}}{4}
+                s_3\right) \dot{u}_{6} + 2 c_3
+                \dot{u}_{4}\\- \frac{m r^{2}}{4}
                 \dot{u}_{3}\end{matrix}\right]}_{\mathbf{f}_2} + \notag\\
     &\underbrace{\left[\begin{matrix}
-                m r \left( g \sin(q_2) - u_1 u_4 \sin(q_3) + u_1 u_6 \cos(q_3)
-                    + (\frac{r u_2}{2} - \frac{r u_3}{4} \tan(q_2) - u_4
-                \cos(q_3) - u_6 \sin(q_3)) u_3 \right)
+                m r \left( g s_2 - u_1 u_4 s_3 + u_1 u_6 c_3
+                    + (\frac{r u_2}{2} - \frac{r u_3}{4} t_2 - u_4
+                c_3 - u_6 s_3) u_3 \right)
                 \\m r \left(- u_{2} u_{4}
-                \sin\left(q_{3}\right) + u_{2} u_{6}
-                \cos\left(q_{3}\right) - u_{3} u_{5}\right)\\\frac{m r^{2}}{4}
-                \left(- 2 u_{2} + u_{3} \tan\left(q_{2}\right)\right)
+                s_3 + u_{2} u_{6}
+                c_3 - u_{3} u_{5}\right)\\\frac{m r^{2}}{4}
+                \left(- 2 u_{2} + u_{3} t_2\right)
                 u_{1}\end{matrix}\right]}_{\mathbf{f}_3} 
@@
 In response to the need we have demonstrated, we present a linearization
-procedure that properly accounts for system constraints. Taking the Jacobian
+procedure that properly accounts for system constraints. Computing the Jacobian
 of the right hand side of the ODE's as we did at the end of the previous
@@
 
-We first consider the case of upright steady cruise and begin by choosing the
+We first consider the case of upright steady rolling and begin by choosing the
 independent coordinates to be zero ($q_i^* = 0$, $i = 1,\dots,5$), which
@@
 independent states, the linear relationship between the independent states and
-their time derivatives are formed. The eigenvalues of this coefficient matrix
+their time derivatives are isolated. The eigenvalues of this coefficient matrix
 may be computed symbolically. Six of the eight eigenvalues of this matrix are
@@
 \begin{align}
-  \frac{g}{r}\sin{(q_2)}  + \frac{3}{2}\cos{(q_2)} \dot{q}_3 \dot{q}_1 +
-  \frac{5}{4}\cos{(q_2)}\sin{(q_2)} \dot{q}_1^2
+  \frac{g}{r} s_2  + \frac{3}{2}c_2 \dot{q}_3 \dot{q}_1 +
+  \frac{5}{4}c_2 s_2 \dot{q}_1^2
  &= 0
@@
 \begin{align}
-\dot{q}_1 &= \frac{2}{5\cos\left(q_{2}\right) \sin\left(q_{2}\right)} \left( - \frac{3}{2} \dot{q}_{3}\cos\left(q_{2}\right) \pm \sqrt{- \frac{5 g}{r} \sin^{2}\left(q_{2}\right) \cos\left(q_{2}\right) + \frac{9}{4} \dot{q}_{3}^{2} \cos^{2}\left(q_{2}\right)} \right)
+\dot{q}_1 &= \frac{2}{5 c_2 s_2} \left( - \frac{3}{2} \dot{q}_{3} c_2 \pm
+\sqrt{\frac{9}{4} \dot{q}_{3}^{2} c^2_2 - \frac{5 g}{r} s^2_2 c_2} \right)
 \end{align}
@@
 \begin{align}
-  \dot{q}_3^2 &\geq \frac{20g}{9r}\sin{(q_2)} \tan{(q_2)}
+  \dot{q}_3^2 &\geq \frac{20g}{9r} s_2 t_2
 \end{align}
@@
 \begin{align}
-\lambda_{1,2} = \pm\sqrt{\frac{4}{5} \cos\left(q_{2}\right) - \dot{q}_{1}^{2} -\frac{14}{5} \sin\left(q_{2}\right) \dot{q}_{1} \dot{q}_{3} - \frac{12}{5}\dot{q}_{3}^{2}}
+\lambda_{1,2} = \pm\sqrt{\frac{4}{5} c_2 - \dot{q}_{1}^{2} -\frac{14}{5} s_2 \dot{q}_{1} \dot{q}_{3} - \frac{12}{5}\dot{q}_{3}^{2}}
 \end{align}
 which can be easily shown to reduce to Equation (\ref{eq:upright_evals}) when
-$q_2 = \dot{q}_1 = 0$. For a more detailed stability of steadily rolling disks,
-see~\cite{O'Reilly1996,Neimark1972,Kuleshov2001}.
+$q_2 = \dot{q}_1 = 0$. For a more detailed analysis of the stability of
+steadily rolling disks, see~\cite{O'Reilly1996,Neimark1972,Kuleshov2001}.
 
@@
 \label{sec:discussion}
-Most modern control techniques assume a system can be written as
+Most modern linear control techniques assume a system can be written as
 \begin{align}
@@
 numerical evaluation (as opposed to choosing a particular set of dependent
-speeds during symbolic derivation). There are several reasons why this can be
-desirable: 1) the configuration of the system in question changes significantly
-enough during simulation that no single choice of independent speeds will work
-for all regions of the configuration space; 2) the choice of parameters greatly
-affects which states should be selected to be dependent; and 3) to choose the
-dependent states which minimize the effect of numerical round off. This
-approach can also be applied to computation of the linearized dynamics. A task
-where the ability to switch the choice of dependent states ``on the fly'' is
-extremely helpful is in the computation of Lyapunov characteristic exponents
-(LCE's). When computing LCE's, time simulation of the nonlinear dynamics
-\textit{and} linearization at each point along the trajectory a
-required~\cite{Benettin1980a,Benettin1980b,Udwadia2001};
-having a computer code that makes it easy to switch choice of dependent states
-at the time of numerical evaluation is of great benefit in this case. The
-matrices which must be used to determine the best choice of coordinates and
-speeds are $\nabla_{\bm{q}} \bm{f}_c P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_v P_{ud}$;
-these matrices appear in equations (\ref{eq:delta_q}) and (\ref{eq:delta_u}).
-If for some choice of dependent coordinates or dependent speeds either of these
-matrices are singular or nearly singular, a different set of dependent
-coordinates and dependent speeds should be chosen. These matrices generally
-depend only upon parameters and coordinates (not speeds). While some systems
-may permit a choice of independent state variables which are valid for all
-configurations of interest, others may not. Methods for automatically selecting
-the ``best'' choice of independent state variables are discussed in
-\cite{Reckdahl1996}; they involve computing the singular value decomposition
-$\nabla_q \bm{f}_c$ and $\nabla_u \bm{f}_v$  to determine a set of independent
-states which will ensure the non-singularity of the aforementioned matrices.
+speeds during symbolic derivation).
+
+There are several reasons why this can be desirable: 1) the configuration of
+the system in question changes significantly enough during simulation that no
+single choice of independent speeds will work for all regions of the
+configuration space; 2) the choice of parameters greatly affects which states
+should be selected to be dependent; and 3) to choose the dependent states which
+minimize the effect of numerical round off. This approach can also be applied
+to computation of the linearized dynamics.
+
+A task where the ability to switch the choice of dependent states ``on the
+fly'' is extremely helpful is in the computation of Lyapunov characteristic
+exponents (LCE's). When computing LCE's, time simulation of the nonlinear
+dynamics \textit{and} linearization at each point along the trajectory a
+required~\cite{Benettin1980a,Benettin1980b,Udwadia2001}; having a computer code
+that makes it easy to switch choice of dependent states at the time of
+numerical evaluation is of great benefit in this case.
+
+The matrices which must be used to determine the best choice of coordinates and
+speeds are $\nabla_{\bm{q}} \bm{f}_c P_{qd}$ and $\nabla_{\bm{u}} \bm{f}_v
+P_{ud}$; these matrices appear in equations (\ref{eq:delta_q}) and
+(\ref{eq:delta_u}).  If for some choice of dependent coordinates or dependent
+speeds either of these matrices are singular or nearly singular, a different
+set of dependent coordinates and dependent speeds should be chosen. These
+matrices generally depend only upon parameters and coordinates (not speeds).
+While some systems may permit a choice of independent state variables which are
+valid for all configurations of interest, others may not. Methods for
+automatically selecting the ``best'' choice of independent state variables are
+discussed in \cite{Reckdahl1996}; they involve computing the singular value
+decomposition $\nabla_q \bm{f}_c$ and $\nabla_u \bm{f}_v$  to determine a set
+of independent states which will ensure the non-singularity of the
+aforementioned matrices.
 

### 2014-03-30 PetersonGedeHubbard2012.tex
Implement Mont's comments on page 5 and 7.

@@
 \end{align}
-where $\mathbf{f}_v$ and $\mathbf{0}$ have shape $m \times 1$,
-and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
-velocity constraints include the time-differentiated configuration constraints
-or equivalent constraints which produce the same behavior. Furthermore,
-velocity constraints should be written such that generalized speeds and
-coordinates appear, but $\dot{\mathbf{q}}$ terms do not.  Finally, acceleration
-constraints are typically time-differentiated velocity constraints (or
-similarly formed from kinematic considerations) and are assumed to have the
-form
+where $\mathbf{f}_v$ and $\mathbf{0}$ have shape $m \times 1$, and
+$\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is assumed that
+$\mathbf{f}_v$ includes time-differentiated configuration constraints as well
+as truly nonholonomic constraints. Further, it is assumed that the kinematic
+differential equations have been used to eliminate $\dot{\mathbf{q}}$ from
+$\mathbf{f}_v$. Finally, acceleration constraints are time-differentiated
+velocity constraints and are assumed to have the form
 \begin{align}
@@
 With these definitions established, Kane's method can be applied. In vector
-form, Kane's dynamical equations are
+form, Kane's holonomic dynamical equations are
 \begin{align}
@@
 force applied to body $B$ at point $Q$, then $\bar{R}_{B^*}=0$ and
-$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The generalized active forces are
+$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The holonomic generalized active forces are
 \begin{equation}
@@
 Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
-$B$ about its mass center $B^*$ \cite{Kane1985}. The generalized inertia forces
+$B$ about its mass center $B^*$ \cite{Kane1985}. The holonomic generalized inertia forces
 are
@@
 
-All the terms in (\ref{eq:kanes_eq}) are now defined, but this equation is
-valid only if there are no velocity constraints. If there are velocity
-constraints, the equations must be transformed. As we present them, Kane's
-nonholonomic dynamical equations will be of dimension $o - m$, where $o$ is the
+If there are velocity constraints, we must used Kane's nonholonomic dynamical
+equations
+%which allows Kane's nonholonomic dynamical equations to be written as
+\begin{align}
+\label{eq:kanes_eq_nonholonomic}
+\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
+\end{align}
+where $\tilde{\mathbf{F}}$ and $\tilde{\mathbf{F}}^*$ are the nonholonomic
+generalized active forces and noholonomic generalized inertia forces.  Equation
+(\ref{eq:kanes_eq_nonholonomic}) will be of dimension $o - m$, where $o$ is the
 total number of generalized speeds and $m$ is the number of velocity
-constraints. However, these $o-m$ equations generally involve all $o$ speeds
-(we assume the dependent speeds and their time derivatives have \textit{not}
-been algebraically eliminated). To obtain the nonholonomic dynamical equations,
-however, a choice must be made as to which speeds will be dependent.  The
-methods for doing this are beyond the scope of this paper, but it is important
-that whatever selection is made avoids singularities which can occur in certain
-configurations or sets of system parameters. For a careful treatment , we
-encourage the reader to examine the work of Reckdahl \cite{Reckdahl1996}.
+constraints.
 
@@
 \end{align}
-which allows Kane's nonholonomic dynamical equations to be written as
-\begin{align}
-\label{eq:kanes_eq_nonholonomic}
-\tilde{\mathbf{F}} + \tilde{\mathbf{F}}^* = \mathbf{0}
-\end{align}
+
 This process of constraining the equations of motion reduces the number of
@@
 \end{align}
+
+The $o-m$ equations in (\ref{eq:kanes_eq_nonholonomic}) generally involve $o$
+speeds (we assume the dependent speeds and their time derivatives have
+\textit{not} been algebraically eliminated). To obtain
+(\ref{eq:kanes_eq_nonholonomic}) one must choose which speeds will be
+dependent.  The methods for doing this are beyond the scope of this paper, but
+it is important that whatever selection is made avoids singularities which can
+occur in certain configurations or sets of system parameters. For a careful
+treatment, we encourage the reader to examine the work of Reckdahl
+\cite{Reckdahl1996}.
 

### 2014-03-30 PetersonGedeHubbard2012.tex
Implement Mont's page 6 comments.

@@
 follows. Consider a system composed of particles $P_1,...,P_h$, rigid bodies
-$B_1,...,B_g$, with points of force application $Q_1,...,Q_k$, and reference
-frames of torque application $E_1,...,E_c$, all defined relative to the
-inertial reference frame $N$. This system has generalized coordinates
-$q_1,...,q_n$ and generalized speeds $u_1,...,u_o$.  While it common that $n =
-o$, this is not a requirement when using Kane's method. Additionally, there are
-$l$ configuration constraints and $m$ velocity constraints.
+$B_1,...,B_g$, with mass centers $B_1^*,...,B_g^*$ and points of force
+application $Q_1,...,Q_k$, and reference frames of torque application
+$E_1,...,E_c$, all defined relative to the inertial reference frame $N$. This
+system has generalized coordinates $q_1,...,q_n$ and generalized speeds
+$u_1,...,u_o$.  While it common that $n = o$, this is not a requirement when
+using Kane's method. Additionally, there are $l$ configuration constraints and
+$m$ velocity constraints.
 
@@
 $\mathbf{F}$ is the multibody generalization of $f$ and $\mathbf{F}^*$ is the
-multibody generalization of $-ma$. As will be shown, these forces are obtained
-by taking the dot product of the partial velocities/angular-velocities, as
-defined in (\ref{eq:particle_translational}) - (\ref{eq:rb_rotational_frames}),
-with the resultant forces/torques ($\bar{R}$/ $\bar{T}$) and with inertial
+multibody generalization of $-ma$. As will be shown, these forces are the dot
+product of the partial velocities/angular-velocities, as defined in
+(\ref{eq:particle_translational}) - (\ref{eq:rb_rotational_frames}), with the
+resultant forces/torques ($\bar{R}$/ $\bar{T}$) and with inertial
 forces/torques ($\bar{R}^*$/$\bar{T}^*$).
@@
 force applied to body $B$ at point $Q$, then $\bar{R}_{B^*}=0$ and
-$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The holonomic generalized active forces are
+$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The $o$ holonomic generalized active forces are
 \begin{equation}
@@
 
-We now define the inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$. We
-leave out the inertial frame $N$ from the expressions; it is assumed to be $N$.
-The inertia force for particle $P$ is
+We now define inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$.
+For an inertial frame $N$, the inertia force for particle $P$ is
 \begin{align}
@@
 Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
-$B$ about its mass center $B^*$ \cite{Kane1985}. The holonomic generalized inertia forces
-are
+$B$ about its mass center $B^*$ \cite{Kane1985}. The $o$ holonomic generalized
+inertia forces are
 \begin{equation}
@@
 
+For brevity, we use $c_i=\cos{q_i}$ and $s_i=\sin{q_i}$.
+
 \begin{figure}[!htb]
@@
   \label{v_qdot}
-  \bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
+  {^N}\bar{v}^{C^*} = \dot{q}_4 \hat{n}_x + \dot{q}_5 \hat{n}_y + \dot{q}_6 \hat{n}_z
 \end{equation}

### 2014-08-15 response_to_reviewers.txt
Add response to reviewers

To all reviewers:
The manuscript has been drastically revised and polished. In particular, the
order of presentation and the motivation for the work has been expanded upon
greatly. Additionally, the example has been expanded upon significantly and the
steps shown more explicitly.

The electronic supplementary material should now be available and should be
executable by installing Python and SymPy (both free and open source, no other
dependencies). The simplest way to do this is to install one of the many
free scientific Python distributions which include both of these (and a number
of other useful tools). We recommend any of the following:
1) Enthought's Canopy: https://www.enthought.com/products/canopy/
2) Python XY: https://code.google.com/p/pythonxy/
3) Anacond Python: https://store.continuum.io/cshop/anaconda/
All are open source, well documented, and have free versions (and paid ones as
well with support), and have a userbase in the 10's-100's of thousands so
finding help online is generally quite easy.

To reviewer #1:
The new organization of the paper should address the concern you raised about
the presentation of linearized equations of motion without any deeper
motivation or introduction to the terms used.

Regarding the use of both independent states and dependent states, we address
this more in the introduction, the example, and the discussion of the revised
manuscript. Briefly, we formulate the equations which include both independent
and dependent states to allow for the possibility that which states are chosen
to be dependent/independent can be changed without having to redo the whole
symbolic formulation.

To reviewer #3:
Thank you for the references. We have since read all of them and highlighted in
the introduction the distinction between those works and our own.
Interestingly, we contacted Arend Schwab during the preparation of our
manuscript regarding references to linearization of constrained systems; his
response helpful, but did not include his own 2003 paper which you reference.
That aside, none of references you list employ Kane's method, nor do any
include discussion of allowing for non-trivial generalized speeds to be
selected (i.e., something other than d/dt(q_i) = u_i), nor are any of the
software packages to which the various authors refer open source -- with the
exception of the work by Minaker. The work by Minaker uses Lagrange's method,
and hence we feel is significantly distinct from our own; further he gives no
reference to the software which he refers to in his paper (Section 6),
presumably it is the software "EoM": http://137.207.14.230/vdc/index.html, but
we cannot be sure. If this is the software, it has no mailing list, uses no
source version control with a history of commits and contributions, and has no
automated test suite to verify it's correctness; it is essentially the project
of a single lab group and has no mechanism for participation by a wider
community. In contrast, the software we use to derive the equations of motion,
SymPy (sympy.org), is open source, has an extremely active mailing with
contributors from all over the world, and has been funded (through paying
students a small summer stipend) by Google for the last 6 years under the
Google Summer of Code program; it also has automated software tests for unit
tests a number example systems including the rolling disk.

Regarding the choice of independent speeds, the motivation for allowing the
choice of dependent speeds to be changed in the linearization procedure (even
though the dynamic equations, as you astutely observed, must be formed with
respect to a particular choice of dependent speeds) is so that the choice of
dependent speeds can be changed at the time of numerical evaluation rather than
at the time of symbolic derivation. We expand upon this more in the last
paragraph of the discussion, as well as provide an example of when it is useful
(computation of Lyapunov characteristic exponents for a system which a single
choice of dependent coordinates/speeds may be infeasible). Essentially, rather
than symbolically outputting the constrained equations of motion for every
possible choice of dependent states, it is possible to only output the
unconstrained equations of motion and the constraint Jacobian matrices, and
compute the "best" choice of dependent states (at runtime) and then
appropriately form the constrained dynamic equations (again, at runtime).

### 2014-08-15 response_to_reviewers_rev2.txt
Add response to reviewers

Response to Reviewer #1 comments:

1. Change in revised abstract.

2. Changed in revised abstract to read: "(i.e. no DAE’s, dependent variables
are not substituted away)". We simply mean that algebraic elimination has not
been used to eliminate dependent variables.

3. We have removed the word "familiar".

4. We erased the duplicated word "system". The constraint equations are
involved as the reviewer mentions.

5. We replaced the word "types" with "levels".  Clarified
holonomic/nonholonomic and the use of time differentiated constraints at the
velocity and acceleration level.

6. We have clarified the text here.  Lagrange's method can be used for
nonoholonomic systems.  See Nonholonomic Mechanics and Control by Anthony Bloch
for extensive discussion of Lagrange's method applied to nonholonomic systems.
The two approaches (Kane's vs. Lagrange) are comparable but Kane's method
offers more flexibility on the choice of generalized speeds (i.e., u_i != dq_i
/ dt). The quoted sentence has been rewritten for clarity.

7. We have changed this equation for clarity.

8. Quoted text rewritten for clarity.  n and o are not required to be equal in
our approach, though this is commonly the case for simple holonomic systems.

9. ...

10. We have added a figure.

11.  The text has been clarified.

12.  We have revised this section to fix the reviewers comments and make the
section more clear. We do not consider the implementation/performance of a
symbolic matrix inversion.  The Jacobian of the right hand side of (44) can be
performed symbolically or numerically.  In this example, we obtained (44) by
symbolically separating and inverting terms in (43).  We symbolically obtained
(45) by computing the Jacobian using SymPy (www.sympy.org). The "steady rolling
condition" is exactly what it sounds like: the disk is rolling at a constant
spin rate, either in a straight line or along a circular path of constant
radius. See reference [14] by Oliver O'Reilly for a more detailed treatment of
rolling disks.

13. ...

14. ...

Response to Reviewer #3

1. The notation has been fixed in these equations.

2. The input description was clarified in the text following the table.

3. We have clarified this in the revised manuscript in the second paragraph
after Table 1.

4. Figure added.

5. Unnecessary step removed.

6. We have improved the discussion in the text. Having any incorrect
eigenvalues will result in an incorrect linear system response.  We do not
explore these incorrect results due to space limitations.

7. Thank you for catching this. The equation has been corrected.

8. This discrepancy has been clarified.

9. Thank you for catching this.  We added a definition for matrix M.

### 2014-08-16 PetersonGedeHubbard2012.tex
Fixed references, defined c_i, s_i
Implement changes based on reviewer 1 comments

@@
 \documentclass[smallcondensed,final]{svjour3}                     % onecolumn (standard format)
-%\documentclass{article}
 \RequirePackage{fix-cm}
@@
 \usepackage{bm}
-%\usepackage{pgf}
-%\usepackage[polutonikogreek]{babel}
 
@@
 and motion equations and their subsequent linearization.  This procedure was
-developed specifically for equations of motion generated by Kane's method, but
-is compatible with any method as long as only ordinary differential equations
-are required to describe the system, and independent and dependent states
-remain in the final equations (i.e. no DAE's, dependent variables are not
-substituted away).
+developed for equations of motion generated by Kane's method where dependent
+states have not been algebraically eliminated; it is compatible with any method
+as long as only ordinary differential equations are required to describe the
+system (i.e., no differential algebraic equations).
 % Object of the document
@@
 Design of linear state feedback controllers and first order stability analysis
-of equilibria are two of the most common uses of multibody system dynamics
-models. Both require linearized equations of motion. For unconstrained systems,
-obtaining the linearized equations of motion is as simple as arranging the
-equations of motion in first order form ($\dot{x} = f(x, t)$) and calculating
-the Jacobian of the right hand side function, with respect to the states,
-evaluated at the equilibria point of interest.  However, when the system in
-questions has constraints, i.e., not all the states are independent, this
-process is not as straightforward.
+are common uses of multibody system dynamics models. Both require linearized
+equations of motion. For unconstrained systems, obtaining the linearized
+equations of motion is as simple as arranging them  in first order form
+($\dot{x} = f(x, t)$) and evaluating $\frac{\partial f}{\partial x}}$ at the
+point of interest.  However, when the system has constraints not all the states
+are independent, and this process is not as straightforward.
 
 For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
-first order ODE's (in the generalized speeds) which include both the
+first order ODEs (in the generalized speeds) which include both the
 independent and dependent states (assuming the dependent states have not been
@@
 We consider three levels of constraints: configuration, velocity, and
-acceleration. Configuration (holonomic) constraints limit the location or
-orientation of parts of the system, relative to the external world or other
-parts of the system. Velocity (nonholonomic or time differentiated holonomic)
-constraints limit the speeds at which the configuration can change.
-Velocity constraints often appear where there is rolling without slip or there
-are closed kinematic loops. For this procedure, acceleration level constraints
-are assumed to be time differentiated velocity constraints.
+acceleration. Constraints at the configuration level (holonomic) limit the
+location or orientation of parts of the system. Configuration constraints may
+arise from closed kinematic loops. Constraints at the velocity level limit
+permissible velocities of the system. Velocity constraints may arise from
+rolling without slip (nonholonomic) or closed kinematic loops (time
+differentiated holonomic).  Constraints at the acceleration level are assumed
+to be time differentiated velocity level constraints.
 
@@
 equations formulated using other methods, it is restricted to systems which can
-be completely described by a set of ODE's (i.e., DAE's needn't be solved).
+be completely described by a set of ODEs (i.e., DAEs needn't be solved).
 
@@
 
-By assumption, the velocity constraints are linear with respect to
+The velocity constraints are assumed linear with respect to
 $\mathbf{u}$, so we introduce
@@
 \end{align}
-which implies
+where $\mathbf{B}$ is implicitly dependent upon $\mathbf{q}$ and $t$.  The velocity
+constraints can then be expressed as
 \begin{align}
@@
 where $\tilde{\mathbf{F}}$ and $\tilde{\mathbf{F}}^*$ are the nonholonomic
-generalized active forces and noholonomic generalized inertia forces.  Equation
+generalized active forces and nonholonomic generalized inertia forces.  Equation
 (\ref{eq:kanes_eq_nonholonomic}) will be of dimension $o - m$, where $o$ is the
@@
                      &= (r u_2 c_3 + u_4) \hat{c}_x + (-r
-                        u_1 + u_5)\hat{c}_y + r u_2 s_3 +
-                        u6) \hat{c}_z
+                        u_1 + u_5)\hat{c}_y + (r u_2 s_3 +
+                        u_6) \hat{c}_z
 \end{align*}
@@
 
-At this point, the reader familiar with the rolling disk problem is certainly
-wondering why we have chosen such a diabolical set of coordinates and speeds.
-In this simple example, there is an {\bf \textit{obvious and minimal}} choice of
-coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this choice
-will be valid for all parameters and configurations; further, the derivation of
-the equations of motion can be done without ever introducing these quantities),
-for some systems this is not the case (e.g., Stewart's platform, bicycle). The
-numeric values of system parameters and the configurations in which the system
-will operate will determine which coordinates and speeds can be taken to be
-dependent (and hence potentially eliminated by algebraic manipulations).
-However, it may not be possible \textit{apriori} to determine easily which
-coordinates and speeds should be taken as dependent; further, some systems may
-operate in distinct enough regions of the configuration space that one
-\textit{must} use different choices of dependent coordinates and/or speeds as
-the system moves from one region to another. The purpose of our complex
-choice of coordinates and speeds is to illustrate, in the context of a familiar
-example, how constraints must be taken into account when linearizing equations
-of motion. It is our hope that by doing this for a well known and relatively
-simple example, readers can apply the same techniques to more complicated
-systems where its use is more appropriate or necessary.
+At this point, the reader familiar with the rolling disk problem may wondering
+why we have purposefully chosen a non-minimal set of coordinates and speeds.
+In this simple example, there is an {\bf \textit{obvious and minimal}} choice
+of coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this
+choice will be valid for all parameters and configurations; further, the
+derivation of the equations of motion can be done without ever introducing
+these quantities), for some systems this is not the case (e.g., Stewart's
+platform, bicycle). The numeric values of system parameters and the
+configurations in which the system will operate will determine which
+coordinates and speeds can be taken to be dependent (and hence potentially
+eliminated by algebraic manipulations).  However, it may not be possible
+\textit{a priori} to determine easily which coordinates and speeds should be
+taken as dependent; further, some systems may operate in distinct enough
+regions of the configuration space that one \textit{must} use different choices
+of dependent coordinates and/or speeds as the system moves from one region to
+another. The purpose of our complex choice of coordinates and speeds is to
+illustrate, in the context of a familiar example, how constraints must be taken
+into account when linearizing equations of motion. It is our hope that by doing
+this for a well known and relatively simple example, readers can apply the same
+techniques to more complicated systems where its use is more appropriate or
+necessary.
 
@@
 acceleration constraint equations provide the final three equations which allow
-for all twelve unkowns to be solved. Combining (\ref{rd:f_0_f_1}),
+for all twelve unknowns to be solved. Combining (\ref{rd:f_0_f_1}),
 (\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete set of motion equations
@@
 where $\mathbf{x}$ are the system states and $\mathbf{u}$ are the control
-inputs (the $\mathbf{u}$ used in control literature is analagous to the
+inputs (the $\mathbf{u}$ used in control literature is analogous to the
 $\mathbf{r}$ we use in this paper).  To transform equation
@@
 fly'' is extremely helpful is in the computation of Lyapunov characteristic
-exponents (LCE's). When computing LCE's, time simulation of the nonlinear
+exponents (LCEs). When computing LCEs, time simulation of the nonlinear
 dynamics \textit{and} linearization at each point along the trajectory a

### 2014-08-16 response_to_reviewers_aug_2014.txt

Response to Editor.

Dear Editor(s),
We feel that in this revision, as in the previous revisions, we have made an
effort to address as many of Reviewer #1's comments as possible. We have valued
a number of the comments provided by Reviewer #1, agreeing that the suggested
changes help to clarify the motivation and methods in the paper.

Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
reasonable to address, comments 7 and 12.

Comment 7 simply states that the reviewer finds our presentation frustrating,
without offering specific faults, mistakes, unclear sections, areas that could
use improvement, etc.  Additionally, some of the other comments (comment 3) are
the reason we have included such a thorough description of Kane's method. There
seems to be a breakdown where the reviewer is not understanding how we are
applying this method, and it is not clear on our end which part of Section 2
needs to be improved to address the reviewer's difficulties.

Comment 12 states that the reviewer found the parts of the manuscript that they
understood confusing, and then admits to being unable to follow the derivations
we have presented. While we cannot simplify the math in the derivations, one
can always attempt to provide more clarification, but only if the reviewer
states which areas they had trouble with or where throughout the derivation
they became lost. This comment as given to us, does not allow us to do
anything.

In summary, we have addressed as many of the comments as possible, but feel
that the other comments are not reasonable, and suggest that Reviewer #1 has
not made a good faith effort to either comprehend the paper, or communicate
where they had difficulty comprehending the paper. As Reviewer #2 seems to
approve of the manuscript, we would ask that you overlook Reviewer #1's
concerns when making the decision to publish or reject this work.

Thank you,
Luke Peterson
Gilbert Gede
Mont Hubbard

Response to Reviewer #1.

1. Abstract. “...but is compatible with any method as long as only ordinary
    differential equations are required to describe the system, and independent
    and dependent states remain in the final equations (i.e. no DAE's,
    dependent variables are not substituted away)”. The sentence, and
    especially its end, is still not clear to me.

Author Response: We have changed the sentence for clarity. It now more clearly
specifies the types of differential equations which can be used.  It also
clarifies that it was developed for use with Kane's Method when dependent
quantities have not been algebraically eliminated.

2. 1. 36-37 (page 1, lines 36-37). “Design of linear state feedback controllers
    and first order stability analysis of equilibria are two of the most common
    uses of multibody system dynamics models”. I am not sure about these are
    the most common uses.

Author Response: We removed the words which suggested that these were the most
common uses, instead stating that they are just common uses.

3. 2. 21-19. The paragraph is still unclear to me. Kane’s equations are the
   projections of the initial (expressed in dependent velocities and as such
   constraint reaction-induced) dynamic equations into the directions of
   generalized speeds, which are by assumption independent for a constrained
   system, and with these projections the constraint reactions are excluded
   from evidence. If the projected dynamic equations are still expressed in the
   dependent velocities (not in the generalized speeds), the number of the
   projected dynamic equations is smaller than the number of the velocities. In
   this context I do not understand the remark that “The kinematical and
   dynamical differential equations can be solved for the time derivatives of
   the generalized coordinates and generalized speeds, ...” as well as most of
   the content of the paragraph. My reservations raised in the previous review
   were not clarified.

Author Response: We believe you have some confusion regarding the different
possible ways to apply Kane's method, which is why we Section 2 clarifies how
we use Kane's method. One possible way to apply the method is when a
non-minimal set of speeds are used to define the velocities of the system; in
this case, the velocity constraint matrix B (equation 11 in the paper) can be
reworked into matrix A (equation 21). This A matrix is used to reduce the
number of equations for generalized speeds by the number of constrained speeds.
This is also outlined by Kane in [Kane 1985], in equations 4.4.3 and 4.4.11,
which shows the reduction in size of the equations for the generalized speeds
and that _all_ the speeds (dependent and independent) are still present.

4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
    the Authors. For example, the remark “Velocity constraints often appear
    where there is rolling without slip or there are closed kinematic loops” is
    unclear (if correct at all). Velocity constraints, or, better, constraint
    equations at the velocity level, are either nonholonomic constraints (often
    consequent to the rolling without slip) or differentiated with respect to
    time holonomic constraint equations (one source of such constraints are
    closed kinematic loops).

Author Response: We have rewritten this entire paragraph to ensure that the
description of each level of constraints is consistent, and more clear. We have
reworked the sentence in question to more clearly state that things like
rolling without slip lead to nonholonomic constraints and closed kinematic
loops lead to time differentiated holonomic constraints, both of which are
velocity level constraints.

5. 5. 25-50. The velocity constraints are introduced in Eq. (8) in a general
    (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
    “by assumption”?), the velocity constraints are linear with respect to u,
    ... which implies Bu + fvt(q,t) = 0”. Firstly, in general, the constraint
    matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
    constraints were not introduced from the beginning in this “linear” form?

Author Response: We have clarified that the B matrix is a function of the
coordinates and time, even though it is written as only a single symbol (a
format which we use throughout the paper).

6. 8. 15 and 831. The meaning of the “exogenous inputs” is not well
    motivated/explained.

Author Response: We believe that the sentence "The s exogenous inputs include
control system inputs and/or external disturbance forces or torques applied to
the system." explains that the exogenous inputs include forces/torques which
are not internal to the system (e.g. a reaction force), but instead applied
from outside the system (a motor driving a body in the system).

7. A general remark to Section 2. While Kane’s method is generally well
    understood, the Authors presentation of its formulation is rather
    “frustrating”.

8. 10. 32-34. One of the velocity constraint equations introduced in Eq. (30)
    should be a time differentiated form of the configuration constraint of Eq.
    (25). It seems that it is Eq. (30c).  If so, it should rather read 
    r u2 s2 + u6  = 0 (not r u2 s3 + u6 = 0). Can you explain this?

Author Response: While it could be written this way, what we have presented is
correct. The method you are suggesting requires ^N v ^P* to be expressed in N.
Then, the N_x and N_y components would be 2 of the motion constraints, and the
time differentiated configuration constraint will equal the N_z component. This
can be simply verified by time differentiating the configuration constraint and
utilizing equation 34 to elimante the q_dots, so only u's are present. In
comparing this to the projection of ^N v ^P* in the N_z direction, you will
find these are equivalent. We believe what we have presented is simpler.

9. 10. 36-38. It is written: “These equations are linear in the ui terms,
    nonlinear in the qi terms, involve geometric system parameters (but not
    mass or inertial parameters), and do not explicitly involve \dot{qi}
    terms”. The remark that constraint equations do not involve inertial
    parameters is a cliché. According to Eq. (11), \dot{qi} terms are neither
    involved.

Author Response: Although it might be cliché, we feel explicitness is valuable,
and we do not wish to remove a sentence which might help a reader avoid
confusion.

10. 10. 39-118. The paragraph is unintelligible. The “diabolical set of
    coordinates and speeds” is just a set of dependent coordinates and speeds,
    resulted in one configuration constraint of Eq. (25) and three velocity
    constraints of Eq. (30) (one holonomic constraint of Eq.  (25) and two
    nonholonomic constraints of Eq. (30a) and (30b)). There are then five
    independent coordinates and three independent speeds. By choosing the
    independent coordinates and speeds, which may in general not be subsets of
    qi and ui, respectively, the constraint equations (25) and (30) are
    satisfied by identity.

Author Response: We have changed "diabolical" to "purposefully nonminimal".
That should clairfy that the example exists to show the technique on a
tractable system with a nonminimal set of coordinates and speeds, and to
highlight the approach, rather than suggest that we should be using these
coordinates and speeds on the rolling disk example.

11. 11. 14. Is Eq. (31c) correct? (see Remark 8).

Author Response: Yes it is correct, our response to comment 8 addresses this.
The code in the electronic supplementary material can be used to verify this.

12. A general remark. The complexity of the followed parts of the manuscript is
    not appealing to me. In fact, I was unable to follow the derivations. I may
    suppose that the same difficulties will have most of the potential readers
    of the paper, not mentioning those who would like to follow the procedures.

### 2014-08-29 response_to_reviewers_aug_2014.txt
Implement Mont's feedback from 3rd round of revisions

@@
 
-Comment 12 states that the reviewer found the parts of the manuscript that they
-understood confusing, and then admits to being unable to follow the derivations
+Comment 12 states that the reviewer is confused by the parts of the manuscript
+that they understood, and then admits to being unable to follow the derivations
 we have presented. While we cannot simplify the math in the derivations, one
 can always attempt to provide more clarification, but only if the reviewer
-states which areas they had trouble with or where throughout the derivation
-they became lost. This comment as given to us, does not allow us to do
-anything.
+states which areas they had trouble with or where in the derivation they became
+lost. This comment as given to us, does not allow us to do anything.
+
+We would also like to add that many of Reviewer #1's comments seem to indicate
+a shortage of effort to understand the paper. As an example, one of the authors
+has advised a masters student in a summer project to implement this method in a
+symbolic computer algebra system and within 2 months the student had
+implemented the entirety of the algorithm, complete with examples, software
+tests to verify mathematical correctness, and documentation. This student was
+able to understand the mathematics presented in the paper with very little
+guidance; fundamentally there is nothing in the derivation that is more
+complicated than a partial derivative and a few matrix multiplications.
 
 In summary, we have addressed as many of the comments as possible, but feel
-that the other comments are not reasonable, and suggest that Reviewer #1 has
-not made a good faith effort to either comprehend the paper, or communicate
-where they had difficulty comprehending the paper. As Reviewer #2 seems to
+that the other comments not addressable, and suggest that Reviewer #1 has
+not made enough of an effort to either comprehend the paper, or communicate
+where they had difficulty comprehending it. As Reviewer #2 seems to
 approve of the manuscript, we would ask that you overlook Reviewer #1's
@@
 Thank you,
-Luke Peterson
+Dale Peterson
 Gilbert Gede
@@
 
-3. 2. 21-19. The paragraph is still unclear to me. Kane’s equations are the
+3. 2. 1-19. The paragraph is still unclear to me. Kane’s equations are the
    projections of the initial (expressed in dependent velocities and as such
@@
 
-Author Response: We believe you have some confusion regarding the different
-possible ways to apply Kane's method, which is why we Section 2 clarifies how
-we use Kane's method. One possible way to apply the method is when a
-non-minimal set of speeds are used to define the velocities of the system; in
-this case, the velocity constraint matrix B (equation 11 in the paper) can be
-reworked into matrix A (equation 21). This A matrix is used to reduce the
+We do not use 'generalized coordinates' or 'generalized speeds' to imply that
+either set is minimal. In other words 'generalized' only implies that the
+coordinates/speeds are for the entire system and are not Cartesian or
+spherical coordinates necessarily.  The number of Kane's dynamical equations in our
+formulation is o - m, where o is the number of generalized speeds and m is the
+number of velocity (nonholonomic and time differentiated holonomic)
+constraints. These o - m dynamical equations will in general involve 1) all n
+coordinates and their time derivatives and 2) all o speeds and their time
+derivatives. So despite the fact that the Newton-Euler equations have been
+projected in the direction of the nonholonomic partial velocities (what you
+refer to as 'the direction of the generalized speeds'), there are still more
+unknowns (o generalized speed time derivatives) than there are equations (o -
+m).  This issue can be addressed by 1) algebraically eliminating dependent
+speed time derivatives to obtain o-m equations in o-m unknowns, or 2)
+augmenting the o-m dynamical equations with the m acceleration constraint
+equations to obtain o equations in o unknowns. Either way is mathematically
+equivalent but dependent upon the specific problem at hand one may be
+preferable to the other.
+
+We have revised manuscript to be more explicit about what we mean by solving
+the kinematical and dynamical differential equations for the time derivatives
+of the coordinates and time derivatives of the generalize speeds. For this
+dynamical differential equation:
+
+-m*du/dt + f = 0
+
+then by "solve the dynamical differential equations for the time derivatives of
+the generalized speeds" we mean:
+
+du/dt = f / m
+
+We welcome specific suggestions about how this can be made more clear.
+
+Section 2 was introduced specifically to clarify how we use Kane's method since
+some of our notation deviates from his presentation in his 1985 book. When the
+system has velocity constraints (i.e., a non-minimal set of generalized speeds
+are introducted, or the system has time differentiated holonomic constraints),
+the velocity constraint matrix B (equation 11 in the paper) can manipulated to
+form matrix A (equation 21 in the paper). This A matrix is used to reduce the
 number of equations for generalized speeds by the number of constrained speeds.
@@
 can be simply verified by time differentiating the configuration constraint and
-utilizing equation 34 to elimante the q_dots, so only u's are present. In
+utilizing equation 34 to eliminate the q_dots, so only u's are present. In
 comparing this to the projection of ^N v ^P* in the N_z direction, you will
@@
 
-Author Response: Although it might be cliché, we feel explicitness is valuable,
-and we do not wish to remove a sentence which might help a reader avoid
-confusion.
+Author Response: Although it might be a cliché, we feel explicitness is
+valuable, and we do not wish to remove a sentence which might help a reader
+avoid confusion.
 
@@
 Author Response: We have changed "diabolical" to "purposefully nonminimal".
-That should clairfy that the example exists to show the technique on a
-tractable system with a nonminimal set of coordinates and speeds, and to
+That should clarfy that the example exists to show the technique on a
+tractable system with a non-minimal set of coordinates and speeds, and to
 highlight the approach, rather than suggest that we should be using these

### 2014-08-29 PetersonGedeHubbard2012.tex
Implement Mont's feedback from 3rd round of revisions

@@
 are common uses of multibody system dynamics models. Both require linearized
-equations of motion. For unconstrained systems, obtaining the linearized
-equations of motion is as simple as arranging them  in first order form
-($\dot{x} = f(x, t)$) and evaluating $\frac{\partial f}{\partial x}}$ at the
-point of interest.  However, when the system has constraints not all the states
-are independent, and this process is not as straightforward.
+equations of motion.  We define a constrained system to be one whose
+configuration has been described by a non-minimal set of generalized
+coordinates or whose motion has been described by a non-minimal set of
+generalized speeds.  For unconstrained systems (with a minimal set of
+generalized coordinates and speeds), obtaining the linearized equations of
+motion is as simple as arranging them in first order form ($\dot{x} = f(x, t)$)
+and evaluating $\frac{\partial f}{\partial x}}$ at the point of interest.
+However, when the system has constraints not all the states (coordinates and
+speeds) are independent, and this process is not as straightforward.
 
 For a system with constraints, Kane's dynamical equations\cite{Kane1985} are
-first order ODEs (in the generalized speeds) which include both the
-independent and dependent states (assuming the dependent states have not been
-algebraically eliminated). The kinematical and dynamical differential equations
-can be solved for the time derivatives of the generalized coordinates and
-generalized speeds, respectively. Taken together, it is tempting to think that
-this system of first order differential equations could be linearized in the
-same fashion as an unconstrained system.  Unfortunately, this is not the case.
-Linearizing these equations in the same manner as one would for an
-unconstrained system (by directly computing the Jacobian) is incorrect and will
-yield incorrect linearized equations (and subsequently incorrect eigenvalues,
-in the case of a stability analysis). We illustrate this by means of a familiar
-example, the rolling disk, formulated with a non-traditional set of coordinates
-and generalized speeds which are not all independent. We then provide a
-systematic method to correctly generate the linear equations of motion for
-constrained systems which have been found using Kane's method (or any other
-method that results in equations with the same form). We apply this method to
-the rolling disk example to illustrate that it yields eigenvalues that match
-previously published results.
+first order ODEs (in the generalized speeds) which include both the independent
+and dependent states (assuming the dependent states have not been algebraically
+eliminated). The kinematical differential equations can be solved for the time
+derivatives of the generalized coordinates and the dynamical differential
+equations can be solved for the time derivatives of the generalized speeds.
+Taken together, it is tempting to think that this system of first order
+differential equations could be linearized in the same fashion as an
+unconstrained system.  Unfortunately, this is not the case.  Linearizing these
+equations in the same manner as one would for an unconstrained system (by
+directly computing the Jacobian) is incorrect and will yield incorrect
+linearized equations (and subsequently incorrect eigenvalues, in the case of a
+stability analysis). We illustrate this by means of a familiar example, the
+rolling disk, formulated with a non-traditional set of coordinates and
+generalized speeds which are not all independent. We then provide a systematic
+method to correctly generate the linear equations of motion for constrained
+systems which have been found using Kane's method (or any other method that
+results in equations with the same form). We apply this method to the rolling
+disk example to illustrate that it yields eigenvalues that match previously
+published results.
 
@@
 linearization, but their methods have been developed for equations of motion
-found using Lagrange's method, which contain Lagrange multipliers ( not present
+found using Lagrange's method, which contain Lagrange multipliers (not present
 in equations of motion generated using Kane's method). Further, both of these
@@
 coordinates; how these generalized speeds can be used to project permissible
-motions of the system, and how this removes the need to consider
+motions of the system; and how this removes the need to consider
 non-contributing (internal constraint) forces.  Then, the mathematical steps
@@
 system while time derivatives of the generalized coordinates characterize the
-velocities of a system.  Kane's method allows velocities to be written in terms
+velocities.  Kane's method allows velocities to be written in terms
 of generalized speeds, which are not required to be the time derivatives of
@@
 (non-contributing forces) imposed by objects such as pin or sliding joints no
-longer need to be considered. The reason for this is that partial velocities
+longer need be considered. The reason for this is that partial velocities
 are \textit{by construction} perpendicular to forces of constraint. Constraint
@@
 
-In summary, Kane's method allows for velocities to be defined in terms of
-generalized speeds, allowing for simpler velocity expressions.  The dot product
-of each partial velocity (the direction of motion associated with each
-generalized speed) with Newton's second law (or Euler's equations) removes all
-terms which are not related to the rate of change of each generalized speed.
-This generally results in simple equations of motion, which are easier to form.
+In summary, Kane's method allows velocities to be defined in terms of
+generalized speeds, leading to simpler velocity expressions and equations of
+motion expressed in first order form.  The dot product of each partial velocity
+(the direction of motion associated with each generalized speed) with Newton's
+second law (or Euler's equations) removes all terms which are not related to
+the rate of change of each generalized speed.  This generally results in simple
+equations of motion, which are easier to form.
 
@@
 aligned with the position vector from the disk center $C^*$ to the lowest point
-of the disk $P^*$. Figure \ref{fig:rolling_disc} shows how the generalized
-coordinates are used to configure the disc.
+of the disk $P^*$. Figure \ref{fig:rolling_disk} shows how the generalized
+coordinates are used to configure the disk.
 
@@
     \includegraphics[width=0.8\textwidth]{disk_drawing_cropped.pdf}
-    \caption{Configuration of rolling disc using non-minimal coordinates.}
-    \label{fig:rolling_disc}
+    \caption{Configuration of rolling disk using non-minimal coordinates.}
+    \label{fig:rolling_disk}
 \end{figure}
@@
 
-At this point, the reader familiar with the rolling disk problem may wondering
-why we have purposefully chosen a non-minimal set of coordinates and speeds.
-In this simple example, there is an {\bf \textit{obvious and minimal}} choice
-of coordinates and speeds ($q_6, u_4, u_5, u_6$, should be dependent; this
-choice will be valid for all parameters and configurations; further, the
-derivation of the equations of motion can be done without ever introducing
-these quantities), for some systems this is not the case (e.g., Stewart's
-platform, bicycle). The numeric values of system parameters and the
-configurations in which the system will operate will determine which
+At this point, the reader familiar with the rolling disk problem may be
+wondering why we have purposefully chosen a non-minimal set of coordinates and
+speeds.  In this simple example, there is an {\bf \textit{obvious and minimal}}
+choice of coordinates and speeds ($q_1, q_2, q_3, q_4, q_5, u_1, u_2, u_3$
+should be independent as this choice will be valid for all parameters and
+configurations.  The equations of motion can be formed without ever introducing
+$q_6, u_4, u_5, u_6$).
+
+For some systems there is not an obvious and minimal choice of coordinates and
+speeds (e.g., Stewart's platform, bicycle).  The values of system parameters
+and the configurations in which the system will operate determine which
 coordinates and speeds can be taken to be dependent (and hence potentially
@@
 \textit{a priori} to determine easily which coordinates and speeds should be
-taken as dependent; further, some systems may operate in distinct enough
+taken as dependent. Further, some systems may operate in distinct enough
 regions of the configuration space that one \textit{must} use different choices
@@
 \end{equation}
-Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot}),
-and resolving these vector equations into the $B$ frame, and rearranging so
+Equating (\ref{w_u}) with (\ref{w_qdot}) and (\ref{v_u}) with (\ref{v_qdot})
+and expressing these vector equations in the $B$ frame, and rearranging so
 that all terms appear to the left of the equality sign, we obtain
@@
 
-The generalized active forces for this example are
+The generalized active forces and torques for this example are
 \begin{align}
@@
 acceleration constraint equations provide the final three equations which allow
-for all twelve unknowns to be solved. Combining (\ref{rd:f_0_f_1}),
+all twelve unknowns to be determined. Combining (\ref{rd:f_0_f_1}),
 (\ref{rd:f_2_f_3}), and (\ref{rd:f_a}), the complete set of motion equations
@@
 \end{align}
-and take the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
-\mathbf{u}]^T$. However, this will give incorrect results. As discussed in
-\cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity, and the system
-is linearized about the upright steady rolling condition, the critical speed is
-$v \triangleq -r\dot{q}_3 =\pm\frac{1}{\sqrt{3}}$. When following the naive approach, and
-evaluating the Jacobian at the same operating conditions and parameters, eight
-of the twelve eigenvalues are identically zero while the remaining four are:
+and calculate the Jacobian of $\mathbf{f}$ with respect to $[\mathbf{q} \quad
+\mathbf{u}]^T$. However, this will give incorrect results.
+
+As discussed in \cite{Schwab2003}, when $m, r,$ and $g$ are taken to be unity,
+and the system is linearized about the upright steady rolling condition, the
+critical speed is $v \triangleq -r\dot{q}_3 =\pm\frac{1}{\sqrt{3}}$. When
+following the naive approach, and evaluating the Jacobian at the same operating
+conditions and parameters, eight of the twelve eigenvalues are identically zero
+while the remaining four are:
 \begin{align}
@@
 Further, the fact that twelve eigenvalues are obtained should be an alert that
-something is incorrect, since the number of independent quantities is only
-eight (five coordinates and three speeds). We now outline a linearization
-procedure which addresses this issue; we revisit this example in Section
+something is procedurally incorrect, since the number of independent quantities
+is only eight (five coordinates and three speeds). We now outline a
+linearization procedure which addresses this issue; we revisit this example in
+Section
 \ref{sec:example_revisited}.
@@
 procedure that properly accounts for system constraints. Computing the Jacobian
-of the right hand side of the ODE's as we did at the end of the previous
+of the right hand side of the ODEs as we did at the end of the previous
 section is incorrect because it fails to apply the chain rule and thereby
@@
 acceleration constraints. That it isn't immediately obvious that the chain rule
-needs to be applied is a byproduct of the commonly use notation which doesn't
+needs to be applied is a byproduct of the commonly used notation which doesn't
 make it explicitly clear that dependent coordinates and dependent speeds are not
@@
 implies (by appealing to the configuration constraint) that $q_6^* = -r
-\cos{q_2} = -r$.  This corresponds to the disk upright, the lowest point of the
-disc in contact with the ground, and the disk heading aligned with the
-$\hat{n}_x$ unit vector.  Next, we solve the velocity constraint equations for
-the dependent speeds in terms of the independent ones, and substitute these
-into the kinematic differential equations. This yields six equations with nine
-unknowns: $u_i (i=1,2,3)$ and $\dot{q}_i (i=1,\dots6)$. We \textit{choose} yaw
-rate $\dot{q}_1^* = 0$, lean rate $\dot{q}_2^* = 0$, let spin rate
-$\dot{q}_3^*$ be a free variable, and solve for the remaining six unknowns.
-This yields $u_1^* = u_3^* = 0$, $u_2^* = \dot{q}_3^*$, $\dot{q}_4^* =
--r\dot{q}_3^*$, $\dot{q}_5^* = \dot{q}_6^* = 0$. Back substituting the first
-three of these results into the velocity constraint equations yields $u_4^* =
+\cos{q_2^*} = -r$.  This corresponds to the disk in a vertical plane, the
+lowest point of the disk in contact with the ground, and the disk plane
+parallel to the $\hat{n}_x$ unit vector.  Next, we solve the velocity
+constraint equations for the dependent speeds in terms of the independent ones,
+and substitute these into the kinematic differential equations. This yields six
+equations with nine unknowns: $u_i (i=1,2,3)$ and $\dot{q}_i (i=1,\dots6)$. We
+\textit{choose} yaw rate $\dot{q}_1^* = 0$, lean rate $\dot{q}_2^* = 0$, let
+spin rate $\dot{q}_3^*$ be a free variable, and solve for the remaining six
+unknowns.  This yields $u_1^* = u_3^* = 0$, $u_2^* = \dot{q}_3^*$, $\dot{q}_4^*
+= -r\dot{q}_3^*$, $\dot{q}_5^* = \dot{q}_6^* = 0$. Substituting the first three
+of these results into the velocity constraint equations yields $u_4^* =
 -r\dot{q}_3^*$, $u_5^* = u_6^* = 0$. Finally, by evaluating the acceleration
@@
 -r \dot{q}_3^{*2}$. The upright equilibrium conditions are thus parameterized
-in terms of the disc spin rate $\dot{q}_3$; for this upright configuration
+in terms of the disk spin rate $\dot{q}_3$; for this upright configuration
 ($q_2 = 0$) it is convenient to introduce the forward speed $v \triangleq

### 2014-08-29 response_to_reviewers_aug_2014.txt
Change quotation marks

@@
 
-1. Abstract. “...but is compatible with any method as long as only ordinary
+1. Abstract. "...but is compatible with any method as long as only ordinary
     differential equations are required to describe the system, and independent
     and dependent states remain in the final equations (i.e. no DAE's,
-    dependent variables are not substituted away)”. The sentence, and
+    dependent variables are not substituted away)". The sentence, and
     especially its end, is still not clear to me.
@@
 
-2. 1. 36-37 (page 1, lines 36-37). “Design of linear state feedback controllers
+2. 1. 36-37 (page 1, lines 36-37). "Design of linear state feedback controllers
     and first order stability analysis of equilibria are two of the most common
-    uses of multibody system dynamics models”. I am not sure about these are
+    uses of multibody system dynamics models". I am not sure about these are
     the most common uses.
@@
    projected dynamic equations is smaller than the number of the velocities. In
-   this context I do not understand the remark that “The kinematical and
+   this context I do not understand the remark that "The kinematical and
    dynamical differential equations can be solved for the time derivatives of
-   the generalized coordinates and generalized speeds, ...” as well as most of
+   the generalized coordinates and generalized speeds, ..." as well as most of
    the content of the paragraph. My reservations raised in the previous review
@@
 4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
-    the Authors. For example, the remark “Velocity constraints often appear
-    where there is rolling without slip or there are closed kinematic loops” is
+    the Authors. For example, the remark "Velocity constraints often appear
+    where there is rolling without slip or there are closed kinematic loops" is
     unclear (if correct at all). Velocity constraints, or, better, constraint
@@
     (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
-    “by assumption”?), the velocity constraints are linear with respect to u,
-    ... which implies Bu + fvt(q,t) = 0”. Firstly, in general, the constraint
+    "by assumption”?), the velocity constraints are linear with respect to u,
+    ... which implies Bu + fvt(q,t) = 0". Firstly, in general, the constraint
     matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
-    constraints were not introduced from the beginning in this “linear” form?
+    constraints were not introduced from the beginning in this "linear" form?
 
@@
     understood, the Authors presentation of its formulation is rather
-    “frustrating”.
+    "frustrating".
 

### 2014-08-29 response_to_editors_aug_2014.txt
Separate responses to review and editor

--- 2014-08-29 response_to_reviewers_aug_2014.txt
@@
 Mont Hubbard
-
-Response to Reviewer #1.
-
-1. Abstract. "...but is compatible with any method as long as only ordinary
-    differential equations are required to describe the system, and independent
-    and dependent states remain in the final equations (i.e. no DAE's,
-    dependent variables are not substituted away)". The sentence, and
-    especially its end, is still not clear to me.
-
-Author Response: We have changed the sentence for clarity. It now more clearly
-specifies the types of differential equations which can be used.  It also
-clarifies that it was developed for use with Kane's Method when dependent
-quantities have not been algebraically eliminated.
-
-2. 1. 36-37 (page 1, lines 36-37). "Design of linear state feedback controllers
-    and first order stability analysis of equilibria are two of the most common
-    uses of multibody system dynamics models". I am not sure about these are
-    the most common uses.
-
-Author Response: We removed the words which suggested that these were the most
-common uses, instead stating that they are just common uses.
-
-3. 2. 1-19. The paragraph is still unclear to me. Kane’s equations are the
-   projections of the initial (expressed in dependent velocities and as such
-   constraint reaction-induced) dynamic equations into the directions of
-   generalized speeds, which are by assumption independent for a constrained
-   system, and with these projections the constraint reactions are excluded
-   from evidence. If the projected dynamic equations are still expressed in the
-   dependent velocities (not in the generalized speeds), the number of the
-   projected dynamic equations is smaller than the number of the velocities. In
-   this context I do not understand the remark that "The kinematical and
-   dynamical differential equations can be solved for the time derivatives of
-   the generalized coordinates and generalized speeds, ..." as well as most of
-   the content of the paragraph. My reservations raised in the previous review
-   were not clarified.
-
-We do not use 'generalized coordinates' or 'generalized speeds' to imply that
-either set is minimal. In other words 'generalized' only implies that the
-coordinates/speeds are for the entire system and are not Cartesian or
-spherical coordinates necessarily.  The number of Kane's dynamical equations in our
-formulation is o - m, where o is the number of generalized speeds and m is the
-number of velocity (nonholonomic and time differentiated holonomic)
-constraints. These o - m dynamical equations will in general involve 1) all n
-coordinates and their time derivatives and 2) all o speeds and their time
-derivatives. So despite the fact that the Newton-Euler equations have been
-projected in the direction of the nonholonomic partial velocities (what you
-refer to as 'the direction of the generalized speeds'), there are still more
-unknowns (o generalized speed time derivatives) than there are equations (o -
-m).  This issue can be addressed by 1) algebraically eliminating dependent
-speed time derivatives to obtain o-m equations in o-m unknowns, or 2)
-augmenting the o-m dynamical equations with the m acceleration constraint
-equations to obtain o equations in o unknowns. Either way is mathematically
-equivalent but dependent upon the specific problem at hand one may be
-preferable to the other.
-
-We have revised manuscript to be more explicit about what we mean by solving
-the kinematical and dynamical differential equations for the time derivatives
-of the coordinates and time derivatives of the generalize speeds. For this
-dynamical differential equation:
-
--m*du/dt + f = 0
-
-then by "solve the dynamical differential equations for the time derivatives of
-the generalized speeds" we mean:
-
-du/dt = f / m
-
-We welcome specific suggestions about how this can be made more clear.
-
-Section 2 was introduced specifically to clarify how we use Kane's method since
-some of our notation deviates from his presentation in his 1985 book. When the
-system has velocity constraints (i.e., a non-minimal set of generalized speeds
-are introducted, or the system has time differentiated holonomic constraints),
-the velocity constraint matrix B (equation 11 in the paper) can manipulated to
-form matrix A (equation 21 in the paper). This A matrix is used to reduce the
-number of equations for generalized speeds by the number of constrained speeds.
-This is also outlined by Kane in [Kane 1985], in equations 4.4.3 and 4.4.11,
-which shows the reduction in size of the equations for the generalized speeds
-and that _all_ the speeds (dependent and independent) are still present.
-
-4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
-    the Authors. For example, the remark "Velocity constraints often appear
-    where there is rolling without slip or there are closed kinematic loops" is
-    unclear (if correct at all). Velocity constraints, or, better, constraint
-    equations at the velocity level, are either nonholonomic constraints (often
-    consequent to the rolling without slip) or differentiated with respect to
-    time holonomic constraint equations (one source of such constraints are
-    closed kinematic loops).
-
-Author Response: We have rewritten this entire paragraph to ensure that the
-description of each level of constraints is consistent, and more clear. We have
-reworked the sentence in question to more clearly state that things like
-rolling without slip lead to nonholonomic constraints and closed kinematic
-loops lead to time differentiated holonomic constraints, both of which are
-velocity level constraints.
-
-5. 5. 25-50. The velocity constraints are introduced in Eq. (8) in a general
-    (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
-    "by assumption”?), the velocity constraints are linear with respect to u,
-    ... which implies Bu + fvt(q,t) = 0". Firstly, in general, the constraint
-    matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
-    constraints were not introduced from the beginning in this "linear" form?
-
-Author Response: We have clarified that the B matrix is a function of the
-coordinates and time, even though it is written as only a single symbol (a
-format which we use throughout the paper).
-
-6. 8. 15 and 831. The meaning of the “exogenous inputs” is not well
-    motivated/explained.
-
-Author Response: We believe that the sentence "The s exogenous inputs include
-control system inputs and/or external disturbance forces or torques applied to
-the system." explains that the exogenous inputs include forces/torques which
-are not internal to the system (e.g. a reaction force), but instead applied
-from outside the system (a motor driving a body in the system).
-
-7. A general remark to Section 2. While Kane’s method is generally well
-    understood, the Authors presentation of its formulation is rather
-    "frustrating".
-
-8. 10. 32-34. One of the velocity constraint equations introduced in Eq. (30)
-    should be a time differentiated form of the configuration constraint of Eq.
-    (25). It seems that it is Eq. (30c).  If so, it should rather read
-    r u2 s2 + u6  = 0 (not r u2 s3 + u6 = 0). Can you explain this?
-
-Author Response: While it could be written this way, what we have presented is
-correct. The method you are suggesting requires ^N v ^P* to be expressed in N.
-Then, the N_x and N_y components would be 2 of the motion constraints, and the
-time differentiated configuration constraint will equal the N_z component. This
-can be simply verified by time differentiating the configuration constraint and
-utilizing equation 34 to eliminate the q_dots, so only u's are present. In
-comparing this to the projection of ^N v ^P* in the N_z direction, you will
-find these are equivalent. We believe what we have presented is simpler.
-
-9. 10. 36-38. It is written: “These equations are linear in the ui terms,
-    nonlinear in the qi terms, involve geometric system parameters (but not
-    mass or inertial parameters), and do not explicitly involve \dot{qi}
-    terms”. The remark that constraint equations do not involve inertial
-    parameters is a cliché. According to Eq. (11), \dot{qi} terms are neither
-    involved.
-
-Author Response: Although it might be a cliché, we feel explicitness is
-valuable, and we do not wish to remove a sentence which might help a reader
-avoid confusion.
-
-10. 10. 39-118. The paragraph is unintelligible. The “diabolical set of
-    coordinates and speeds” is just a set of dependent coordinates and speeds,
-    resulted in one configuration constraint of Eq. (25) and three velocity
-    constraints of Eq. (30) (one holonomic constraint of Eq.  (25) and two
-    nonholonomic constraints of Eq. (30a) and (30b)). There are then five
-    independent coordinates and three independent speeds. By choosing the
-    independent coordinates and speeds, which may in general not be subsets of
-    qi and ui, respectively, the constraint equations (25) and (30) are
-    satisfied by identity.
-
-Author Response: We have changed "diabolical" to "purposefully nonminimal".
-That should clarfy that the example exists to show the technique on a
-tractable system with a non-minimal set of coordinates and speeds, and to
-highlight the approach, rather than suggest that we should be using these
-coordinates and speeds on the rolling disk example.
-
-11. 11. 14. Is Eq. (31c) correct? (see Remark 8).
-
-Author Response: Yes it is correct, our response to comment 8 addresses this.
-The code in the electronic supplementary material can be used to verify this.
-
-12. A general remark. The complexity of the followed parts of the manuscript is
-    not appealing to me. In fact, I was unable to follow the derivations. I may
-    suppose that the same difficulties will have most of the potential readers
-    of the paper, not mentioning those who would like to follow the procedures.

### 2014-08-29 response_to_reviewers_aug_2014.txt
Separate responses to review and editor

@@
-Response to Editor.
-
-Dear Editor(s),
-We feel that in this revision, as in the previous revisions, we have made an
-effort to address as many of Reviewer #1's comments as possible. We have valued
-a number of the comments provided by Reviewer #1, agreeing that the suggested
-changes help to clarify the motivation and methods in the paper.
-
-Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
-reasonable to address, comments 7 and 12.
-
-Comment 7 simply states that the reviewer finds our presentation frustrating,
-without offering specific faults, mistakes, unclear sections, areas that could
-use improvement, etc.  Additionally, some of the other comments (comment 3) are
-the reason we have included such a thorough description of Kane's method. There
-seems to be a breakdown where the reviewer is not understanding how we are
-applying this method, and it is not clear on our end which part of Section 2
-needs to be improved to address the reviewer's difficulties.
-
-Comment 12 states that the reviewer is confused by the parts of the manuscript
-that they understood, and then admits to being unable to follow the derivations
-we have presented. While we cannot simplify the math in the derivations, one
-can always attempt to provide more clarification, but only if the reviewer
-states which areas they had trouble with or where in the derivation they became
-lost. This comment as given to us, does not allow us to do anything.
-
-We would also like to add that many of Reviewer #1's comments seem to indicate
-a shortage of effort to understand the paper. As an example, one of the authors
-has advised a masters student in a summer project to implement this method in a
-symbolic computer algebra system and within 2 months the student had
-implemented the entirety of the algorithm, complete with examples, software
-tests to verify mathematical correctness, and documentation. This student was
-able to understand the mathematics presented in the paper with very little
-guidance; fundamentally there is nothing in the derivation that is more
-complicated than a partial derivative and a few matrix multiplications.
-
-In summary, we have addressed as many of the comments as possible, but feel
-that the other comments not addressable, and suggest that Reviewer #1 has
-not made enough of an effort to either comprehend the paper, or communicate
-where they had difficulty comprehending it. As Reviewer #2 seems to
-approve of the manuscript, we would ask that you overlook Reviewer #1's
-concerns when making the decision to publish or reject this work.
-
-Thank you,
-Dale Peterson
-Gilbert Gede
-Mont Hubbard
-
 Response to Reviewer #1.

### 2014-09-29 PetersonGedeHubbard2012.tex
Remove extra '}' in math mode

@@
 motion is as simple as arranging them in first order form ($\dot{x} = f(x, t)$)
-and evaluating $\frac{\partial f}{\partial x}}$ at the point of interest.
+and evaluating $\frac{\partial f}{\partial x}$ at the point of interest.
 However, when the system has constraints not all the states (coordinates and
